### 2013-09-19 article/article.tex
Initial setup: tex templates for presentation and article

\documentclass[a4paper,10pt, twocolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[usenames,dvipsnames]{color}
\usepackage{comment}
\usepackage{tikz}
\usepackage{verbatim}
\usetikzlibrary{arrows,shapes}
\usepackage{listings}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{tabularx}
\usepackage{enumitem}
\setlist{nolistsep}
\setlength{\parindent}{0cm}

%page boarders
\usepackage[top=3cm, bottom=3cm, left=2cm, right=2cm]{geometry}
\usepackage[style=numeric,backend=bibtex]{biblatex}
\addbibresource{../sources}

\newtheorem{lemma}{Lemma}
\newtheorem{corollary}[lemma]{Corollary}
\newtheorem{theorem}[lemma]{Theorem}

\title{The Power of Two Random Choices \\ 
\large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Träff and Francesco Versaci, lecture "Parallel Algorithms", 2012 winter term at TU Wien.}}
\author{Martin Kalany, 0825673}

\begin{document}
\maketitle

\section{Abstract}
\label{sec:abstract}
\cite{mitzen01}

\section{Introduction}
\label{sec:intro}

%\pagebreak
\printbibliography
\end{document}

### 2013-09-19 presentation/presentation.tex
Initial setup: tex templates for presentation and article

\documentclass{beamer}
\let\Tiny=\tiny %to avoid warnings related to font size and beamer 
\usetheme{Amsterdam}
\usecolortheme{dolphin}
\usepackage{color}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{verbatim}
\usetikzlibrary{arrows,shapes}
\usepackage{hyperref}
\usepackage[style=numeric-comp,backend=bibtex]{biblatex}
\usepackage{comment}
\usepackage{listings}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{tabularx}
\usepackage{color, colortbl}
\addbibresource{../sources}

\let\oldfootnotesize\footnotesize
\renewcommand*{\footnotesize}{\oldfootnotesize\tiny}

\title{The Power of Two Random Choices}
\subtitle{A follow up seminar to the Parallel Algorithms lecture}
\author{Martin Kalany}
\institute
{

  Graduate student in Computer Science\\
  Vienna University of Technology\\
}
\date{\today}

\begin{document}

\frame{\titlepage}

\begin{frame}
\cite{mitzen01}
\end{frame}

\begin{frame}[allowframebreaks]
\frametitle<presentation>{Literature}    
\printbibliography
\end{frame} 	 
\end{document}

### 2013-11-07 article/article.tex
use acm template for article; fix footnote issue with acm template and missing bracket

@@
-\documentclass[a4paper,10pt, twocolumn]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage[usenames,dvipsnames]{color}
-\usepackage{comment}
-\usepackage{tikz}
-\usepackage{verbatim}
-\usetikzlibrary{arrows,shapes}
-\usepackage{listings}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{hyperref}
-\usepackage{graphicx}
-\usepackage{epstopdf}
-\usepackage{tabularx}
-\usepackage{enumitem}
-\setlist{nolistsep}
-\setlength{\parindent}{0cm}
+\documentclass{acm_proc_article-sp}
 
-%page boarders
-\usepackage[top=3cm, bottom=3cm, left=2cm, right=2cm]{geometry}
-\usepackage[style=numeric,backend=bibtex]{biblatex}
-\addbibresource{../sources}
-
-\newtheorem{lemma}{Lemma}
-\newtheorem{corollary}[lemma]{Corollary}
-\newtheorem{theorem}[lemma]{Theorem}
-
-\title{The Power of Two Random Choices \\ 
-\large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Träff and Francesco Versaci, lecture "Parallel Algorithms", 2012 winter term at TU Wien.}}
-\author{Martin Kalany, 0825673}
+%\newtheorem{lemma}{Lemma}
+%\newtheorem{corollary}[lemma]{Corollary}
+%\newtheorem{theorem}[lemma]{Theorem}
 
 \begin{document}
+
+\title{The Power of Two Random Choices}
+\subtitle{\large A follow up seminar to the lecture "Parallel Algorithms"\\ by Jesper Larsson Tr\"aff and Francesco Versaci, 2012 winter term at TU Wien.}
+
+\numberofauthors{1}
+\author{
+\alignauthor
+Martin Kalany\\
+       0825673
+}    
+\date{\today}
+
 \maketitle
+\begin{abstract}
 
-\section{Abstract}
-\label{sec:abstract}
-\cite{mitzen01}
+\end{abstract}
 
 \section{Introduction}
+\cite{mitzen01}
 \label{sec:intro}
 
-%\pagebreak
-\printbibliography
+%bibliography
+\bibliographystyle{abbrv}
+\bibliography{../sources} 
 \end{document}

### 2013-12-28 article/article.tex
introduction

@@
 \documentclass{acm_proc_article-sp}
+
+\usepackage{comment}
+\usepackage{hyperref}
+\usepackage{amssymb}
+
+\newcommand*{\PH}{\makebox[1ex]{\textbf{$\cdot$}}}
 
@@
 \section{Introduction}
-\cite{mitzen01}
 \label{sec:intro}
+
+\subsection{Notation}
+\label{sec:notation}
+Throughout this paper the number of balls is denoted by $m$ and the number of bins by $n$.
+An event $\mathcal E$ occurs with high probability if $\Pr\left[\mathcal E \right] = 1 - o(1)$.
+Asymptotic notations ($O(\PH)$, $o(\PH)$ and $\omega(\PH)$) are always with respect to $n$.
+Let $f$, $g$ be arbitrary functions.
+Then, $f \ll g$ means $f = o(g)$ and $ f \gg g$ means $f = \omega(g)$.
+
+\subsection{Balls into Bins}
+\label{sec:ballsIntoBins}
+Suppose that $m$ balls are sequentially thrown into $n$ bins, where each ball is placed in an independently and uniformly chosen random bin.
+We are interested in the \emph{number of balls in the fullest bin}, or equivalently the \emph{gap} of the allocation, that is the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
+
+These \emph{balls into bins games} or \emph{allocation problems} have been studied extensively in the probability literature (see e.g., \cite{JK77}).
+For the case of $m = n$, the fullest bin contains $\Theta \left( \frac{\log n}{\log \log n} \right)$ \cite{RS98} with high probability.
+Since the average load of the bins is 1, this is also a bound for the gap.
+%More accurately, the number of balls in the bin with the most balls is $\Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$ \cite{G91}.
+%TODO: really include Gonnet?
+For the case $m \gg n$ the number of balls in the fullest bin is $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ with high probability.
+
+Consider a variation of the above \emph{balls-into-bins} problem, which we will call \emph{d-choice balls into bins}: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}\footnote{Assuming a fixed ordering of the bins, a ball will be put into the left-most bin in case of a tie}.
+In this variant, the number of balls in the fullest bin is $\frac{\log \log n}{log d} \left( 1 + O\left(1\right)\right)$ for all $m$ with high probability.
+
+\subsection{Applications}
+\label{sec:applications}
 

### 2014-01-04 article/article.tex
outline for next paragraphs

@@
 Throughout this paper the number of balls is denoted by $m$ and the number of bins by $n$.
-An event $\mathcal E$ occurs with high probability if $\Pr\left[\mathcal E \right] = 1 - o(1)$.
+We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] = 1 - o(1)$, except when stated otherwise.
 Asymptotic notations ($O(\PH)$, $o(\PH)$ and $\omega(\PH)$) are always with respect to $n$.
@@
 Suppose that $m$ balls are sequentially thrown into $n$ bins, where each ball is placed in an independently and uniformly chosen random bin.
-We are interested in the \emph{number of balls in the fullest bin}, or equivalently the \emph{gap} of the allocation, that is the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
+We are interested in the \emph{number of balls in the fullest bin}, also called the \emph{maximum load}.
+Equivalently, one can state the \emph{gap} of the allocation, that is the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
 
 These \emph{balls into bins games} or \emph{allocation problems} have been studied extensively in the probability literature (see e.g., \cite{JK77}).
-For the case of $m = n$, the fullest bin contains $\Theta \left( \frac{\log n}{\log \log n} \right)$ \cite{RS98} with high probability.
+When using the strategy as outlined above (that is, choose a bin independently and uniformly at random for each ball), the fullest bin contains $\Theta \left( \frac{\log n}{\log \log n} \right)$ for the case of $m = n$ with high probability \cite{RS98}.
 Since the average load of the bins is 1, this is also a bound for the gap.
@@
 %TODO: really include Gonnet?
-For the case $m \gg n$ the number of balls in the fullest bin is $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ with high probability.
+For the \emph{heavily loaded case}, where $m \gg n \log n$ the maximum load is $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ with high probability \cite{RS98}.
+In the remainder of this paper, "heavily loaded case" will denote the case where $m = \omega\left(n\right)$ and the discussed strategy will be called \emph{once-choice balls into bins}.
 
-Consider a variation of the above \emph{balls-into-bins} problem, which we will call \emph{d-choice balls into bins}: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}\footnote{Assuming a fixed ordering of the bins, a ball will be put into the left-most bin in case of a tie}.
-In this variant, the number of balls in the fullest bin is $\frac{\log \log n}{log d} \left( 1 + O\left(1\right)\right)$ for all $m$ with high probability.
+Consider a variation of the above \emph{once-choice balls into bins} strategy, which we will call \emph{multiple-choice balls into bins}: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}\footnote{If a tie occurs, chose one of the least full bins arbitrarily.}.
+Note that the $d$ selected bins are not necessarily distinct.
+In this variant, the maximum load is $\frac{m}{n} + \frac{\log \log n}{log d} \left( 1 + O\left(1\right)\right)$ for all $m$ with high probability.
+This was first proven by Azar et al.
+\cite{ABKU99} for the case $m = n$.
+(However, the case $d = 2$ was implicitly proven by Karp et al.
+\cite{KLM92}).
+See also \cite{RMS01} for a simplified proof and an overview of proof techniques for the given problem.
+Unfortunately, the analysis breaks down for the heavily loaded case, where $m = \omega\left(n\right)$.
+Berenbrink at.
+all \cite{BFZR08} were the first to proof that the same bound holds for the heavily loaded case\footnote{However, they say that an event $\mathcal E$ occurs with high probability if $\Pr\left[\mathcal E \right]  \geq 1- n^{-\alpha}$, for an arbitrarily chosen constant $\alpha \geq 1$.}.
+A simplified proof may be found in \cite{TW13}.
+
+The result stated above implies a gap of $\Theta\left(\log \log n \right)$ for arbitrarily large $m$. This is remarkable in several ways:
+\begin{itemize}
+\item The apparently small change made for the \emph{multiple-choice balls into bins} strategy results in an exponential decrease of the gap between the fullest box and the average load, even for $d=2$.
+\item Each additional choice ($d > 2$) decreases the gap by only a constant factor \cite{RMS01}. 
+\item For the \emph{heavily loaded case}, the resulting gap of the \emph{multiple-choice balls into bins} strategy does not depend on the number of balls $m$.
+In contrast, the bound for the \emph{one-choice} diverges with $m$.
+\item The given bounds are \emph{tight}, meaning that no other strategy that places each ball into one of $d$ randomly selected bins achieves a gap that is asymptotically lower.
+\end{itemize}
 
@@
 \label{sec:applications}
+\begin{enumerate}
+\item Hashing
+\item On-line load balancing
+\item Dynamic resource allocation
+\item Shared memory emulation on distributed memory machines
+\end{enumerate}
 

### 2014-01-07 article/article.tex
some additions to introdcution

@@
 \label{sec:ballsIntoBins}
-Suppose that $m$ balls are sequentially thrown into $n$ bins, where each ball is placed in an independently and uniformly chosen random bin.
+Suppose that $m$ balls are sequentially thrown into $n$ initially empty bins, where each ball is placed in an independently and uniformly chosen random bin.
 We are interested in the \emph{number of balls in the fullest bin}, also called the \emph{maximum load}.
-Equivalently, one can state the \emph{gap} of the allocation, that is the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
+Equivalently, one can look at the \emph{gap} of the allocation, that is the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
 
 These \emph{balls into bins games} or \emph{allocation problems} have been studied extensively in the probability literature (see e.g., \cite{JK77}).
-When using the strategy as outlined above (that is, choose a bin independently and uniformly at random for each ball), the fullest bin contains $\Theta \left( \frac{\log n}{\log \log n} \right)$ for the case of $m = n$ with high probability \cite{RS98}.
+Two basic variants of this problem exists (\cite{ABKU99}, \cite{RMS01}):
+\begin{itemize}
+\item In a \emph{static problem}, the balls are placed into the bins and no balls are ever removed.
+\item In a \emph{dynamic problem}, balls can also be removed from the system.
+In a simple dynamic model, for each newly inserted ball a ball choosen uniformly and independently at random from among the balls currently in the bins is removed from the system.
+\end{itemize}
+
+When using the \emph{one-choice balls into bins} strategy as outlined above (that is, choose a bin independently and uniformly at random for each ball), the fullest bin contains $\Theta \left( \frac{\log n}{\log \log n} \right)$ for the case of $m = n$ w.h.p.
+\cite{RS98}.
 Since the average load of the bins is 1, this is also a bound for the gap.
@@
 For the \emph{heavily loaded case}, where $m \gg n \log n$ the maximum load is $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ with high probability \cite{RS98}.
-In the remainder of this paper, "heavily loaded case" will denote the case where $m = \omega\left(n\right)$ and the discussed strategy will be called \emph{once-choice balls into bins}.
+In the remainder of this paper, "heavily loaded case" will denote the case where $m = \omega\left(n\right)$.
 
@@
 Note that the $d$ selected bins are not necessarily distinct.
+
+A \emph{uniform balls into bins problem} is a \emph{multiple choice balls and bins problem} where the $d$ random choices for one ball are independent and uniform.
+\emph{Non-uniform balls into bins problems} can reduce the maximum load and where first studied by Vöcking \cite{}.
+
 In this variant, the maximum load is $\frac{m}{n} + \frac{\log \log n}{log d} \left( 1 + O\left(1\right)\right)$ for all $m$ with high probability.

### 2014-01-10 article/article.tex
re-work introduction; split into different sections; outline for multiple-choice schemes

--- 2014-01-04 article/article.tex
@@
 \title{The Power of Two Random Choices}
-\subtitle{\large A follow up seminar to the lecture "Parallel Algorithms"\\ by Jesper Larsson Tr\"aff and Francesco Versaci, 2012 winter term at TU Wien.}
+\subtitle{\large A follow up seminar to the lecture ``Parallel Algorithms''\\ by Jesper Larsson Tr\"aff and Francesco Versaci, 2012 winter term at TU Wien.}
 
@@
 \label{sec:intro}
+Suppose that $m$ balls are thrown into $n$ bins, according to some strategy.
+We are interested in the \emph{number of balls in the fullest bin}, also called the \emph{maximum load}.
+Equivalently, one can look at the \emph{addative gap} of the allocation, that is the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
+These \emph{balls into bins games} or \emph{allocation problems} have been studied extensively in the probability literature (see e.g., \cite{JK77}).
+In the context of Computer Science, the balls into bins model is useful to study e.g., strategies for distributing \emph{tasks} (balls) to \emph{processors} (bins) as evenly as possible without having to query all processors for the number of tasks currently assigned to them.
 
-\subsection{Notation}
-\label{sec:notation}
-Throughout this paper the number of balls is denoted by $m$ and the number of bins by $n$.
-We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] = 1 - o(1)$, except when stated otherwise.
-Asymptotic notations ($O(\PH)$, $o(\PH)$ and $\omega(\PH)$) are always with respect to $n$.
-Let $f$, $g$ be arbitrary functions.
-Then, $f \ll g$ means $f = o(g)$ and $ f \gg g$ means $f = \omega(g)$.
+We will first introduce different variations of the \emph{balls into bins problem} in Section \ref{sec:classification} to provide some overview of different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}.
+Sections \ref{sec:strategies} to TODO will deal with the most basic variation of the balls into bins problem, namely the sequential and static variant.
+Thus, the balls are assumed to arrive sequentially and no strategy for deleting balls in the bins has to be considered.
+TODO: finite or infinite process?
 
-\subsection{Balls into Bins}
-\label{sec:ballsIntoBins}
-Suppose that $m$ balls are sequentially thrown into $n$ bins, where each ball is placed in an independently and uniformly chosen random bin.
-We are interested in the \emph{number of balls in the fullest bin}, also called the \emph{maximum load}.
-Equivalently, one can state the \emph{gap} of the allocation, that is the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
+A simple strategy for the balls into bins problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice balls into bins} strategy.
+In Section \ref{sec:strategies}, we will provide asymptotic bounds for the addative gap of this strategy, which we will then use to devise more complex strategies that significantly reduce these bounds.
+These result may sometimes seem surprising, which is why we will denote a significant part of this article to demonstrate a proof technique that may be useful for devising asymptotic bounds for the additive gap of different balls into bins scenarios and strategies.
+TODO
 
-These \emph{balls into bins games} or \emph{allocation problems} have been studied extensively in the probability literature (see e.g., \cite{JK77}).
-When using the strategy as outlined above (that is, choose a bin independently and uniformly at random for each ball), the fullest bin contains $\Theta \left( \frac{\log n}{\log \log n} \right)$ for the case of $m = n$ with high probability \cite{RS98}.
+\section{Classification}
+\label{sec:classification}
+Several different variations of this problem exist, which may be classified by the following properties:
+\begin{itemize}
+\item In a \emph{sequential process}, the balls arrive one after another.
+A  ball is dealt with only after the previous ball has been placed.
+Contrary, in a \emph{parallel process}, balls arrive in sets and the balls in one set are dealt with simultaneously \cite{ABS98}.
+\item In a \emph{static problem}, the balls are placed into the bins and no balls are ever removed.
+In a \emph{dynamic problem}, balls can also be removed from the system.
+In a simple \emph{random deletions} model, for each newly inserted ball a ball choosen uniformly and independently at random from among the balls currently in the bins is removed from the system \cite{ABKU99} \cite{RMS01}.
+More elaborate shemes exists: An \emph{oblivious adversery} specifies the sequence of insertions and deletions of balls in advance, without any knowledge of the placement of the balls \cite{CFM+98}.
+Another approach was studied by Adler et.
+al \cite{ABS98} in the context of the parallel balls into bins problem: In each bin, the balls are stored in FIFO-order and at every time step the first ball in each of the bins is deleted.
+\item A \emph{finite process} is analyzed for a given time interval that is known beforehand, whereas for an \emph{infinite process}, a placement and a deletion process is defined over an infinite time span \cite{ABS98}.
+\end{itemize}
+
+\section{Applications}
+\label{sec:applications}
+\subsection{Hashing}
+\subsection{On-line load balancing}
+\subsection{Dynamic resource allocation}
+\subsection{Shared memory emulation on distributed memory machines}
+
+\section{Strategies}
+\label{sec:strategies}
+We will now consider the most basic variant of the balls into bins problem: Namely the static and sequential variant.
+Thus, the balls are assumed to arrived one after the other and we do not delete any balls from the bins.
+Additionally, all bins are empty at the start of the placement process.
+
+We will distinguish two cases with respect to the number of balls $m$.
+In the \emph{lightly loaded case}, $m$ is in the order of $n$, whereas in the \emph{heavily loaded case}, $m$ is aribtrarily large.
+For clarity, we will assume that $m = n$ in the slightly loaded case and that $m \ll n$\footnote{Let $f$, $g$ be arbitrary functions.
+Then, $f \ll g$ means $f = o(g)$ and $ f \gg g$ means $f = \omega(g)$} in the heavily loaded case.
+
+\subsection{Single-choice balls into bins}
+\label{sec:single-choice}
+In the \emph{single-choice balls into bins} strategy, each ball is placed into a bin choosen uniformly and independently at random.
+
+The fullest bin contains $\Theta \left( \frac{\log n}{\log \log n} \right)$ for the lightly loaded case with high probability\footnote{We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] = 1 - o(1)$, except when stated otherwise.} \cite{RS98}.
 Since the average load of the bins is 1, this is also a bound for the gap.
@@
 %TODO: really include Gonnet?
-For the \emph{heavily loaded case}, where $m \gg n \log n$ the maximum load is $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ with high probability \cite{RS98}.
-In the remainder of this paper, "heavily loaded case" will denote the case where $m = \omega\left(n\right)$ and the discussed strategy will be called \emph{once-choice balls into bins}.
+For the \emph{heavily loaded case} the maximum load is $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ w.h.p.
 
-Consider a variation of the above \emph{once-choice balls into bins} strategy, which we will call \emph{multiple-choice balls into bins}: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}\footnote{If a tie occurs, chose one of the least full bins arbitrarily.}.
+TODO: proof? Maybe just outline the $m = n$ case as in \cite{RS98}?
+
+\subsection{Multiple-choice balls into bins}
+\label{sec:multiple-choice}
+The \emph{single-choice balls into bins strategy} (Section \ref{sec:single-choice}) chooses one bin for each ball.
+A \emph{multiple-choice balls into bins strategy} generalizes this idea by choosing $d$ bins and placing the ball into the least full bin.
+We will discuss two different multiple-choice schemes: The \emph{Greedy$[d]$} scheme (Section \ref{sec:greedy}) places the ball into the least full bin among $d$ bins chosen uniformly and independently at random whereas the \emph{Always-Go-Left} scheme (Section \ref{sec:AlwaysGoLeft}) uses a nonuniform selection of $d$ bins.
+
+\subsubsection{The Greedy$[d]$ scheme}
+\label{sec:greedy}
+
+Consider a variation of the above \emph{one-choice balls into bins} strategy, which we will call the \emph{Greedy$[d]$ scheme}: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}.
+If there are multiple bins with the same minimal load, chose one among them arbitrarily.
 Note that the $d$ selected bins are not necessarily distinct.
-In this variant, the maximum load is $\frac{m}{n} + \frac{\log \log n}{log d} \left( 1 + O\left(1\right)\right)$ for all $m$ with high probability.
+
+The \emph{Greedy$[d]$ scheme} achieves a maximum load of $\frac{m}{n} + \frac{\log \log n}{\log d} \left( 1 + O\left(1\right)\right)$ for all $m$ w.h.p. (see \cite{ABKU99}, \cite{BCSV06}).
+
 This was first proven by Azar et al.
@@
 
-\subsection{Applications}
-\label{sec:applications}
-\begin{enumerate}
-\item Hashing
-\item On-line load balancing
-\item Dynamic resource allocation
-\item Shared memory emulation on distributed memory machines
-\end{enumerate}
+\subsubsection{The Always-Go-Left scheme}
+\label{sec:AlwaysGoLeft}
+The multiple-choice balls into bins strategy presented in Section \ref{sec:multiple-choice} assumes that $d$ bins are chosen uniformly and independently at random.
+V\"ocking \cite{VOC03} was the first to study the impact of non-uniform and dependent selections of the $d$ bins.
+Surprisingly, a multiple-choice balls into bins strategy with a \emph{nonuniform and dependent} selection of the $d$ bins, such as the \emph{Always-Go-Left} algorithm introduced in \cite{VOC03}, achieves better\cite{BCSV06} load balancing than the \emph{Greedy} scheme (Section \ref{sec:multiple-choice}).
+
+TODO: algorithm
+TODO: analysis
 

### 2014-01-13 article/article.tex
move paragraph about heavily loaded case to introduction
applications: hashing and PRAM emulation on DMM

@@
 In a \emph{dynamic problem}, balls can also be removed from the system.
-In a simple \emph{random deletions} model, for each newly inserted ball a ball choosen uniformly and independently at random from among the balls currently in the bins is removed from the system \cite{ABKU99} \cite{RMS01}.
+In a simple \emph{random deletions} model, for each newly inserted ball a ball choosen uniformly and independently at random from among the balls currently in the bins is removed from the system \cite{ABKU99} \cite{MRS01}.
 More elaborate shemes exists: An \emph{oblivious adversery} specifies the sequence of insertions and deletions of balls in advance, without any knowledge of the placement of the balls \cite{CFM+98}.
@@
 
+We will distinguish two cases with respect to the number of balls $m$.
+In the \emph{lightly loaded case}, $m$ is in the order of $n$, whereas in the \emph{heavily loaded case}, $m$ is aribtrarily large.
+For clarity, we will assume that $m = n$ in the slightly loaded case and that $m \ll n$\footnote{Let $f$, $g$ be arbitrary functions.
+Then, $f \ll g$ means $f = o(g)$ and $ f \gg g$ means $f = \omega(g)$} in the heavily loaded case.
+
 \section{Applications}
@@
 \subsection{Hashing}
+A simple hash table implementation typically uses a single hash function to map elements to entries in a table.
+If more than one element is mapped to the same entry, all elements mapped to this table entry are stored in a \emph{chain}, usually implemented as a linked list.
+The maximum time to search for an element is then proportional to the length of the longest chain in the hash table.
+
+It is easy to see how the balls-into-bins idea can be applied: Assuming a perfect hash function\footnote{i.e., each of the $n$ elements is mapped to one of the $n$ table entries uniformly and independently at random}, the elements are balls to be mapped to table entries or bins, implying a worst case look-up time of $\Theta\left(\frac{\log n}{\log \log n}\right)$ w.h.p.
+Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries implies a bound of $\Theta\left(\log \log n\right)$ w.h.p.
+This approach is easy to parallelize, it does not require re-hashing of data and requires only two hash functions \cite{MRS01} \cite{ABKU99}.
+
 \subsection{On-line load balancing}
+
 \subsection{Dynamic resource allocation}
+
 \subsection{Shared memory emulation on distributed memory machines}
+The \emph{Parallel Random Access Machine} model\footnote{as defined in the lecture or in~\cite{P03}}, or \emph{PRAM} for short, is a high level abstraction of a shared memory machine, in which memory access by any processor to any memory location is assumed to require constant time.
+It is especially useful for the study of parallel algorithms.
+However, a PRAM is much harder to realize in hardware than a \emph{Distributed Memory Machine} (DMM).
+Thus, ways to emulate a PRAM on a DMM have been studied extensivly.
+Such an emulation distributes the processors and memory cells of the PRAM to the processors and memory modules (typically one per processor) of the DMM with the goal to reduce the slowdown of the emulation, which is the time the DMM requires to simulate one step of the PRAM.
+Each memory modul can handle only one memory access at a time step.
+More accesses to the same module will result in memory contention.
+The balls-into-bins idea is useful for managing communication between the processors and the memory modules such that memory contention is minimal.
+For details, see Karp et.
+al \cite{KLM92} and the papers referenced there.
 
@@
 Additionally, all bins are empty at the start of the placement process.
-
-We will distinguish two cases with respect to the number of balls $m$.
-In the \emph{lightly loaded case}, $m$ is in the order of $n$, whereas in the \emph{heavily loaded case}, $m$ is aribtrarily large.
-For clarity, we will assume that $m = n$ in the slightly loaded case and that $m \ll n$\footnote{Let $f$, $g$ be arbitrary functions.
-Then, $f \ll g$ means $f = o(g)$ and $ f \gg g$ means $f = \omega(g)$} in the heavily loaded case.
 
@@
 \cite{KLM92}).
-See also \cite{RMS01} for a simplified proof and an overview of proof techniques for the given problem.
+See also \cite{MRS01} for a simplified proof and an overview of proof techniques for the given problem.
 Unfortunately, the analysis breaks down for the heavily loaded case, where $m = \omega\left(n\right)$.
@@
 \item The apparently small change made for the \emph{multiple-choice balls into bins} strategy results in an exponential decrease of the gap between the fullest box and the average load, even for $d=2$.
-\item Each additional choice ($d > 2$) decreases the gap by only a constant factor \cite{RMS01}. 
+\item Each additional choice ($d > 2$) decreases the gap by only a constant factor \cite{MRS01}. 
 \item For the \emph{heavily loaded case}, the resulting gap of the \emph{multiple-choice balls into bins} strategy does not depend on the number of balls $m$.

### 2014-01-13 article/article.tex
application: online load balancing

@@
 
-\subsection{On-line load balancing}
+\subsection{Online load balancing}
+The following simple scenario is a typical application of online load balancing: Consider $n$ servers\footnote{Such as file-, database- or network servers} and $m$ requests issued by clients that need to be handled by these servers.
+For simplicity, assume that each task can be handled by any server and that all tasks are of the same size, i.e., require the same amount of execution time.
+
+To maximize the throughput of the servers, a uniform distribution of tasks to servers is desireable.
+This can easily be achieved by a central dispatcher.
+However, such a central dispatcher poses a bottleneck within a distributed system and querying $m$ servers for their current load involves significant overhead, e.g., sending a message to each server and waiting for the reply and more efficient solutions are neccessary.
+As with hashing, the balls-into-bins model can be applied easily: we need to assign tasks (the balls) to servers (the bins).
+Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends issues its request to the least loaded server, the overhead is comperatively small and the difference in load on the $n$ servers is bounded by $\Theta\left(\log \log n\right)$ w.h.p.
+\cite{RS98} \cite{MRS01} \cite{KLM92}.
 

### 2014-01-13 article/article.tex
fix: citations in ascending order
Application: Low congestions circuit routing

@@
 Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries implies a bound of $\Theta\left(\log \log n\right)$ w.h.p.
-This approach is easy to parallelize, it does not require re-hashing of data and requires only two hash functions \cite{MRS01} \cite{ABKU99}.
+This approach is easy to parallelize, it does not require re-hashing of data and requires only two hash functions \cite{ABKU99} \cite{MRS01}.
 
@@
 Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends issues its request to the least loaded server, the overhead is comperatively small and the difference in load on the $n$ servers is bounded by $\Theta\left(\log \log n\right)$ w.h.p.
-\cite{RS98} \cite{MRS01} \cite{KLM92}.
-
-\subsection{Dynamic resource allocation}
+\cite{KLM92} \cite{RS98} \cite{MRS01}.
 
@@
 al \cite{KLM92} and the papers referenced there.
+
+\subsection{Low congestion circuit routing}
+In multi-stage interconnection networks, virtual circuit-switching is used to route communication.
+For simplicity, we consider the \emph{permutation routing problem}, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes and assume a \emph{butterfly network} as the underlying interconnection network (see \cite{CLR09} for a detailed definition).
+Then, congestions is the maximum number of paths that are routed through a network link.
+The goal of a \emph{circuit routing algorithm} is to to allocate paths for all $n$ communication requests such that congestion is minimal.
+
+In this context, \emph{Valiant's paradigm} \cite{V82} states that any such permutation routing problem can be solved by transforming it into two problems: First, find a path for each request to an intermediate destination chosen uniformly and independently at random.
+Next, find a path from the intermediate to the actual destination.
+This routing technique is analogous to the single-choice balls-into-bins problem where each communication request (or ball) chooses an intermediate destination (a bin) at random.
+In this analogy, the congestion achieved by Valiant's paradigm corresponds to the maximum load in the balls-into-bins model, thus implying a bound of $\Theta\left(\log \log n\right)$ on the congestion w.h.p.
+\cite{CMM+98} \cite{MRS01}.
+
+Applying the two-choice paradigm in this context is somewhat more involved and beyond the scope of this paper. The interested reader may find more details in \cite{CMM+98}.
 

### 2014-01-13 article/article.tex
introduction for applications

@@
 \label{sec:applications}
+The two-choice paradigm applied to the balls-into-bins model has several useful applications, some of which we will outline here.
+The applications in \emph{Hashing} (Section \ref{sec:hashing}), \emph{Online Load Balancing} (Section \ref{sec:loadbalancing}) and \emph{Emulation of shared memory machines on distributed memory machines} \ref{sec:DMM}) share a close relation to load balancing.
+Section \ref{sec:circuitrouting} applies the balls-into-bins model and the two-choice paradigm in a different context, namely that of routing circuits in multi-stage interconnection networks with minimal congestion.
+
 \subsection{Hashing}
+\label{sec:hashing}
 A simple hash table implementation typically uses a single hash function to map elements to entries in a table.
@@
 \subsection{Online load balancing}
+\label{sec:loadbalancing}
 The following simple scenario is a typical application of online load balancing: Consider $n$ servers\footnote{Such as file-, database- or network servers} and $m$ requests issued by clients that need to be handled by these servers.
@@
 
-\subsection{Shared memory emulation on distributed memory machines}
+\subsection{Emulation of shared memory machines on distributed memory machines}
+\label{sec:DMM}
 The \emph{Parallel Random Access Machine} model\footnote{as defined in the lecture or in~\cite{P03}}, or \emph{PRAM} for short, is a high level abstraction of a shared memory machine, in which memory access by any processor to any memory location is assumed to require constant time.
@@
 \subsection{Low congestion circuit routing}
+\label{sec:circuitrouting}
 In multi-stage interconnection networks, virtual circuit-switching is used to route communication.

### 2014-01-14 article/article.tex
balls into bins -> balls-into-bins

@@
 Equivalently, one can look at the \emph{addative gap} of the allocation, that is the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
-These \emph{balls into bins games} or \emph{allocation problems} have been studied extensively in the probability literature (see e.g., \cite{JK77}).
-In the context of Computer Science, the balls into bins model is useful to study e.g., strategies for distributing \emph{tasks} (balls) to \emph{processors} (bins) as evenly as possible without having to query all processors for the number of tasks currently assigned to them.
+These \emph{balls-into-bins games} or \emph{allocation problems} have been studied extensively in the probability literature (see e.g., \cite{JK77}).
+In the context of Computer Science, the balls-into-bins model is useful to study e.g., strategies for distributing \emph{tasks} (balls) to \emph{processors} (bins) as evenly as possible without having to query all processors for the number of tasks currently assigned to them.
 
-We will first introduce different variations of the \emph{balls into bins problem} in Section \ref{sec:classification} to provide some overview of different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}.
-Sections \ref{sec:strategies} to TODO will deal with the most basic variation of the balls into bins problem, namely the sequential and static variant.
+We will first introduce different variations of the \emph{balls-into-bins problem} in Section \ref{sec:classification} to provide some overview of different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}.
+Sections \ref{sec:strategies} to TODO will deal with the most basic variation of the balls-into-bins problem, namely the sequential and static variant.
 Thus, the balls are assumed to arrive sequentially and no strategy for deleting balls in the bins has to be considered.
@@
 
-A simple strategy for the balls into bins problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice balls into bins} strategy.
+A simple strategy for the balls-into-bins problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice balls-into-bins} strategy.
 In Section \ref{sec:strategies}, we will provide asymptotic bounds for the addative gap of this strategy, which we will then use to devise more complex strategies that significantly reduce these bounds.
-These result may sometimes seem surprising, which is why we will denote a significant part of this article to demonstrate a proof technique that may be useful for devising asymptotic bounds for the additive gap of different balls into bins scenarios and strategies.
+These result may sometimes seem surprising, which is why we will denote a significant part of this article to demonstrate a proof technique that may be useful for devising asymptotic bounds for the additive gap of different balls-into-bins scenarios and strategies.
 TODO
@@
 Another approach was studied by Adler et.
-al \cite{ABS98} in the context of the parallel balls into bins problem: In each bin, the balls are stored in FIFO-order and at every time step the first ball in each of the bins is deleted.
+al \cite{ABS98} in the context of the parallel balls-into-bins problem: In each bin, the balls are stored in FIFO-order and at every time step the first ball in each of the bins is deleted.
 \item A \emph{finite process} is analyzed for a given time interval that is known beforehand, whereas for an \emph{infinite process}, a placement and a deletion process is defined over an infinite time span \cite{ABS98}.
@@
 \label{sec:strategies}
-We will now consider the most basic variant of the balls into bins problem: Namely the static and sequential variant.
+We will now consider the most basic variant of the balls-into-bins problem: Namely the static and sequential variant.
 Thus, the balls are assumed to arrived one after the other and we do not delete any balls from the bins.
@@
 
-\subsection{Single-choice balls into bins}
+\subsection{Single-choice balls-into-bins}
 \label{sec:single-choice}
-In the \emph{single-choice balls into bins} strategy, each ball is placed into a bin choosen uniformly and independently at random.
+In the \emph{single-choice balls-into-bins} strategy, each ball is placed into a bin choosen uniformly and independently at random.
 
@@
 
-\subsection{Multiple-choice balls into bins}
+\subsection{Multiple-choice balls-into-bins}
 \label{sec:multiple-choice}
-The \emph{single-choice balls into bins strategy} (Section \ref{sec:single-choice}) chooses one bin for each ball.
-A \emph{multiple-choice balls into bins strategy} generalizes this idea by choosing $d$ bins and placing the ball into the least full bin.
+The \emph{single-choice balls-into-bins strategy} (Section \ref{sec:single-choice}) chooses one bin for each ball.
+A \emph{multiple-choice balls-into-bins strategy} generalizes this idea by choosing $d$ bins and placing the ball into the least full bin.
 We will discuss two different multiple-choice schemes: The \emph{Greedy$[d]$} scheme (Section \ref{sec:greedy}) places the ball into the least full bin among $d$ bins chosen uniformly and independently at random whereas the \emph{Always-Go-Left} scheme (Section \ref{sec:AlwaysGoLeft}) uses a nonuniform selection of $d$ bins.
@@
 
-Consider a variation of the above \emph{one-choice balls into bins} strategy, which we will call the \emph{Greedy$[d]$ scheme}: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}.
+Consider a variation of the above \emph{one-choice balls-into-bins} strategy, which we will call the \emph{Greedy$[d]$ scheme}: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}.
 If there are multiple bins with the same minimal load, chose one among them arbitrarily.
@@
 \begin{itemize}
-\item The apparently small change made for the \emph{multiple-choice balls into bins} strategy results in an exponential decrease of the gap between the fullest box and the average load, even for $d=2$.
+\item The apparently small change made for the \emph{multiple-choice balls-into-bins} strategy results in an exponential decrease of the gap between the fullest box and the average load, even for $d=2$.
 \item Each additional choice ($d > 2$) decreases the gap by only a constant factor \cite{MRS01}. 
-\item For the \emph{heavily loaded case}, the resulting gap of the \emph{multiple-choice balls into bins} strategy does not depend on the number of balls $m$.
+\item For the \emph{heavily loaded case}, the resulting gap of the \emph{multiple-choice balls-into-bins} strategy does not depend on the number of balls $m$.
 In contrast, the bound for the \emph{one-choice} diverges with $m$.
@@
 \label{sec:AlwaysGoLeft}
-The multiple-choice balls into bins strategy presented in Section \ref{sec:multiple-choice} assumes that $d$ bins are chosen uniformly and independently at random.
+The multiple-choice balls-into-bins strategy presented in Section \ref{sec:multiple-choice} assumes that $d$ bins are chosen uniformly and independently at random.
 V\"ocking \cite{VOC03} was the first to study the impact of non-uniform and dependent selections of the $d$ bins.
-Surprisingly, a multiple-choice balls into bins strategy with a \emph{nonuniform and dependent} selection of the $d$ bins, such as the \emph{Always-Go-Left} algorithm introduced in \cite{VOC03}, achieves better\cite{BCSV06} load balancing than the \emph{Greedy} scheme (Section \ref{sec:multiple-choice}).
+Surprisingly, a multiple-choice balls-into-bins strategy with a \emph{nonuniform and dependent} selection of the $d$ bins, such as the \emph{Always-Go-Left} algorithm introduced in \cite{VOC03}, achieves better\cite{BCSV06} load balancing than the \emph{Greedy} scheme (Section \ref{sec:multiple-choice}).
 

### 2014-01-14 article/article.tex
rewise introduction

@@
 
-We will first introduce different variations of the \emph{balls-into-bins problem} in Section \ref{sec:classification} to provide some overview of different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}.
-Sections \ref{sec:strategies} to TODO will deal with the most basic variation of the balls-into-bins problem, namely the sequential and static variant.
-Thus, the balls are assumed to arrive sequentially and no strategy for deleting balls in the bins has to be considered.
-TODO: finite or infinite process?
+A simple strategy for the balls-into-bins problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice balls-into-bins} strategy.
+Using this strategy, the maximum load of the fullest bin will be $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ w.h.p.~\cite{RS98}.
 
-A simple strategy for the balls-into-bins problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice balls-into-bins} strategy.
+Consider a variation of the above \emph{single-choice balls-into-bins} strategy: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}\footnote{If there are multiple bins with the same minimal load, chose one among them arbitrarily.}.
+This simple \emph{multiple-choice balls-into-bins} scheme, which we will call the \emph{Greedy} algorithm, achieves a maximum load of $\frac{m}{n} + \Theta\left(\frac{\log \log n}{\log d}\right)$ w.h.p.~\cite{ABKU99}~\cite{BCSV06}.
+This apperantly minor change introduces a small amount of choice which, compared to the single-choice scheme, leads to a significant reduction in the maximum load.
+Curiously, the exponential decrease in the gap is already achieved having just two choices (that is, $d=2$).
+Each additional choice decreases the bound for the maximum gap by just a constant factor~\cite{MRS01}.
+Extensive research following the influental paper by Azar et.
+al \cite{ABKU99}, who were the first to explicitly prove this phenomenon, suggests that this \emph{two-choice paradigm} is not just an artificat of the simple balls-into-bins model, but a general principle applicable to a variety of problems.
+
+The paper is organized as follows: We will first introduce different variations of the \emph{balls-into-bins problem} in Section \ref{sec:classification} to provide some overview of different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}.
+Sections \ref{sec:strategies} to TODO will deal with the most basic variation of the balls-into-bins problem, namely the sequential, finite and static variant (see Section \ref{sec:classification} for details).
+For comparison and as a starting point, Section \ref{sec:single-choice} provides the results for the single-choice balls-into-bins problem.
+Section \ref{sec:multiple-choice} provides an in-depth presentation and analysis of the a basic \emph{multiple-choice balls-into-bins} scheme, the \emph{Greedy} algorithm.
+
+Surprisingly, V\"ocking \cite{VOC03} showed that \emph{asymetric tie-breaking} in combination with \emph{partitioning} the $n$ bins into $d$ groups can further reduce the bound for the gap.
+Since the scheme is somewhat techniqual, we refer the reader to Section \ref{sec:AlwaysGoLeft} for details.
+TODO: more here?
+
+\begin{comment}
 In Section \ref{sec:strategies}, we will provide asymptotic bounds for the addative gap of this strategy, which we will then use to devise more complex strategies that significantly reduce these bounds.
-These result may sometimes seem surprising, which is why we will denote a significant part of this article to demonstrate a proof technique that may be useful for devising asymptotic bounds for the additive gap of different balls-into-bins scenarios and strategies.
-TODO
+\end{comment}
+
+\begin{comment}
+These result may sometimes seem surprising, which is why we will denote a significant part of this article to demonstrate proofs in detail.
+This proof technique may be useful for devising asymptotic bounds for the additive gap of different balls-into-bins scenarios and strategies.
+\end{comment}
 
@@
 \label{sec:strategies}
-We will now consider the most basic variant of the balls-into-bins problem: Namely the static and sequential variant.
-Thus, the balls are assumed to arrived one after the other and we do not delete any balls from the bins.
-Additionally, all bins are empty at the start of the placement process.
+We will now consider the most basic variant of the balls-into-bins problem: Namely the static, finite and sequential variant.
+Thus, the balls are assumed to arrive sequentially, no strategy for deleting balls in the bins has to be considered and we are interested in a bound for the maximum load or gap only after all $m$ balls have arrived.
 

### 2014-01-14 article/article.tex
fix intro to Always-go-left

@@
 V\"ocking \cite{VOC03} was the first to study the impact of non-uniform and dependent selections of the $d$ bins.
-Surprisingly, a multiple-choice balls-into-bins strategy with a \emph{nonuniform and dependent} selection of the $d$ bins, such as the \emph{Always-Go-Left} algorithm introduced in \cite{VOC03}, achieves better\cite{BCSV06} load balancing than the \emph{Greedy} scheme (Section \ref{sec:multiple-choice}).
+Surprisingly, a multiple-choice balls-into-bins strategy with a \emph{nonuniform and independent} selection of the $d$ bins, such as the \emph{Always-Go-Left} algorithm introduced in \cite{VOC03}, achieves better\cite{BCSV06} load balancing than the \emph{Greedy} scheme (Section \ref{sec:greedy}).
+Furthermore, V\"ocking proved that the \emph{Always-Go-Left} algorithm is almost optimal (TODO: be more precise).
 

### 2014-01-14 article/article.tex
revise sections

@@
 The paper is organized as follows: We will first introduce different variations of the \emph{balls-into-bins problem} in Section \ref{sec:classification} to provide some overview of different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}.
-Sections \ref{sec:strategies} to TODO will deal with the most basic variation of the balls-into-bins problem, namely the sequential, finite and static variant (see Section \ref{sec:classification} for details).
+Sections \ref{sec:single-choice} to \ref{sec:multiple-choice} will deal with the most basic variation of the balls-into-bins problem, namely the sequential, finite and static variant (see Section \ref{sec:classification} for details).
 For comparison and as a starting point, Section \ref{sec:single-choice} provides the results for the single-choice balls-into-bins problem.
@@
 
-\section{Strategies}
-\label{sec:strategies}
+\section{Single-choice balls-into-bins}
+\label{sec:single-choice}
 We will now consider the most basic variant of the balls-into-bins problem: Namely the static, finite and sequential variant.
@@
 
-\subsection{Single-choice balls-into-bins}
-\label{sec:single-choice}
 In the \emph{single-choice balls-into-bins} strategy, each ball is placed into a bin choosen uniformly and independently at random.
@@
 
-\subsection{Multiple-choice balls-into-bins}
+\section{Multiple-choice balls-into-bins}
 \label{sec:multiple-choice}
@@
 
-\subsubsection{The Greedy$[d]$ scheme}
+\subsection{The Greedy$[d]$ scheme}
 \label{sec:greedy}
@@
 In contrast, the bound for the \emph{one-choice} diverges with $m$.
+\begin{comment}
+TODO: this is not true, I think
 \item The given bounds are \emph{tight}, meaning that no other strategy that places each ball into one of $d$ randomly selected bins achieves a gap that is asymptotically lower.
+\end{comment}
 \end{itemize}
 
-\subsubsection{The Always-Go-Left scheme}
+\subsection{The Always-Go-Left scheme}
 \label{sec:AlwaysGoLeft}
@@
 
+\section{Conclusion}
+\label{sec:conclusion}
+
 %bibliography

### 2014-01-14 article/article.tex
change layout

@@
-\documentclass{acm_proc_article-sp}
+\documentclass[a4paper,10pt, twocolumn]{article}
+\usepackage[utf8]{inputenc}
 
@@
 \usepackage{amssymb}
+%page boarders
+\usepackage[top=3cm, bottom=3cm, left=2cm, right=2cm]{geometry}
+\usepackage{enumitem}
+\usepackage{float}
+\floatstyle{plaintop}
+
+\restylefloat{table}
+\setlist{nolistsep}
 
@@
 
-\title{The Power of Two Random Choices}
-\subtitle{\large A follow up seminar to the lecture ``Parallel Algorithms''\\ by Jesper Larsson Tr\"aff and Francesco Versaci, 2012 winter term at TU Wien.}
-
-\numberofauthors{1}
-\author{
-\alignauthor
-Martin Kalany\\
-       0825673
-}    
-\date{\today}
+\title{The Power of Two Random Choices\\ 
+\large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Träff, lecture ``Parallel Algorithms'', 2012 winter term at TU Wien.}}
+\author{Martin Kalany, 0825673}
 

### 2014-01-15 presentation/presentation.tex
use acm template for article; fix footnote issue with acm template and missing bracket
order new packages suggested by francesco!
clean up presentation template
drawing balls in (invisible) bins
drawing balls and bins

@@
 \usecolortheme{dolphin}
-\usepackage{color}
-\usepackage{amsmath}
+
+\usepackage{stringstrings}
+
+%graphs, pictures
 \usepackage{tikz}
-\usepackage{verbatim}
-\usetikzlibrary{arrows,shapes}
-\usepackage{hyperref}
+\usepackage{pgf}
+\usepackage{xxcolor}
+\usetikzlibrary{positioning}
+
+\usepackage{comment}
+
 \usepackage[style=numeric-comp,backend=bibtex]{biblatex}
-\usepackage{comment}
-\usepackage{listings}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{tabularx}
-\usepackage{color, colortbl}
 \addbibresource{../sources}
 
+%reduce font size of footnotes
 \let\oldfootnotesize\footnotesize
@@
 
+%---------------------------------------------------------------------
+%Front matter
+%---------------------------------------------------------------------
 \title{The Power of Two Random Choices}
-\subtitle{A follow up seminar to the Parallel Algorithms lecture}
+\subtitle{ A follow up seminar to the lecture "Parallel Algorithms"}
 \author{Martin Kalany}
@@
 {
-
   Graduate student in Computer Science\\
@@
 
+%---------------------------------------------------------------------
+%parameters for balls into bins graphs
+%---------------------------------------------------------------------
+%scale factor
+\newcommand\scalefac{0.5}
+%diameter of balls
+\newcommand\ballsize{5mm}
+%number of bins
+\newcommand\nrbins{6}
+%space between ball and bin
+\newcommand\padding{0.1*\ballsize}
+%height of a bin
+\newcommand\binheight{6*\balldiameter}
+%width of a bin
+\newcommand\binwidth{\balldiameter}
+%gap from bin to bin
+\newcommand\bingap{1.6*\balldiameter}
+%color of balls
+
+\tikzstyle{ballstyle} = [ball color=black!30!red]
+
+%---------------------------------------------------------------------
+%tikz styles and commands
+%---------------------------------------------------------------------
+\newcommand\balldiameter{2*\ballsize}
+
+%bin style
+\tikzstyle{topflat} = [
+	minimum width=(\binwidth+2*\padding)*\scalefac, 
+	minimum height=(\binheight+3*\padding)*\scalefac,
+	append after command={
+    		\pgfextra
+        		\fill[fill=blue!30] 
+        			(\tikzlastnode.north east) [rounded corners] |-
+        			(\tikzlastnode.south)  -| 
+        			(\tikzlastnode.west) [sharp corners] |- 
+        			(\tikzlastnode.north) -- cycle;
+	         \draw[rounded corners] 
+	         	(\tikzlastnode.north east) |- 
+	         	(\tikzlastnode.south) -| 
+	         	(\tikzlastnode.north west);
+    		\endpgfextra
+    }
+]
+
+%draw i'th bin
+\newcommand\bin[1]{
+	\path node[topflat, xshift=#1*\bingap*\scalefac, above, yshift=-\padding*\scalefac]  {};
+}
+
+%draw all bins
+\newcommand\bins{
+	\foreach \ibin in {1,...,\nrbins}
+		\bin{\ibin};
+}
+
+%paint a ball in bin #1, position #2 
+\newcommand\ball[2]{
+	\shade[ballstyle] (#1*\bingap,#2*\balldiameter-\ballsize) circle (\ballsize)
+}
+
+%put #2 balls into bin 1 <= #1 <= 6
+\newcommand\putinbin[2]{
+	\ifnum #2 > 0
+		\foreach \nrballs in {1,...,#2}
+ 			\ball{#1}{\nrballs};
+ 	\fi
+}
+
+%put balls into bins. #i: number of balls to put in i'th bin
+\newcounter{index}
+\newcommand\balls[1]{%
+	\getargs{#1}%
+  	\setcounter{index}{0}%
+  	\whiledo{\theindex < \narg}{%
+    		\stepcounter{index}%
+    		\putinbin{\theindex}{\csname arg\romannumeral\theindex\endcsname}%
+  	}%
+}
+
+%draw balls and bins. #i: number of balls to put in i'th bin
+\newcommand\bab[1]{%
+	\bins
+	\balls{#1}
+}
+
+%---------------------------------------------------------------------
+%document
+%---------------------------------------------------------------------
 \begin{document}
-
 \frame{\titlepage}
@@
 \begin{frame}
-\cite{mitzen01}
+\begin{tikzpicture}[scale=\scalefac]
+
+	\bab{1 2 3 4 5 6}
+
+\end{tikzpicture}
 \end{frame}
 
+%---------------------------------------------------------------------
+%bibliography
+%---------------------------------------------------------------------
 \begin{frame}[allowframebreaks]

### 2014-01-15 presentation/presentation.tex
minor adaption for style

@@
 	minimum width=(\binwidth+2*\padding)*\scalefac, 
-	minimum height=(\binheight+3*\padding)*\scalefac,
+	minimum height=(\binheight+5*\padding)*\scalefac,
 	append after command={
     		\pgfextra
-        		\fill[fill=blue!30] 
+        		\fill[fill=blue!40] 
         			(\tikzlastnode.north east) [rounded corners] |-

### 2014-01-22 article/article.tex
style changes by francesco
order new packages suggested by francesco!
highlight todos

@@
-\documentclass[a4paper,10pt, twocolumn]{article}
+\documentclass[a4paper,12pt]{article}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{hyperref}
-\usepackage{amssymb}
 %page boarders
-\usepackage[top=3cm, bottom=3cm, left=2cm, right=2cm]{geometry}
-\usepackage{enumitem}
+\usepackage[vmargin=3cm, hmargin=3cm]{geometry}
 \usepackage{float}
 \floatstyle{plaintop}
+\usepackage{paralist} %compactitem
+\usepackage{fixltx2e} %some latex fixes
+\usepackage{microtype} %Subliminal refinements towards typographical perfection
+\setlength{\emergencystretch}{2em}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage{mathrsfs} %Support use of the Raph Smith’s Formal Script font in mathematics
+\usepackage[sc]{mathpazo} %mathematical fonts
+\usepackage{amsfonts}
+\usepackage{booktabs}
+\usepackage[margin=10pt, font=small, labelfont=bf]{caption}
 
-\restylefloat{table}
-\setlist{nolistsep}
-
-\newcommand*{\PH}{\makebox[1ex]{\textbf{$\cdot$}}}
-
-%\newtheorem{lemma}{Lemma}
-%\newtheorem{corollary}[lemma]{Corollary}
-%\newtheorem{theorem}[lemma]{Theorem}
+\usepackage{color}
+\newcommand\todo[1]{\textcolor{red}{(#1)}}
 
@@
 Since the scheme is somewhat techniqual, we refer the reader to Section \ref{sec:AlwaysGoLeft} for details.
-TODO: more here?
+\todo{more here?}
 
@@
 Several different variations of this problem exist, which may be classified by the following properties:
-\begin{itemize}
+\begin{compactitem}
 \item In a \emph{sequential process}, the balls arrive one after another.
@@
 \item A \emph{finite process} is analyzed for a given time interval that is known beforehand, whereas for an \emph{infinite process}, a placement and a deletion process is defined over an infinite time span \cite{ABS98}.
-\end{itemize}
+\end{compactitem}
 
@@
 In the \emph{lightly loaded case}, $m$ is in the order of $n$, whereas in the \emph{heavily loaded case}, $m$ is aribtrarily large.
-For clarity, we will assume that $m = n$ in the slightly loaded case and that $m \ll n$\footnote{Let $f$, $g$ be arbitrary functions.
+For clarity, we will assume that $m = n$ in the slightly loaded case and that $m \ll n$\todo{TODO} \footnote{Let $f$, $g$ be arbitrary functions.
 Then, $f \ll g$ means $f = o(g)$ and $ f \gg g$ means $f = \omega(g)$} in the heavily loaded case.
@@
 %More accurately, the number of balls in the bin with the most balls is $\Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$ \cite{G91}.
-%TODO: really include Gonnet?
+%\todo{really include Gonnet?}
 For the \emph{heavily loaded case} the maximum load is $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ w.h.p.
 
-TODO: proof? Maybe just outline the $m = n$ case as in \cite{RS98}?
+\todo{proof? Maybe just outline the $m = n$ case as in \cite{RS98}?}
 
@@
 The result stated above implies a gap of $\Theta\left(\log \log n \right)$ for arbitrarily large $m$. This is remarkable in several ways:
-\begin{itemize}
+\begin{compactitem}
 \item The apparently small change made for the \emph{multiple-choice balls-into-bins} strategy results in an exponential decrease of the gap between the fullest box and the average load, even for $d=2$.
@@
 \begin{comment}
-TODO: this is not true, I think
+\todo{this is not true, I think}
 \item The given bounds are \emph{tight}, meaning that no other strategy that places each ball into one of $d$ randomly selected bins achieves a gap that is asymptotically lower.
 \end{comment}
-\end{itemize}
+\end{compactitem}
 
@@
 Surprisingly, a multiple-choice balls-into-bins strategy with a \emph{nonuniform and independent} selection of the $d$ bins, such as the \emph{Always-Go-Left} algorithm introduced in \cite{VOC03}, achieves better\cite{BCSV06} load balancing than the \emph{Greedy} scheme (Section \ref{sec:greedy}).
-Furthermore, V\"ocking proved that the \emph{Always-Go-Left} algorithm is almost optimal (TODO: be more precise).
+Furthermore, V\"ocking proved that the \emph{Always-Go-Left} algorithm is almost optimal \todo{be more precise}.
 
-TODO: algorithm
-TODO: analysis
+\todo{algorithm}
+
+\todo{analysis}
 
@@
 %bibliography
-\bibliographystyle{abbrv}
+\bibliographystyle{acm}
 \bibliography{../sources} 

### 2014-01-24 article/article.tex
some review

@@
 This apperantly minor change introduces a small amount of choice which, compared to the single-choice scheme, leads to a significant reduction in the maximum load.
+\begin{comment}
 Curiously, the exponential decrease in the gap is already achieved having just two choices (that is, $d=2$).
 Each additional choice decreases the bound for the maximum gap by just a constant factor~\cite{MRS01}.
+\end{comment}
 Extensive research following the influental paper by Azar et.
 al \cite{ABKU99}, who were the first to explicitly prove this phenomenon, suggests that this \emph{two-choice paradigm} is not just an artificat of the simple balls-into-bins model, but a general principle applicable to a variety of problems.
+
+Condsider a slightly more involved strategy, which we will call the \emph{Always-go-left} scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size $\lfloor \frac{n}{d}\rfloor$ or $\lceil\frac{n}{d}\rceil$ \todo{simplify to n/d?}.
+We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
+In case a tie occurs, put the ball into the \emph{left-most} bin.
+That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with same minimal load, we put the ball into the bin with smallest index $i$.
+Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing \emph{asymetric tie-breaking} in combination with \emph{partitioning}\todo{check} reduces the bound for the gap to $\Theta \left( \frac{\log\log n}{d \log \Phi_d}\right)$ \todo{$\Phi_d$ is undefined}.
 
@@
 For comparison and as a starting point, Section \ref{sec:single-choice} provides the results for the single-choice balls-into-bins problem.
-Section \ref{sec:multiple-choice} provides an in-depth presentation and analysis of the a basic \emph{multiple-choice balls-into-bins} scheme, the \emph{Greedy} algorithm.
-
-Surprisingly, V\"ocking \cite{VOC03} showed that \emph{asymetric tie-breaking} in combination with \emph{partitioning} the $n$ bins into $d$ groups can further reduce the bound for the gap.
-Since the scheme is somewhat techniqual, we refer the reader to Section \ref{sec:AlwaysGoLeft} for details.
-\todo{more here?}
+\todo{TODO: multiple-choice, analysis, mention that witness tree method very helpful}
 
@@
 
-In the \emph{single-choice balls-into-bins} strategy, each ball is placed into a bin choosen uniformly and independently at random.
+In the \emph{single-choice balls-into-bins} strategy, each ball is placed into a bin choosen uniformly and independently at random. \todo{more specific?}
 
@@
 A \emph{multiple-choice balls-into-bins strategy} generalizes this idea by choosing $d$ bins and placing the ball into the least full bin.
+
+\todo{discuss the 3 classes}
+
 We will discuss two different multiple-choice schemes: The \emph{Greedy$[d]$} scheme (Section \ref{sec:greedy}) places the ball into the least full bin among $d$ bins chosen uniformly and independently at random whereas the \emph{Always-Go-Left} scheme (Section \ref{sec:AlwaysGoLeft}) uses a nonuniform selection of $d$ bins.
@@
 
-\todo{analysis}
-
 \section{Conclusion}

### 2014-01-26 article/article.tex
add fibonacci and golden ration
classes of placement algorithms

@@
 \label{sec:multiple-choice}
-The \emph{single-choice balls-into-bins strategy} (Section \ref{sec:single-choice}) chooses one bin for each ball.
-A \emph{multiple-choice balls-into-bins strategy} generalizes this idea by choosing $d$ bins and placing the ball into the least full bin.
+The \emph{single-choice balls-into-bins strategy} (Section \ref{sec:single-choice}) chooses one bin uniformly and independently at random for each ball.
+A \emph{multiple-choice balls-into-bins strategy} generalizes this idea by  choosing $d \geq 2$ bins and placing the ball into the least full bin.
+Depending on how the $d$ locations are sampled, we distinguish three classes of placement algorithms \cite{VOC03}.
 
-\todo{discuss the 3 classes}
+Thus, let $B$ be the set of bins $\{1,2...n\}$, $\Omega = B^{d}$ be the sample space and $\mathcal{F} = 2^{\Omega}$ be the set of measurable events.
+We are only interested in events that represent a set of bins of size $d$.
+Thus, let $\mathcal{F}_d = \{\mathcal{E} \in \mathcal{F}: \left\vert \mathcal{E} \right\vert = d\}$ and set $\Pr\left(\mathcal{E}  \right) = 0$  $\forall \mathcal{E} \in \mathcal{F} \setminus \mathcal{F}_d $.
+The three classes of algorithms are distinguished by the probability function $\Pr\left(\mathcal{E}\right)$ for events $\mathcal{E}_d \in \mathcal{F}_d$:
+\begin{itemize}
+\item \emph{Uniform and independent.} Choose each of the $d$ bins for a ball uniformly and independently at random from the set of bins $B$. This implies 
+\[
+\forall \mathcal{E} \in \mathcal{F}_d: \Pr\left(\mathcal{E}\right) = \frac{1}{n ^{d}}
+\]
+\item \emph{Non-uniform and independent.} Choose the $i$th bin from $B$ independently at random according to some probability function
+\[
+\Pr: B \rightarrow \left[0,1\right]
+\]
+
+\item \emph{Non-uniform and dependent.} Choose the $d$ bins for a ball from $\Omega$ at random:
+\[
+\Pr: B^{d} \rightarrow \left[0,1\right]
+\]
+
+\end{itemize}
 
@@
 
+\subsection{Analysis}
+\label{sec:analysis}
+
+\subsubsection{Preliminaries}
+\label{sec:preliminaries}
+\todo{where to put this? required for  bound of always-go-left in intro}
+We define $d$-ary Fibonacci numbers as follows. For $k \leq0$, $F_d(k) = 0$, $F_d(1) = 1$ and for $k \geq 1$,
+
+\[
+F_d(k) = \sum_{i=1}^{d}F_d(k-i)
+\]
+Furthermore, define 
+\[
+\Phi_d = \lim_{k \rightarrow \infty} \sqrt[k]{F_d(k)}
+\]
+Notice that $F_2$ corresponds to the usual Fibonacci numbers, while $\Phi_2 $ corresponds to the golden ratio \cite{Knuth73}, which is generalized to $\Phi_d$, for which 
+\[
+i, j \in \mathbb{N}, i < j \Rightarrow \Phi_i < \Phi_j
+\]
+and 
+\[
+\lim_{d\rightarrow \infty} \Phi_d = 2
+\]
+hold.
+
 \section{Conclusion}

### 2014-01-26 article/article.tex
intro for greedy and always-go-left

@@
 Thus, let $\mathcal{F}_d = \{\mathcal{E} \in \mathcal{F}: \left\vert \mathcal{E} \right\vert = d\}$ and set $\Pr\left(\mathcal{E}  \right) = 0$  $\forall \mathcal{E} \in \mathcal{F} \setminus \mathcal{F}_d $.
-The three classes of algorithms are distinguished by the probability function $\Pr\left(\mathcal{E}\right)$ for events $\mathcal{E}_d \in \mathcal{F}_d$:
+The three classes of algorithms are distinguished by the probability function $\Pr\left(\mathcal{E}\right)$ for events $\mathcal{E} \in \mathcal{F}_d$:
 \begin{itemize}
@@
 \]
-\item \emph{Non-uniform and independent.} Choose the $i$th bin from $B$ independently at random according to some probability function
+\item Class 2: \emph{Non-uniform and independent.} Choose the $i$th bin from $B$ independently at random according to some probability function
 \[
@@
 
-\item \emph{Non-uniform and dependent.} Choose the $d$ bins for a ball from $\Omega$ at random:
+\item Class 3: \emph{Non-uniform and dependent.} Choose the $d$ bins for a ball from $\Omega$ at random:
 \[
@@
 
-We will discuss two different multiple-choice schemes: The \emph{Greedy$[d]$} scheme (Section \ref{sec:greedy}) places the ball into the least full bin among $d$ bins chosen uniformly and independently at random whereas the \emph{Always-Go-Left} scheme (Section \ref{sec:AlwaysGoLeft}) uses a nonuniform selection of $d$ bins.
+We will discuss two different multiple-choice placement strategies: The \emph{Greedy} algorithm (Section \ref{sec:greedy}) belongs to class 1, whereas the \emph{Always-Go-Left} scheme (Section \ref{sec:AlwaysGoLeft}) uses a non-uniform and independent selection of $d$ bins and thus belongs to class 2.
 

### 2014-01-26 article/article.tex
rework greedy

@@
 
-\subsection{The Greedy$[d]$ scheme}
+\subsection{The Greedy scheme}
 \label{sec:greedy}
 
-Consider a variation of the above \emph{one-choice balls-into-bins} strategy, which we will call the \emph{Greedy$[d]$ scheme}: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}.
+Consider a variation of the above \emph{one-choice balls-into-bins} strategy, which we will call the \emph{Greedy scheme}: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}.
 If there are multiple bins with the same minimal load, chose one among them arbitrarily.
@@
 
-The \emph{Greedy$[d]$ scheme} achieves a maximum load of $\frac{m}{n} + \frac{\log \log n}{\log d} \left( 1 + O\left(1\right)\right)$ for all $m$ w.h.p. (see \cite{ABKU99}, \cite{BCSV06}).
+The \emph{Greedy scheme} achieves a maximum load of 
+\[
+\frac{m}{n} + \Theta\left(\frac{\log \log n}{\log d} \right)
+\]
+for all $m$ w.h.p. \cite{ABKU99} \cite{BCSV06}.
 
+This implies a gap of $\Theta\left(\log \log n \right)$ for arbitrarily large $m$, which is remarkable in several ways:
+\begin{compactitem}
+\item The apparently small change made for the \emph{Greedy} algorithm compared to the \emph{single-choice balls-into-bins} scheme results in an exponential decrease of the gap, even for $d=2$. 
+\item Each additional choice ($d > 2$) decreases the gap by only a constant factor \cite{MRS01}. 
+\item For the \emph{heavily loaded case}, the resulting gap of the \emph{multiple-choice balls-into-bins} strategy does not depend on the number of balls $m$.
+In contrast, the bound for the \emph{one-choice balls-into-bins} scheme diverges with $m$.
+\begin{comment}
+\todo{the following is only true for class 1 algorithms}
+\item The given bounds are \emph{tight}, meaning that no other strategy that places each ball into one of $d$ randomly selected bins achieves a gap that is asymptotically lower.
+\end{comment}
+\end{compactitem}
+
+We will provide a detailed proof of this property in Section \ref{sec:analysis}.
+
+\subsubsection{History}
+\label{sec:historyOfGreedy}
 This was first proven by Azar et al.
 \cite{ABKU99} for the case $m = n$.
-(However, the case $d = 2$ was implicitly proven by Karp et al.
-\cite{KLM92}).
+However, the case $d = 2$ was implicitly proven by Karp et al.
+\cite{KLM92}.
 See also \cite{MRS01} for a simplified proof and an overview of proof techniques for the given problem.
@@
 A simplified proof may be found in \cite{TW13}.
-
-The result stated above implies a gap of $\Theta\left(\log \log n \right)$ for arbitrarily large $m$. This is remarkable in several ways:
-\begin{compactitem}
-\item The apparently small change made for the \emph{multiple-choice balls-into-bins} strategy results in an exponential decrease of the gap between the fullest box and the average load, even for $d=2$.
-\item Each additional choice ($d > 2$) decreases the gap by only a constant factor \cite{MRS01}. 
-\item For the \emph{heavily loaded case}, the resulting gap of the \emph{multiple-choice balls-into-bins} strategy does not depend on the number of balls $m$.
-In contrast, the bound for the \emph{one-choice} diverges with $m$.
-\begin{comment}
-\todo{this is not true, I think}
-\item The given bounds are \emph{tight}, meaning that no other strategy that places each ball into one of $d$ randomly selected bins achieves a gap that is asymptotically lower.
-\end{comment}
-\end{compactitem}
 

### 2014-01-26 article/article.tex
history of 2-choice paradigm

@@
 For comparison and as a starting point, Section \ref{sec:single-choice} provides the results for the single-choice balls-into-bins problem.
-\todo{TODO: multiple-choice, analysis, mention that witness tree method very helpful}
+\todo{TODO: multiple-choice, analysis, mention that witness tree method very helpful}  See also \cite{MRS01} an overview of proof techniques for the given problem.
 
@@
 \label{sec:historyOfGreedy}
-This was first proven by Azar et al.
+The history of proving this remarkable property of the two-choice paradigm is in itself quite interesting.
+It was first proven by Azar et al.
 \cite{ABKU99} for the case $m = n$.
-However, the case $d = 2$ was implicitly proven by Karp et al.
-\cite{KLM92}.
-See also \cite{MRS01} for a simplified proof and an overview of proof techniques for the given problem.
-Unfortunately, the analysis breaks down for the heavily loaded case, where $m = \omega\left(n\right)$.
+In the same paper, they claimed that the bound holds for arbitrary $m$.
+Unfortunately, their analysis breaks down for the heavily loaded case, where $m = \omega\left(n\right)$.
 Berenbrink at.
-all \cite{BFZR08} were the first to proof that the same bound holds for the heavily loaded case\footnote{However, they say that an event $\mathcal E$ occurs with high probability if $\Pr\left[\mathcal E \right]  \geq 1- n^{-\alpha}$, for an arbitrarily chosen constant $\alpha \geq 1$.}.
+all \cite{BFZR08} were the first to proof that indeed the same bound holds for the heavily loaded case\footnote{However, they use a slightly different definition for w.h.p.: An event $\mathcal E$ occurs with high probability if $\Pr\left[\mathcal E \right]  \geq 1- n^{-\alpha}$, for an arbitrarily chosen constant $\alpha \geq 1$.}.
 A simplified proof may be found in \cite{TW13}.
+Curiously, the case $d = 2$ was implicitly proven by Karp et al.
+in the context of simulating PRAMs on distributed memory machines\cite{KLM92}.
+As we will see in the next section, V\"ocking was able to slightly improve the bound as well as to show that the always-go-left strategy that he introduces is optimal up to additive constants \cite{VOC03}.
 

### 2014-01-26 article/article.tex
rework single choice balls into bins

@@
 A simple strategy for the balls-into-bins problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice balls-into-bins} strategy.
-Using this strategy, the maximum load of the fullest bin will be $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ w.h.p.~\cite{RS98}.
+Using this strategy, the maximum load of the fullest bin will be $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ with high probability\footnote{We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] = 1 - o(1)$, except when stated otherwise.}~\cite{RS98}.
 
@@
 This simple \emph{multiple-choice balls-into-bins} scheme, which we will call the \emph{Greedy} algorithm, achieves a maximum load of $\frac{m}{n} + \Theta\left(\frac{\log \log n}{\log d}\right)$ w.h.p.~\cite{ABKU99}~\cite{BCSV06}.
-This apperantly minor change introduces a small amount of choice which, compared to the single-choice scheme, leads to a significant reduction in the maximum load.
+This apparently minor change introduces a small amount of choice which, compared to the single-choice scheme, leads to a significant reduction in the maximum load.
 \begin{comment}
@@
 
-In the \emph{single-choice balls-into-bins} strategy, each ball is placed into a bin choosen uniformly and independently at random. \todo{more specific?}
+In the \emph{single-choice balls-into-bins} strategy, each ball is placed into a bin chosen uniformly and independently at random.
+In other words, a ball gets placed into a certain bin with probability $\frac{1}{n}$.
 
-The fullest bin contains $\Theta \left( \frac{\log n}{\log \log n} \right)$ for the lightly loaded case with high probability\footnote{We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] = 1 - o(1)$, except when stated otherwise.} \cite{RS98}.
-Since the average load of the bins is 1, this is also a bound for the gap.
+The fullest bin contains
+\[
+\Theta \left( \frac{\log n}{\log \log n} \right)
+\] 
+balls in the lightly loaded case w.h.p. Since the average load of the bins is 1, this is also a bound for the gap. 
 %More accurately, the number of balls in the bin with the most balls is $\Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$ \cite{G91}.
 %\todo{really include Gonnet?}
-For the \emph{heavily loaded case} the maximum load is $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ w.h.p.
-
-\todo{proof? Maybe just outline the $m = n$ case as in \cite{RS98}?}
+In the \emph{heavily loaded case} the maximum load is 
+\[
+\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)
+\] 
+w.h.p.~\cite{RS98}.
 

### 2014-01-26 article/article.tex
rework always-go-left

@@
 
-We will discuss two different multiple-choice placement strategies: The \emph{Greedy} algorithm (Section \ref{sec:greedy}) belongs to class 1, whereas the \emph{Always-Go-Left} scheme (Section \ref{sec:AlwaysGoLeft}) uses a non-uniform and independent selection of $d$ bins and thus belongs to class 2.
+We will discuss two different multiple-choice placement strategies: The \emph{Greedy} algorithm (Section \ref{sec:greedy}) belongs to class 1, whereas the \emph{Always-go-left} scheme (Section \ref{sec:AlwaysGoLeft}) uses a non-uniform and independent selection of $d$ bins and thus belongs to class 2.
 
@@
 
-Consider a variation of the above \emph{one-choice balls-into-bins} strategy, which we will call the \emph{Greedy scheme}: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}.
+Consider a variation of the above \emph{one-choice balls-into-bins} strategy, which we will call the \emph{Greedy} scheme: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}.
 If there are multiple bins with the same minimal load, chose one among them arbitrarily.
@@
 
-The \emph{Greedy scheme} achieves a maximum load of 
+The \emph{Greedy} scheme achieves a maximum load of 
 \[
@@
 in the context of simulating PRAMs on distributed memory machines\cite{KLM92}.
-As we will see in the next section, V\"ocking was able to slightly improve the bound as well as to show that the always-go-left strategy that he introduces is optimal up to additive constants \cite{VOC03}.
+As we will see in the next section, V\"ocking was able to slightly improve the bound as well as to show that the \emph{Always-go-left} strategy that he introduces is optimal up to additive constants \cite{VOC03}.
 
-\subsection{The Always-Go-Left scheme}
+\subsection{The Always-go-left scheme}
 \label{sec:AlwaysGoLeft}
-The multiple-choice balls-into-bins strategy presented in Section \ref{sec:multiple-choice} assumes that $d$ bins are chosen uniformly and independently at random.
+The \emph{Greedy} strategy presented in Section \ref{sec:multiple-choice} assumes that $d$ bins are chosen uniformly and independently at random.
 V\"ocking \cite{VOC03} was the first to study the impact of non-uniform and dependent selections of the $d$ bins.
-Surprisingly, a multiple-choice balls-into-bins strategy with a \emph{nonuniform and independent} selection of the $d$ bins, such as the \emph{Always-Go-Left} algorithm introduced in \cite{VOC03}, achieves better\cite{BCSV06} load balancing than the \emph{Greedy} scheme (Section \ref{sec:greedy}).
-Furthermore, V\"ocking proved that the \emph{Always-Go-Left} algorithm is almost optimal \todo{be more precise}.
+Surprisingly, a multiple-choice balls-into-bins strategy with a \emph{non-uniform} and independent selection of the $d$ bins, such as the \emph{Always-go-left} algorithm introduced in \cite{VOC03}, achieves better\cite{BCSV06} load balancing than the \emph{Greedy} scheme (Section \ref{sec:greedy}).
+Furthermore, V\"ocking proved that the \emph{Always-go-left} algorithm is optimal up to additive constants.
 
-\todo{algorithm}
+The \emph{Always-go-left} algorithm works as follows.
+Partition the $n$ bins into $d$ groups, each of size $\lfloor \frac{n}{d}\rfloor$ or $\lceil\frac{n}{d}\rceil$ .
+We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
+In case a tie occurs, put the ball into the \emph{left-most} bin.
+That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with the same minimal load, put the ball into the bin with smallest index $i$.
 

### 2014-01-27 article/article.tex
analysis is it's own major section
state results more clearly using theorems

@@
 \newcommand\todo[1]{\textcolor{red}{(#1)}}
+
+\newtheorem{lemma}{Lemma}
+\newtheorem{corollary}[lemma]{Corollary}
+\newtheorem{theorem}[lemma]{Theorem}
 
@@
 
-The \emph{Greedy} scheme achieves a maximum load of 
+\todo{specify deletions; fix intro text since we now do use deletions}
+
+\todo{use theorem from V\"ocking?}
+
+\begin{theorem}
+Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time.
+If the balls are placed into bins by the Greedy algorithm, at any time $t$ the maximum load is at most
 \[
@@
 \]
-for all $m$ w.h.p. \cite{ABKU99} \cite{BCSV06}.
+w.h.p.~at any time $t$ \cite{ABKU99} \cite{BCSV06}.
+
+\end{theorem}
 
@@
 
-\subsection{Analysis}
+\begin{theorem}
+If $n$ balls are placed into $n$ bins using the Always-go-left scheme, the number of balls in the fullest bin will be 
+\[
+\frac{\ln\ln n}{d \ln \Phi_d} \pm \Theta(1)
+\]
+w.h.p.
+\end{theorem}
+
+\begin{theorem}
+Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time.
+If the balls are placed into bins by the Greedy algorithm, at any time $t$ the maximum load is at most
+\[
+\frac{\ln\ln n}{d \ln \Phi_d} + O(h)
+\]
+w.h.p.
+\end{theorem}
+
+\section{Analysis}
 \label{sec:analysis}
 
-\subsubsection{Preliminaries}
+\subsection{Preliminaries}
 \label{sec:preliminaries}

### 2014-01-27 article/article.tex
intro and structure of analysis

@@
 \begin{theorem}
+\label{theorem:greedy}
 Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time.
@@
 \begin{theorem}
+\label{theorem:agln}
 If $n$ balls are placed into $n$ bins using the Always-go-left scheme, the number of balls in the fullest bin will be 
@@
 \begin{theorem}
+\label{theorem:algm}
 Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time.
@@
 \label{sec:analysis}
+In this section, we will prove Theorems \ref{theorem:greedy}, \ref{theorem:agln} and \ref{theorem:algm} in detail, using \emph{witness trees} to provide an upper bound for the event that a bin contains too many balls.
+We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections \ref{sec:witnessGreedy} and \ref{sec:AlwaysGoLeft} respectively.
+However, we use some simplifying assumptions in those two sections to provide a more clear presentation of the proof techniuqe.
+Section \ref{sec:nondistinctBalls} will extend these proofs to remove the simplifying assumptions we made initially.
+Section \ref{sec:moreBalls} will extend the proof to cover the heavily loaded case where $m > n$ as well.
 
@@
 
+\subsection{Witness tree for Greedy}
+\label{sec:witnessGreedy}
+
+\subsection{Witness tree for Always-go-left}
+\label{sec:witnessAlg}
+
+\subsection{Nondistinct balls}
+\label{sec:nondistinctBalls}
+
+\subsection{More balls than bins}
+\label{sec:moreBalls}
+
 \section{Conclusion}

### 2014-01-27 article/article.tex
witness trees

@@
 In this section, we will prove Theorems \ref{theorem:greedy}, \ref{theorem:agln} and \ref{theorem:algm} in detail, using \emph{witness trees} to provide an upper bound for the event that a bin contains too many balls.
+
+When the maximum load exceeds some given threshold value we say that a \emph{bad event} occurs and construct an \emph{activated witness tree}, i.e., the activation of a witness tree is implied by a bad event.
+The probability for the existance of an activated witness tree upper-bounds the probability that a bad event occurs.
+
 We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections \ref{sec:witnessGreedy} and \ref{sec:AlwaysGoLeft} respectively.

### 2014-02-07 article/article.tex
analysis of the symmetric case

--- 2014-01-27 article/article.tex
@@
 \newtheorem{lemma}{Lemma}
-\newtheorem{corollary}[lemma]{Corollary}
-\newtheorem{theorem}[lemma]{Theorem}
-
+\newtheorem{corollary}{Corollary}
+\newtheorem{theorem}{Theorem}
+\newtheorem{assumption}{Assumption}
 \begin{document}
@@
 The three classes of algorithms are distinguished by the probability function $\Pr\left(\mathcal{E}\right)$ for events $\mathcal{E} \in \mathcal{F}_d$:
-\begin{itemize}
+\begin{compactitem}
 \item  \emph{Uniform and independent.} Choose each of the $d$ bins for a ball uniformly and independently at random from the set of bins $B$. This implies 
@@
 \]
-
-\end{itemize}
+\end{compactitem}
 
@@
 In this section, we will prove Theorems \ref{theorem:greedy}, \ref{theorem:agln} and \ref{theorem:algm} in detail, using \emph{witness trees} to provide an upper bound for the event that a bin contains too many balls.
-We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections \ref{sec:witnessGreedy} and \ref{sec:AlwaysGoLeft} respectively.
-However, we use some simplifying assumptions in those two sections to provide a more clear presentation of the proof techniuqe.
-Section \ref{sec:nondistinctBalls} will extend these proofs to remove the simplifying assumptions we made initially.
-Section \ref{sec:moreBalls} will extend the proof to cover the heavily loaded case where $m > n$ as well.
+Our analysis closely follows the proofs in \cite{VOC03}.
+\todo{is it enough to cite this once for the whole section?}
+
+When the maximum load exceeds some given threshold value we say that a \emph{bad event} occurs and construct an \emph{activated witness tree}.
+In other words, the activation of a witness tree is implied by a bad event.
+The probability for the existance of such an activated witness tree is an upper-bound for the probability that a bad event occurs.
+
+We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections \ref{sec:analysis:greedy} and \ref{sec:AlwaysGoLeft}, respectively.
+However, we use a simplifying assumption \ref{assumption:independence} in those two sections to provide a more clear presentation of the proof technique.
+Section \ref{sec:analysis:nondistinctBalls} will extend these proofs to remove the simplifying assumption we make initially.
+Section \ref{sec:analysis:moreBalls} will extend the proof to cover the heavily loaded case where $m > n$ as well.
 
@@
 
-\subsection{Witness tree for Greedy}
-\label{sec:witnessGreedy}
+\subsection{Analysis of the Greedy-scheme}
+\label{sec:analysis:greedy}
+A \emph{symmetric witness tree} of order $L$ is a complete $d$-ary tree with $d^{L}$ leaves.
+Each node $v$ represents a ball $b_v = \mathrm{ball}(v)$ .
+The same ball may be represented by several nodes.
+The nodes and edges of the witness tree represent events that may occur or not.
+For simplicity, we assume that all events represented by a witness tree are stochastically independent.
 
-\subsection{Witness tree for Always-go-left}
-\label{sec:witnessAlg}
+In the following, we donate the $d$ bins that are selected for a ball as the \emph{locations} of a ball.
+That is, before inserting a ball $d$ locations are generated and the ball is then inserted into one of those.
+We define edge and leaf events:
+\begin{compactitem}
+\item \emph{Edge event:} Consider some edge $e = (u,v)$ of the witness tree, s.t.~$v$ is the $i$th child of $u$.
+Then $e$ represents the event that the $i$'th location of the ball $b_u = \mathrm{ball}(u)$ points to the same bin as one of the locations of the ball $b_v = \mathrm{ball}(v)$.
+\todo{illustrate}
+\item \emph{Leaf event:} A leaf node $v$ of the witness tree represents the event that each of the $d$ locations of the ball $b_v = \mathrm{ball}(v)$ points to a bin that contains at least three additional balls (i.e., balls that are not represented by a node of the tree) at the time of the insertion of the ball $b_v$.
+\todo{illustrate}
+\end{compactitem}
+
+The following constraints have to be fullfilled by an assignment.
+The ball $b_r$ that is represented by the root node $r$ has to exist at timestep\footnote{As defined in Theorem \ref{theorem:greedy}} $t$ and each ball $b_v$ represented by a node $v \neq r$ and parent node $u$ has to exist when the ball $b_u = \mathrm{ball}(u)$ is inserted.
+
+The definition for leaf events refers to bins containing some number of balls at some specified time.
+To avoid dependencies, the crucial trick is that edge events are defined in terms of the alternative locations of balls, not by their finally assigned bin.
+
+\subsubsection{Symmetric witness tree: construction}
+\label{sec:constructionSymWT}
+Assume the bad event that some bin contains more than $L+3$ balls at time $t$ occurs.
+We will use an activated symmetric witness tree of order $L$ to witness this event and provide an upper-bound for the occurance of the event.
+Thus, we have to show that the existance of a bin $x$ containing at least $L+4$ balls implies the existance of an activated witness tree of order $L$.
+We construct the witness tree as follows:
+\begin{compactenum}
+\item The topmost ball $b_1$ in bin $x$ is assigned to the root node $v_r$. Note that $b_1$ is the last ball that was inserted into the bin $x$.
+\item Since the Greedy-scheme puts ball $b_1$ into bin $x$ only if each of the $d$ sampled locations contain at least the same number of balls at bin $x$, we can conclude that each of the $d$ locations contain at least $L+3$ balls at the time ball $b_1$ was inserted.
+Thus, the $d$ children of the root node are assigned the topmost balls of these $d$ bins.
+\item Consider these child nodes and continue the assignment of the previous step recursively until the leaf nodes are reached.  
+\end{compactenum}
+
+By construction, all edge events of the witness tree are activated.
+Note that the ball assigned to the root node is on top of at least $L+3$ balls.
+The balls assigned to the root's child node are each on top of $L+2$ or more balls and so on.
+It follows that each ball assigned to a leaf node is on top of at least 3 other balls, implying that the $d$ locations of each leaf node point to bins with at least 3 balls in each.
+This implies the activation of all leaf nodes of the witness tree.
+
+\subsubsection{Symmetric witness tree: probability of activation}
+\label{sec:probabilitySymWT}
+In this section, we derive the probability that some witness tree of order $L$ is activated, which gives an upper bound for the probability that the maximum load is $L+3$ at some time $t$.
+The probability $p_a$ that some witness tree of order $L$ is activated is bounded by the number of witness trees $w$ times the upper bound on the probability $p_1$ that a specific witness tree is activated.
+
+We will make use of Assumption \ref{assumption:independence} in the analysis. 
+\begin{assumption}
+\label{assumption:independence}
+All events represented by the nodes and edges of a witness tree are stochastically independent. 
+\end{assumption}
+In other words, we assume that no ball is referenced by more than one node of a witness tree.
+Obviously, a correct analysis will have to consider balls assigned to more than one node, but this simplifying assumption is useful to highlight the differences between the symmetric and asymmetric allocation schemes.
+Section \ref{sec:analysis:nondistinctBalls} will extend the analysis to witness trees with reduntant balls.
+
+The number of different witness trees $w$ is given by the number of ways to assign balls to its nodes.
+The ball assigned to the root node can be chosen from at most $n$ balls, since by definition (Theorem \ref{theorem:greedy}), at most $n$ balls are in the bins at any time $t$ and a ball assigned to a node has to be in some bin at time $t$.
+Each child node of the root is assigned one out of at most $n$ balls too by the same argument.
+Applying this argument for each level of the witness tree, there are at most $w = n^k$ possibilities to assign balls to a witness tree, with $k = d^{L+1}-1$ being the number of nodes in a $d$-ary witness tree of order $L$.
+
+An edge\footnote{where $v$ is the $i$'th child of $u$} $(u, v)$ is activated with probability at most $d/n$, which is the probability that the $i$'th location of the ball $b_u = \mathrm{ball}(u)$ points to a particular location of the ball $b_v = \mathrm{ball}(v)$ times the number of locations $d$.
+This implies a probability of at most
+\[
+p_e = \left(\frac{d}{n}\right)^{k-1}
+\] 
+or the event that all $k-1$ edges of a witness tree are activated.
+Note that  the assumption that all balls are distinct (Assumption \ref{assumption:independence}) implies that all edge events are independent.
+
+The probability that a leaf event is activated is bounded above by $3^{-d}$.
+Recall that by the definition of leaf events, each of the $d$ locations of the ball $b_{vl} = \mathrm{ball}(v_l)$ has to point to a bin with at least 3 additional balls in it.
+The number of those bins is bounded by $\lfloor n/3 \rfloor$ at all time steps $t$, implying a probability of at most
+\[
+\frac{n/3}{n} = \frac{1}{3}
+\end{align}
+that one such bin is chosen.
+This implies a probability of at most $3^{-d}$ that all $d$ bins are chosen among those with at least 3 additional bins in them.
+Thus, the probability that all leaf events are activated is at most $p_l = 3^{-d\cdot l}$, where $l = d^{L}$ is the number of leaves in a witness tree.
+
+Thus, the probability that an activated symmetric witness tree exists at time $t$ is at most 
+\begin{align}
+p_{swt} &= n^{k} \cdot \left(\frac{d}{n}\right)^{k-1} \cdot 3^{-d \cdot l} \\
+		&= n \cdot d^{k-1} \cdot 3^{-d \cdot l}
+\end{align}
+
+Using $k \leq 2l$ we get
+\begin{align}
+p_{swt} &\leq n \cdot d^{2\cdot l} \cdot 3^{-d \cdot l} \\
+		&= n \cdot \left(d^2\right)^l \cdot \left( 3^{-d} \right)^l
+\end{align}
+which we simplify by using $2d^2 \leq 3^d$
+\begin{align}
+p_{swt} &\leq n \cdot \left(\frac{3^d}{2}\right)^l \cdot \left( 3^{-d}\right)^l \\
+		&= n \cdot 2^{-l} \cdot 3^{d\cdot l} \cdot 3^{-d \cdot l} \\
+		&= n\cdot 2^{-d^L}		
+\end{align}
+
+Finally, we can show that the activation of a symmetric witness tree with distinct balls is unlikely, if the order of the tree $L$ is large enough. For any constant $\alpha > 0 $, let
+\begin{align}
+L \geq \log_d\log_2 n + \log_d\left(1+\alpha\right)
+\end{align}
+Then, 
+\begin{align}
+p_{swt} &\leq n \cdot 2^{{-d}^{\log_d\log_2 n + \log_d\left(1+\alpha\right)}} \\
+		&= n \cdot 2^{-{d}^{\log_d\log_2 n}\cdot d^{\log_d \cdot\left(1+\alpha\right)}} \\
+		&= n \cdot \left( 2^{-\log_2 n \cdot \left(1+ \alpha \right)} \right) \\
+		&= n \cdot n^{-\left(1+\alpha\right)} \\
+		&= n^{-\alpha}
+\end{align}
+
+\subsection{Analysis of the Always-go-left scheme}
+\label{sec:analysis:alg}
 
 \subsection{Nondistinct balls}
-\label{sec:nondistinctBalls}
+\label{sec:analysis:nondistinctBalls}
 
-\subsection{More balls than bins}
-\label{sec:moreBalls}
+\subsection{The heavily loaded case}
+\label{sec:analysis:moreBalls}
 

### 2014-02-09 article/article.tex
Use align environment for formulas

@@
 The fullest bin contains
-\[
+\begin{align}
 \Theta \left( \frac{\log n}{\log \log n} \right)
-\] 
+\end{align} 
 balls in the lightly loaded case w.h.p. Since the average load of the bins is 1, this is also a bound for the gap. 
@@
 In the \emph{heavily loaded case} the maximum load is 
-\[
+\begin{align}
 \frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)
-\] 
+\end{align} 
 w.h.p.~\cite{RS98}.
@@
 \item  \emph{Uniform and independent.} Choose each of the $d$ bins for a ball uniformly and independently at random from the set of bins $B$. This implies 
-\[
+\begin{align}
 \forall \mathcal{E} \in \mathcal{F}_d: \Pr\left(\mathcal{E}\right) = \frac{1}{n ^{d}}
-\]
+\end{align}
 \item Class 2: \emph{Non-uniform and independent.} Choose the $i$th bin from $B$ independently at random according to some probability function
-\[
+\begin{align}
 \Pr: B \rightarrow \left[0,1\right]
-\]
+\end{align}
 
 \item Class 3: \emph{Non-uniform and dependent.} Choose the $d$ bins for a ball from $\Omega$ at random:
-\[
+\begin{align}
 \Pr: B^{d} \rightarrow \left[0,1\right]
-\]
+\end{align}
 \end{compactitem}
@@
 If the balls are placed into bins by the Greedy algorithm, at any time $t$ the maximum load is at most
-\[
+\begin{align}
 \frac{m}{n} + \Theta\left(\frac{\log \log n}{\log d} \right)
-\]
+\end{align}
 w.h.p.~at any time $t$ \cite{ABKU99} \cite{BCSV06}.
@@
 If $n$ balls are placed into $n$ bins using the Always-go-left scheme, the number of balls in the fullest bin will be 
-\[
+\begin{align}
 \frac{\ln\ln n}{d \ln \Phi_d} \pm \Theta(1)
-\]
+\end{align}
 w.h.p.
@@
 If the balls are placed into bins by the Greedy algorithm, at any time $t$ the maximum load is at most
-\[
+\begin{align}
 \frac{\ln\ln n}{d \ln \Phi_d} + O(h)
-\]
+\end{align}
 w.h.p.
@@
 
-\[
+\begin{align}
 F_d(k) = \sum_{i=1}^{d}F_d(k-i)
-\]
+\end{align}
 Furthermore, define 
-\[
+\begin{align}
 \Phi_d = \lim_{k \rightarrow \infty} \sqrt[k]{F_d(k)}
-\]
+\end{align}
 Notice that $F_2$ corresponds to the usual Fibonacci numbers, while $\Phi_2 $ corresponds to the golden ratio \cite{Knuth73}, which is generalized to $\Phi_d$, for which 
-\[
+\begin{align}
 i, j \in \mathbb{N}, i < j \Rightarrow \Phi_i < \Phi_j
-\]
+\end{align}
 and 
-\[
+\begin{align}
 \lim_{d\rightarrow \infty} \Phi_d = 2
-\]
+\end{align}
 hold.
@@
 This implies a probability of at most
-\[
+\begin{align}
 p_e = \left(\frac{d}{n}\right)^{k-1}
-\] 
+\end{align} 
 or the event that all $k-1$ edges of a witness tree are activated.
@@
 The number of those bins is bounded by $\lfloor n/3 \rfloor$ at all time steps $t$, implying a probability of at most
-\[
+\begin{align}
 \frac{n/3}{n} = \frac{1}{3}

### 2014-02-09 article/article.tex
use \cdot for multiplication

@@
 A simple strategy for the balls-into-bins problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice balls-into-bins} strategy.
-Using this strategy, the maximum load of the fullest bin will be $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ with high probability\footnote{We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] = 1 - o(1)$, except when stated otherwise.}~\cite{RS98}.
+Using this strategy, the maximum load of the fullest bin will be $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \cdot \log n}{n}}\right)$ with high probability\footnote{We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] = 1 - o(1)$, except when stated otherwise.}~\cite{RS98}.
 
@@
 That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with same minimal load, we put the ball into the bin with smallest index $i$.
-Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing \emph{asymetric tie-breaking} in combination with \emph{partitioning}\todo{check} reduces the bound for the gap to $\Theta \left( \frac{\log\log n}{d \log \Phi_d}\right)$ \todo{$\Phi_d$ is undefined}.
+Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing \emph{asymetric tie-breaking} in combination with \emph{partitioning}\todo{check} reduces the bound for the gap to $\Theta \left( \frac{\log\log n}{d  \cdot \log \Phi_d}\right)$ \todo{$\Phi_d$ is undefined}.
 
@@
 \begin{align}
-\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)
+\frac{m}{n} + \Theta\left(\sqrt{\frac{m \cdot \log n}{n}}\right)
 \end{align} 
@@
 \begin{align}
-\frac{\ln\ln n}{d \ln \Phi_d} \pm \Theta(1)
+\frac{\ln\ln n}{d \cdot \ln \Phi_d} \pm \Theta(1)
 \end{align}
@@
 \begin{align}
-\frac{\ln\ln n}{d \ln \Phi_d} + O(h)
+\frac{\ln\ln n}{d \cdot \ln \Phi_d} + O(h)
 \end{align}

### 2014-02-09 article/article.tex
fibonacci tree definition

@@
 \label{sec:analysis:alg}
+Similarly to Section \ref{sec:analysis:greedy}, where we used symmetric witness trees to upper-bound the probability that a bin contains too many balls when using the \emph{Greedy} scheme, we will use asymmetric witness trees to get an analogous upper bound for the \emph{Always-go-left} scheme.
+
+With one exception, the definition of an \emph{asymmetric witness tree} is the same as the definition of a symmetric witness tree, including the definition of activation.
+Only its structure differs from that of a symmetric tree.
+An asymmetric witness tree has the same structure as a $d$-ary Fibonacci tree $T_d\left(k \right)$, which is defined as follows.
+\begin{compactitem}
+\item $T_d(1)$ is a single node
+\item $T_d(2) = T_d(1)$  
+\item For $3\leq k \leq d$, $T_d(k)$ is a tree with $k-1$ children, which are the roots of the trees $T_d(k-1),\ldots,T_d(1)$.
+\item For $k>d$, $T_d(k)$ is a tree with $d$ children, which are the roots of the trees $T_d(k-1),\ldots,T_d(k-d)$.
+\end{compactitem}
+\todo{illustrate}
+
+A $d$-ary asymmetric witness tree of order $L$ has the structure of the Fibonacci tree $T_d(d \cdot L+1)$ and thus has $F_d(d\cdot L + 1) \geq \Phi_d^{d\cdot L-1}$ \todo{How to get this inequality?} leaves.
 

### 2014-02-09 article/article.tex
revise section titles

@@
 \maketitle
+
 \begin{abstract}
@@
 \end{abstract}
+
+%\tableofcontents
 
@@
 
-We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections \ref{sec:analysis:greedy} and \ref{sec:AlwaysGoLeft}, respectively.
+We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections \ref{sec:analysis:greedy} and \ref{sec:analysis:alg}, respectively.
 However, we use a simplifying assumption \ref{assumption:independence} in those two sections to provide a more clear presentation of the proof technique.
@@
 
-\subsection{Analysis of the Greedy-scheme}
+\subsection{Greedy}
 \label{sec:analysis:greedy}
@@
 
-\subsubsection{Symmetric witness tree: construction}
+Assume the bad event that some bin contains more than $L+3$ balls at time $t$ occurs.
+We will use an activated symmetric witness tree of order $L$ to witness this event and provide an upper-bound for the occurrence of the event.
+Thus, we have to show that the existence of a bin $x$ containing at least $L+4$ balls implies the existence of an activated witness tree of order $L$.
+
+\subsubsection{Witness tree construction}
 \label{sec:constructionSymWT}
-Assume the bad event that some bin contains more than $L+3$ balls at time $t$ occurs.
-We will use an activated symmetric witness tree of order $L$ to witness this event and provide an upper-bound for the occurance of the event.
-Thus, we have to show that the existance of a bin $x$ containing at least $L+4$ balls implies the existance of an activated witness tree of order $L$.
-We construct the witness tree as follows:
+We construct a witness tree for the event that some bin contains $L+4$ balls as follows:
 \begin{compactenum}
@@
 
-\subsubsection{Symmetric witness tree: probability of activation}
+\subsubsection{Probability of activation}
 \label{sec:probabilitySymWT}
@@
 
-\subsection{Analysis of the Always-go-left scheme}
+\subsection{Always-go-left}
 \label{sec:analysis:alg}
@@
 
-\subsection{Nondistinct balls}
+\subsubsection{Witness tree construction}
+\label{sec:analysis:constructionAsymWT}
+
+\subsubsection{Probability of activation}
+\label{sec:analysis:probabilityAsymWT}
+
+\subsection{Non-distinct balls}
 \label{sec:analysis:nondistinctBalls}

### 2014-02-09 article/article.tex
asymmetric witness trees

@@
 \newtheorem{assumption}{Assumption}
+\newtheorem{invariant}{Invariant}
 \begin{document}
@@
 
-Condsider a slightly more involved strategy, which we will call the \emph{Always-go-left} scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size $\lfloor \frac{n}{d}\rfloor$ or $\lceil\frac{n}{d}\rceil$ \todo{simplify to n/d?}.
+Condsider a slightly more involved strategy, which we will call the \emph{Always-go-left} scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$ \todo{simplify to n/d?}.
 We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
@@
 The \emph{Always-go-left} algorithm works as follows.
-Partition the $n$ bins into $d$ groups, each of size $\lfloor \frac{n}{d}\rfloor$ or $\lceil\frac{n}{d}\rceil$ .
+Partition the $n$ bins into $d$ groups, each of size $\left\lfloor \frac{n}{d}\right\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$ .
 We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
@@
 
+Throughout this section, we donate the $d$ bins that are selected for a ball as the \emph{locations} of a ball.
+That is, before inserting a ball $d$ locations are generated and the ball is then inserted into one of those.
+
 We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections \ref{sec:analysis:greedy} and \ref{sec:analysis:alg}, respectively.
@@
 \label{sec:analysis:greedy}
+\subsubsection{Symmetric witness trees}
+\label{sec:analysis:definitionSymWT}
 A \emph{symmetric witness tree} of order $L$ is a complete $d$-ary tree with $d^{L}$ leaves.
@@
 
-In the following, we donate the $d$ bins that are selected for a ball as the \emph{locations} of a ball.
-That is, before inserting a ball $d$ locations are generated and the ball is then inserted into one of those.
 We define edge and leaf events:
@@
 
-\subsubsection{Witness tree construction}
-\label{sec:constructionSymWT}
+\subsubsection{Construction}
+\label{sec:analysis:constructionSymWT}
 We construct a witness tree for the event that some bin contains $L+4$ balls as follows:
@@
 \subsubsection{Probability of activation}
-\label{sec:probabilitySymWT}
+\label{sec:analysis:probabilitySymWT}
 In this section, we derive the probability that some witness tree of order $L$ is activated, which gives an upper bound for the probability that the maximum load is $L+3$ at some time $t$.
@@
 \begin{align}
-p_{swt} &= n^{k} \cdot \left(\frac{d}{n}\right)^{k-1} \cdot 3^{-d \cdot l} \\
+p_{swt} &= w \cdot p_e \cdot p_l \\
+		&= n^{k} \cdot \left(\frac{d}{n}\right)^{k-1} \cdot 3^{-d \cdot l} \\
 		&= n \cdot d^{k-1} \cdot 3^{-d \cdot l}
@@
 		&= n \cdot 2^{-l} \cdot 3^{d\cdot l} \cdot 3^{-d \cdot l} \\
+		&= n\cdot 2^{-l}		\\
 		&= n\cdot 2^{-d^L}		
@@
 
-Finally, we can show that the activation of a symmetric witness tree with distinct balls is unlikely, if the order of the tree $L$ is large enough. For any constant $\alpha > 0 $, let
+For any constant $\alpha > 0 $, let
 \begin{align}
 L \geq \log_d\log_2 n + \log_d\left(1+\alpha\right)
+\end{align}
+and $n$ be large enough such that 
+\begin{align}
+L \in \Theta \left(\ln \ln n \right)
 \end{align}
@@
 
+\todo{add the asymptotic bound}
+
 \subsection{Always-go-left}
@@
 Similarly to Section \ref{sec:analysis:greedy}, where we used symmetric witness trees to upper-bound the probability that a bin contains too many balls when using the \emph{Greedy} scheme, we will use asymmetric witness trees to get an analogous upper bound for the \emph{Always-go-left} scheme.
+We will make the same simplification as stated in Assumption \ref{assumption:independence}, namely that all the events represented by the edges and nodes of a witness tree are stochastically independent.
 
+\subsubsection{Asymmetric witness trees}
+\label{sec:analysis:definitionAsymWT}
 With one exception, the definition of an \emph{asymmetric witness tree} is the same as the definition of a symmetric witness tree, including the definition of activation.
 Only its structure differs from that of a symmetric tree.
-An asymmetric witness tree has the same structure as a $d$-ary Fibonacci tree $T_d\left(k \right)$, which is defined as follows.
+An asymmetric witness tree has the structure of a $d$-ary Fibonacci tree $T_d\left(k \right)$, which is defined as follows.
 \begin{compactitem}
-\item $T_d(1)$ is a single node
-\item $T_d(2) = T_d(1)$  
+\item $T_d(1)$ is a single node.
+\item $T_d(2) = T_d(1)$.
 \item For $3\leq k \leq d$, $T_d(k)$ is a tree with $k-1$ children, which are the roots of the trees $T_d(k-1),\ldots,T_d(1)$.
@@
 
-\subsubsection{Witness tree construction}
+For each node we define a label $(h, i)$, where $h, i \in \mathbb{N}$, $0 \leq h \leq L$ and  $1\leq i \leq d$. 
+\begin{compactitem}
+\item A node that is the root of a Fibonacci tree $T_d(1)$ is assigned the label $(0,1)$.
+\item Analogously, a node that is the root of a Fibonacci tree $T_d(2)$ is assigned the label $(0,2)$. Thus, nodes labelled $(0,1)$ or $(0,2)$ do not have children.
+\item A node with label $(0, i)$, where $i>2$, has $i-1$ children with labels $(0,i-1)\dots (0,1)$
+\item A node with label $(h,i)$ , where $h>0$, has $d$ children with the labels $(h,i-1)\dots(h,1)$,$(h-1,d)\dots(h-1, i)$.
+\item The root of a Fibonacci tree $T_d(d\cdot h+i)$ has the label $(h, i)$. 
+\end{compactitem}
+Thus, our the root node of our witness tree will be labelled $(L, 1)$. Note that the labels are unique within the set of children of a node $v$, but not for all nodes of the tree.
+
+\subsubsection{Construction}
 \label{sec:analysis:constructionAsymWT}
+The labels defined in the previous section are useful to state the following invariant, which will be maintained while assigning balls to the nodes of the witness tree.
+
+\begin{invariant}
+Ball $b_v$ referenced by a node $v$ with label $(h,i)$ was placed in a bin belonging to group $i$. The bin contained at least $h+3$ other balls at the time of insertion of the ball $b_v$.
+\end{invariant}
+
+Let $x$ be the bin that contains $L+4$ balls.
+The root node gets assigned the topmost ball in bin $x$.
+Each of the $d$ locations of this ball points to a bin with at least $L+3$ balls in it, at the time of its insertion.
+These are the $d$ bins that were sampled by the Always-go-left scheme and thus bin $i$ belongs to the $i$'th group of bins, where each group consists of $n/d$ bins\footnote{For the sake of simplicity, we assume that $d$ divides $n$ evenly, i.e., that all groups have the exact same size.
+It can be shown that the analysis still holds if each group is of size $\lfloor\frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$.}.
+The child node of the root with label $(L-1, i)$ is assigned the topmost ball of the bin the $i$' location points to.
+The assignment of balls to nodes than proceeds recursively.
+Given a node $v$ with label $(h, i)$ whose ball $b_v$ was put into some bin belonging to group $i$ and containing at least $l+3$ other balls, we can conclude the following.
+\begin{compactitem}
+\item For $1\leq j < i$, the $j$'th location of the ball $b_v$ points to a bin containing at least $l+4$ balls at the insertion time of the ball $b_v$.
+If the $j$'th location contained less balls than the $i$'th location, the Always-go-left scheme would have put the ball $b_v$ into the bin pointed to by location $j$.
+The topmost ball in location $j$ is assigned to the  child $u$ of $v$ with label $(l,j)$.
+\item For $i < j \leq d$, the $j$'th location contains at least $l+3$ balls.
+If location $j$ contained less, the Always-go-left scheme would have put the ball $b_v$ into the bin it points to.
+The topmost ball in the bin that location $j$ points to is assigned to the child node $u$ of $v$ labelled with $(l-1, j)$.
+\todo{the paper claims that the locations $j$ point to bins with $l+2$ balls.
+This can not be true, since the topmost, i.e., the $l+2$ ball - call it $b_t$ - gets assigned to a node with label $(l'=l-1, j)$.
+Since there are $l+1=l'+2$ balls below it, at the time of the placement of ball $b_t$, the bin $j$ it gets put into has only $l'+2$ balls in it, violating the invariant.
+Same with the case $j<i$.}
+\end{compactitem}
 
@@
 \label{sec:analysis:probabilityAsymWT}
+Analogously to the symmetric witness tree, the possible number of asymmetric witness trees is upper-bounded by $w=n^k$ and the probability that all edge events occur is\footnote{As in Section \ref{sec:analysis:greedy}, $k$ denotes the number of nodes of the witness tree and $l$ the number of leaves.}
+\begin{align}
+p_e = \left(\frac{d}{n}\right)^{k-1}
+\end{align}
+
+The probability that all leaf events occur too is the same as for the symmetric witness tree, but its derivation differs.
+Again, at any time at most $b'=n/3$ bins contain 3 or more balls, but we do not know how those bins are distributed among the groups.
+Let $b_i$ be the number of bins in group $i$ that contain 3 or more balls.
+Then $\beta_i = b_i\cdot d/ n$ is the fraction of the bins $b_i$ in group $i$.
+Since no more than $n/3$ bins can contain 3 or more balls, the values $b_i$ are subject to the constraint
+\begin{align}
+\frac{n}{d}\sum_{i=1}^d \beta_i \leq \frac{n}{3}
+\end{align}
+and thus 
+\begin{align}
+\sum_{i=1}^d \beta_i \leq \frac{d}{3}
+\end{align}
+
+The probability that all $d$ locations of a ball point to a bin containing 3 or more balls is then 
+\begin{align}
+p_{l_1} = \prod_{i=1}^{d} \beta_i
+\end{align}
+In other words, a leaf event occurs with probability $p_{l_1}$, which is maximized when all $b_i$ are equal.
+We thus set $b_i \leq 1/3$ for all $i$ and get as upper bound for the probability that all leaf events occur
+\begin{align}
+p_l \leq 3^{-d}
+\end{align}
+
+The above results imply the same probability for the activation of an asymmetric witness tree as we derived in Section \ref{sec:analysis:probabilitySymWT} for the activation of a symmetric witness tree, that is
+\begin{align}
+p_{awt} &\leq n 2^{-l}
+\end{align}
+However, the number of leaves is larger, since 
+\begin{align}
+l \geq \Phi_d^{d\cdot L -1}
+\end{align}
+holds\footnote{see Section \ref{sec:analysis:definitionAsymWT} for details}.
+
+For any constant $\alpha > 0$, let 
+\begin{align}
+L = \left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d}}\right  \rceil +1
+\end{align}
+and $n$ be large enough such that
+\begin{align}
+L \in \Theta\left(\frac{\ln\ln n}{d\cdot \ln \Phi_d}\right)
+\end{align}
 

### 2014-02-10 article/article.tex
add asymptotic bounds for L
analysis of always go left

@@
 Furthermore, define 
-\begin{align}
+\begin{align}\label{eqn:goldenFib}
 \Phi_d = \lim_{k \rightarrow \infty} \sqrt[k]{F_d(k)}
@@
 \begin{align}
-L \geq \log_d\log_2 n + \log_d\left(1+\alpha\right)
+L &\geq \log_d\log_2 n + \log_d\left(1+\alpha\right)
 \end{align}
-and $n$ be large enough such that 
+and note that
 \begin{align}
-L \in \Theta \left(\ln \ln n \right)
+L &\geq  \frac{\ln\log_2 n}{\ln d} + \frac{\ln(1+\alpha)}{\ln d} \\
+   &= \frac{\ln \log_2 n}{\ln d} \cdot \left(1 + \frac{\ln\left(1+\alpha\right)}{\ln \log_2 n}\right) \\
+  &\geq  \frac{\ln \ln n}{\ln d} \cdot \left( 1+ o\left( 1\right)\right)
 \end{align}
+
 Then, 
@@
 		&= n \cdot \left( 2^{-\log_2 n \cdot \left(1+ \alpha \right)} \right) \\
-		&= n \cdot n^{-\left(1+\alpha\right)} \\
-		&= n^{-\alpha}
+		&= n \cdot n^{-\left(1+\alpha\right)} = n^{-\alpha}
 \end{align}
-
-\todo{add the asymptotic bound}
 
@@
 
-A $d$-ary asymmetric witness tree of order $L$ has the structure of the Fibonacci tree $T_d(d \cdot L+1)$ and thus has $F_d(d\cdot L + 1) \geq \Phi_d^{d\cdot L-1}$ \todo{How to get this inequality?} leaves.
+A $d$-ary asymmetric witness tree of order $L$ has the structure of the Fibonacci tree $T_d(d \cdot L+1)$ and by Equation \ref{eqn:goldenFib} has $F_d(d\cdot L + 1) \geq \Phi_d^{d\cdot L-1}$ leaves.
 
@@
 \begin{align}
-p_{awt} &\leq n 2^{-l}
+p_{awt} &\leq n \cdot 2^{-l}
 \end{align}
@@
 \begin{align}
-L = \left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d}}\right  \rceil +1
+L &\geq \left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d}}\right\rceil+\frac{1}{d}
 \end{align}
-and $n$ be large enough such that
+\todo{the paper uses $+1$, instead of $+1/d$, but I can't make workout the below result then} and note that 
 \begin{align}
-L \in \Theta\left(\frac{\ln\ln n}{d\cdot \ln \Phi_d}\right)
+L &\geq \frac{\ln\log_2 n}{d\cdot \ln \Phi_d} + \frac{\ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d} + \frac{1}{d}\\
+  &\geq \frac{\ln\ln n}{d\cdot \ln \Phi_d} + O\left(1\right)
+\end{align}
+
+Then, 
+\begin{align}
+p_{awt} &\leq n \cdot 2^{-\Phi_d^{d\cdot L -1}} \\
+        &= n \cdot 2^{-\Phi_d^{d\cdot \left(\left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d}}\right\rceil+\frac{1}{d}\right) -1}} \\
+        &\leq n \cdot 2^{-\Phi_d^{d\cdot \left({\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d}}+\frac{1}{d}\right) -1}} \\
+        &= n \cdot 2^{-\Phi_d^{{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d}}}} \\
+        &= n \cdot 2^{-\Phi_d^{\log_{\Phi_d}\log_2 n + \log_{\Phi_d}\left(1+\alpha\right)}} \\
+         &= n \cdot 2^{-\Phi_d^{\log_{\Phi_d}\log_2 n}\cdot \Phi_d^{ \log_{\Phi_d}\left(1+\alpha\right)}} \\
+         &= n \cdot 2^{-\log_2 n \cdot\left(1+\alpha\right)} \\
+         &= n \cdot n ^{-(1+\alpha)} = n^{-\alpha}
 \end{align}

### 2014-02-10 article/article.tex
nondistinct balls

@@
 \begin{align}
-p_l \leq 3^{-d}
+p_l \leq 3^{-d\cdot l}
 \end{align}
@@
 \label{sec:analysis:nondistinctBalls}
+So far, we have assumed that all balls represented by nodes in a witness tree are stochastically independent (Assumption \ref{assumption:independence}).
+Obviously, to get a valid upper bound on the probability for the activation of a witness tree, the possibility of non-distinct balls has to be accounted for.
+In other words, the edge and leaf events of the witness tree are not independent anymore.
+Will start with a witness tree referencing non-distinct balls and prune it such that dependencies between edge and leaf events are removed.
+Note that pruning will reduce the size of the witness tree and thus increase the probability of its activation.
+To compensate for this loss in size, we start with a larger witness tree, which we call a \emph{full witness tree} and will show that, speaking informaly, the pruning of many events is unlikely.
+Note that the technique presented in this section is applicable to both symmetric and asymmetric witness trees.
 
-\subsection{The heavily loaded case}
-\label{sec:analysis:moreBalls}
+\subsubsection{Full witness trees}
+\label{sec:analysis:FullWT}
+Let $gamma \geq 2$ be a suitable constant.
+The root node of a \emph{full witness tree} has $\gamma$ children, denoted by $v_{c_i}$ with $1\leq i \leq \gamma$.
+The nodes $v_{c_i}$ have exactly one child each and the balls assigned to them are pairwise distinct.
+The root thus has $\gamma$ grandchildren, each of which is the root of a witness tree of order $L$.
+The time constraints activation of a full witness tree is defined analogously to the previous variants.
+
+\subsubsection{Construction}
+\label{sec:analysis:constructionFullWT}
+Assume a bin $x$ contains at least $L+3+\gamma$ balls at time $t$. We construct a witness tree as follows.
+\begin{compactitem}
+\item The root is not assigned any ball.
+\item Each child of the root is assigned one of the topmost $\gamma$ balls in bin $x$, which are neccessarily distinct.
+\item For the assignment of nodes to the grandchildren of the root, consider a ball $b_c$ that is assigned to a child node $v_c$.
+Let the child node of $v_c$ be denoted by $v_g$.
+Since the ball $b$ was placed into bin $x$, at least one of its locations points to this bin $x$, which we call location $i$.
+Let $j = i+1 mod d$ be the ball's $j$'th location.
+Then, at the time of insertion of ball $b_c$, its $j$'th location points to a bin $y$ with at least $L+3$ other balls in it.
+We assign the topmost ball in bin $y$ to node $v_g$.
+\end{compactitem}
+Notice that the $\gamma$ additional balls in bin $x$ are assigned to the child nodes of the root only.
+
+\subsubsection{Pruning}
+\label{sec:analysis:pruningFullWT}
+To remove all dependencies in the witness tree, we have to find nodes that represent the same ball and prune the tree such that each ball is referenced only once.
+Thus, traverse the witness tree in breadth-frist order.
+When we reach a node $v$ that represents a ball $b$ already encountered before, let $e = (v,u)$ be the edge connecting $v$ and its parent node $u$.
+We call $e$ a \emph{cutoff edge} and use it as a witness that one of the $d$ locations of the ball $b_u$ is the same as one of the locations of some other ball $b_{u'}$, represented by a node $u'$, which was already traversed.
+Futhermore, we remove the whole tree rooted at node $v$ from our witness tree.
+We continue this pruning until either the breath-first traversal terminates or until we have obtained $\gamma$ cut-off edges.
+Note that more than $\gamma$ cut-off edges cannot exist, \todo{why?}.
+When finished, the \emph{pruned witness tree} does not contain any ball twice and thus all its events are independent.
+
+\subsubsection{Probability of activation}
+\label{sec:analysis:probabilityFullWT}
+If there are less than $\gamma$ cut-off edges, in at least one of the $\gamma$ symmetric or asymmetric witness trees contained in the full witness tree the referenced balls are distinct.
+Let this symmetric or asymmetric witness tree be denoted by $T$.
+Then, the probability for the activation of the full witness tree is upper-bounded by the probability for the activiation of $T$, implying a probability for activation of this witnees tree of $p = n^{- \alpha}$, as shown in Sections \ref{sec:analysis:probabilitySymWT} and \ref{sec:analysis:probabilityAsymWT}, respectively.
+
+Thus, it remains to find an upper-bound for the activiation of a full witness tree with $\gamma$ cut-off edges.
+Let $q$ be the number of references to balls by the full witness tree\footnote{Since no ball was assigned to the root node, the full witness tree consists of $q+1$ nodes} and assume that $q  \leq 2\cdot \gamma\left(\alpha+1\right)\cdot \log_2 n$, for a suitable value of $\gamma$.
+Then, there are at most $w_p=q^\gamma$ possibilities to transform a full witness tree with $q+1$ nodes to a pruned witness tree with $\gamma$ cut-off edges.
+Fix the shape of the tree to be one of those.
+Let $k$ be the number of balls referenced by the pruned witness tree and $l$ be the number of its leaf nodes.
+
+To bound the probability for the activation of all edge events, we need to specify which event is represented by an edge $e_i = (v_r, v_i)$, where $v_i$ with $1 \leq i \leq \gamma$ is a child of the root node $v_r$.
+Assume that the root node $v_r$ represents an arbitrary bin.
+Then the edge $e_i$ represents the event that one of the locations of the ball represented by node $v_i$ points to this bin.
+This implies that the probability of activation for an each $e_i$ is $p_{e_i} = 1/n \leq d/n$.
+
+By the same arguments as in Sections \ref{sec:analysis:probabilitySymWT} and \ref{sec:analysis:probabilityAsymWT}, we can conclude that there are at most $w=n^k$ assignments of balls to the nodes of the witness tree.
+Furthermore, the probability for the activation of all edge events is bounded from above by
+\begin{align}
+p_e = \left(\frac{d}{n}\right)^{k}
+\end{align}
+while the probability for the activation of all leaf events is bounded by 
+\begin{align}
+p_l \leq 3^{-d\cdot l}
+\end{align}
+
+The number of edges $k$ and the number of leaves $l$ in a pruned witness tree may be small compared to a symmetric or asymmetric witness tree, but we will use the cut-off edges to compensate.
+By definition (Section \ref{sec:analysis:pruningFullWT}), a cut-off edge $e=(v, u)$ represents the event that the ball $b_u$ represented by node $u$ shares a common location $i$ with another ball $b_{u'}$, represented by node $u'$.
+Since $u'$ has already been visited by the time the breadth-first traversal reaches node $v$, it is part of the pruned witness tree and was assigned one of $k\leq q$ balls.
+The probability that location $i$ of ball $b_u$ points to one of the $d$ locations of ball $b_{u'}$ is at most $d/n$.
+The probability that there are $\gamma$ cut-off edges is thus at most
+\begin{align}
+p_c &= \left(k\cdot \frac{d}{n}\right)^\gamma \\
+    & \leq \left(q\cdot \frac{d}{n}\right)^\gamma
+\end{align}
+
+Putting it all together, we get 
+\begin{align}
+p_{pwt} &= w_p\cdot w \cdot p_e \cdot p_l \cdot p_c \\
+		&= q\gamma\cdot n^{k+1}\cdot \left(\frac{d}{n}\right)^k\cdot3^{-d\cdot l} \cdot \left(q \frac{d}{n}\right)^\gamma\\
+		&= n \cdot d^k\cdot 3^{-d\cdot l} \cdot \left(q^2 \frac{d}{n} \right)^\gamma 
+\end{align}
+
+Using $k \leq 2l$, $d^2 \leq 3^d$ and $q \leq 2\gamma \cdot \left( \alpha +1 \right) \cdot \log_2 n$ this becomes
+\begin{align}
+p_{pwt} &\leq n\cdot \left(\frac{\left(2\gamma \cdot \left(\alpha +1 \right) \cdot \log_2 n\right)^2\cdot d}{n}\right)^\gamma \\
+        &= n^{\gamma +1 1} \cdot n ^{\log_n\left( \left(2\gamma\cdot \left(\alpha+1\right)\cdot \log_2n \right)^{2\gamma} \cdot d^\gamma \right)}\\
+        &= n^{-\gamma+1+o\left(1\right)}
+\end{align}
+
+%^{\log_n \left(2\gamma \cdot \left( \alpha +1 \right) \cdot \log_2 n\right)^{2\gamma}\cdot d^\gamma \right)}
 

### 2014-02-11 article/article.tex
review

@@
 \usepackage{color}
-\newcommand\todo[1]{\textcolor{red}{(#1)}}
+\newcommand\todo[1]{\textcolor{red}{(TODO: #1)}}
+\newcommand\load{L_{\mathrm{max}}}
 
@@
 \title{The Power of Two Random Choices\\ 
-\large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Träff, lecture ``Parallel Algorithms'', 2012 winter term at TU Wien.}}
+\large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Tr\"aff and Francesco Versaci, lecture ``Parallel Algorithms'', 2012 winter term at TU Wien.}}
 \author{Martin Kalany, 0825673}
@@
 \maketitle
-
 \begin{abstract}
@@
 
-%\tableofcontents
-
 \section{Introduction}
 \label{sec:intro}
-Suppose that $m$ balls are thrown into $n$ bins, according to some strategy.
+Suppose that $m$ balls are thrown into $n$ bins, according to some arbitrary placement strategy.
 We are interested in the \emph{number of balls in the fullest bin}, also called the \emph{maximum load}.
-Equivalently, one can look at the \emph{addative gap} of the allocation, that is the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
+Equivalently, one can look at the \emph{additive gap} of the allocation, that is the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
 These \emph{balls-into-bins games} or \emph{allocation problems} have been studied extensively in the probability literature (see e.g., \cite{JK77}).
-In the context of Computer Science, the balls-into-bins model is useful to study e.g., strategies for distributing \emph{tasks} (balls) to \emph{processors} (bins) as evenly as possible without having to query all processors for the number of tasks currently assigned to them.
+In the context of Computer Science, the balls-into-bins model is useful to study, e.g., strategies for distributing \emph{tasks} (balls) to \emph{processors} (bins) as evenly as possible without having to query all processors for the number of tasks currently assigned to them.
 
+\subsection{Balls-into-bins model}
+\label{sec:model}
+Although several different variants of this simple balls-into-bins game exist, we will assume a very basic model to highlight the crucial ideas in both placement strategies as well as the analysis of them.
+In our model, we start with $n$ bins and put balls into them, one per time step $t$.
+We assume that at any time step $t$ at most $m = h n$ balls are in the bins, where $h$ is some constant greater or equal to $1$.
+At each time step $t$, we delete one randomly chosen ball from the bins.
+While most of the paper and especially the analysis focuses on the lightly loaded case where $m=n$, we will give results for the case $m > n$ as well.
+Throughout the paper, we will denote the maximum load by $\load$ and the additive gap by $\delta$.
+Obviously,
+\begin{align}
+\delta = \load - \frac{m}{n}.
+\end{align}
+
+\subsection{Single-choice balls-into-bins}
+\label{sec:single-choice}
 A simple strategy for the balls-into-bins problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice balls-into-bins} strategy.
-Using this strategy, the maximum load of the fullest bin will be $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \cdot \log n}{n}}\right)$ with high probability\footnote{We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] = 1 - o(1)$, except when stated otherwise.}~\cite{RS98}.
+Using this strategy, the maximum load of the fullest bin will be
+\begin{comment}
+\footnote{A more accurate result exists: The maximum load is $\load = \Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$ \cite{G91}.}
+\end{comment}
+\begin{align}
+\load = \frac{\ln n}{\ln \ln n} \cdot \left(1 + o\left(1\right)\right)
+\end{align}
+for the case $m=n$, with high probability\footnote{We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] \geq 1 - n^{-\alpha}$ for an arbitrary constant $\alpha > 0$.}~\cite{RS98}.
+This implies\footnote{For a detailed proof, see Theorem \ref{theorem:asymptotic} in Appendix \ref{sec:asymptotic}.}
+\begin{align}
+\load\in \Theta\left(\frac{\ln n}{\ln \ln n}\right).
+\end{align}
+Note that since the average load is 1, $\delta = L_{max}$ in this case. For the heavily loaded case, w.h.p.~the maximum gap is at most~\cite{RS98}.
+\begin{align}
+\delta = \Theta\left(\sqrt{\frac{m \cdot \log n}{n}}\right).
+\end{align}
 
+\subsection{Multiple-choice balls-into-bins}
 Consider a variation of the above \emph{single-choice balls-into-bins} strategy: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}\footnote{If there are multiple bins with the same minimal load, chose one among them arbitrarily.}.
-This simple \emph{multiple-choice balls-into-bins} scheme, which we will call the \emph{Greedy} algorithm, achieves a maximum load of $\frac{m}{n} + \Theta\left(\frac{\log \log n}{\log d}\right)$ w.h.p.~\cite{ABKU99}~\cite{BCSV06}.
+This simple \emph{multiple-choice balls-into-bins} scheme, which we will call the \emph{Greedy} algorithm\footnote{A detailed definition will follow in Section \ref{sec:greedy}}, achieves a gap
+\begin{align}
+\delta = \frac{\log \log n}{\log d} + O(1)
+\end{align}
+ w.h.p.~for all $m$~\cite{ABKU99}~\cite{BCSV06}.
+
 This apparently minor change introduces a small amount of choice which, compared to the single-choice scheme, leads to a significant reduction in the maximum load. 
@@
 \end{comment}
-Extensive research following the influental paper by Azar et.
-al \cite{ABKU99}, who were the first to explicitly prove this phenomenon, suggests that this \emph{two-choice paradigm} is not just an artificat of the simple balls-into-bins model, but a general principle applicable to a variety of problems.
+Extensive research following the influential paper by Azar et al.~\cite{ABKU99}, who were the first to explicitly prove this phenomenon\footnote{For a brief history of the proof, which is quite interesting by itself, refer to Appendix \ref{sec:historyOfGreedy}.}, suggests that this \emph{two-choice paradigm} is not just an artifact of the simple balls-into-bins model, but a general principle applicable to a variety of problems.
 
-Condsider a slightly more involved strategy, which we will call the \emph{Always-go-left} scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$ \todo{simplify to n/d?}.
+Consider a slightly more involved strategy, which we will call the \emph{Always-go-left} scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size\footnote{For the sake of a clear presentation, we will assume that $d$ divides $n$ evenly in the remainder of this paper.} $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$.
 We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
 In case a tie occurs, put the ball into the \emph{left-most} bin.
-That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with same minimal load, we put the ball into the bin with smallest index $i$.
-Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing \emph{asymetric tie-breaking} in combination with \emph{partitioning}\todo{check} reduces the bound for the gap to $\Theta \left( \frac{\log\log n}{d  \cdot \log \Phi_d}\right)$ \todo{$\Phi_d$ is undefined}.
+That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with the same minimal load, we put the ball into the bin with smallest index $i$.
+Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing \emph{asymmetric tie-breaking} in combination with \emph{non-uniform} bin selection reduces the bound for the gap to\footnote{The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers.
+Refer to Appendix \ref{sec:fibonacci} for the definition.}
+\begin{align}
+\delta = \frac{\ln\ln n}{d \cdot \ln \Phi_d} + O(1)
+\end{align}
+w.h.p. Not only is this a significant reduction of the asymptotic bound, V\"ocking was able to show that the Always-go-left strategy is optimal up to additive constants \cite{VOC03}.
 
-The paper is organized as follows: We will first introduce different variations of the \emph{balls-into-bins problem} in Section \ref{sec:classification} to provide some overview of different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}.
-Sections \ref{sec:single-choice} to \ref{sec:multiple-choice} will deal with the most basic variation of the balls-into-bins problem, namely the sequential, finite and static variant (see Section \ref{sec:classification} for details).
-For comparison and as a starting point, Section \ref{sec:single-choice} provides the results for the single-choice balls-into-bins problem.
-\todo{TODO: multiple-choice, analysis, mention that witness tree method very helpful}  See also \cite{MRS01} an overview of proof techniques for the given problem.
+This paper is organized as follows: We give a basic classification of balls-into-bins games in Section \ref{sec:classification}, to provide some overview of the different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}.
+In the remaining sections of this paper we will consider the model given in Section \ref{sec:model}, which can be classified as a \emph{sequential, finite and dynamic process with random deletions}.
+Section \ref{sec:multiple-choice} starts with a brief definition of three different classes of random placement strategies and then continues with the presentation of the two placement algorithms outlines in the introduction: The  \emph{Greedy} algorithm (Section \ref{sec:greedy}) and the \emph{Always-go-left} algorithm (Section \ref{sec:AlwaysGoLeft}).
+Section \ref{sec:analysis} is the core of this paper and will give a detailed analysis of both strategies.
+We will use a proof technique called \emph{witness trees}, which allows for a clean analysis of both problems.
+Note that the technique is applicable to a wide variety of balls-into-bins scenarios, making it the most useful of several proof techniques applied to balls-into-bins problems.
+Other proof techniques such as \emph{layered induction} and a \emph{differential equations approach} have been used.
+They are however not part of this paper.
+See, e.g., \cite{MRS01} for an overview of different proof techniques for the balls-into-bins problem.
 
 \begin{comment}
-In Section \ref{sec:strategies}, we will provide asymptotic bounds for the addative gap of this strategy, which we will then use to devise more complex strategies that significantly reduce these bounds.
+In Section \ref{sec:strategies}, we will provide asymptotic bounds for the ad dative gap of this strategy, which we will then use to devise more complex strategies that significantly reduce these bounds.
 \end{comment}
@@
 In a \emph{dynamic problem}, balls can also be removed from the system.
-In a simple \emph{random deletions} model, for each newly inserted ball a ball choosen uniformly and independently at random from among the balls currently in the bins is removed from the system \cite{ABKU99} \cite{MRS01}.
-More elaborate shemes exists: An \emph{oblivious adversery} specifies the sequence of insertions and deletions of balls in advance, without any knowledge of the placement of the balls \cite{CFM+98}.
-Another approach was studied by Adler et.
-al \cite{ABS98} in the context of the parallel balls-into-bins problem: In each bin, the balls are stored in FIFO-order and at every time step the first ball in each of the bins is deleted.
+In a simple \emph{random deletions} model, for each newly inserted ball a ball chosen uniformly and independently at random from among the balls currently in the bins is removed from the system \cite{ABKU99} \cite{MRS01}.
+More elaborate schemes exists: An \emph{oblivious adversary} specifies the sequence of insertions and deletions of balls in advance, without any knowledge of the placement of the balls \cite{CFM+98}.
+Another approach was studied by Adler et al.~\cite{ABS98} in the context of the parallel balls-into-bins problem: In each bin, the balls are stored in FIFO-order and at every time step the first ball in each of the bins is deleted.
 \item A \emph{finite process} is analyzed for a given time interval that is known beforehand, whereas for an \emph{infinite process}, a placement and a deletion process is defined over an infinite time span \cite{ABS98}.
+\item In the \emph{lightly loaded case}, $m$ is in the order of $n$, whereas in the \emph{heavily loaded case}, $m$ is arbitrarily large.
+For clarity, we will assume that $m = n$ in the lightly loaded case and that $m > n$ in the heavily loaded case.
 \end{compactitem}
-
-We will distinguish two cases with respect to the number of balls $m$.
-In the \emph{lightly loaded case}, $m$ is in the order of $n$, whereas in the \emph{heavily loaded case}, $m$ is aribtrarily large.
-For clarity, we will assume that $m = n$ in the slightly loaded case and that $m \ll n$\todo{TODO} \footnote{Let $f$, $g$ be arbitrary functions.
-Then, $f \ll g$ means $f = o(g)$ and $ f \gg g$ means $f = \omega(g)$} in the heavily loaded case.
 
@@
 
-To maximize the throughput of the servers, a uniform distribution of tasks to servers is desireable.
+To maximize the throughput of the servers, a uniform distribution of tasks to servers is desirable.
 This can easily be achieved by a central dispatcher.
-However, such a central dispatcher poses a bottleneck within a distributed system and querying $m$ servers for their current load involves significant overhead, e.g., sending a message to each server and waiting for the reply and more efficient solutions are neccessary.
+However, such a central dispatcher poses a bottleneck within a distributed system and querying $m$ servers for their current load involves significant overhead, e.g., sending a message to each server and waiting for the reply and more efficient solutions are necessary.
 As with hashing, the balls-into-bins model can be applied easily: we need to assign tasks (the balls) to servers (the bins).
-Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends issues its request to the least loaded server, the overhead is comperatively small and the difference in load on the $n$ servers is bounded by $\Theta\left(\log \log n\right)$ w.h.p.
+Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends issues its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is bounded by $\Theta\left(\log \log n\right)$ w.h.p.
 \cite{KLM92} \cite{RS98} \cite{MRS01}.
@@
 However, a PRAM is much harder to realize in hardware than a \emph{Distributed Memory Machine} (DMM).
-Thus, ways to emulate a PRAM on a DMM have been studied extensivly.
+Thus, ways to emulate a PRAM on a DMM have been studied extensively.
 Such an emulation distributes the processors and memory cells of the PRAM to the processors and memory modules (typically one per processor) of the DMM with the goal to reduce the slowdown of the emulation, which is the time the DMM requires to simulate one step of the PRAM.
-Each memory modul can handle only one memory access at a time step.
+Each memory module can handle only one memory access at a time step.
 More accesses to the same module will result in memory contention.
 The balls-into-bins idea is useful for managing communication between the processors and the memory modules such that memory contention is minimal.
-For details, see Karp et.
-al \cite{KLM92} and the papers referenced there.
+For details, see Karp et al.~\cite{KLM92} and the papers referenced there.
 
@@
 For simplicity, we consider the \emph{permutation routing problem}, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes and assume a \emph{butterfly network} as the underlying interconnection network (see \cite{CLR09} for a detailed definition).
-Then, congestions is the maximum number of paths that are routed through a network link.
+Then, congestion is the maximum number of paths that are routed through a network link.
 The goal of a \emph{circuit routing algorithm} is to to allocate paths for all $n$ communication requests such that congestion is minimal.
@@
 
-\section{Single-choice balls-into-bins}
-\label{sec:single-choice}
-We will now consider the most basic variant of the balls-into-bins problem: Namely the static, finite and sequential variant.
-Thus, the balls are assumed to arrive sequentially, no strategy for deleting balls in the bins has to be considered and we are interested in a bound for the maximum load or gap only after all $m$ balls have arrived.
-
-In the \emph{single-choice balls-into-bins} strategy, each ball is placed into a bin chosen uniformly and independently at random.
-In other words, a ball gets placed into a certain bin with probability $\frac{1}{n}$.
-
-The fullest bin contains
-\begin{align}
-\Theta \left( \frac{\log n}{\log \log n} \right)
-\end{align} 
-balls in the lightly loaded case w.h.p. Since the average load of the bins is 1, this is also a bound for the gap. 
-%More accurately, the number of balls in the bin with the most balls is $\Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$ \cite{G91}.
-%\todo{really include Gonnet?}
-In the \emph{heavily loaded case} the maximum load is 
-\begin{align}
-\frac{m}{n} + \Theta\left(\sqrt{\frac{m \cdot \log n}{n}}\right)
-\end{align} 
-w.h.p.~\cite{RS98}.
-
 \section{Multiple-choice balls-into-bins}
@@
 
-Thus, let $B$ be the set of bins $\{1,2...n\}$, $\Omega = B^{d}$ be the sample space and $\mathcal{F} = 2^{\Omega}$ be the set of measurable events.
+\section{Classes of random placement strategies}
+\label{sec:classesOfPlacement}
+Let $B$ be the set of bins $\{1,2...n\}$, $\Omega = B^{d}$ be the sample space and $\mathcal{F} = 2^{\Omega}$ be the set of measurable events.
 We are only interested in events that represent a set of bins of size $d$.
@@
 \end{align}
-\item Class 2: \emph{Non-uniform and independent.} Choose the $i$th bin from $B$ independently at random according to some probability function
+\item Class 2: \emph{Non-uniform and independent.} Choose the i-th bin from $B$ independently at random according to some probability function
 \begin{align}
@@
 
-\todo{specify deletions; fix intro text since we now do use deletions}
-
-\todo{use theorem from V\"ocking?}
-
 \begin{theorem}
 \label{theorem:greedy}
-Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time.
-If the balls are placed into bins by the Greedy algorithm, at any time $t$ the maximum load is at most
+Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, the maximum load is at most 
 \begin{align}
-\frac{m}{n} + \Theta\left(\frac{\log \log n}{\log d} \right)
+\frac{m}{n} + \frac{\log \log n}{\log d}+ O(h)
 \end{align}
 w.h.p.~at any time $t$ \cite{ABKU99} \cite{BCSV06}.
-
 \end{theorem}
@@
 
-\subsubsection{History}
-\label{sec:historyOfGreedy}
-The history of proving this remarkable property of the two-choice paradigm is in itself quite interesting.
-It was first proven by Azar et al.
-\cite{ABKU99} for the case $m = n$.
-In the same paper, they claimed that the bound holds for arbitrary $m$.
-Unfortunately, their analysis breaks down for the heavily loaded case, where $m = \omega\left(n\right)$.
-Berenbrink at.
-all \cite{BFZR08} were the first to proof that indeed the same bound holds for the heavily loaded case\footnote{However, they use a slightly different definition for w.h.p.: An event $\mathcal E$ occurs with high probability if $\Pr\left[\mathcal E \right]  \geq 1- n^{-\alpha}$, for an arbitrarily chosen constant $\alpha \geq 1$.}.
-A simplified proof may be found in \cite{TW13}.
-Curiously, the case $d = 2$ was implicitly proven by Karp et al.
-in the context of simulating PRAMs on distributed memory machines\cite{KLM92}.
-As we will see in the next section, V\"ocking was able to slightly improve the bound as well as to show that the \emph{Always-go-left} strategy that he introduces is optimal up to additive constants \cite{VOC03}.
-
 \subsection{The Always-go-left scheme}
@@
 V\"ocking \cite{VOC03} was the first to study the impact of non-uniform and dependent selections of the $d$ bins.
-Surprisingly, a multiple-choice balls-into-bins strategy with a \emph{non-uniform} and independent selection of the $d$ bins, such as the \emph{Always-go-left} algorithm introduced in \cite{VOC03}, achieves better\cite{BCSV06} load balancing than the \emph{Greedy} scheme (Section \ref{sec:greedy}).
+Surprisingly, a multiple-choice balls-into-bins strategy with a \emph{non-uniform} and independent selection of the $d$ bins, such as the \emph{Always-go-left} algorithm introduced in \cite{VOC03}, achieves better \cite{BCSV06} load balancing than the \emph{Greedy} scheme (Section \ref{sec:greedy}).
 Furthermore, V\"ocking proved that the \emph{Always-go-left} algorithm is optimal up to additive constants.
@@
 \begin{align}
-\frac{\ln\ln n}{d \cdot \ln \Phi_d} \pm \Theta(1)
+\frac{\ln\ln n}{d \cdot \ln \Phi_d} + O(1)
 \end{align}
@@
 
-When the maximum load exceeds some given threshold value we say that a \emph{bad event} occurs and construct an \emph{activated witness tree}.
+We say that a \emph{bad event} occurs when the maximum load exceeds some given threshold value.
+When such a bad event occurs we construct an \emph{activated witness tree}.
 In other words, the activation of a witness tree is implied by a bad event.
-The probability for the existance of such an activated witness tree is an upper-bound for the probability that a bad event occurs.
+The probability for the existence of such an activated witness tree is an upper-bound for the probability that the associated bad event occurs.
 
-Throughout this section, we donate the $d$ bins that are selected for a ball as the \emph{locations} of a ball.
-That is, before inserting a ball $d$ locations are generated and the ball is then inserted into one of those.
+Throughout this section, \emph{location} will denote the a specific place in one of the $n$ bins.
+A location of a ball $b$ refers not only to a bin, but also to the number of balls already in this bin at the time of insertion of ball $b$.
 
@@
 Section \ref{sec:analysis:nondistinctBalls} will extend these proofs to remove the simplifying assumption we make initially.
-Section \ref{sec:analysis:moreBalls} will extend the proof to cover the heavily loaded case where $m > n$ as well.
-
-\subsection{Preliminaries}
-\label{sec:preliminaries}
-\todo{where to put this? required for  bound of always-go-left in intro}
-We define $d$-ary Fibonacci numbers as follows. For $k \leq0$, $F_d(k) = 0$, $F_d(1) = 1$ and for $k \geq 1$,
-
-\begin{align}
-F_d(k) = \sum_{i=1}^{d}F_d(k-i)
-\end{align}
-Furthermore, define 
-\begin{align}\label{eqn:goldenFib}
-\Phi_d = \lim_{k \rightarrow \infty} \sqrt[k]{F_d(k)}
-\end{align}
-Notice that $F_2$ corresponds to the usual Fibonacci numbers, while $\Phi_2 $ corresponds to the golden ratio \cite{Knuth73}, which is generalized to $\Phi_d$, for which 
-\begin{align}
-i, j \in \mathbb{N}, i < j \Rightarrow \Phi_i < \Phi_j
-\end{align}
-and 
-\begin{align}
-\lim_{d\rightarrow \infty} \Phi_d = 2
-\end{align}
-hold.
+The witness tree method can be used to show that the same bounds hold for the heavily loaded case where $m> n$ too, but we do not give such a proof in this paper.
+Instead, we refer the interested reader to V\"ocking \cite{VOC03}.
+Berenbrink et al.
+~\cite{BCSV06}, Talwar et al.~\cite{TW13} and Mitzenmacher \cite{MRS01} provide proofs the heavily loaded case using different approaches as well.
 
@@
 \label{sec:analysis:greedy}
+In this Section, we will give a detailed proof for Theorem \ref{theorem:greedy}.
+However, we will restrict ourselves to the case where $m = n$ and will make use of the symplifying assumption \ref{assumption:independence}.
+
 \subsubsection{Symmetric witness trees}
@@
 \begin{compactitem}
-\item \emph{Edge event:} Consider some edge $e = (u,v)$ of the witness tree, s.t.~$v$ is the $i$th child of $u$.
-Then $e$ represents the event that the $i$'th location of the ball $b_u = \mathrm{ball}(u)$ points to the same bin as one of the locations of the ball $b_v = \mathrm{ball}(v)$.
+\item \emph{Edge event:} Consider some edge $e = (u,v)$ of the witness tree, s.t.~$v$ is the i-th child of $u$.
+Then $e$ represents the event that the i-th location of the ball $b_u = \mathrm{ball}(u)$ points to the same bin as one of the locations of the ball $b_v = \mathrm{ball}(v)$.
 \todo{illustrate}
@@
 
-The following constraints have to be fullfilled by an assignment.
-The ball $b_r$ that is represented by the root node $r$ has to exist at timestep\footnote{As defined in Theorem \ref{theorem:greedy}} $t$ and each ball $b_v$ represented by a node $v \neq r$ and parent node $u$ has to exist when the ball $b_u = \mathrm{ball}(u)$ is inserted.
+The following constraints have to be fulfilled by an assignment.
+The ball $b_r$ that is represented by the root node $r$ has to exist at time step\footnote{As defined in Theorem \ref{theorem:greedy}} $t$ and each ball $b_v$ represented by a node $v \neq r$ and parent node $u$ has to exist when the ball $b_u = \mathrm{ball}(u)$ is inserted.
 
@@
 
-We will make use of Assumption \ref{assumption:independence} in the analysis. 
+Initially, we will make use of the following assumption in the analysis. 
 \begin{assumption}
@@
 Obviously, a correct analysis will have to consider balls assigned to more than one node, but this simplifying assumption is useful to highlight the differences between the symmetric and asymmetric allocation schemes.
-Section \ref{sec:analysis:nondistinctBalls} will extend the analysis to witness trees with reduntant balls.
+Section \ref{sec:analysis:nondistinctBalls} will extend the analysis to witness trees with redundant balls.
 
@@
 
-An edge\footnote{where $v$ is the $i$'th child of $u$} $(u, v)$ is activated with probability at most $d/n$, which is the probability that the $i$'th location of the ball $b_u = \mathrm{ball}(u)$ points to a particular location of the ball $b_v = \mathrm{ball}(v)$ times the number of locations $d$.
+An edge\footnote{where $v$ is the i-th child of $u$} $(u, v)$ is activated with probability at most $d/n$, which is the probability that the i-th location of the ball $b_u = \mathrm{ball}(u)$ points to a particular location of the ball $b_v = \mathrm{ball}(v)$ times the number of locations $d$.
 This implies a probability of at most
@@
 \end{align}
-
 Using $k \leq 2l$ we get
@@
    &= \frac{\ln \log_2 n}{\ln d} \cdot \left(1 + \frac{\ln\left(1+\alpha\right)}{\ln \log_2 n}\right) \\
-  &\geq  \frac{\ln \ln n}{\ln d} \cdot \left( 1+ o\left( 1\right)\right)
+  &=  \frac{\ln \ln n}{\ln d} \cdot \left( 1+ o\left( 1\right)\right)
 \end{align}
@@
 \end{compactitem}
-\todo{illustrate}
+\todo{illustrate and note difference to usual Fibonacci trees}
 
@@
 \item A node that is the root of a Fibonacci tree $T_d(1)$ is assigned the label $(0,1)$.
-\item Analogously, a node that is the root of a Fibonacci tree $T_d(2)$ is assigned the label $(0,2)$. Thus, nodes labelled $(0,1)$ or $(0,2)$ do not have children.
+\item Analogously, a node that is the root of a Fibonacci tree $T_d(2)$ is assigned the label $(0,2)$. Thus, nodes labeled $(0,1)$ or $(0,2)$ do not have children.
 \item A node with label $(0, i)$, where $i>2$, has $i-1$ children with labels $(0,i-1)\dots (0,1)$
@@
 \end{compactitem}
-Thus, our the root node of our witness tree will be labelled $(L, 1)$. Note that the labels are unique within the set of children of a node $v$, but not for all nodes of the tree.
+Thus, our the root node of our witness tree will be labeled $(L, 1)$. Note that the labels are unique within the set of children of a node $v$, but not for all nodes of the tree.
 
@@
 If location $j$ contained less, the Always-go-left scheme would have put the ball $b_v$ into the bin it points to.
-The topmost ball in the bin that location $j$ points to is assigned to the child node $u$ of $v$ labelled with $(l-1, j)$.
+The topmost ball in the bin that location $j$ points to is assigned to the child node $u$ of $v$ labeled with $(l-1, j)$.
 \todo{the paper claims that the locations $j$ point to bins with $l+2$ balls.
@@
 Note that pruning will reduce the size of the witness tree and thus increase the probability of its activation.
-To compensate for this loss in size, we start with a larger witness tree, which we call a \emph{full witness tree} and will show that, speaking informaly, the pruning of many events is unlikely.
+To compensate for this loss in size, we start with a larger witness tree, which we call a \emph{full witness tree} and will show that, speaking informally, the pruning of many events is unlikely.
 Note that the technique presented in this section is applicable to both symmetric and asymmetric witness trees.
@@
 \item The root is not assigned any ball.
-\item Each child of the root is assigned one of the topmost $\gamma$ balls in bin $x$, which are neccessarily distinct.
+\item Each child of the root is assigned one of the topmost $\gamma$ balls in bin $x$, which are necessarily distinct.
 \item For the assignment of nodes to the grandchildren of the root, consider a ball $b_c$ that is assigned to a child node $v_c$.
@@
 To remove all dependencies in the witness tree, we have to find nodes that represent the same ball and prune the tree such that each ball is referenced only once.
-Thus, traverse the witness tree in breadth-frist order.
+Thus, traverse the witness tree in breadth-first order.
 When we reach a node $v$ that represents a ball $b$ already encountered before, let $e = (v,u)$ be the edge connecting $v$ and its parent node $u$.
 We call $e$ a \emph{cutoff edge} and use it as a witness that one of the $d$ locations of the ball $b_u$ is the same as one of the locations of some other ball $b_{u'}$, represented by a node $u'$, which was already traversed.
-Futhermore, we remove the whole tree rooted at node $v$ from our witness tree.
+Furthermore, we remove the whole tree rooted at node $v$ from our witness tree.
 We continue this pruning until either the breath-first traversal terminates or until we have obtained $\gamma$ cut-off edges.
@@
 Let this symmetric or asymmetric witness tree be denoted by $T$.
-Then, the probability for the activation of the full witness tree is upper-bounded by the probability for the activiation of $T$, implying a probability for activation of this witnees tree of $p = n^{- \alpha}$, as shown in Sections \ref{sec:analysis:probabilitySymWT} and \ref{sec:analysis:probabilityAsymWT}, respectively.
+Then, the probability for the activation of the full witness tree is upper-bounded by the probability for the activation of $T$, implying a probability for activation of this witness tree of $p = n^{- \alpha}$, as shown in Sections \ref{sec:analysis:probabilitySymWT} and \ref{sec:analysis:probabilityAsymWT}, respectively.
 
-Thus, it remains to find an upper-bound for the activiation of a full witness tree with $\gamma$ cut-off edges.
+Thus, it remains to find an upper-bound for the activation of a full witness tree with $\gamma$ cut-off edges.
 Let $q$ be the number of references to balls by the full witness tree\footnote{Since no ball was assigned to the root node, the full witness tree consists of $q+1$ nodes} and assume that $q  \leq 2\cdot \gamma\left(\alpha+1\right)\cdot \log_2 n$, for a suitable value of $\gamma$.
@@
 
+\appendix
+\section{Generalized Fibonacci numbers}
+\label{sec:fibonacci}
+We define $d$-ary Fibonacci numbers as follows. For $k \leq0$, $F_d(k) = 0$, $F_d(1) = 1$ and for $k \geq 1$,
+
+\begin{align}
+F_d(k) = \sum_{i=1}^{d}F_d(k-i)
+\end{align}
+Furthermore, define 
+\begin{align}\label{eqn:goldenFib}
+\Phi_d = \lim_{k \rightarrow \infty} \sqrt[k]{F_d(k)}
+\end{align}
+Notice that $F_2$ corresponds to the usual Fibonacci numbers, while $\Phi_2 $ corresponds to the golden ratio \cite{Knuth73}, which is generalized to $\Phi_d$, for which 
+\begin{align}
+i, j \in \mathbb{N}, i < j \Rightarrow \Phi_i < \Phi_j
+\end{align}
+and 
+\begin{align}
+\lim_{d\rightarrow \infty} \Phi_d = 2
+\end{align}
+hold.
+
+\section{Asymptotic notations}
+\label{sec:asymptotic}
+We use the following definitions for the asymptotic notations $O(\cdot)$, $o(\cdot)$ and $\Theta(\cdot)$.
+
+\begin{align*}
+f(n) \in o(g(n)) &\Leftrightarrow \forall \epsilon > 0 \exists n_o\forall n> n_0: |f(n)|\leq \epsilon \cdot |g(n)|\\
+f(n) \in O(g(n)) &\Leftrightarrow \exists c > 0 \exists n_o\forall n> n_0: |f(n)|\leq c \cdot |g(n)|\\
+f(n) \in \Omega(g(n)) &\Leftrightarrow \exists c > 0 \exists n_o\forall n> n_0: c \cdot |g(n)| \leq |f(n)| \\
+f(n) \in \Theta(g(n)) &\Leftrightarrow f(n) \in \Omega(g(n)) \land f(n)\in O(g(n))
+\end{align*}
+
+The following theorem is useful for understanding the asymptotic bounds presented throughout the paper.
+\begin{theorem}\label{theorem:asymptotic}
+\begin{align*}
+f(n) = g(n) \cdot (1+o(1)) \Rightarrow f(n) \in \Theta(g(n)) 
+\end{align*}
+\end{theorem}
+The proof is straight forward:
+\begin{align*}
+f(n) &= g(n)+o(1)\cdot g(n) \\
+&\Rightarrow f(n) \in \Omega(g(n)) \land f(n) \in O(g(n)) \\
+&\Rightarrow f(n) \in \Theta(g(n))
+\end{align*}
+However, the reverse
+\begin{align*}
+f(n) \in \Theta(g(n)) \Rightarrow f(n) = g(n) \cdot (1+o(1)) 
+\end{align*}
+is not true. As a counter example, consider $f(n) = n$ and $g(n) = 2n$.
+
+\section{A bit of history}
+\label{sec:historyOfGreedy}
+The history of proving the remarkable property of the two-choice paradigm (see Theorem \ref{theorem:greedy}) is in itself quite interesting.
+It was first proven by Azar et al.
+\cite{ABKU99} for the case $m = n$.
+In the same paper, they claimed that the bound holds for arbitrary $m$.
+Unfortunately, their analysis breaks down for the heavily loaded case, where $m \in \Omega\left(n  \log n\right)$.
+Berenbrink at.
+all \cite{BFZR08} were the first to proof that the same bound indeed holds for the heavily loaded case.
+Their proof is fairly complex and a simpler analysis may be found in \cite{TW13}, which also provides a tight lower bound for the heavily loaded case.
+Curiously, the case $d = 2$ was implicitly proven by Karp et al.
+in the context of simulating PRAMs on distributed memory machines\cite{KLM92}.
+
+As discussed in Section \ref{sec:AlwaysGoLeft}, V\"ocking \cite{VOC03} was able to slightly improve the bound as well as to show that the \emph{Always-go-left} strategy that he introduces is optimal up to additive constants.
+
 %bibliography

### 2014-02-14 article/article.tex
use align* instead of align

@@
 \newcommand\load{L_{\mathrm{max}}}
+\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
+% within an align* environment, number and label the equation. #1 is the label text to be appended to eqn:
+\newcommand\neqn[1]{\numberthis\label{eqn:#1}}
 
@@
 Obviously,
-\begin{align}
+\begin{align*}
 \delta = \load - \frac{m}{n}.
-\end{align}
+\end{align*}
 
@@
 \end{comment}
-\begin{align}
+\begin{align*}
 \load = \frac{\ln n}{\ln \ln n} \cdot \left(1 + o\left(1\right)\right)
-\end{align}
+\end{align*}
 for the case $m=n$, with high probability\footnote{We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] \geq 1 - n^{-\alpha}$ for an arbitrary constant $\alpha > 0$.}~\cite{RS98}.
 This implies\footnote{For a detailed proof, see Theorem \ref{theorem:asymptotic} in Appendix \ref{sec:asymptotic}.}
-\begin{align}
+\begin{align*}
 \load\in \Theta\left(\frac{\ln n}{\ln \ln n}\right).
-\end{align}
+\end{align*}
 Note that since the average load is 1, $\delta = L_{max}$ in this case. For the heavily loaded case, w.h.p.~the maximum gap is at most~\cite{RS98}.
-\begin{align}
+\begin{align*}
 \delta = \Theta\left(\sqrt{\frac{m \cdot \log n}{n}}\right).
-\end{align}
+\end{align*}
 
@@
 This simple \emph{multiple-choice balls-into-bins} scheme, which we will call the \emph{Greedy} algorithm\footnote{A detailed definition will follow in Section \ref{sec:greedy}}, achieves a gap
-\begin{align}
+\begin{align*}
 \delta = \frac{\log \log n}{\log d} + O(1)
-\end{align}
+\end{align*}
  w.h.p.~for all $m$~\cite{ABKU99}~\cite{BCSV06}.
@@
 Refer to Appendix \ref{sec:fibonacci} for the definition.}
-\begin{align}
+\begin{align*}
 \delta = \frac{\ln\ln n}{d \cdot \ln \Phi_d} + O(1)
-\end{align}
+\end{align*}
 w.h.p. Not only is this a significant reduction of the asymptotic bound, V\"ocking was able to show that the Always-go-left strategy is optimal up to additive constants \cite{VOC03}.
@@
 \item  \emph{Uniform and independent.} Choose each of the $d$ bins for a ball uniformly and independently at random from the set of bins $B$. This implies 
-\begin{align}
+\begin{align*}
 \forall \mathcal{E} \in \mathcal{F}_d: \Pr\left(\mathcal{E}\right) = \frac{1}{n ^{d}}
-\end{align}
+\end{align*}
 \item Class 2: \emph{Non-uniform and independent.} Choose the i-th bin from $B$ independently at random according to some probability function
-\begin{align}
+\begin{align*}
 \Pr: B \rightarrow \left[0,1\right]
-\end{align}
+\end{align*}
 
 \item Class 3: \emph{Non-uniform and dependent.} Choose the $d$ bins for a ball from $\Omega$ at random:
-\begin{align}
+\begin{align*}
 \Pr: B^{d} \rightarrow \left[0,1\right]
-\end{align}
+\end{align*}
 \end{compactitem}
@@
 Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, the maximum load is at most 
-\begin{align}
+\begin{align*}
 \frac{m}{n} + \frac{\log \log n}{\log d}+ O(h)
-\end{align}
+\end{align*}
 w.h.p.~at any time $t$ \cite{ABKU99} \cite{BCSV06}.
@@
 If $n$ balls are placed into $n$ bins using the Always-go-left scheme, the number of balls in the fullest bin will be 
-\begin{align}
+\begin{align*}
 \frac{\ln\ln n}{d \cdot \ln \Phi_d} + O(1)
-\end{align}
+\end{align*}
 w.h.p.
@@
 If the balls are placed into bins by the Greedy algorithm, at any time $t$ the maximum load is at most
-\begin{align}
+\begin{align*}
 \frac{\ln\ln n}{d \cdot \ln \Phi_d} + O(h)
-\end{align}
+\end{align*}
 w.h.p.
@@
 We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections \ref{sec:analysis:greedy} and \ref{sec:analysis:alg}, respectively.
-However, we use a simplifying assumption \ref{assumption:independence} in those two sections to provide a more clear presentation of the proof technique.
+However, we use a simplifying Assumption \ref{assumption:independence} in those two sections to provide a more clear presentation of the proof technique.
 Section \ref{sec:analysis:nondistinctBalls} will extend these proofs to remove the simplifying assumption we make initially.
@@
 \label{sec:analysis:greedy}
-In this Section, we will give a detailed proof for Theorem \ref{theorem:greedy}.
-However, we will restrict ourselves to the case where $m = n$ and will make use of the symplifying assumption \ref{assumption:independence}.
 
@@
 Each node $v$ represents a ball $b_v = \mathrm{ball}(v)$ .
+The nodes and edges of the witness tree represent events that may occur or not.
 The same ball may be represented by several nodes.
-The nodes and edges of the witness tree represent events that may occur or not.
-For simplicity, we assume that all events represented by a witness tree are stochastically independent.
+However, initially we will make use of the following assumption in the analysis.
+\begin{assumption}
+\label{assumption:independence}
+All events represented by the nodes and edges of a witness tree are stochastically independent. 
+\end{assumption}
+In other words, we assume that no ball is referenced by more than one node of a witness tree.
+Obviously, a correct analysis will have to consider balls assigned to more than one node, but this simplifying assumption is useful to highlight the differences between the symmetric and asymmetric allocation schemes.
+Section \ref{sec:analysis:nondistinctBalls} will extend the analysis to witness trees with redundant balls.
 
@@
 \end{compactitem}
+The definition for leaf events refers to bins containing some number of balls at some specified time.
+To avoid dependencies, the crucial trick is that edge events are defined in terms of the alternative locations of balls, not by their finally assigned bin.
+
+We say that an edge or leaf event is \emph{activated} if the associated event occurs. A witness tree is activated when all its edge and leaf events are activated.
 
@@
 The ball $b_r$ that is represented by the root node $r$ has to exist at time step\footnote{As defined in Theorem \ref{theorem:greedy}} $t$ and each ball $b_v$ represented by a node $v \neq r$ and parent node $u$ has to exist when the ball $b_u = \mathrm{ball}(u)$ is inserted.
-
-The definition for leaf events refers to bins containing some number of balls at some specified time.
-To avoid dependencies, the crucial trick is that edge events are defined in terms of the alternative locations of balls, not by their finally assigned bin.
 
@@
 
-\subsubsection{Construction}
-\label{sec:analysis:constructionSymWT}
+\begin{lemma}
+When using the Greedy algorithm to put balls into bins and arbitrary $L > 0$, a bin containing more than $L+3$ balls implies the activation of a symmetric witness tree of order $L$.
+\end{lemma}
+
+\begin{proof}
 We construct a witness tree for the event that some bin contains $L+4$ balls as follows:
@@
 This implies the activation of all leaf nodes of the witness tree.
+\end{proof}
 
@@
 In this section, we derive the probability that some witness tree of order $L$ is activated, which gives an upper bound for the probability that the maximum load is $L+3$ at some time $t$.
-The probability $p_a$ that some witness tree of order $L$ is activated is bounded by the number of witness trees $w$ times the upper bound on the probability $p_1$ that a specific witness tree is activated.
 
-Initially, we will make use of the following assumption in the analysis. 
-\begin{assumption}
-\label{assumption:independence}
-All events represented by the nodes and edges of a witness tree are stochastically independent. 
-\end{assumption}
-In other words, we assume that no ball is referenced by more than one node of a witness tree.
-Obviously, a correct analysis will have to consider balls assigned to more than one node, but this simplifying assumption is useful to highlight the differences between the symmetric and asymmetric allocation schemes.
-Section \ref{sec:analysis:nondistinctBalls} will extend the analysis to witness trees with redundant balls.
+\begin{lemma}\label{lemma:swt:activation}
+For some arbitrary $\alpha$ and large enough $L$, the activation of a symmetric witness tree of order $L$ occurs with probability at most $n^{-\alpha}$.
+\end{lemma}
+Speaking less formaly, Lemma \ref{lemma:swt:activation} states that the for large enough $L$, the activation of a symmetric witness tree is unlikely.
+
+\begin{proof}
+The probability $p_{swt}$ that some symmetric witness tree of order $L$ is activated is bounded by the number of witness trees $w$ times the upper bound on the probability $p_1$ that a specific witness tree is activated.
 
@@
 This implies a probability of at most
-\begin{align}
+\begin{align*}
 p_e = \left(\frac{d}{n}\right)^{k-1}
-\end{align} 
+\end{align*} 
 or the event that all $k-1$ edges of a witness tree are activated.
@@
 The number of those bins is bounded by $\lfloor n/3 \rfloor$ at all time steps $t$, implying a probability of at most
-\begin{align}
+\begin{align*}
 \frac{n/3}{n} = \frac{1}{3}
-\end{align}
+\end{align*}
 that one such bin is chosen.
@@
 Thus, the probability that an activated symmetric witness tree exists at time $t$ is at most 
-\begin{align}
+\begin{align*}
 p_{swt} &= w \cdot p_e \cdot p_l \\
@@
 		&= n \cdot d^{k-1} \cdot 3^{-d \cdot l}
-\end{align}
+\end{align*}
 Using $k \leq 2l$ we get
-\begin{align}
+\begin{align*}
 p_{swt} &\leq n \cdot d^{2\cdot l} \cdot 3^{-d \cdot l} \\
 		&= n \cdot \left(d^2\right)^l \cdot \left( 3^{-d} \right)^l
-\end{align}
+\end{align*}
 which we simplify by using $2d^2 \leq 3^d$
-\begin{align}
+\begin{align*}
 p_{swt} &\leq n \cdot \left(\frac{3^d}{2}\right)^l \cdot \left( 3^{-d}\right)^l \\
@@
 		&= n\cdot 2^{-d^L}		
-\end{align}
+\end{align*}
 
 For any constant $\alpha > 0 $, let
-\begin{align}
-L &\geq \log_d\log_2 n + \log_d\left(1+\alpha\right)
-\end{align}
-and note that
-\begin{align}
-L &\geq  \frac{\ln\log_2 n}{\ln d} + \frac{\ln(1+\alpha)}{\ln d} \\
-   &= \frac{\ln \log_2 n}{\ln d} \cdot \left(1 + \frac{\ln\left(1+\alpha\right)}{\ln \log_2 n}\right) \\
-  &=  \frac{\ln \ln n}{\ln d} \cdot \left( 1+ o\left( 1\right)\right)
-\end{align}
+\begin{align*}
+L &\geq \log_d\log_2 n + \log_d\left(1+\alpha\right) \neqn{swt:chooseL}
+\end{align*}
 
 Then, 
-\begin{align}
+\begin{align*}
 p_{swt} &\leq n \cdot 2^{{-d}^{\log_d\log_2 n + \log_d\left(1+\alpha\right)}} \\
@@
 		&= n \cdot \left( 2^{-\log_2 n \cdot \left(1+ \alpha \right)} \right) \\
-		&= n \cdot n^{-\left(1+\alpha\right)} = n^{-\alpha}
-\end{align}
+		&= n \cdot n^{-\left(1+\alpha\right)} = n^{-\alpha} \neqn{swt:result}
+\end{align*}
+Thus, for any constant $\alpha$ and large enough $L$ as specified in Equation \ref{eqn:swt:chooseL} the probability that a symmetric witness tree of order $L$ is activated is bounded above by $p_{swt }\leq n^{-\alpha}$.
+\end{proof}
+
+Since the probabilty for the activation of a symmetric witness tree of order $L$ upper-bounds the probability of the bad event that some bin contains more than $L+3$ balls, Equation \ref{eqn:swt:result} also is an upper bound for the latter.
+Note that Equation \ref{eqn:swt:chooseL} can be written as
+\begin{align*}
+L &\geq  \frac{\ln\log_2 n}{\ln d} + \frac{\ln(1+\alpha)}{\ln d} \\
+   &= \frac{\ln \log_2 n}{\ln d} \cdot \left(1 + \frac{\ln\left(1+\alpha\right)}{\ln \log_2 n}\right) \\
+  &=  \frac{\ln \ln n}{\ln d} \cdot \left( 1+ o\left( 1\right)\right) \neqn{swt:boundL}
+\end{align*}
+This proves Theorem \ref{theorem:greedy} for the case $m=n$, under the assumption that all events are stochastically independent (Assumption \ref{assumption:independence}). \qed
 
@@
 
+\begin{lemma}
+When using the Always-go-left algorithm to put balls into bins and arbitrary $L > 0$, a bin containing more than $L+3$ balls implies the activation of an asymmetric witness tree of order $L$.
+\end{lemma}
+\begin{proof}
 For each node we define a label $(h, i)$, where $h, i \in \mathbb{N}$, $0 \leq h \leq L$ and  $1\leq i \leq d$. 
@@
 \end{compactitem}
-Thus, our the root node of our witness tree will be labeled $(L, 1)$. Note that the labels are unique within the set of children of a node $v$, but not for all nodes of the tree.
-
-\subsubsection{Construction}
-\label{sec:analysis:constructionAsymWT}
-The labels defined in the previous section are useful to state the following invariant, which will be maintained while assigning balls to the nodes of the witness tree.
+Thus, our the root node of our witness tree will be labeled $(L, 1)$.
+Note that the labels are unique within the set of children of a node $v$, but not for all nodes of the tree.
+The labels are useful to state the following invariant, which will be maintained while assigning balls to the nodes of the witness tree.
 
@@
 \end{compactitem}
+\end{proof}
 
@@
 \label{sec:analysis:probabilityAsymWT}
+\begin{lemma}\label{lemma:awt:activation}
+For some arbitrary $\alpha$ and large enough $L$, the activation of an asymmetric witness tree of order $L$ occurs with probability at most $n^{-\alpha}$.
+\end{lemma}
+
+\begin{proof}
 Analogously to the symmetric witness tree, the possible number of asymmetric witness trees is upper-bounded by $w=n^k$ and the probability that all edge events occur is\footnote{As in Section \ref{sec:analysis:greedy}, $k$ denotes the number of nodes of the witness tree and $l$ the number of leaves.}
-\begin{align}
+\begin{align*}
 p_e = \left(\frac{d}{n}\right)^{k-1}
-\end{align}
+\end{align*}
 
@@
 Since no more than $n/3$ bins can contain 3 or more balls, the values $b_i$ are subject to the constraint
-\begin{align}
+\begin{align*}
 \frac{n}{d}\sum_{i=1}^d \beta_i \leq \frac{n}{3}
-\end{align}
+\end{align*}
 and thus 
-\begin{align}
+\begin{align*}
 \sum_{i=1}^d \beta_i \leq \frac{d}{3}
-\end{align}
+\end{align*}
 
 The probability that all $d$ locations of a ball point to a bin containing 3 or more balls is then 
-\begin{align}
+\begin{align*}
 p_{l_1} = \prod_{i=1}^{d} \beta_i
-\end{align}
+\end{align*}
 In other words, a leaf event occurs with probability $p_{l_1}$, which is maximized when all $b_i$ are equal.
 We thus set $b_i \leq 1/3$ for all $i$ and get as upper bound for the probability that all leaf events occur
-\begin{align}
+\begin{align*}
 p_l \leq 3^{-d\cdot l}
-\end{align}
+\end{align*}
 
 The above results imply the same probability for the activation of an asymmetric witness tree as we derived in Section \ref{sec:analysis:probabilitySymWT} for the activation of a symmetric witness tree, that is
-\begin{align}
+\begin{align*}
 p_{awt} &\leq n \cdot 2^{-l}
-\end{align}
+\end{align*}
 However, the number of leaves is larger, since 
-\begin{align}
+\begin{align*}
 l \geq \Phi_d^{d\cdot L -1}
-\end{align}
+\end{align*}
 holds\footnote{see Section \ref{sec:analysis:definitionAsymWT} for details}.
@@
 For any constant $\alpha > 0$, let 
-\begin{align}
-L &\geq \left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d}}\right\rceil+\frac{1}{d}
-\end{align}
-\todo{the paper uses $+1$, instead of $+1/d$, but I can't make workout the below result then} and note that 
-\begin{align}
-L &\geq \frac{\ln\log_2 n}{d\cdot \ln \Phi_d} + \frac{\ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d} + \frac{1}{d}\\
-  &\geq \frac{\ln\ln n}{d\cdot \ln \Phi_d} + O\left(1\right)
-\end{align}
-
+\begin{align*}
+L &\geq \left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d}}\right\rceil+\frac{1}{d} \neqn{awt:chooseL}
+\end{align*}
+\todo{the paper uses $+1$, instead of $+1/d$, but I can't make workout the below result then} 
 Then, 
-\begin{align}
+\begin{align*}
 p_{awt} &\leq n \cdot 2^{-\Phi_d^{d\cdot L -1}} \\
@@
          &= n \cdot 2^{-\log_2 n \cdot\left(1+\alpha\right)} \\
-         &= n \cdot n ^{-(1+\alpha)} = n^{-\alpha}
-\end{align}
+         &= n \cdot n ^{-(1+\alpha)} = n^{-\alpha} \neqn{awt:result}
+\end{align*}
+Thus, for any constant $\alpha$ and large enough $L$ as specified in Equation \ref{eqn:awt:chooseL} the probability that a symmetric witness tree of order $L$ is activated is bounded above by $p_{awt }\leq n^{-\alpha}$.
+\end{proof}
+
+Since the probabilty for the activation of an asymmetric witness tree of order $L$ upper-bounds the probability of the bad event that some bin contains more than $L+3$ balls, Equation \ref{eqn:awt:result} also is an upper bound for the latter.
+Note that Equation \ref{eqn:awt:chooseL} can be written as
+\begin{align*}
+L &\geq \frac{\ln\log_2 n}{d\cdot \ln \Phi_d} + \frac{\ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d} + \frac{1}{d}\\
+  &\geq \frac{\ln\ln n}{d\cdot \ln \Phi_d} + O\left(1\right)
+\end{align*}
+This proves Theorem \ref{theorem:agln}, under the assumption that all events are stochastically independent (Assumption \ref{assumption:independence}). \qed
 
@@
 \label{sec:analysis:FullWT}
-Let $gamma \geq 2$ be a suitable constant.
+Let $\gamma \geq 2$ be a suitable constant.
 The root node of a \emph{full witness tree} has $\gamma$ children, denoted by $v_{c_i}$ with $1\leq i \leq \gamma$.
@@
 The root thus has $\gamma$ grandchildren, each of which is the root of a witness tree of order $L$.
-The time constraints activation of a full witness tree is defined analogously to the previous variants.
+The time constraints and activation of a full witness tree is defined analogously to the previous variants.
 
@@
 \label{sec:analysis:probabilityFullWT}
+\begin{lemma}\label{lemma:fwt:activation}
+For some suitable $\gamma \geq 2$ and large enough $L$, the activation of a prunded witness tree occurs with probability at most $n^{-\gamma +1 +o(1)}$.
+\end{lemma}
 If there are less than $\gamma$ cut-off edges, in at least one of the $\gamma$ symmetric or asymmetric witness trees contained in the full witness tree the referenced balls are distinct.
@@
 Thus, it remains to find an upper-bound for the activation of a full witness tree with $\gamma$ cut-off edges.
-Let $q$ be the number of references to balls by the full witness tree\footnote{Since no ball was assigned to the root node, the full witness tree consists of $q+1$ nodes} and assume that $q  \leq 2\cdot \gamma\left(\alpha+1\right)\cdot \log_2 n$, for a suitable value of $\gamma$.
-Then, there are at most $w_p=q^\gamma$ possibilities to transform a full witness tree with $q+1$ nodes to a pruned witness tree with $\gamma$ cut-off edges.
+Let $K$ be the number of references to balls by the full witness tree\footnote{Since no ball was assigned to the root node, the full witness tree consists of $K+1$ nodes} and assume that $K  \leq 2\cdot \gamma\left(\alpha+1\right)\cdot \log_2 n$, for some arbitrary $\alpha > 0$.
+Then, there are at most $w_p=K^\gamma$ possibilities to transform a full witness tree with $K+1$ nodes to a pruned witness tree with $\gamma$ cut-off edges.
 Fix the shape of the tree to be one of those.
-Let $k$ be the number of balls referenced by the pruned witness tree and $l$ be the number of its leaf nodes.
 
@@
 
-By the same arguments as in Sections \ref{sec:analysis:probabilitySymWT} and \ref{sec:analysis:probabilityAsymWT}, we can conclude that there are at most $w=n^k$ assignments of balls to the nodes of the witness tree.
+Let $k$ be the number of balls referenced by the pruned witness tree and $l$ be the number of its leaf nodes.
+Then, by the same arguments as in Sections \ref{sec:analysis:probabilitySymWT} and \ref{sec:analysis:probabilityAsymWT}, we can conclude that there are at most $w=n^k$ assignments of balls to the nodes of the witness tree.
 Furthermore, the probability for the activation of all edge events is bounded from above by
-\begin{align}
+\begin{align*}
 p_e = \left(\frac{d}{n}\right)^{k}
-\end{align}
+\end{align*}
 while the probability for the activation of all leaf events is bounded by 
-\begin{align}
+\begin{align*}
 p_l \leq 3^{-d\cdot l}
-\end{align}
+\end{align*}
 
@@
 By definition (Section \ref{sec:analysis:pruningFullWT}), a cut-off edge $e=(v, u)$ represents the event that the ball $b_u$ represented by node $u$ shares a common location $i$ with another ball $b_{u'}$, represented by node $u'$.
-Since $u'$ has already been visited by the time the breadth-first traversal reaches node $v$, it is part of the pruned witness tree and was assigned one of $k\leq q$ balls.
+Since $u'$ has already been visited by the time the breadth-first traversal reaches node $v$, it is part of the pruned witness tree and was assigned one of $k\leq K$ balls.
 The probability that location $i$ of ball $b_u$ points to one of the $d$ locations of ball $b_{u'}$ is at most $d/n$.
 The probability that there are $\gamma$ cut-off edges is thus at most
-\begin{align}
+\begin{align*}
 p_c &= \left(k\cdot \frac{d}{n}\right)^\gamma \\
-    & \leq \left(q\cdot \frac{d}{n}\right)^\gamma
-\end{align}
+    & \leq \left(K\cdot \frac{d}{n}\right)^\gamma
+\end{align*}
 
 Putting it all together, we get 
-\begin{align}
+\begin{align*}
 p_{pwt} &= w_p\cdot w \cdot p_e \cdot p_l \cdot p_c \\
-		&= q\gamma\cdot n^{k+1}\cdot \left(\frac{d}{n}\right)^k\cdot3^{-d\cdot l} \cdot \left(q \frac{d}{n}\right)^\gamma\\
-		&= n \cdot d^k\cdot 3^{-d\cdot l} \cdot \left(q^2 \frac{d}{n} \right)^\gamma 
-\end{align}
+		&= K\gamma\cdot n^{k+1}\cdot \left(\frac{d}{n}\right)^k\cdot3^{-d\cdot l} \cdot \left(K \frac{d}{n}\right)^\gamma\\
+		&= n \cdot d^k\cdot 3^{-d\cdot l} \cdot \left(K^2 \frac{d}{n} \right)^\gamma 
+\end{align*}
 
-Using $k \leq 2l$, $d^2 \leq 3^d$ and $q \leq 2\gamma \cdot \left( \alpha +1 \right) \cdot \log_2 n$ this becomes
-\begin{align}
+Using $K \leq 2l$, $d^2 \leq 3^d$ and $K \leq 2\gamma \cdot \left( \alpha +1 \right) \cdot \log_2 n$ this becomes
+\begin{align*}
 p_{pwt} &\leq n\cdot \left(\frac{\left(2\gamma \cdot \left(\alpha +1 \right) \cdot \log_2 n\right)^2\cdot d}{n}\right)^\gamma \\
-        &= n^{\gamma +1 1} \cdot n ^{\log_n\left( \left(2\gamma\cdot \left(\alpha+1\right)\cdot \log_2n \right)^{2\gamma} \cdot d^\gamma \right)}\\
+        &= n^{-\gamma +1} \cdot n ^{\log_n\left( \left(2\gamma\cdot \left(\alpha+1\right)\cdot \log_2n \right)^{2\gamma} \cdot d^\gamma \right)}\\
         &= n^{-\gamma+1+o\left(1\right)}
-\end{align}
-
-%^{\log_n \left(2\gamma \cdot \left( \alpha +1 \right) \cdot \log_2 n\right)^{2\gamma}\cdot d^\gamma \right)}
+\end{align*}
+\todo{Where are our $L$ and $\alpha$ here?}
+\qed
 
@@
 
-\begin{align}
+\begin{align*}
 F_d(k) = \sum_{i=1}^{d}F_d(k-i)
-\end{align}
+\end{align*}
 Furthermore, define 
-\begin{align}\label{eqn:goldenFib}
-\Phi_d = \lim_{k \rightarrow \infty} \sqrt[k]{F_d(k)}
-\end{align}
+\begin{align*}
+\Phi_d = \lim_{k \rightarrow \infty} \sqrt[k]{F_d(k)} \neqn{goldenFib}
+\end{align*}
 Notice that $F_2$ corresponds to the usual Fibonacci numbers, while $\Phi_2 $ corresponds to the golden ratio \cite{Knuth73}, which is generalized to $\Phi_d$, for which 
-\begin{align}
+\begin{align*}
 i, j \in \mathbb{N}, i < j \Rightarrow \Phi_i < \Phi_j
-\end{align}
+\end{align*}
 and 
-\begin{align}
+\begin{align*}
 \lim_{d\rightarrow \infty} \Phi_d = 2
-\end{align}
+\end{align*}
 hold.
@@
 \end{theorem}
-The proof is straight forward:
+\begin{proof}
 \begin{align*}
@@
 is not true. As a counter example, consider $f(n) = n$ and $g(n) = 2n$.
+\end{proof}
 

### 2014-02-14 article/article.tex
conclusion

@@
 \begin{align*}
-\delta = \Theta\left(\sqrt{\frac{m \cdot \log n}{n}}\right).
+\delta = \Theta\left(\sqrt{\frac{m \cdot \ln n}{n}}\right).
 \end{align*}
@@
 \label{sec:conclusion}
+Balls-into-bins games have many applications in Computer Science and have been studied extensively.
+The single-choice balls-into-bins strategy chooses a bin uniformly and independently at random for each ball.
+The classical result for this scheme upper-bounds the maximum load with
+\begin{align*}
+\load = \frac{m}{n} + \Theta\left(\sqrt{\frac{m \cdot \log n}{n}}\right).
+\end{align*}
+w.h.p.
+In this paper we have presented two different miltiple-choice balls-into-bins placement algorithms: The Greedy-scheme due to Azar et al.\cite{ABKU99}, which achieves a maximum load of 
+\begin{align*}
+\load = \frac{m}{n} + \frac{\ln\ln n}{\ln d} + O(1)
+\end{align*}
+w.h.p.~and the Always-go-left scheme due to V\"ocking \cite{VOC03}, which achieves a slightly better maximum load of
+\begin{align*}
+\load = \frac{m}{n} + \frac{\ln \ln n}{d\cdot \ln \Phi_d} + O(1)
+\end{align*}
+w.h.p.~for all $m$. We used \emph{witness trees}, a very useful technique in the balls-into-bins context, to prove these upper-bounds in detail.
 

### 2014-02-14 article/article.tex
abstract

@@
 \begin{abstract}
-
+Balls-into-bins is a classical allocation problem, where $m$ balls have to be distributed among $n$ bins as evenly as possible.
+Using ranomized allocations schemes, the placement algorithm does not need to know the current load of all bins when the next ball has to be allocated.
+A classical and well studied approach is to select a bin uniformly and independently at random for each ball.
+Another strategy is to select $d\geq 2$ bins for each ball and put the ball into the bin with the least load.
+Somewhat surprisingly, the maximum load achieved by such a multiple-choice scheme is significantly smaller, even for $d=2$.
+In this paper, we will present two different multiple-choice allocation strategies for the balls-into-bins problem and provide a detailed analysis for the upper bounds on the maximum load that they achieve.
+We will use a technique called witness trees that has proven useful in a wide variaty of balls-into-bins problems.
 \end{abstract}

### 2014-02-14 article/article.tex
some more fixes

--- 2014-02-14 article/article.tex
@@
 \begin{abstract}
-
+Balls-into-bins is a classical allocation problem, where $m$ balls have to be distributed among $n$ bins as evenly as possible.
+A well studied approach is to select a bin uniformly and independently at random for each ball.
+Another strategy is to select $d\geq 2$ bins for each ball and put the ball into the bin with the least load.
+Somewhat surprisingly, the maximum load achieved by such a multiple-choice scheme is significantly smaller, even for $d=2$.
+In this paper, we will present two different multiple-choice allocation strategies for the balls-into-bins problem and provide a detailed analysis of the upper bounds on the maximum load that they achieve.
+We will use a technique called witness trees, which has proven useful in a wide variety of balls-into-bins problems.
 \end{abstract}
@@
 \label{sec:intro}
-Suppose that $m$ balls are thrown into $n$ bins, according to some arbitrary placement strategy.
-We are interested in the \emph{number of balls in the fullest bin}, also called the \emph{maximum load}.
-Equivalently, one can look at the \emph{additive gap} of the allocation, that is the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
-These \emph{balls-into-bins games} or \emph{allocation problems} have been studied extensively in the probability literature (see e.g., \cite{JK77}).
-In the context of Computer Science, the balls-into-bins model is useful to study, e.g., strategies for distributing \emph{tasks} (balls) to \emph{processors} (bins) as evenly as possible without having to query all processors for the number of tasks currently assigned to them.
+Suppose that $m$ balls are to be thrown into $n$ bins, according to some placement strategy.
+We are interested in the number of balls in the fullest bin, also called the \emph{maximum load}.
+Equivalently, one can look at the \emph{additive gap} of the allocation, i.e., the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
+These balls-into-bins games (BiB) have been studied extensively (see, e.g., \cite{JK77}).
+In the context of Computer Science, the balls-into-bins model is useful to study, e.g., strategies for distributing tasks (balls) to processors (bins) as evenly as possible without having to query all the processors for the number of tasks currently assigned to them.
 
@@
 \label{sec:model}
-Although several different variants of this simple balls-into-bins game exist, we will assume a very basic model to highlight the crucial ideas in both placement strategies as well as the analysis of them.
+Although several different variants of this balls-into-bins game exist, we will focus on a very basic model to highlight the crucial ideas in the placement strategies aand their analysis.
 In our model, we start with $n$ bins and put balls into them, one per time step $t$.
 We assume that at any time step $t$ at most $m = h n$ balls are in the bins, where $h$ is some constant greater or equal to $1$.
-At each time step $t$, we delete one randomly chosen ball from the bins.
-While most of the paper and especially the analysis focuses on the lightly loaded case where $m=n$, we will give results for the case $m > n$ as well.
-Throughout the paper, we will denote the maximum load by $\load$ and the additive gap by $\delta$.
-Obviously,
+At each time step $t$, we delete one ball from the bins.
+While most of the paper and the analyses focuses on the lightly loaded case where $m=n$, we will give results for the case $m > n$ as well.
+Throughout the paper, we will denote the maximum load by $\load$ and the additive gap by $\delta$:
 \begin{align*}
@@
 
-\subsection{Single-choice balls-into-bins}
+\subsection{Single-choice BiB}
 \label{sec:single-choice}
@@
 
-\subsection{Multiple-choice balls-into-bins}
+\subsection{Multiple-choice BiB}
 Consider a variation of the above \emph{single-choice balls-into-bins} strategy: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}\footnote{If there are multiple bins with the same minimal load, chose one among them arbitrarily.}.
@@
 Then $e$ represents the event that the i-th location of the ball $b_u = \mathrm{ball}(u)$ points to the same bin as one of the locations of the ball $b_v = \mathrm{ball}(v)$.
-\todo{illustrate}
+%$\todo{illustrate}
 \item \emph{Leaf event:} A leaf node $v$ of the witness tree represents the event that each of the $d$ locations of the ball $b_v = \mathrm{ball}(v)$ points to a bin that contains at least three additional balls (i.e., balls that are not represented by a node of the tree) at the time of the insertion of the ball $b_v$.
-\todo{illustrate}
+%\todo{illustrate}
 \end{compactitem}
@@
 \end{compactitem}
-\todo{illustrate and note difference to usual Fibonacci trees}
+%\todo{illustrate and note difference to usual Fibonacci trees}
 
@@
 \item For $1\leq j < i$, the $j$'th location of the ball $b_v$ points to a bin containing at least $l+4$ balls at the insertion time of the ball $b_v$.
-If the $j$'th location contained less balls than the $i$'th location, the Always-go-left scheme would have put the ball $b_v$ into the bin pointed to by location $j$.
+If the j-th location contained less balls than the i-th location, the Always-go-left scheme would have put the ball $b_v$ into the bin pointed to by location $j$.
 The topmost ball in location $j$ is assigned to the  child $u$ of $v$ with label $(l,j)$.
-\item For $i < j \leq d$, the $j$'th location contains at least $l+3$ balls.
+\item For $i < j \leq d$, the j-th location contains at least $l+3$ balls.
 If location $j$ contained less, the Always-go-left scheme would have put the ball $b_v$ into the bin it points to.
@@
 \end{align*}
-\todo{the paper uses $+1$, instead of $+1/d$, but I can't make workout the below result then} 
+\todo{the paper uses $+1$, instead of $+1/d$, but I can't workout the below result then} 
 Then, 
@@
 Since the ball $b$ was placed into bin $x$, at least one of its locations points to this bin $x$, which we call location $i$.
-Let $j = i+1 mod d$ be the ball's $j$'th location.
-Then, at the time of insertion of ball $b_c$, its $j$'th location points to a bin $y$ with at least $L+3$ other balls in it.
+Let $j = i+1 mod d$ be the ball's j-th location.
+Then, at the time of insertion of ball $b_c$, its j-th location points to a bin $y$ with at least $L+3$ other balls in it.
 We assign the topmost ball in bin $y$ to node $v_g$.
@@
 We continue this pruning until either the breath-first traversal terminates or until we have obtained $\gamma$ cut-off edges.
-Note that more than $\gamma$ cut-off edges cannot exist, \todo{why?}.
+Note that more than $\gamma$ cut-off edges cannot exist.
 When finished, the \emph{pruned witness tree} does not contain any ball twice and thus all its events are independent.
@@
 \bibliographystyle{acm}
-\bibliography{../sources} 
+\bibliography{sources} 
 \end{document}

### 2014-02-21 article/article.tex
use less \emph

@@
 These balls-into-bins games (BiB) have been studied extensively (see, e.g., \cite{JK77}).
-In the context of Computer Science, the balls-into-bins model is useful to study, e.g., strategies for distributing tasks (balls) to processors (bins) as evenly as possible without having to query all the processors for the number of tasks currently assigned to them.
+In the context of Computer Science, the BiB model is useful to study, e.g., strategies for distributing tasks (balls) to processors (bins) as evenly as possible without having to query all the processors for the number of tasks currently assigned to them.
 
-\subsection{Balls-into-bins model}
+\subsection{BiB model}
 \label{sec:model}
-Although several different variants of this balls-into-bins game exist, we will focus on a very basic model to highlight the crucial ideas in the placement strategies aand their analysis.
-In our model, we start with $n$ bins and put balls into them, one per time step $t$.
-We assume that at any time step $t$ at most $m = h n$ balls are in the bins, where $h$ is some constant greater or equal to $1$.
+Although several different variants BiB exist, we will focus on a very basic model to highlight the crucial ideas in the placement strategies and their analyses.
+In our model, we start with $n$ \todo{empty?}  bins and put balls into them, one per time step $t$.
+We assume that at any time $t$ at most $m = h n$ balls are in the bins, where $h$ is some constant greater or equal to $1$.
 At each time step $t$, we delete one ball from the bins.
@@
 \label{sec:single-choice}
-A simple strategy for the balls-into-bins problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice balls-into-bins} strategy.
+A simple strategy for the BiB problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice} strategy.
 Using this strategy, the maximum load of the fullest bin will be
@@
 \begin{align*}
-\load = \frac{\ln n}{\ln \ln n} \cdot \left(1 + o\left(1\right)\right)
+\load = \frac{\ln n}{\ln \ln n}  \left(1 + o\left(1\right)\right)
 \end{align*}
-for the case $m=n$, with high probability\footnote{We say that an event $\mathcal E$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E \right] \geq 1 - n^{-\alpha}$ for an arbitrary constant $\alpha > 0$.}~\cite{RS98}.
-This implies\footnote{For a detailed proof, see Theorem \ref{theorem:asymptotic} in Appendix \ref{sec:asymptotic}.}
+for the case $m=n$, with high probability\footnote{We say that an event $\mathcal E_n$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E_n \right] \geq 1 - n^{-\alpha}$ for some arbitrary constant $\alpha > 0$.}~\cite{RS98}.
+This implies (see Theorem \ref{theorem:asymptotic} in Appendix \ref{sec:asymptotic})
 \begin{align*}
@@
 \end{align*}
-Note that since the average load is 1, $\delta = L_{max}$ in this case. For the heavily loaded case, w.h.p.~the maximum gap is at most~\cite{RS98}.
+Note that, since the average load is 1, $\delta = L_{max}$ in this case. For the heavily loaded case, w.h.p.~the maximum gap is ~\cite{RS98}
 \begin{align*}
-\delta = \Theta\left(\sqrt{\frac{m \cdot \ln n}{n}}\right).
+\delta = \Theta\left(\sqrt{\frac{m  \ln n}{n}}\right).
 \end{align*}
@@
 \subsection{Multiple-choice BiB}
-Consider a variation of the above \emph{single-choice balls-into-bins} strategy: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}\footnote{If there are multiple bins with the same minimal load, chose one among them arbitrarily.}.
-This simple \emph{multiple-choice balls-into-bins} scheme, which we will call the \emph{Greedy} algorithm\footnote{A detailed definition will follow in Section \ref{sec:greedy}}, achieves a gap
+Consider a variation of the above single-choice BiB strategy: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin.
+If there are multiple bins with the same minimal load, chose one among them uniformly and independently at random.
+This simple multiple-choice scheme, which we will call the Greedy algorithm, achieves a gap
 \begin{align*}
@@
 \end{comment}
-Extensive research following the influential paper by Azar et al.~\cite{ABKU99}, who were the first to explicitly prove this phenomenon\footnote{For a brief history of the proof, which is quite interesting by itself, refer to Appendix \ref{sec:historyOfGreedy}.}, suggests that this \emph{two-choice paradigm} is not just an artifact of the simple balls-into-bins model, but a general principle applicable to a variety of problems.
+Extensive research following the influential paper by Azar et al.~\cite{ABKU99}, who were the first to explicitly prove this phenomenon\footnote{For a brief history of the proof, which is quite interesting by itself, refer to Appendix \ref{sec:historyOfGreedy}.}, suggests that this two-choice paradigm is not just an artifact of the simple BiB model, but a general principle applicable to a variety of problems.
 
-Consider a slightly more involved strategy, which we will call the \emph{Always-go-left} scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size\footnote{For the sake of a clear presentation, we will assume that $d$ divides $n$ evenly in the remainder of this paper.} $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$.
+Consider a slightly more involved strategy, which we will call the Always-go-left scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size\footnote{For the sake of a clear presentation, we will assume that $d$ divides $n$ evenly in the remainder of this paper.} $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$.
 We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
-In case a tie occurs, put the ball into the \emph{left-most} bin.
+In case a tie occurs, put the ball into the left-most bin.
 That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with the same minimal load, we put the ball into the bin with smallest index $i$.
-Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing \emph{asymmetric tie-breaking} in combination with \emph{non-uniform} bin selection reduces the bound for the gap to\footnote{The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers.
+Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing asymmetric tie-breaking in combination with non-uniform bin selection reduces the bound for the gap to\footnote{The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers.
 Refer to Appendix \ref{sec:fibonacci} for the definition.}
 \begin{align*}
-\delta = \frac{\ln\ln n}{d \cdot \ln \Phi_d} + O(1)
+\delta = \frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
 \end{align*}
@@
 
-This paper is organized as follows: We give a basic classification of balls-into-bins games in Section \ref{sec:classification}, to provide some overview of the different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}.
-In the remaining sections of this paper we will consider the model given in Section \ref{sec:model}, which can be classified as a \emph{sequential, finite and dynamic process with random deletions}.
-Section \ref{sec:multiple-choice} starts with a brief definition of three different classes of random placement strategies and then continues with the presentation of the two placement algorithms outlines in the introduction: The  \emph{Greedy} algorithm (Section \ref{sec:greedy}) and the \emph{Always-go-left} algorithm (Section \ref{sec:AlwaysGoLeft}).
+This paper is organized as follows: We give a basic classification of BiB games in Section \ref{sec:classification}, to provide some overview of the different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}.
+In the remaining sections of this paper we will consider the model given in Section \ref{sec:model}, which can be classified as a sequential, finite and dynamic process with random deletions.
+Section \ref{sec:multiple-choice} starts with a brief definition of three different classes of random placement strategies and then continues with the presentation of the two placement algorithms outlines in the introduction: The  Greedy algorithm (Section \ref{sec:greedy}) and the Always-go-left algorithm (Section \ref{sec:AlwaysGoLeft}).
 Section \ref{sec:analysis} is the core of this paper and will give a detailed analysis of both strategies.
-We will use a proof technique called \emph{witness trees}, which allows for a clean analysis of both problems.
-Note that the technique is applicable to a wide variety of balls-into-bins scenarios, making it the most useful of several proof techniques applied to balls-into-bins problems.
-Other proof techniques such as \emph{layered induction} and a \emph{differential equations approach} have been used.
+We will use a proof technique called witness trees, which allows for a clean analysis of both problems.
+Note that the technique is applicable to a wide variety of BiB scenarios, making it the most useful of several proof techniques applied to BiB problems.
+Other proof techniques such as layered induction and a differential equations approach have been used.
 They are however not part of this paper.
-See, e.g., \cite{MRS01} for an overview of different proof techniques for the balls-into-bins problem.
+See, e.g., \cite{MRS01} for an overview of different proof techniques for the BiB problem.
 
@@
 These result may sometimes seem surprising, which is why we will denote a significant part of this article to demonstrate proofs in detail.
-This proof technique may be useful for devising asymptotic bounds for the additive gap of different balls-into-bins scenarios and strategies.
+This proof technique may be useful for devising asymptotic bounds for the additive gap of different BiB scenarios and strategies.
 \end{comment}
@@
 \begin{compactitem}
-\item In a \emph{sequential process}, the balls arrive one after another.
+\item In a \emph{sequential} process, the balls arrive one after another.
 A  ball is dealt with only after the previous ball has been placed.
-Contrary, in a \emph{parallel process}, balls arrive in sets and the balls in one set are dealt with simultaneously \cite{ABS98}.
-\item In a \emph{static problem}, the balls are placed into the bins and no balls are ever removed.
-In a \emph{dynamic problem}, balls can also be removed from the system.
-In a simple \emph{random deletions} model, for each newly inserted ball a ball chosen uniformly and independently at random from among the balls currently in the bins is removed from the system \cite{ABKU99} \cite{MRS01}.
-More elaborate schemes exists: An \emph{oblivious adversary} specifies the sequence of insertions and deletions of balls in advance, without any knowledge of the placement of the balls \cite{CFM+98}.
-Another approach was studied by Adler et al.~\cite{ABS98} in the context of the parallel balls-into-bins problem: In each bin, the balls are stored in FIFO-order and at every time step the first ball in each of the bins is deleted.
-\item A \emph{finite process} is analyzed for a given time interval that is known beforehand, whereas for an \emph{infinite process}, a placement and a deletion process is defined over an infinite time span \cite{ABS98}.
-\item In the \emph{lightly loaded case}, $m$ is in the order of $n$, whereas in the \emph{heavily loaded case}, $m$ is arbitrarily large.
+Contrary, in a \emph{parallel} process, balls arrive in sets and the balls in one set are dealt with simultaneously \cite{ABS98}.
+\item In a \emph{static} problem, the balls are placed into the bins and no balls are ever removed.
+In a \emph{dynamic} problem, balls can also be removed from the system.
+In a simple random deletions model, for each newly inserted ball a ball chosen uniformly and independently at random from among the balls currently in the bins is removed from the system \cite{ABKU99} \cite{MRS01}.
+More elaborate schemes exists: An oblivious adversary specifies the sequence of insertions and deletions of balls in advance, without any knowledge of the placement of the balls \cite{CFM+98}.
+Another approach was studied by Adler et al.~\cite{ABS98} in the context of the parallel BiB problem: In each bin, the balls are stored in FIFO-order and at every time step the first ball in each of the bins is deleted.
+\item A \emph{finite} process is analyzed for a given time interval that is known beforehand, whereas for an \emph{infinite} process, a placement and a deletion process is defined over an infinite time span \cite{ABS98}.
+\item In the lightly loaded case, $m$ is in the order of $n$, whereas in the heavily loaded case, $m$ is arbitrarily large.
 For clarity, we will assume that $m = n$ in the lightly loaded case and that $m > n$ in the heavily loaded case.
@@
 \label{sec:applications}
-The two-choice paradigm applied to the balls-into-bins model has several useful applications, some of which we will outline here.
-The applications in \emph{Hashing} (Section \ref{sec:hashing}), \emph{Online Load Balancing} (Section \ref{sec:loadbalancing}) and \emph{Emulation of shared memory machines on distributed memory machines} \ref{sec:DMM}) share a close relation to load balancing.
-Section \ref{sec:circuitrouting} applies the balls-into-bins model and the two-choice paradigm in a different context, namely that of routing circuits in multi-stage interconnection networks with minimal congestion.
+The two-choice paradigm applied to the BiB model has several useful applications, some of which we will outline here.
+The applications in Hashing (Section \ref{sec:hashing}), Online Load Balancing (Section \ref{sec:loadbalancing}) and Emulation of shared memory machines on distributed memory machines \ref{sec:DMM}) share a close relation to load balancing.
+Section \ref{sec:circuitrouting} applies the BiB model and the two-choice paradigm in a different context, namely that of routing circuits in multi-stage interconnection networks with minimal congestion.
 
@@
 A simple hash table implementation typically uses a single hash function to map elements to entries in a table.
-If more than one element is mapped to the same entry, all elements mapped to this table entry are stored in a \emph{chain}, usually implemented as a linked list.
+If more than one element is mapped to the same entry, all elements mapped to this table entry are stored in a chain, usually implemented as a linked list.
 The maximum time to search for an element is then proportional to the length of the longest chain in the hash table.
 
-It is easy to see how the balls-into-bins idea can be applied: Assuming a perfect hash function\footnote{i.e., each of the $n$ elements is mapped to one of the $n$ table entries uniformly and independently at random}, the elements are balls to be mapped to table entries or bins, implying a worst case look-up time of $\Theta\left(\frac{\log n}{\log \log n}\right)$ w.h.p.
+It is easy to see how the BiB idea can be applied: Assuming a perfect hash function\footnote{i.e., each of the $n$ elements is mapped to one of the $n$ table entries uniformly and independently at random}, the elements are balls to be mapped to table entries or bins, implying a worst case look-up time of $\Theta\left(\frac{\log n}{\log \log n}\right)$ w.h.p.
 Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries implies a bound of $\Theta\left(\log \log n\right)$ w.h.p.
@@
 However, such a central dispatcher poses a bottleneck within a distributed system and querying $m$ servers for their current load involves significant overhead, e.g., sending a message to each server and waiting for the reply and more efficient solutions are necessary.
-As with hashing, the balls-into-bins model can be applied easily: we need to assign tasks (the balls) to servers (the bins).
+As with hashing, the BiB model can be applied easily: we need to assign tasks (the balls) to servers (the bins).
 Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends issues its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is bounded by $\Theta\left(\log \log n\right)$ w.h.p.
@@
 \label{sec:DMM}
-The \emph{Parallel Random Access Machine} model\footnote{as defined in the lecture or in~\cite{P03}}, or \emph{PRAM} for short, is a high level abstraction of a shared memory machine, in which memory access by any processor to any memory location is assumed to require constant time.
+The Parallel Random Access Machine model\footnote{as defined in the lecture or in~\cite{P03}}, or PRAM for short, is a high level abstraction of a shared memory machine, in which memory access by any processor to any memory location is assumed to require constant time.
 It is especially useful for the study of parallel algorithms.
-However, a PRAM is much harder to realize in hardware than a \emph{Distributed Memory Machine} (DMM).
+However, a PRAM is much harder to realize in hardware than a Distributed Memory Machine (DMM).
 Thus, ways to emulate a PRAM on a DMM have been studied extensively.
@@
 More accesses to the same module will result in memory contention.
-The balls-into-bins idea is useful for managing communication between the processors and the memory modules such that memory contention is minimal.
+The BiB idea is useful for managing communication between the processors and the memory modules such that memory contention is minimal.
 For details, see Karp et al.~\cite{KLM92} and the papers referenced there.
@@
 In multi-stage interconnection networks, virtual circuit-switching is used to route communication.
-For simplicity, we consider the \emph{permutation routing problem}, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes and assume a \emph{butterfly network} as the underlying interconnection network (see \cite{CLR09} for a detailed definition).
+For simplicity, we consider the permutation routing problem, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes and assume a butterfly network as the underlying interconnection network (see \cite{CLR09} for a detailed definition).
 Then, congestion is the maximum number of paths that are routed through a network link.
-The goal of a \emph{circuit routing algorithm} is to to allocate paths for all $n$ communication requests such that congestion is minimal.
+The goal of a circuit routing algorithm is to to allocate paths for all $n$ communication requests such that congestion is minimal.
 
-In this context, \emph{Valiant's paradigm} \cite{V82} states that any such permutation routing problem can be solved by transforming it into two problems: First, find a path for each request to an intermediate destination chosen uniformly and independently at random.
+In this context, Valiant's paradigm \cite{V82} states that any such permutation routing problem can be solved by transforming it into two problems: First, find a path for each request to an intermediate destination chosen uniformly and independently at random.
 Next, find a path from the intermediate to the actual destination.
-This routing technique is analogous to the single-choice balls-into-bins problem where each communication request (or ball) chooses an intermediate destination (a bin) at random.
-In this analogy, the congestion achieved by Valiant's paradigm corresponds to the maximum load in the balls-into-bins model, thus implying a bound of $\Theta\left(\log \log n\right)$ on the congestion w.h.p.
+This routing technique is analogous to the single-choice BiB problem where each communication request (or ball) chooses an intermediate destination (a bin) at random.
+In this analogy, the congestion achieved by Valiant's paradigm corresponds to the maximum load in the BiB model, thus implying a bound of $\Theta\left(\log \log n\right)$ on the congestion w.h.p.
 \cite{CMM+98} \cite{MRS01}.
@@
 
-\section{Multiple-choice balls-into-bins}
+\section{Multiple-choice BiB}
 \label{sec:multiple-choice}
-The \emph{single-choice balls-into-bins strategy} (Section \ref{sec:single-choice}) chooses one bin uniformly and independently at random for each ball.
-A \emph{multiple-choice balls-into-bins strategy} generalizes this idea by  choosing $d \geq 2$ bins and placing the ball into the least full bin.
+The single-choice BiB strategy (Section \ref{sec:single-choice}) chooses one bin uniformly and independently at random for each ball.
+A multiple-choice BiB strategy generalizes this idea by  choosing $d \geq 2$ bins and placing the ball into the least full bin.
 Depending on how the $d$ locations are sampled, we distinguish three classes of placement algorithms \cite{VOC03}.
@@
 \begin{compactitem}
-\item  \emph{Uniform and independent.} Choose each of the $d$ bins for a ball uniformly and independently at random from the set of bins $B$. This implies 
+\item  Class 1: \emph{Uniform and independent.} Choose each of the $d$ bins for a ball uniformly and independently at random from the set of bins $B$. This implies 
 \begin{align*}
@@
 
-We will discuss two different multiple-choice placement strategies: The \emph{Greedy} algorithm (Section \ref{sec:greedy}) belongs to class 1, whereas the \emph{Always-go-left} scheme (Section \ref{sec:AlwaysGoLeft}) uses a non-uniform and independent selection of $d$ bins and thus belongs to class 2.
+We will discuss two different multiple-choice placement strategies: The Greedy algorithm (Section \ref{sec:greedy}) belongs to class 1, whereas the Always-go-left scheme (Section \ref{sec:AlwaysGoLeft}) uses a non-uniform and independent selection of $d$ bins and thus belongs to class 2.
 
@@
 
-Consider a variation of the above \emph{one-choice balls-into-bins} strategy, which we will call the \emph{Greedy} scheme: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}.
+Consider a variation of the above one-choice BiB strategy, which we will call the Greedy scheme: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin.
 If there are multiple bins with the same minimal load, chose one among them arbitrarily.
@@
 \begin{compactitem}
-\item The apparently small change made for the \emph{Greedy} algorithm compared to the \emph{single-choice balls-into-bins} scheme results in an exponential decrease of the gap, even for $d=2$. 
+\item The apparently small change made for the Greedy algorithm compared to the single-choice BiB scheme results in an exponential decrease of the gap, even for $d=2$. 
 \item Each additional choice ($d > 2$) decreases the gap by only a constant factor \cite{MRS01}. 
-\item For the \emph{heavily loaded case}, the resulting gap of the \emph{multiple-choice balls-into-bins} strategy does not depend on the number of balls $m$.
-In contrast, the bound for the \emph{one-choice balls-into-bins} scheme diverges with $m$.
+\item For the heavily loaded case, the resulting gap of the multiple-choice BiB strategy does not depend on the number of balls $m$.
+In contrast, the bound for the one-choice BiB scheme diverges with $m$.
 \begin{comment}
 \todo{the following is only true for class 1 algorithms}
-\item The given bounds are \emph{tight}, meaning that no other strategy that places each ball into one of $d$ randomly selected bins achieves a gap that is asymptotically lower.
+\item The given bounds are tight, meaning that no other strategy that places each ball into one of $d$ randomly selected bins achieves a gap that is asymptotically lower.
 \end{comment}
@@
 \label{sec:AlwaysGoLeft}
-The \emph{Greedy} strategy presented in Section \ref{sec:multiple-choice} assumes that $d$ bins are chosen uniformly and independently at random.
+The Greedy strategy presented in Section \ref{sec:multiple-choice} assumes that $d$ bins are chosen uniformly and independently at random.
 V\"ocking \cite{VOC03} was the first to study the impact of non-uniform and dependent selections of the $d$ bins.
-Surprisingly, a multiple-choice balls-into-bins strategy with a \emph{non-uniform} and independent selection of the $d$ bins, such as the \emph{Always-go-left} algorithm introduced in \cite{VOC03}, achieves better \cite{BCSV06} load balancing than the \emph{Greedy} scheme (Section \ref{sec:greedy}).
-Furthermore, V\"ocking proved that the \emph{Always-go-left} algorithm is optimal up to additive constants.
+Surprisingly, a multiple-choice BiB strategy with a non-uniform and independent selection of the $d$ bins, such as the Always-go-left algorithm introduced in \cite{VOC03}, achieves better \cite{BCSV06} load balancing than the Greedy scheme (Section \ref{sec:greedy}).
+Furthermore, V\"ocking proved that the Always-go-left algorithm is optimal up to additive constants.
 
-The \emph{Always-go-left} algorithm works as follows.
+The Always-go-left algorithm works as follows.
 Partition the $n$ bins into $d$ groups, each of size $\left\lfloor \frac{n}{d}\right\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$ .
 We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
-In case a tie occurs, put the ball into the \emph{left-most} bin.
+In case a tie occurs, put the ball into the left-most bin.
 That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with the same minimal load, put the ball into the bin with smallest index $i$.
@@
 \begin{align*}
-\frac{\ln\ln n}{d \cdot \ln \Phi_d} + O(1)
+\frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
 \end{align*}
@@
 \begin{align*}
-\frac{\ln\ln n}{d \cdot \ln \Phi_d} + O(h)
+\frac{\ln\ln n}{d  \ln \Phi_d} + O(h)
 \end{align*}
@@
 \label{sec:analysis}
-In this section, we will prove Theorems \ref{theorem:greedy}, \ref{theorem:agln} and \ref{theorem:algm} in detail, using \emph{witness trees} to provide an upper bound for the event that a bin contains too many balls.
+In this section, we will prove Theorems \ref{theorem:greedy}, \ref{theorem:agln} and \ref{theorem:algm} in detail, using witness trees to provide an upper bound for the event that a bin contains too many balls.
 Our analysis closely follows the proofs in \cite{VOC03}.
@@
 
-We say that a \emph{bad event} occurs when the maximum load exceeds some given threshold value.
-When such a bad event occurs we construct an \emph{activated witness tree}.
+We say that a bad event occurs when the maximum load exceeds some given threshold value.
+When such a bad event occurs we construct an activated witness tree.
 In other words, the activation of a witness tree is implied by a bad event.
@@
 
-Throughout this section, \emph{location} will denote the a specific place in one of the $n$ bins.
+Throughout this section, location will denote the a specific place in one of the $n$ bins.
 A location of a ball $b$ refers not only to a bin, but also to the number of balls already in this bin at the time of insertion of ball $b$.
@@
 \label{sec:analysis:definitionSymWT}
-A \emph{symmetric witness tree} of order $L$ is a complete $d$-ary tree with $d^{L}$ leaves.
+A symmetric witness tree of order $L$ is a complete $d$-ary tree with $d^{L}$ leaves.
 Each node $v$ represents a ball $b_v = \mathrm{ball}(v)$ .
@@
 
-We say that an edge or leaf event is \emph{activated} if the associated event occurs. A witness tree is activated when all its edge and leaf events are activated.
+We say that an edge or leaf event is activated if the associated event occurs. A witness tree is activated when all its edge and leaf events are activated.
 
@@
 This implies a probability of at most $3^{-d}$ that all $d$ bins are chosen among those with at least 3 additional bins in them.
-Thus, the probability that all leaf events are activated is at most $p_l = 3^{-d\cdot l}$, where $l = d^{L}$ is the number of leaves in a witness tree.
+Thus, the probability that all leaf events are activated is at most $p_l = 3^{-d l}$, where $l = d^{L}$ is the number of leaves in a witness tree.
 
@@
 \begin{align*}
-p_{swt} &= w \cdot p_e \cdot p_l \\
-		&= n^{k} \cdot \left(\frac{d}{n}\right)^{k-1} \cdot 3^{-d \cdot l} \\
-		&= n \cdot d^{k-1} \cdot 3^{-d \cdot l}
+p_{swt} &= w  p_e  p_l \\
+		&= n^{k}  \left(\frac{d}{n}\right)^{k-1}  3^{-d  l} \\
+		&= n  d^{k-1}  3^{-d  l}
 \end{align*}
@@
 \begin{align*}
-p_{swt} &\leq n \cdot d^{2\cdot l} \cdot 3^{-d \cdot l} \\
-		&= n \cdot \left(d^2\right)^l \cdot \left( 3^{-d} \right)^l
+p_{swt} &\leq n  d^{2 l}  3^{-d  l} \\
+		&= n  \left(d^2\right)^l  \left( 3^{-d} \right)^l
 \end{align*}
@@
 \begin{align*}
-p_{swt} &\leq n \cdot \left(\frac{3^d}{2}\right)^l \cdot \left( 3^{-d}\right)^l \\
-		&= n \cdot 2^{-l} \cdot 3^{d\cdot l} \cdot 3^{-d \cdot l} \\
-		&= n\cdot 2^{-l}		\\
-		&= n\cdot 2^{-d^L}		
+p_{swt} &\leq n  \left(\frac{3^d}{2}\right)^l  \left( 3^{-d}\right)^l \\
+		&= n  2^{-l}  3^{d l}  3^{-d  l} \\
+		&= n 2^{-l}		\\
+		&= n 2^{-d^L}		
 \end{align*}
@@
 \begin{align*}
-p_{swt} &\leq n \cdot 2^{{-d}^{\log_d\log_2 n + \log_d\left(1+\alpha\right)}} \\
-		&= n \cdot 2^{-{d}^{\log_d\log_2 n}\cdot d^{\log_d \cdot\left(1+\alpha\right)}} \\
-		&= n \cdot \left( 2^{-\log_2 n \cdot \left(1+ \alpha \right)} \right) \\
-		&= n \cdot n^{-\left(1+\alpha\right)} = n^{-\alpha} \neqn{swt:result}
+p_{swt} &\leq n  2^{{-d}^{\log_d\log_2 n + \log_d\left(1+\alpha\right)}} \\
+		&= n  2^{-{d}^{\log_d\log_2 n} d^{\log_d \left(1+\alpha\right)}} \\
+		&= n  \left( 2^{-\log_2 n  \left(1+ \alpha \right)} \right) \\
+		&= n  n^{-\left(1+\alpha\right)} = n^{-\alpha} \neqn{swt:result}
 \end{align*}
@@
 L &\geq  \frac{\ln\log_2 n}{\ln d} + \frac{\ln(1+\alpha)}{\ln d} \\
-   &= \frac{\ln \log_2 n}{\ln d} \cdot \left(1 + \frac{\ln\left(1+\alpha\right)}{\ln \log_2 n}\right) \\
-  &=  \frac{\ln \ln n}{\ln d} \cdot \left( 1+ o\left( 1\right)\right) \neqn{swt:boundL}
+   &= \frac{\ln \log_2 n}{\ln d}  \left(1 + \frac{\ln\left(1+\alpha\right)}{\ln \log_2 n}\right) \\
+  &=  \frac{\ln \ln n}{\ln d}  \left( 1+ o\left( 1\right)\right) \neqn{swt:boundL}
 \end{align*}
@@
 \label{sec:analysis:alg}
-Similarly to Section \ref{sec:analysis:greedy}, where we used symmetric witness trees to upper-bound the probability that a bin contains too many balls when using the \emph{Greedy} scheme, we will use asymmetric witness trees to get an analogous upper bound for the \emph{Always-go-left} scheme.
+Similarly to Section \ref{sec:analysis:greedy}, where we used symmetric witness trees to upper-bound the probability that a bin contains too many balls when using the Greedy scheme, we will use asymmetric witness trees to get an analogous upper bound for the Always-go-left scheme.
 We will make the same simplification as stated in Assumption \ref{assumption:independence}, namely that all the events represented by the edges and nodes of a witness tree are stochastically independent.
@@
 \label{sec:analysis:definitionAsymWT}
-With one exception, the definition of an \emph{asymmetric witness tree} is the same as the definition of a symmetric witness tree, including the definition of activation.
+With one exception, the definition of an asymmetric witness tree is the same as the definition of a symmetric witness tree, including the definition of activation.
 Only its structure differs from that of a symmetric tree.
@@
 
-A $d$-ary asymmetric witness tree of order $L$ has the structure of the Fibonacci tree $T_d(d \cdot L+1)$ and by Equation \ref{eqn:goldenFib} has $F_d(d\cdot L + 1) \geq \Phi_d^{d\cdot L-1}$ leaves.
+A $d$-ary asymmetric witness tree of order $L$ has the structure of the Fibonacci tree $T_d(d  L+1)$ and by Equation \ref{eqn:goldenFib} has $F_d(d L + 1) \geq \Phi_d^{d L-1}$ leaves.
 
@@
 \item A node with label $(h,i)$ , where $h>0$, has $d$ children with the labels $(h,i-1)\dots(h,1)$,$(h-1,d)\dots(h-1, i)$.
-\item The root of a Fibonacci tree $T_d(d\cdot h+i)$ has the label $(h, i)$. 
+\item The root of a Fibonacci tree $T_d(d h+i)$ has the label $(h, i)$. 
 \end{compactitem}
@@
 Let $b_i$ be the number of bins in group $i$ that contain 3 or more balls.
-Then $\beta_i = b_i\cdot d/ n$ is the fraction of the bins $b_i$ in group $i$.
+Then $\beta_i = b_i d/ n$ is the fraction of the bins $b_i$ in group $i$.
 Since no more than $n/3$ bins can contain 3 or more balls, the values $b_i$ are subject to the constraint
@@
 \begin{align*}
-p_l \leq 3^{-d\cdot l}
+p_l \leq 3^{-d l}
 \end{align*}
@@
 \begin{align*}
-p_{awt} &\leq n \cdot 2^{-l}
+p_{awt} &\leq n  2^{-l}
 \end{align*}
@@
 \begin{align*}
-l \geq \Phi_d^{d\cdot L -1}
+l \geq \Phi_d^{d L -1}
 \end{align*}
@@
 \begin{align*}
-L &\geq \left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d}}\right\rceil+\frac{1}{d} \neqn{awt:chooseL}
+L &\geq \left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}\right\rceil+\frac{1}{d} \neqn{awt:chooseL}
 \end{align*}
@@
 \begin{align*}
-p_{awt} &\leq n \cdot 2^{-\Phi_d^{d\cdot L -1}} \\
-        &= n \cdot 2^{-\Phi_d^{d\cdot \left(\left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d}}\right\rceil+\frac{1}{d}\right) -1}} \\
-        &\leq n \cdot 2^{-\Phi_d^{d\cdot \left({\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d}}+\frac{1}{d}\right) -1}} \\
-        &= n \cdot 2^{-\Phi_d^{{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d}}}} \\
-        &= n \cdot 2^{-\Phi_d^{\log_{\Phi_d}\log_2 n + \log_{\Phi_d}\left(1+\alpha\right)}} \\
-         &= n \cdot 2^{-\Phi_d^{\log_{\Phi_d}\log_2 n}\cdot \Phi_d^{ \log_{\Phi_d}\left(1+\alpha\right)}} \\
-         &= n \cdot 2^{-\log_2 n \cdot\left(1+\alpha\right)} \\
-         &= n \cdot n ^{-(1+\alpha)} = n^{-\alpha} \neqn{awt:result}
+p_{awt} &\leq n  2^{-\Phi_d^{d L -1}} \\
+        &= n  2^{-\Phi_d^{d \left(\left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}\right\rceil+\frac{1}{d}\right) -1}} \\
+        &\leq n  2^{-\Phi_d^{d \left({\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}+\frac{1}{d}\right) -1}} \\
+        &= n  2^{-\Phi_d^{{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}}} \\
+        &= n  2^{-\Phi_d^{\log_{\Phi_d}\log_2 n + \log_{\Phi_d}\left(1+\alpha\right)}} \\
+         &= n  2^{-\Phi_d^{\log_{\Phi_d}\log_2 n} \Phi_d^{ \log_{\Phi_d}\left(1+\alpha\right)}} \\
+         &= n  2^{-\log_2 n \left(1+\alpha\right)} \\
+         &= n  n ^{-(1+\alpha)} = n^{-\alpha} \neqn{awt:result}
 \end{align*}
@@
 \begin{align*}
-L &\geq \frac{\ln\log_2 n}{d\cdot \ln \Phi_d} + \frac{\ln\left(1+\alpha\right)}{d\cdot \ln \Phi_d} + \frac{1}{d}\\
-  &\geq \frac{\ln\ln n}{d\cdot \ln \Phi_d} + O\left(1\right)
+L &\geq \frac{\ln\log_2 n}{d \ln \Phi_d} + \frac{\ln\left(1+\alpha\right)}{d \ln \Phi_d} + \frac{1}{d}\\
+  &\geq \frac{\ln\ln n}{d \ln \Phi_d} + O\left(1\right)
 \end{align*}
@@
 Note that pruning will reduce the size of the witness tree and thus increase the probability of its activation.
-To compensate for this loss in size, we start with a larger witness tree, which we call a \emph{full witness tree} and will show that, speaking informally, the pruning of many events is unlikely.
+To compensate for this loss in size, we start with a larger witness tree, which we call a full witness tree and will show that, speaking informally, the pruning of many events is unlikely.
 Note that the technique presented in this section is applicable to both symmetric and asymmetric witness trees.
@@
 Let $\gamma \geq 2$ be a suitable constant.
-The root node of a \emph{full witness tree} has $\gamma$ children, denoted by $v_{c_i}$ with $1\leq i \leq \gamma$.
+The root node of a full witness tree has $\gamma$ children, denoted by $v_{c_i}$ with $1\leq i \leq \gamma$.
 The nodes $v_{c_i}$ have exactly one child each and the balls assigned to them are pairwise distinct.
@@
 When we reach a node $v$ that represents a ball $b$ already encountered before, let $e = (v,u)$ be the edge connecting $v$ and its parent node $u$.
-We call $e$ a \emph{cutoff edge} and use it as a witness that one of the $d$ locations of the ball $b_u$ is the same as one of the locations of some other ball $b_{u'}$, represented by a node $u'$, which was already traversed.
+We call $e$ a cutoff edge and use it as a witness that one of the $d$ locations of the ball $b_u$ is the same as one of the locations of some other ball $b_{u'}$, represented by a node $u'$, which was already traversed.
 Furthermore, we remove the whole tree rooted at node $v$ from our witness tree.
@@
 Note that more than $\gamma$ cut-off edges cannot exist.
-When finished, the \emph{pruned witness tree} does not contain any ball twice and thus all its events are independent.
+When finished, the pruned witness tree does not contain any ball twice and thus all its events are independent.
 
@@
 Thus, it remains to find an upper-bound for the activation of a full witness tree with $\gamma$ cut-off edges.
-Let $K$ be the number of references to balls by the full witness tree\footnote{Since no ball was assigned to the root node, the full witness tree consists of $K+1$ nodes} and assume that $K  \leq 2\cdot \gamma\left(\alpha+1\right)\cdot \log_2 n$, for some arbitrary $\alpha > 0$.
+Let $K$ be the number of references to balls by the full witness tree\footnote{Since no ball was assigned to the root node, the full witness tree consists of $K+1$ nodes} and assume that $K  \leq 2 \gamma\left(\alpha+1\right) \log_2 n$, for some arbitrary $\alpha > 0$.
 Then, there are at most $w_p=K^\gamma$ possibilities to transform a full witness tree with $K+1$ nodes to a pruned witness tree with $\gamma$ cut-off edges.
@@
 \begin{align*}
-p_c &= \left(k\cdot \frac{d}{n}\right)^\gamma \\
-    & \leq \left(K\cdot \frac{d}{n}\right)^\gamma
+p_c &= \left(k \frac{d}{n}\right)^\gamma \\
+    & \leq \left(K \frac{d}{n}\right)^\gamma
 \end{align*}
@@
 \begin{align*}
-p_{pwt} &= w_p\cdot w \cdot p_e \cdot p_l \cdot p_c \\
-		&= K\gamma\cdot n^{k+1}\cdot \left(\frac{d}{n}\right)^k\cdot3^{-d\cdot l} \cdot \left(K \frac{d}{n}\right)^\gamma\\
-		&= n \cdot d^k\cdot 3^{-d\cdot l} \cdot \left(K^2 \frac{d}{n} \right)^\gamma 
+p_{pwt} &= w_p w  p_e  p_l  p_c \\
+		&= K\gamma n^{k+1} \left(\frac{d}{n}\right)^k3^{-d l}  \left(K \frac{d}{n}\right)^\gamma\\
+		&= n  d^k 3^{-d l}  \left(K^2 \frac{d}{n} \right)^\gamma 
 \end{align*}
 
-Using $K \leq 2l$, $d^2 \leq 3^d$ and $K \leq 2\gamma \cdot \left( \alpha +1 \right) \cdot \log_2 n$ this becomes
+Using $K \leq 2l$, $d^2 \leq 3^d$ and $K \leq 2\gamma  \left( \alpha +1 \right)  \log_2 n$ this becomes
 \begin{align*}
-p_{pwt} &\leq n\cdot \left(\frac{\left(2\gamma \cdot \left(\alpha +1 \right) \cdot \log_2 n\right)^2\cdot d}{n}\right)^\gamma \\
-        &= n^{-\gamma +1} \cdot n ^{\log_n\left( \left(2\gamma\cdot \left(\alpha+1\right)\cdot \log_2n \right)^{2\gamma} \cdot d^\gamma \right)}\\
+p_{pwt} &\leq n \left(\frac{\left(2\gamma  \left(\alpha +1 \right)  \log_2 n\right)^2 d}{n}\right)^\gamma \\
+        &= n^{-\gamma +1}  n ^{\log_n\left( \left(2\gamma \left(\alpha+1\right) \log_2n \right)^{2\gamma}  d^\gamma \right)}\\
         &= n^{-\gamma+1+o\left(1\right)}
@@
 \label{sec:conclusion}
-Balls-into-bins games have many applications in Computer Science and have been studied extensively.
-The single-choice balls-into-bins strategy chooses a bin uniformly and independently at random for each ball.
+BiB games have many applications in Computer Science and have been studied extensively.
+The single-choice BiB strategy chooses a bin uniformly and independently at random for each ball.
 The classical result for this scheme upper-bounds the maximum load with
 \begin{align*}
-\load = \frac{m}{n} + \Theta\left(\sqrt{\frac{m \cdot \log n}{n}}\right).
+\load = \frac{m}{n} + \Theta\left(\sqrt{\frac{m  \log n}{n}}\right).
 \end{align*}
 w.h.p.
-In this paper we have presented two different miltiple-choice balls-into-bins placement algorithms: The Greedy-scheme due to Azar et al.\cite{ABKU99}, which achieves a maximum load of 
+In this paper we have presented two different miltiple-choice BiB placement algorithms: The Greedy-scheme due to Azar et al.\cite{ABKU99}, which achieves a maximum load of 
 \begin{align*}
@@
 \begin{align*}
-\load = \frac{m}{n} + \frac{\ln \ln n}{d\cdot \ln \Phi_d} + O(1)
+\load = \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
 \end{align*}
-w.h.p.~for all $m$. We used \emph{witness trees}, a very useful technique in the balls-into-bins context, to prove these upper-bounds in detail.
+w.h.p.~for all $m$. We used witness trees, a very useful technique in the BiB context, to prove these upper-bounds in detail.
 
@@
 \label{sec:asymptotic}
-We use the following definitions for the asymptotic notations $O(\cdot)$, $o(\cdot)$ and $\Theta(\cdot)$.
+We use the following definitions for the asymptotic notations $O()$, $o()$ and $\Theta()$.
 
 \begin{align*}
-f(n) \in o(g(n)) &\Leftrightarrow \forall \epsilon > 0 \exists n_o\forall n> n_0: |f(n)|\leq \epsilon \cdot |g(n)|\\
-f(n) \in O(g(n)) &\Leftrightarrow \exists c > 0 \exists n_o\forall n> n_0: |f(n)|\leq c \cdot |g(n)|\\
-f(n) \in \Omega(g(n)) &\Leftrightarrow \exists c > 0 \exists n_o\forall n> n_0: c \cdot |g(n)| \leq |f(n)| \\
+f(n) \in o(g(n)) &\Leftrightarrow \forall \epsilon > 0 \exists n_o\forall n> n_0: |f(n)|\leq \epsilon  |g(n)|\\
+f(n) \in O(g(n)) &\Leftrightarrow \exists c > 0 \exists n_o\forall n> n_0: |f(n)|\leq c  |g(n)|\\
+f(n) \in \Omega(g(n)) &\Leftrightarrow \exists c > 0 \exists n_o\forall n> n_0: c  |g(n)| \leq |f(n)| \\
 f(n) \in \Theta(g(n)) &\Leftrightarrow f(n) \in \Omega(g(n)) \land f(n)\in O(g(n))
@@
 \begin{align*}
-f(n) = g(n) \cdot (1+o(1)) \Rightarrow f(n) \in \Theta(g(n)) 
+f(n) = g(n)  (1+o(1)) \Rightarrow f(n) \in \Theta(g(n)) 
 \end{align*}
@@
 \begin{align*}
-f(n) &= g(n)+o(1)\cdot g(n) \\
+f(n) &= g(n)+o(1) g(n) \\
 &\Rightarrow f(n) \in \Omega(g(n)) \land f(n) \in O(g(n)) \\
@@
 \begin{align*}
-f(n) \in \Theta(g(n)) \Rightarrow f(n) = g(n) \cdot (1+o(1)) 
+f(n) \in \Theta(g(n)) \Rightarrow f(n) = g(n)  (1+o(1)) 
 \end{align*}
@@
 
-As discussed in Section \ref{sec:AlwaysGoLeft}, V\"ocking \cite{VOC03} was able to slightly improve the bound as well as to show that the \emph{Always-go-left} strategy that he introduces is optimal up to additive constants.
+As discussed in Section \ref{sec:AlwaysGoLeft}, V\"ocking \cite{VOC03} was able to slightly improve the bound as well as to show that the Always-go-left strategy that he introduces is optimal up to additive constants.
 

### 2014-02-21 article/article.tex
continued fixing stuff

@@
 \subsection{Multiple-choice BiB}
-Consider a variation of the above single-choice BiB strategy: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin.
+Consider a variation of the  single-choice BiB strategy: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin.
 If there are multiple bins with the same minimal load, chose one among them uniformly and independently at random.
@@
 
-This apparently minor change introduces a small amount of choice which, compared to the single-choice scheme, leads to a significant reduction in the maximum load. 
+This apparently minor change introduces a small amount of choice which leads to a significant reduction in the maximum load. 
 \begin{comment}
@@
 \end{comment}
-Extensive research following the influential paper by Azar et al.~\cite{ABKU99}, who were the first to explicitly prove this phenomenon\footnote{For a brief history of the proof, which is quite interesting by itself, refer to Appendix \ref{sec:historyOfGreedy}.}, suggests that this two-choice paradigm is not just an artifact of the simple BiB model, but a general principle applicable to a variety of problems.
+Extensive research following the influential paper by Azar et al.~\cite{ABKU99}, who were the first ones to explicitly prove this phenomenon\footnote{For a brief history of the proof, which is quite interesting by itself, refer to Appendix \ref{sec:historyOfGreedy}.}, suggests that this two-choice paradigm is not just an artifact of the simple BiB model, but a general principle applicable to a variety of problems.
 
-Consider a slightly more involved strategy, which we will call the Always-go-left scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size\footnote{For the sake of a clear presentation, we will assume that $d$ divides $n$ evenly in the remainder of this paper.} $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$.
+Consider the slightly more Always-go-left scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size\footnote{To simplify the presentation, we will assume that $d$ divides $n$ evenly in the remainder of this paper.} $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$.
 We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
-In case a tie occurs, put the ball into the left-most bin.
+In case a tie occurs, we put the ball into the left-most bin.
 That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with the same minimal load, we put the ball into the bin with smallest index $i$.
@@
 
-This paper is organized as follows: We give a basic classification of BiB games in Section \ref{sec:classification}, to provide some overview of the different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}.
+This survey is organized as follows: We give a basic classification of BiB games in Section \ref{sec:classification}, to provide some overview of the different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}.
 In the remaining sections of this paper we will consider the model given in Section \ref{sec:model}, which can be classified as a sequential, finite and dynamic process with random deletions.
 Section \ref{sec:multiple-choice} starts with a brief definition of three different classes of random placement strategies and then continues with the presentation of the two placement algorithms outlines in the introduction: The  Greedy algorithm (Section \ref{sec:greedy}) and the Always-go-left algorithm (Section \ref{sec:AlwaysGoLeft}).
-Section \ref{sec:analysis} is the core of this paper and will give a detailed analysis of both strategies.
-We will use a proof technique called witness trees, which allows for a clean analysis of both problems.
-Note that the technique is applicable to a wide variety of BiB scenarios, making it the most useful of several proof techniques applied to BiB problems.
+Section \ref{sec:analysis} is the core of this survey and will give a detailed analysis of both strategies.
+We will use a proof technique called witness trees, which allows for a clean analysis of both problems and is applicable to a wide variety of BiB scenarios.
 Other proof techniques such as layered induction and a differential equations approach have been used.
-They are however not part of this paper.
 See, e.g., \cite{MRS01} for an overview of different proof techniques for the BiB problem.
@@
 \label{sec:classification}
-Several different variations of this problem exist, which may be classified by the following properties:
+Several different variations of this problem exist, which may be classified as follows:
 \begin{compactitem}
@@
 A  ball is dealt with only after the previous ball has been placed.
-Contrary, in a \emph{parallel} process, balls arrive in sets and the balls in one set are dealt with simultaneously \cite{ABS98}.
-\item In a \emph{static} problem, the balls are placed into the bins and no balls are ever removed.
-In a \emph{dynamic} problem, balls can also be removed from the system.
-In a simple random deletions model, for each newly inserted ball a ball chosen uniformly and independently at random from among the balls currently in the bins is removed from the system \cite{ABKU99} \cite{MRS01}.
-More elaborate schemes exists: An oblivious adversary specifies the sequence of insertions and deletions of balls in advance, without any knowledge of the placement of the balls \cite{CFM+98}.
+Conversely, in a \emph{parallel} process, balls arrive in sets and the balls in one set are dealt with simultaneously \cite{ABS98}.
+\item In a \emph{static} problem, the balls are placed into the bins and no balls are ever removed whereas in a \emph{dynamic} problem balls can also be removed from the system.
+In the random deletions model, for each newly inserted ball one is chosen uniformly and independently at random and removed from the system \cite{ABKU99} \cite{MRS01}.
+More elaborate schemes exists: An oblivious adversary specifies the sequence of insertions and deletions of balls in advance, without any knowledge of their placement \cite{CFM+98}.
 Another approach was studied by Adler et al.~\cite{ABS98} in the context of the parallel BiB problem: In each bin, the balls are stored in FIFO-order and at every time step the first ball in each of the bins is deleted.
 \item A \emph{finite} process is analyzed for a given time interval that is known beforehand, whereas for an \emph{infinite} process, a placement and a deletion process is defined over an infinite time span \cite{ABS98}.
-\item In the lightly loaded case, $m$ is in the order of $n$, whereas in the heavily loaded case, $m$ is arbitrarily large.
-For clarity, we will assume that $m = n$ in the lightly loaded case and that $m > n$ in the heavily loaded case.
+\item In the lightly loaded case, $m$ is in the order of $n$, whereas in the heavily loaded case, $m$ is arbitrarily large. We will assume that $m = n$ in the lightly loaded case. 
 \end{compactitem}
@@
 The two-choice paradigm applied to the BiB model has several useful applications, some of which we will outline here.
-The applications in Hashing (Section \ref{sec:hashing}), Online Load Balancing (Section \ref{sec:loadbalancing}) and Emulation of shared memory machines on distributed memory machines \ref{sec:DMM}) share a close relation to load balancing.
+Hashing (Section \ref{sec:hashing}), online load balancing (Section \ref{sec:loadbalancing}) and the emulation of shared memory machines on distributed memory machines \ref{sec:DMM}) share a close relation to load balancing.
 Section \ref{sec:circuitrouting} applies the BiB model and the two-choice paradigm in a different context, namely that of routing circuits in multi-stage interconnection networks with minimal congestion.
@@
 
-It is easy to see how the BiB idea can be applied: Assuming a perfect hash function\footnote{i.e., each of the $n$ elements is mapped to one of the $n$ table entries uniformly and independently at random}, the elements are balls to be mapped to table entries or bins, implying a worst case look-up time of $\Theta\left(\frac{\log n}{\log \log n}\right)$ w.h.p.
-Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries implies a bound of $\Theta\left(\log \log n\right)$ w.h.p.
+It is easy to see how the BiB idea can be applied: Assuming a perfect hash function\footnote{i.e., each element is mapped to one of the $n$ table entries uniformly and independently at random}, the elements are balls to be mapped to table entries or bins, implying a worst case look-up time of $\Theta\left(\frac{\log n}{\log \log n}\right)$ w.h.p.
+Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries implies a look-up time of $\Theta\left(\log \log n\right)$ w.h.p.
 This approach is easy to parallelize, it does not require re-hashing of data and requires only two hash functions \cite{ABKU99} \cite{MRS01}.
@@
 \label{sec:loadbalancing}
-The following simple scenario is a typical application of online load balancing: Consider $n$ servers\footnote{Such as file-, database- or network servers} and $m$ requests issued by clients that need to be handled by these servers.
-For simplicity, assume that each task can be handled by any server and that all tasks are of the same size, i.e., require the same amount of execution time.
+The following scenario is a typical application of online load balancing: Consider $n$ servers (such as file-, database- or network servers) and $m$ requests.
+For simplicity, assume that each task can be handled by any server and that all tasks are of the same size, i.e., they require the same execution time.
 
 To maximize the throughput of the servers, a uniform distribution of tasks to servers is desirable.
-This can easily be achieved by a central dispatcher.
-However, such a central dispatcher poses a bottleneck within a distributed system and querying $m$ servers for their current load involves significant overhead, e.g., sending a message to each server and waiting for the reply and more efficient solutions are necessary.
+While a central dispatcher dispatcher can easily achieve this, it also poses a bottleneck within a distributed system since querying $m$ servers for their current load involves significant overhead (e.g., sending a message to each server and waiting for the reply).
 As with hashing, the BiB model can be applied easily: we need to assign tasks (the balls) to servers (the bins).
-Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends issues its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is bounded by $\Theta\left(\log \log n\right)$ w.h.p.
+Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is at most$O\left(\log \log n\right)$ w.h.p.
 \cite{KLM92} \cite{RS98} \cite{MRS01}.
@@
 \label{sec:DMM}
-The Parallel Random Access Machine model\footnote{as defined in the lecture or in~\cite{P03}}, or PRAM for short, is a high level abstraction of a shared memory machine, in which memory access by any processor to any memory location is assumed to require constant time.
+The Parallel Random Access Machine model (PRAM) ~\cite{P03} is a high level abstraction of a shared memory machine, in which memory access by any processor to any memory location is assumed to require constant time.
 It is especially useful for the study of parallel algorithms.
@@
 In multi-stage interconnection networks, virtual circuit-switching is used to route communication.
-For simplicity, we consider the permutation routing problem, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes and assume a butterfly network as the underlying interconnection network (see \cite{CLR09} for a detailed definition).
-Then, congestion is the maximum number of paths that are routed through a network link.
-The goal of a circuit routing algorithm is to to allocate paths for all $n$ communication requests such that congestion is minimal.
+For simplicity, we consider the permutation routing problem, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes and assume a butterfly network as the underlying interconnection network~\cite{CLR09}.
+Then, the congestion is the maximum number of paths that are routed through a network link.
+The goal of a circuit routing algorithm is to to allocate paths for all $n$ communication requests such that congestion is minimized.
 
@@
 
-Applying the two-choice paradigm in this context is somewhat more involved and beyond the scope of this paper. The interested reader may find more details in \cite{CMM+98}.
+Applying the two-choice paradigm in this context is somewhat more involved. The interested reader may find more details in \cite{CMM+98}.
 
@@
 
-Consider a variation of the above one-choice BiB strategy, which we will call the Greedy scheme: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin.
+Consider the Greedy scheme, a variation of the one-choice BiB strategy: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin.
 If there are multiple bins with the same minimal load, chose one among them arbitrarily.
@@
 \label{theorem:greedy}
-Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, the maximum load is at most 
+Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, the maximum load is bounded as 
 \begin{align*}
-\frac{m}{n} + \frac{\log \log n}{\log d}+ O(h)
+\load \leq \frac{m}{n} + \frac{\log \log n}{\log d}+ O(h)
 \end{align*}
-w.h.p.~at any time $t$ \cite{ABKU99} \cite{BCSV06}.
+w.h.p.~for all $t$ \cite{ABKU99} \cite{BCSV06}.
 \end{theorem}
 
-This implies a gap of $\Theta\left(\log \log n \right)$ for arbitrarily large $m$, which is remarkable in several ways:
+This implies a gap of $O\left(\log \log n \right)$ for arbitrarily large $m$, which is remarkable in several ways:
 \begin{compactitem}
@@
 \label{theorem:agln}
-If $n$ balls are placed into $n$ bins using the Always-go-left scheme, the number of balls in the fullest bin will be 
+If $n$ balls are placed into $n$ bins using the Always-go-left scheme, the maximum load will be
 \begin{align*}
-\frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
+\load \leq \frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
 \end{align*}
@@
 \begin{align*}
-\frac{\ln\ln n}{d  \ln \Phi_d} + O(h)
+\load \leq \frac{\ln\ln n}{d  \ln \Phi_d} + O(h)
 \end{align*}
@@
 Our analysis closely follows the proofs in \cite{VOC03}.
-\todo{is it enough to cite this once for the whole section?}
 
@@
 \begin{compactitem}
-\item \emph{Edge event:} Consider some edge $e = (u,v)$ of the witness tree, s.t.~$v$ is the i-th child of $u$.
-Then $e$ represents the event that the i-th location of the ball $b_u = \mathrm{ball}(u)$ points to the same bin as one of the locations of the ball $b_v = \mathrm{ball}(v)$.
+\item \emph{Edge event:} Consider some edge $e = (u,v)$ of the witness tree, s.t.~$v$ is the $i$-th child of $u$.
+Then $e$ represents the event that the $i$-th location of the ball $b_u = \mathrm{ball}(u)$ points to the same bin as one of the locations of the ball $b_v = \mathrm{ball}(v)$.
 %$\todo{illustrate}
@@
 
-An edge\footnote{where $v$ is the i-th child of $u$} $(u, v)$ is activated with probability at most $d/n$, which is the probability that the i-th location of the ball $b_u = \mathrm{ball}(u)$ points to a particular location of the ball $b_v = \mathrm{ball}(v)$ times the number of locations $d$.
+An edge\footnote{where $v$ is the $i$-th child of $u$} $(u, v)$ is activated with probability at most $d/n$, which is the probability that the $i$-th location of the ball $b_u = \mathrm{ball}(u)$ points to a particular location of the ball $b_v = \mathrm{ball}(v)$ times the number of locations $d$.
 This implies a probability of at most
@@
 \item For $1\leq j < i$, the $j$'th location of the ball $b_v$ points to a bin containing at least $l+4$ balls at the insertion time of the ball $b_v$.
-If the j-th location contained less balls than the i-th location, the Always-go-left scheme would have put the ball $b_v$ into the bin pointed to by location $j$.
+If the $j$-th location contained less balls than the $i$-th location, the Always-go-left scheme would have put the ball $b_v$ into the bin pointed to by location $j$.
 The topmost ball in location $j$ is assigned to the  child $u$ of $v$ with label $(l,j)$.
-\item For $i < j \leq d$, the j-th location contains at least $l+3$ balls.
+\item For $i < j \leq d$, the $j$-th location contains at least $l+3$ balls.
 If location $j$ contained less, the Always-go-left scheme would have put the ball $b_v$ into the bin it points to.
@@
 Since the ball $b$ was placed into bin $x$, at least one of its locations points to this bin $x$, which we call location $i$.
-Let $j = i+1 mod d$ be the ball's j-th location.
-Then, at the time of insertion of ball $b_c$, its j-th location points to a bin $y$ with at least $L+3$ other balls in it.
+Let $j = i+1 mod d$ be the ball's $j$-th location.
+Then, at the time of insertion of ball $b_c$, its $j$-th location points to a bin $y$ with at least $L+3$ other balls in it.
 We assign the topmost ball in bin $y$ to node $v_g$.

### 2014-02-27 article/article.tex
fix model

@@
 
-\subsection{BiB model}
-\label{sec:model}
-Although several different variants BiB exist, we will focus on a very basic model to highlight the crucial ideas in the placement strategies and their analyses.
-In our model, we start with $n$ \todo{empty?}  bins and put balls into them, one per time step $t$.
-We assume that at any time $t$ at most $m = h n$ balls are in the bins, where $h$ is some constant greater or equal to $1$.
-At each time step $t$, we delete one ball from the bins.
 While most of the paper and the analyses focuses on the lightly loaded case where $m=n$, we will give results for the case $m > n$ as well.
@@
 \begin{align*}
-\delta = \frac{\log \log n}{\log d} + O(1)
+\delta = \frac{\ln \ln n}{\ln d} + O(1)
 \end{align*}
@@
 
-It is easy to see how the BiB idea can be applied: Assuming a perfect hash function\footnote{i.e., each element is mapped to one of the $n$ table entries uniformly and independently at random}, the elements are balls to be mapped to table entries or bins, implying a worst case look-up time of $\Theta\left(\frac{\log n}{\log \log n}\right)$ w.h.p.
-Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries implies a look-up time of $\Theta\left(\log \log n\right)$ w.h.p.
+It is easy to see how the BiB idea can be applied: Assuming a perfect hash function\footnote{i.e., each element is mapped to one of the $n$ table entries uniformly and independently at random}, the elements are balls to be mapped to table entries or bins, implying a worst case look-up time of $\Theta\left(\frac{\ln n}{\log \ln n}\right)$ w.h.p.
+Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries implies a look-up time of $\Theta\left(\ln \ln n\right)$ w.h.p.
 This approach is easy to parallelize, it does not require re-hashing of data and requires only two hash functions \cite{ABKU99} \cite{MRS01}.
@@
 As with hashing, the BiB model can be applied easily: we need to assign tasks (the balls) to servers (the bins).
-Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is at most$O\left(\log \log n\right)$ w.h.p.
+Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is at most$O\left(\ln \ln n\right)$ w.h.p.
 \cite{KLM92} \cite{RS98} \cite{MRS01}.
@@
 
+\section{BiB model}
+\label{sec:model}
+Although several different variants of the BiB problem exist, we will focus on a very basic model to highlight the crucial ideas in the placement strategies and their analyses.
+We will base our analysis on a sequential, infinite and dynamic process using an oblivious adversary that provides a sequence $\sigma = \sigma_1 \sigma_2 \dots$, where $\sigma_t$ represents a request for a ball $b_d$ to be deleted and a different ball $b_a$ to be added to the bins at time step $t$.
+In our model, the ball $b_d$ associated with some $\sigma_t$ is deleted at the beginning of time step $t$.
+Next, the ball $b_a$ has to be placed into one of the $n$ bins without knowledge of the following requests $\sigma_{t+1}$, $\sigma_{t+2}\dots$.
+This infinite sequence of insertions and deletions is fixed in advance and is thus independent of previously allocated balls and any decisions made by the allocation algorithm.
+We assume that at any time $t$ at most $m = h n$ balls are in the bins, where $h$ is some constant greater or equal to $1$.
+
 \section{Multiple-choice BiB}
@@
 
-\section{Classes of random placement strategies}
+\subsection{Classes of random placement strategies}
 \label{sec:classesOfPlacement}
@@
 \begin{align*}
-\load \leq \frac{m}{n} + \frac{\log \log n}{\log d}+ O(h)
+\load \leq \frac{\ln \ln n}{\ln d}+ O(h)
 \end{align*}
@@
 
-This implies a gap of $O\left(\log \log n \right)$ for arbitrarily large $m$, which is remarkable in several ways:
+This implies a gap of $O\left(\ln \ln n \right)$ for arbitrarily large $m$, which is remarkable in several ways:
 \begin{compactitem}

### 2014-03-01 article/article.tex
make clear that we use m=n for proofs

@@
 This infinite sequence of insertions and deletions is fixed in advance and is thus independent of previously allocated balls and any decisions made by the allocation algorithm.
-We assume that at any time $t$ at most $m = h n$ balls are in the bins, where $h$ is some constant greater or equal to $1$.
+We assume that at any time $t$ at most $m = n$ balls are in the bins.
 
@@
 \label{theorem:greedy}
-Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, the maximum load is bounded as 
+Consider any sequence of insertions and deletions s.t.~at most $n$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, the maximum load is 
 \begin{align*}
-\load \leq \frac{\ln \ln n}{\ln d}+ O(h)
+\load = \frac{\ln \ln n}{\ln d}+ O(1)
 \end{align*}
@@
 
-This implies a gap of $O\left(\ln \ln n \right)$ for arbitrarily large $m$, which is remarkable in several ways:
+This result is remarkable in several ways:
 \begin{compactitem}
@@
 \item Each additional choice ($d > 2$) decreases the gap by only a constant factor \cite{MRS01}.
-\item For the heavily loaded case, the resulting gap of the multiple-choice BiB strategy does not depend on the number of balls $m$.
-In contrast, the bound for the one-choice BiB scheme diverges with $m$.
-\begin{comment}
-\todo{the following is only true for class 1 algorithms}
-\item The given bounds are tight, meaning that no other strategy that places each ball into one of $d$ randomly selected bins achieves a gap that is asymptotically lower.
-\end{comment}
 \end{compactitem}
-
 We will provide a detailed proof of this property in Section \ref{sec:analysis}.
@@
 
+\begin{comment}
 \begin{theorem}
@@
 \end{theorem}
+\end{comment}
 
 \begin{theorem}
-\label{theorem:algm}
-Consider any sequence of insertions and deletions s.t.~at most $hn$ balls are in the bins at any time.
-If the balls are placed into bins by the Greedy algorithm, at any time $t$ the maximum load is at most
+\label{theorem:aglm}
+Consider any sequence of insertions and deletions s.t.~at most $n$ balls are in the bins at any time. If the balls are placed into bins by the Always-go-left algorithm, the maximum load is at most 
 \begin{align*}
-\load \leq \frac{\ln\ln n}{d  \ln \Phi_d} + O(h)
+\load = \frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
 \end{align*}
-w.h.p.
+w.h.p.~for all $t$ \cite{VOC03}.
 \end{theorem}
@@
 \label{sec:analysis}
-In this section, we will prove Theorems \ref{theorem:greedy}, \ref{theorem:agln} and \ref{theorem:algm} in detail, using witness trees to provide an upper bound for the event that a bin contains too many balls.
+In this section, we will prove Theorems \ref{theorem:greedy} and \ref{theorem:aglm} in detail, using witness trees to provide an upper bound for the event that a bin contains too many balls.
 Our analysis closely follows the proofs in \cite{VOC03}.
@@
 Section \ref{sec:analysis:nondistinctBalls} will extend these proofs to remove the simplifying assumption we make initially.
-The witness tree method can be used to show that the same bounds hold for the heavily loaded case where $m> n$ too, but we do not give such a proof in this paper.
+The witness tree method can be used to show that the same bounds hold for case where $m > h n$ for some constant $h > 1$ too, but we do not give such a proof in this survey.
 Instead, we refer the interested reader to V\"ocking \cite{VOC03}.
-Berenbrink et al.
-~\cite{BCSV06}, Talwar et al.~\cite{TW13} and Mitzenmacher \cite{MRS01} provide proofs the heavily loaded case using different approaches as well.
+Berenbrink et al.~\cite{BCSV06} and Talwar et al.~\cite{TW13} provide proofs the heavily loaded case using different approaches.
 
@@
 \end{align*}
-This proves Theorem \ref{theorem:agln}, under the assumption that all events are stochastically independent (Assumption \ref{assumption:independence}). \qed
+This proves Theorem \ref{theorem:aglm}, under the assumption that all events are stochastically independent (Assumption \ref{assumption:independence}). \qed
 

### 2014-03-01 article/article.tex
asymptotic notations are more beautiful now

@@
 \label{sec:asymptotic}
-We use the following definitions for the asymptotic notations $O()$, $o()$ and $\Theta()$.
+We use the following definitions for the asymptotic notations \cite{CLRS01}, where $f(\cdot)$ and $g(\cdot)$ are positive functions with one argument $n\in \mathbb{N}$ and $n_0 \in \mathbb{N}$, $c \in \mathbb{R}_{\geq 0}$ are suitable constants.
 
 \begin{align*}
-f(n) \in o(g(n)) &\Leftrightarrow \forall \epsilon > 0 \exists n_o\forall n> n_0: |f(n)|\leq \epsilon  |g(n)|\\
-f(n) \in O(g(n)) &\Leftrightarrow \exists c > 0 \exists n_o\forall n> n_0: |f(n)|\leq c  |g(n)|\\
-f(n) \in \Omega(g(n)) &\Leftrightarrow \exists c > 0 \exists n_o\forall n> n_0: c  |g(n)| \leq |f(n)| \\
-f(n) \in \Theta(g(n)) &\Leftrightarrow f(n) \in \Omega(g(n)) \land f(n)\in O(g(n))
+f(n) \in O(g(n)) &\Leftrightarrow \exists c, n_0 \;\forall n> n_0: f(n)\leq c  g(n) \\
+f(n) \in \Omega(g(n)) &\Leftrightarrow \exists c, n_0\;\forall n> n_0: cg(n) \leq f(n) \\
+f(n) \in \Theta(g(n)) &\Leftrightarrow f(n) \in \Omega(g(n)) \land f(n)\in O(g(n)) \\
+f(n) \in o(g(n)) &\Leftrightarrow \forall c \; \exists n_0\;\forall n> n_0: f(n) < cg(n)
 \end{align*}
@@
 \bibliographystyle{acm}
-\bibliography{sources} 
+\bibliography{../sources} 
 \end{document}

### 2014-03-01 article/article.tex
improved conclusion

@@
 \end{align*}
-w.h.p. Not only is this a significant reduction of the asymptotic bound, V\"ocking was able to show that the Always-go-left strategy is optimal up to additive constants \cite{VOC03}.
+w.h.p.\ for all $m$. Not only is this a significant reduction of the asymptotic bound, V\"ocking was able to show that the Always-go-left strategy is optimal up to additive constants \cite{VOC03}.
 
@@
 The single-choice BiB strategy chooses a bin uniformly and independently at random for each ball.
-The classical result for this scheme upper-bounds the maximum load with
+The classical result for this scheme upper-bounds the gap with \cite{RS98}
 \begin{align*}
-\load = \frac{m}{n} + \Theta\left(\sqrt{\frac{m  \log n}{n}}\right).
+\delta =  \Theta\left(\sqrt{\frac{m  \log n}{n}}\right).
 \end{align*}
 w.h.p.
-In this paper we have presented two different miltiple-choice BiB placement algorithms: The Greedy-scheme due to Azar et al.\cite{ABKU99}, which achieves a maximum load of 
+In this paper we have presented two different multiple-choice BiB placement algorithms: The Greedy-scheme due to Azar et al.\cite{ABKU99}, which achieves a gap of 
 \begin{align*}
-\load = \frac{m}{n} + \frac{\ln\ln n}{\ln d} + O(1)
+\delta = \frac{\ln\ln n}{\ln d} + O(1)
 \end{align*}
-w.h.p.~and the Always-go-left scheme due to V\"ocking \cite{VOC03}, which achieves a slightly better maximum load of
+w.h.p.~and the Always-go-left scheme due to V\"ocking \cite{VOC03}, which achieves a slightly better gap of
 \begin{align*}
-\load = \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
+\delta = \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
 \end{align*}
-w.h.p.~for all $m$. We used witness trees, a very useful technique in the BiB context, to prove these upper-bounds in detail.
+w.h.p.~for all $m$. We used witness trees, a very useful technique in the BiB context, to prove these upper-bounds for the case $m=n$ in detail.
 

### 2014-03-03 article/article.tex
improved first proof

@@
 
-While most of the paper and the analyses focuses on the lightly loaded case where $m=n$, we will give results for the case $m > n$ as well.
-Throughout the paper, we will denote the maximum load by $\load$ and the additive gap by $\delta$:
+While most of this survey and the analyses focuses on the lightly loaded case where $m=n$, we will give results for the case $m > n$ as well.
+We will denote the maximum load by $\load$ and the additive gap by $\delta$:
 \begin{align*}
@@
 \label{sec:single-choice}
-A simple strategy for the BiB problem is to choose one bin uniformly and independently at random for each ball, which we will call the \emph{single-choice} strategy.
+The \emph{single-choice} strategy is a simple allocation scheme for the BiB problem and chooses one bin uniformly and independently at random for each ball.
 Using this strategy, the maximum load of the fullest bin will be
@@
 If there are multiple bins with the same minimal load, chose one among them uniformly and independently at random.
-This simple multiple-choice scheme, which we will call the Greedy algorithm, achieves a gap
+This simple multiple-choice scheme, called the Greedy algorithm, achieves a gap
 \begin{align*}
@@
 
-This apparently minor change introduces a small amount of choice which leads to a significant reduction in the maximum load. 
+This apparently minor change introduces an additional choice which leads to a significant reduction in the maximum load. 
 \begin{comment}
@@
 
-Consider the slightly more Always-go-left scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size\footnote{To simplify the presentation, we will assume that $d$ divides $n$ evenly in the remainder of this paper.} $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$.
-We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
-In case a tie occurs, we put the ball into the left-most bin.
-That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with the same minimal load, we put the ball into the bin with smallest index $i$.
+Consider the slightly more involved Always-go-left scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size\footnote{To simplify the presentation, we will assume that $d$ divides $n$ evenly in the remainder of this paper.} $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$.
+We select one from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
+For tie-breaking, we assume a fixed ordering of bins $b_i$ for $1 \leq i \leq n$ and put the ball into the bin with smallest index $i$ among those with same minimal load.
 Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing asymmetric tie-breaking in combination with non-uniform bin selection reduces the bound for the gap to\footnote{The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers.
@@
 In the remaining sections of this paper we will consider the model given in Section \ref{sec:model}, which can be classified as a sequential, finite and dynamic process with random deletions.
-Section \ref{sec:multiple-choice} starts with a brief definition of three different classes of random placement strategies and then continues with the presentation of the two placement algorithms outlines in the introduction: The  Greedy algorithm (Section \ref{sec:greedy}) and the Always-go-left algorithm (Section \ref{sec:AlwaysGoLeft}).
-Section \ref{sec:analysis} is the core of this survey and will give a detailed analysis of both strategies.
+Section \ref{sec:multiple-choice} starts with a brief definition of three different classes of random placement strategies and then continues with the presentation of the two placement algorithms outlined in the introduction: The Greedy algorithm (Section \ref{sec:greedy}) and the Always-go-left algorithm (Section \ref{sec:AlwaysGoLeft}).
+Section \ref{sec:analysis} is the core of this survey and will give a detailed analysis of both strategies for the case $m=n$.
 We will use a proof technique called witness trees, which allows for a clean analysis of both problems and is applicable to a wide variety of BiB scenarios.
@@
 See, e.g., \cite{MRS01} for an overview of different proof techniques for the BiB problem.
+The proof for arbitrarily large $m$ is more complicated and we refer the interested reader to \cite{BFZR08} and \cite{TW13}.
 
@@
 Although several different variants of the BiB problem exist, we will focus on a very basic model to highlight the crucial ideas in the placement strategies and their analyses.
-We will base our analysis on a sequential, infinite and dynamic process using an oblivious adversary that provides a sequence $\sigma = \sigma_1 \sigma_2 \dots$, where $\sigma_t$ represents a request for a ball $b_d$ to be deleted and a different ball $b_a$ to be added to the bins at time step $t$.
-In our model, the ball $b_d$ associated with some $\sigma_t$ is deleted at the beginning of time step $t$.
+which we will base on a sequential, infinite and dynamic process using an oblivious adversary.
+The adversary provides a sequence $\sigma = \sigma_1 \sigma_2 \dots$, where $\sigma_t$ represents a request for a ball $b_d$ to be deleted and a different ball $b_a$ to be added to the bins at time step $t$.
+In our model, the ball $b_d$ associated with some request $\sigma_t$ is deleted at the beginning of time step $t$.
 Next, the ball $b_a$ has to be placed into one of the $n$ bins without knowledge of the following requests $\sigma_{t+1}$, $\sigma_{t+2}\dots$.
@@
 \label{sec:classesOfPlacement}
-Let $B$ be the set of bins $\{1,2...n\}$, $\Omega = B^{d}$ be the sample space and $\mathcal{F} = 2^{\Omega}$ be the set of measurable events.
-We are only interested in events that represent a set of bins of size $d$.
-Thus, let $\mathcal{F}_d = \{\mathcal{E} \in \mathcal{F}: \left\vert \mathcal{E} \right\vert = d\}$ and set $\Pr\left(\mathcal{E}  \right) = 0$  $\forall \mathcal{E} \in \mathcal{F} \setminus \mathcal{F}_d $.
-The three classes of algorithms are distinguished by the probability function $\Pr\left(\mathcal{E}\right)$ for events $\mathcal{E} \in \mathcal{F}_d$:
+Let $B$ be the set of bins $\{1,2...n\}$ and $B^{d}$ be the sample space. The three classes of algorithms are distinguished by the way the $d$ candidate bins are chosen for each ball:
 \begin{compactitem}
-\item  Class 1: \emph{Uniform and independent.} Choose each of the $d$ bins for a ball uniformly and independently at random from the set of bins $B$. This implies 
+\item  Class 1: \emph{Uniform and independent.} Choose each of the $d$ bins for a ball uniformly and independently at random from the set of bins $B$.
+This implies the probability distribution for the selection of the i-th balls being
 \begin{align*}
-\forall \mathcal{E} \in \mathcal{F}_d: \Pr\left(\mathcal{E}\right) = \frac{1}{n ^{d}}
+D_i: B \rightarrow \frac{1}{n}
 \end{align*}
-\item Class 2: \emph{Non-uniform and independent.} Choose the i-th bin from $B$ independently at random according to some probability function
+\item Class 2: \emph{Non-uniform and independent.} Choose the i-th bin from $B$ independently at random according to some probability distribution: 
 \begin{align*}
-\Pr: B \rightarrow \left[0,1\right]
+D_i: B \rightarrow \left[0,1\right]
 \end{align*}
 
-\item Class 3: \emph{Non-uniform and dependent.} Choose the $d$ bins for a ball from $\Omega$ at random:
+\item Class 3: \emph{Non-uniform and dependent.} Choose the $d$ bins for a ball from $\Omega$ according to some probability probability distribution:
 \begin{align*}
-\Pr: B^{d} \rightarrow \left[0,1\right]
+D: B^{d} \rightarrow \left[0,1\right]
 \end{align*}
@@
 \label{sec:greedy}
-
 Consider the Greedy scheme, a variation of the one-choice BiB strategy: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin.
@@
 \label{theorem:greedy}
-Consider any sequence of insertions and deletions s.t.~at most $n$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, the maximum load is 
+Consider any sequence of insertions and deletions s.t.~at most $n$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, the maximum load is at most
 \begin{align*}
@@
 \end{compactitem}
-We will provide a detailed proof of this property in Section \ref{sec:analysis}.
 
@@
 \label{sec:AlwaysGoLeft}
-The Greedy strategy presented in Section \ref{sec:multiple-choice} assumes that $d$ bins are chosen uniformly and independently at random.
+The Greedy strategy presented in Section \ref{sec:greedy} assumes that $d$ bins are chosen uniformly and independently at random.
 V\"ocking \cite{VOC03} was the first to study the impact of non-uniform and dependent selections of the $d$ bins.
-Surprisingly, a multiple-choice BiB strategy with a non-uniform and independent selection of the $d$ bins, such as the Always-go-left algorithm introduced in \cite{VOC03}, achieves better \cite{BCSV06} load balancing than the Greedy scheme (Section \ref{sec:greedy}).
-Furthermore, V\"ocking proved that the Always-go-left algorithm is optimal up to additive constants.
+Surprisingly, a multiple-choice BiB strategy with a non-uniform and independent selection of the $d$ bins, such as the Always-go-left algorithm introduced in \cite{VOC03}, achieves better \cite{BCSV06} load balancing than the Greedy scheme.
 
 The Always-go-left algorithm works as follows.
-Partition the $n$ bins into $d$ groups, each of size $\left\lfloor \frac{n}{d}\right\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$ .
-We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
-In case a tie occurs, put the ball into the left-most bin.
-That is, assuming a fixed ordering of bins $b_i$ for $1 \leq i \leq n$, from among the $d' \leq d$ bins with the same minimal load, put the ball into the bin with smallest index $i$.
+Partition the $n$ bins into $d$ groups, each of size $\frac{n}{d}$.
+We select one from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
+For tie-breaking, we assume a fixed ordering of bins $b_i$ for $1 \leq i \leq n$ and put the ball into the bin with smallest index $i$ among those with same minimal load.
 
@@
 
-We say that a bad event occurs when the maximum load exceeds some given threshold value.
-When such a bad event occurs we construct an activated witness tree.
-In other words, the activation of a witness tree is implied by a bad event.
-The probability for the existence of such an activated witness tree is an upper-bound for the probability that the associated bad event occurs.
+When the maximum load exceeds some given threshold value we say that a bad event occurs and construct an activated witness tree.
+Thus, the activation of a witness tree is implied by a bad event.
+The probability for the existence of such an activated witness tree upper-bounds the probability that the associated bad event occurs.
 
-Throughout this section, location will denote the a specific place in one of the $n$ bins.
-A location of a ball $b$ refers not only to a bin, but also to the number of balls already in this bin at the time of insertion of ball $b$.
+Throughout this section, location will denote a specific place in one of the $n$ bins.
+A location of a ball $b$ refers to a specific bin and to the number of balls already in it (at the time of insertion of ball $b$).
 
 We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections \ref{sec:analysis:greedy} and \ref{sec:analysis:alg}, respectively.
-However, we use a simplifying Assumption \ref{assumption:independence} in those two sections to provide a more clear presentation of the proof technique.
+However, we use a simplifying assumption (Assumption \ref{assumption:independence}) in those two sections to provide a more clear presentation of the proof technique.
 Section \ref{sec:analysis:nondistinctBalls} will extend these proofs to remove the simplifying assumption we make initially.
@@
 Instead, we refer the interested reader to V\"ocking \cite{VOC03}.
-Berenbrink et al.~\cite{BCSV06} and Talwar et al.~\cite{TW13} provide proofs the heavily loaded case using different approaches.
+Berenbrink et al.~\cite{BCSV06} and Talwar et al.~\cite{TW13} provide proofs for the heavily loaded case using different approaches.
 
@@
 \label{sec:analysis:definitionSymWT}
-A symmetric witness tree of order $L$ is a complete $d$-ary tree with $d^{L}$ leaves.
-Each node $v$ represents a ball $b_v = \mathrm{ball}(v)$ .
-The nodes and edges of the witness tree represent events that may occur or not.
-The same ball may be represented by several nodes.
+A symmetric witness tree of order $L$ is a complete $d$-ary tree with $d^{L}$ leaves, where each node $v$ references a ball $b_v = \mathrm{ball}(v)$ and the nodes and edges of the witness tree represent events that may occur or not.
+In general, the same ball may be referenced by several nodes.
 However, initially we will make use of the following assumption in the analysis.
@@
 In other words, we assume that no ball is referenced by more than one node of a witness tree.
-Obviously, a correct analysis will have to consider balls assigned to more than one node, but this simplifying assumption is useful to highlight the differences between the symmetric and asymmetric allocation schemes.
+A correct analysis will have to consider balls assigned to more than one node, but this simplifying assumption is useful to highlight the differences between the symmetric and asymmetric allocation schemes.
 Section \ref{sec:analysis:nondistinctBalls} will extend the analysis to witness trees with redundant balls.
@@
 \item \emph{Edge event:} Consider some edge $e = (u,v)$ of the witness tree, s.t.~$v$ is the $i$-th child of $u$.
-Then $e$ represents the event that the $i$-th location of the ball $b_u = \mathrm{ball}(u)$ points to the same bin as one of the locations of the ball $b_v = \mathrm{ball}(v)$.
-%$\todo{illustrate}
-\item \emph{Leaf event:} A leaf node $v$ of the witness tree represents the event that each of the $d$ locations of the ball $b_v = \mathrm{ball}(v)$ points to a bin that contains at least three additional balls (i.e., balls that are not represented by a node of the tree) at the time of the insertion of the ball $b_v$.
-%\todo{illustrate}
+Then $e$ represents the event that the $i$-th location of the ball $b_u$ to the same bin as one of the locations of the ball $b_v$.
+\item \emph{Leaf event:} A leaf node $v$ of the witness tree represents the event that each of the $d$ locations of the ball $b_v$ points to a bin that contains at least three additional balls (i.e., balls that are not represented by a node of the tree), at the time of the insertion of the ball $b_v$.
+
 \end{compactitem}
 The definition for leaf events refers to bins containing some number of balls at some specified time.
-To avoid dependencies, the crucial trick is that edge events are defined in terms of the alternative locations of balls, not by their finally assigned bin.
+To avoid dependencies, the crucial trick is to defone edge events in terms of the possible locations of balls, not by their finally assigned bin.
+
+The following constraints have to be fulfilled by a witness tree.
+The ball $b_r$ that is represented by the root node $r$ has to exist at time step $t$ and each ball $b_v$ represented by a node $v \neq r$ and parent node $u$ has to exist when the ball $b_u$ is inserted.
 
 We say that an edge or leaf event is activated if the associated event occurs. A witness tree is activated when all its edge and leaf events are activated.
-
-The following constraints have to be fulfilled by an assignment.
-The ball $b_r$ that is represented by the root node $r$ has to exist at time step\footnote{As defined in Theorem \ref{theorem:greedy}} $t$ and each ball $b_v$ represented by a node $v \neq r$ and parent node $u$ has to exist when the ball $b_u = \mathrm{ball}(u)$ is inserted.
 
@@
 \item The topmost ball $b_1$ in bin $x$ is assigned to the root node $v_r$. Note that $b_1$ is the last ball that was inserted into the bin $x$.
-\item Since the Greedy-scheme puts ball $b_1$ into bin $x$ only if each of the $d$ sampled locations contain at least the same number of balls at bin $x$, we can conclude that each of the $d$ locations contain at least $L+3$ balls at the time ball $b_1$ was inserted.
+\item Since the Greedy-scheme puts ball $b_1$ into bin $x$ only if each of the $d$ sampled locations containd at least the same number of balls as bin $x$, we can conclude that each of the $d$ locations containd at least $L+3$ balls at the time ball $b_1$ was inserted.
 Thus, the $d$ children of the root node are assigned the topmost balls of these $d$ bins.
@@
 The balls assigned to the root's child node are each on top of $L+2$ or more balls and so on.
-It follows that each ball assigned to a leaf node is on top of at least 3 other balls, implying that the $d$ locations of each leaf node point to bins with at least 3 balls in each.
+It follows that each ball assigned to a leaf node is on top of at least 3 other balls, implying that the $d$ locations of each leaf node point to bins with at least 3 balls in each of them.
 This implies the activation of all leaf nodes of the witness tree.
@@
 \label{sec:analysis:probabilitySymWT}
-In this section, we derive the probability that some witness tree of order $L$ is activated, which gives an upper bound for the probability that the maximum load is $L+3$ at some time $t$.
+In this section, we derive the probability that some witness tree of order $L$ is activated, which gives an upper bound for the probability that the maximum load is greater than $L+3$ at some time $t$.
 
@@
 \end{lemma}
-Speaking less formaly, Lemma \ref{lemma:swt:activation} states that the for large enough $L$, the activation of a symmetric witness tree is unlikely.
+Speaking less formaly, Lemma \ref{lemma:swt:activation} states that for large enough $L$, the activation of a symmetric witness tree is unlikely.
 
@@
 The ball assigned to the root node can be chosen from at most $n$ balls, since by definition (Theorem \ref{theorem:greedy}), at most $n$ balls are in the bins at any time $t$ and a ball assigned to a node has to be in some bin at time $t$.
-Each child node of the root is assigned one out of at most $n$ balls too by the same argument.
+By the same argument, each child node of the root is assigned one out of at most $n$ balls too.
 Applying this argument for each level of the witness tree, there are at most $w = n^k$ possibilities to assign balls to a witness tree, with $k = d^{L+1}-1$ being the number of nodes in a $d$-ary witness tree of order $L$.
 
-An edge\footnote{where $v$ is the $i$-th child of $u$} $(u, v)$ is activated with probability at most $d/n$, which is the probability that the $i$-th location of the ball $b_u = \mathrm{ball}(u)$ points to a particular location of the ball $b_v = \mathrm{ball}(v)$ times the number of locations $d$.
+An edge\footnote{where $v$ is the $i$-th child of $u$} $(u, v)$ is activated with probability at most $d/n$, which is the probability that the $i$-th location of the ball $b_u$ points to a particular location of the ball $b_v$ times the number of locations $d$.
 This implies a probability of at most
@@
 \end{align*} 
-or the event that all $k-1$ edges of a witness tree are activated.
+for the event that all $k-1$ edges of a witness tree are activated.
 Note that  the assumption that all balls are distinct (Assumption \ref{assumption:independence}) implies that all edge events are independent.
@@
 The probability that a leaf event is activated is bounded above by $3^{-d}$.
-Recall that by the definition of leaf events, each of the $d$ locations of the ball $b_{vl} = \mathrm{ball}(v_l)$ has to point to a bin with at least 3 additional balls in it.
+Recall that by the definition of leaf events, each of the $d$ locations of a ball $b_{v_l}$, where $v_l$ is a leaf node, has to point to a bin with at least 3 additional balls in it.
 The number of those bins is bounded by $\lfloor n/3 \rfloor$ at all time steps $t$, implying a probability of at most
@@
 that one such bin is chosen.
-This implies a probability of at most $3^{-d}$ that all $d$ bins are chosen among those with at least 3 additional bins in them.
+This implies a probability of at most $3^{-d}$ that all $d$ bins are chosen from among those with at least 3 additional balls in them.
 Thus, the probability that all leaf events are activated is at most $p_l = 3^{-d l}$, where $l = d^{L}$ is the number of leaves in a witness tree.
@@
 
-Since the probabilty for the activation of a symmetric witness tree of order $L$ upper-bounds the probability of the bad event that some bin contains more than $L+3$ balls, Equation \ref{eqn:swt:result} also is an upper bound for the latter.
+Since the probabilty for the activation of a symmetric witness tree of order $L$ upper-bounds the probability of the bad event that some bin contains more than $L+3$ balls, Equation \ref{eqn:swt:result} is also an upper bound for the latter.
 Note that Equation \ref{eqn:swt:chooseL} can be written as
@@
 \end{align*}
+
 This proves Theorem \ref{theorem:greedy} for the case $m=n$, under the assumption that all events are stochastically independent (Assumption \ref{assumption:independence}). \qed

### 2014-03-11 article/article.tex
more improvements on first few sections

@@
 \usepackage{hyperref}
+\usepackage{cite}
 %page boarders
@@
 \label{sec:single-choice}
-The \emph{single-choice} strategy is a simple allocation scheme for the BiB problem and chooses one bin uniformly and independently at random for each ball.
+The \emph{Single-choice} strategy is a simple allocation scheme for the BiB problem and chooses one bin uniformly and independently at random for each ball.
 Using this strategy, the maximum load of the fullest bin will be
@@
 \end{align*}
-Note that, since the average load is 1, $\delta = L_{max}$ in this case. For the heavily loaded case, w.h.p.~the maximum gap is ~\cite{RS98}
+Note that, since the average load is 1, $\delta \in \Theta\left( L_{max} \right)$ in this case. For the heavily loaded case, w.h.p.~the maximum gap is ~\cite{RS98}
 \begin{align*}
@@
 If there are multiple bins with the same minimal load, chose one among them uniformly and independently at random.
-This simple multiple-choice scheme, called the Greedy algorithm, achieves a gap
+This simple multiple-choice scheme, called the \emph{Greedy} algorithm, achieves a gap
 \begin{align*}
-\delta = \frac{\ln \ln n}{\ln d} + O(1)
+\delta = \frac{\ln \ln n}{\ln d} + O(1),
 \end{align*}
- w.h.p.~for all $m$~\cite{ABKU99}~\cite{BCSV06}.
-
-This apparently minor change introduces an additional choice which leads to a significant reduction in the maximum load. 
+w.h.p., for all $m$~\cite{ABKU99, BCSV06}, i.e., this additional choice leads to an exponential reduction in the maximum load. 
 \begin{comment}
@@
 \end{comment}
-Extensive research following the influential paper by Azar et al.~\cite{ABKU99}, who were the first ones to explicitly prove this phenomenon\footnote{For a brief history of the proof, which is quite interesting by itself, refer to Appendix \ref{sec:historyOfGreedy}.}, suggests that this two-choice paradigm is not just an artifact of the simple BiB model, but a general principle applicable to a variety of problems.
+Extensive research following the influential paper by Azar et al.~\cite{ABKU99}, who were the first ones to explicitly prove this phenomenon\footnote{For a brief history of the proof, which is quite interesting by itself, refer to Appendix \ref{sec:historyOfGreedy}.}, suggests that this \emph{two-choice paradigm} is not just an artifact of the simple BiB model, but a general principle applicable to a variety of problems.
 
-Consider the slightly more involved Always-go-left scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size\footnote{To simplify the presentation, we will assume that $d$ divides $n$ evenly in the remainder of this paper.} $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$.
-We select one from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
-For tie-breaking, we assume a fixed ordering of bins $b_i$ for $1 \leq i \leq n$ and put the ball into the bin with smallest index $i$ among those with same minimal load.
-Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing asymmetric tie-breaking in combination with non-uniform bin selection reduces the bound for the gap to\footnote{The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers.
-Refer to Appendix \ref{sec:fibonacci} for the definition.}
+Consider the slightly more involved \emph{Always-go-left} scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$ (To simplify the presentation, we will assume that $d$ divides $n$ evenly in the remainder of this survey).
+We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
+For tie-breaking, we assume a fixed ordering of bins $i$ for $1 \leq i \leq n$ and put the ball into the bin with smallest index $i$ among those with same minimal load.
+Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing asymmetric tie-breaking in combination with non-uniform bin selection, reduces the bound for the gap to\
 \begin{align*}
-\delta = \frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
+\delta = \frac{\ln\ln n}{d  \ln \Phi_d} + O(1),
 \end{align*}
-w.h.p.\ for all $m$. Not only is this a significant reduction of the asymptotic bound, V\"ocking was able to show that the Always-go-left strategy is optimal up to additive constants \cite{VOC03}.
+w.h.p., for all $m$.
+(The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers.
+Refer to Appendix \ref{sec:fibonacci} for the definition.) Not only is this a significant reduction of the asymptotic bound, V\"ocking was able to show that the Always-go-left strategy is optimal up to additive constants \cite{VOC03}.
 
-This survey is organized as follows: We give a basic classification of BiB games in Section \ref{sec:classification}, to provide some overview of the different directions in research as well as possibilities of abstraction from applications, some of which are listed in Section \ref{sec:applications}.
-In the remaining sections of this paper we will consider the model given in Section \ref{sec:model}, which can be classified as a sequential, finite and dynamic process with random deletions.
-Section \ref{sec:multiple-choice} starts with a brief definition of three different classes of random placement strategies and then continues with the presentation of the two placement algorithms outlined in the introduction: The Greedy algorithm (Section \ref{sec:greedy}) and the Always-go-left algorithm (Section \ref{sec:AlwaysGoLeft}).
-Section \ref{sec:analysis} is the core of this survey and will give a detailed analysis of both strategies for the case $m=n$.
-We will use a proof technique called witness trees, which allows for a clean analysis of both problems and is applicable to a wide variety of BiB scenarios.
-Other proof techniques such as layered induction and a differential equations approach have been used.
-See, e.g., \cite{MRS01} for an overview of different proof techniques for the BiB problem.
-The proof for arbitrarily large $m$ is more complicated and we refer the interested reader to \cite{BFZR08} and \cite{TW13}.
-
-\begin{comment}
-In Section \ref{sec:strategies}, we will provide asymptotic bounds for the ad dative gap of this strategy, which we will then use to devise more complex strategies that significantly reduce these bounds.
-\end{comment}
-
-\begin{comment}
-These result may sometimes seem surprising, which is why we will denote a significant part of this article to demonstrate proofs in detail.
-This proof technique may be useful for devising asymptotic bounds for the additive gap of different BiB scenarios and strategies.
-\end{comment}
+\emph{Outline:}
+In Section \ref{sec:classification}, we give a basic classification of BiB games to provide an overview of the research, followed by some applications in Section \ref{sec:applications}.
+Section \ref{sec:model} gives a detailed definition of the model we use, which can be classified as a sequential, finite and dynamic process with random deletions.
+In Section \ref{sec:multiple-choice} we provide a brief definition of three different classes of random placement strategies and present two placement algorithms outlined in the introduction: The Greedy algorithm (Section \ref{sec:greedy}) and the Always-go-left algorithm (Section \ref{sec:AlwaysGoLeft}).
+Section \ref{sec:analysis} contains a detailed analysis of both strategies for the case $m=n$.
 
@@
 \begin{compactitem}
-\item In a \emph{sequential} process, the balls arrive one after another.
-A  ball is dealt with only after the previous ball has been placed.
+\item In a \emph{sequential} process, the balls arrive one after another and ball is dealt with only after the previous one has been placed.
 Conversely, in a \emph{parallel} process, balls arrive in sets and the balls in one set are dealt with simultaneously \cite{ABS98}.
-\item In a \emph{static} problem, the balls are placed into the bins and no balls are ever removed whereas in a \emph{dynamic} problem balls can also be removed from the system.
-In the random deletions model, for each newly inserted ball one is chosen uniformly and independently at random and removed from the system \cite{ABKU99} \cite{MRS01}.
-More elaborate schemes exists: An oblivious adversary specifies the sequence of insertions and deletions of balls in advance, without any knowledge of their placement \cite{CFM+98}.
-Another approach was studied by Adler et al.~\cite{ABS98} in the context of the parallel BiB problem: In each bin, the balls are stored in FIFO-order and at every time step the first ball in each of the bins is deleted.
-\item A \emph{finite} process is analyzed for a given time interval that is known beforehand, whereas for an \emph{infinite} process, a placement and a deletion process is defined over an infinite time span \cite{ABS98}.
-\item In the lightly loaded case, $m$ is in the order of $n$, whereas in the heavily loaded case, $m$ is arbitrarily large. We will assume that $m = n$ in the lightly loaded case. 
+\item In a \emph{static} problem, the balls are placed into the bins and no balls are ever removed, whereas in a \emph{dynamic} problem balls may also be removed from the system.
+In the random deletions model, for each newly inserted ball another one is chosen uniformly and independently at random and removed from the system \cite{ABKU99, MRS01}.
+More elaborate schemes exists: An oblivious adversary might specify the sequence of insertions and deletions of balls in advance, without any knowledge of their placement \cite{CFM+98}.
+Adler et al.~\cite{ABS98} studied another approach in the context of the parallel BiB problem, where the balls are stored in FIFO-order and at every time step the first ball in each of the bins is deleted.
+\item A \emph{finite} process is analyzed for a given time interval that is known beforehand, whereas for an \emph{infinite} one, placements and deletions are given for an infinite time span \cite{ABS98}.
+\item In the lightly loaded case, $m$ is in the order of $n$, whereas in the heavily loaded case, $m$ is arbitrarily large. W.l.o.g.~we will assume that $m = n$ for the lightly loaded case. 
 \end{compactitem}
@@
 The two-choice paradigm applied to the BiB model has several useful applications, some of which we will outline here.
-Hashing (Section \ref{sec:hashing}), online load balancing (Section \ref{sec:loadbalancing}) and the emulation of shared memory machines on distributed memory machines \ref{sec:DMM}) share a close relation to load balancing.
-Section \ref{sec:circuitrouting} applies the BiB model and the two-choice paradigm in a different context, namely that of routing circuits in multi-stage interconnection networks with minimal congestion.
+Hashing (Section \ref{sec:hashing}), online load balancing (Section \ref{sec:loadbalancing}) and the emulation of shared memory machines on distributed memory machines (Section \ref{sec:DMM}) share a close relation to load balancing.
+Section \ref{sec:circuitrouting} applies the BiB model and the two-choice paradigm to a different context, namely routing circuits in multi-stage interconnection networks with minimal congestion.
 
@@
 
-It is easy to see how the BiB idea can be applied: Assuming a perfect hash function\footnote{i.e., each element is mapped to one of the $n$ table entries uniformly and independently at random}, the elements are balls to be mapped to table entries or bins, implying a worst case look-up time of $\Theta\left(\frac{\ln n}{\log \ln n}\right)$ w.h.p.
-Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries implies a look-up time of $\Theta\left(\ln \ln n\right)$ w.h.p.
-This approach is easy to parallelize, it does not require re-hashing of data and requires only two hash functions \cite{ABKU99} \cite{MRS01}.
+It is easy to see how the BiB model can be used: Assuming a perfect hash function\footnote{I.e., each element is mapped to one of the $n$ table entries uniformly and independently at random}, the elements can be seen as balls to be mapped to table entries (bins), implying a worst case look-up time of $\Theta\left(\frac{\ln n}{\log \ln n}\right)$, w.h.p., where $n$ denotes the number of elements.
+Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries, results in a look-up time of $\Theta\left(\ln \ln n\right)$, w.h.p.
+This approach is easy to parallelize, it does not require re-hashing of data and requires only two hash functions \cite{ABKU99, MRS01}.
 
@@
 To maximize the throughput of the servers, a uniform distribution of tasks to servers is desirable.
-While a central dispatcher dispatcher can easily achieve this, it also poses a bottleneck within a distributed system since querying $m$ servers for their current load involves significant overhead (e.g., sending a message to each server and waiting for the reply).
+While a central dispatcher can easily achieve this, it also poses a bottleneck within a distributed system, since querying $m$ servers for their current load involves significant overhead (e.g., sending a message to each server and waiting for the reply).
 As with hashing, the BiB model can be applied easily: we need to assign tasks (the balls) to servers (the bins).
-Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is at most$O\left(\ln \ln n\right)$ w.h.p.
-\cite{KLM92} \cite{RS98} \cite{MRS01}.
+Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is at most $O\left(\ln \ln n\right)$, w.h.p., \cite{KLM92,RS98, MRS01}.
 
@@
 The Parallel Random Access Machine model (PRAM) ~\cite{P03} is a high level abstraction of a shared memory machine, in which memory access by any processor to any memory location is assumed to require constant time.
-It is especially useful for the study of parallel algorithms.
-However, a PRAM is much harder to realize in hardware than a Distributed Memory Machine (DMM).
+It is extremely useful for the study of parallel algorithms.
+However, a PRAM is much harder to implement in hardware than a Distributed Memory Machine (DMM).
 Thus, ways to emulate a PRAM on a DMM have been studied extensively.
-Such an emulation distributes the processors and memory cells of the PRAM to the processors and memory modules (typically one per processor) of the DMM with the goal to reduce the slowdown of the emulation, which is the time the DMM requires to simulate one step of the PRAM.
-Each memory module can handle only one memory access at a time step.
-More accesses to the same module will result in memory contention.
-The BiB idea is useful for managing communication between the processors and the memory modules such that memory contention is minimal.
-For details, see Karp et al.~\cite{KLM92} and the papers referenced there.
+Such an emulation distributes the processors and memory cells of the PRAM to the processors and memory modules (typically one per processor) of the DMM, with the goal to minimize the slowdown of the emulation, i.e., the time the DMM requires to simulate one step of the PRAM execution.
+Each memory module can handle only one memory access at a time step and more accesses to the same module will result in memory contention.
+The two-choice paradigm is useful for managing communication between the processors and the memory modules such that memory contention is minimal \cite{KLM92}.
 
@@
 In multi-stage interconnection networks, virtual circuit-switching is used to route communication.
-For simplicity, we consider the permutation routing problem, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes and assume a butterfly network as the underlying interconnection network~\cite{CLR09}.
-Then, the congestion is the maximum number of paths that are routed through a network link.
-The goal of a circuit routing algorithm is to to allocate paths for all $n$ communication requests such that congestion is minimized.
+For simplicity, we consider the permutation routing problem, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes and we assume a butterfly network as the underlying interconnection network~\cite{CLR09}.
+The congestion is the maximum number of paths that are routed through a network link and the goal of a circuit routing algorithm is to to allocate paths for all $n$ communication requests such that link congestion is minimized.
 
@@
 This routing technique is analogous to the single-choice BiB problem where each communication request (or ball) chooses an intermediate destination (a bin) at random.
-In this analogy, the congestion achieved by Valiant's paradigm corresponds to the maximum load in the BiB model, thus implying a bound of $\Theta\left(\log \log n\right)$ on the congestion w.h.p.
-\cite{CMM+98} \cite{MRS01}.
-
-Applying the two-choice paradigm in this context is somewhat more involved. The interested reader may find more details in \cite{CMM+98}.
+Extensive research \cite{CMM+98,MRS01} shows that by applying the two-choice paradigm, the link congestion can be bounded by $\Theta\left(\log \log n\right)$ , w.h.p.
 
@@
 \label{sec:model}
-Although several different variants of the BiB problem exist, we will focus on a very basic model to highlight the crucial ideas in the placement strategies and their analyses.
-which we will base on a sequential, infinite and dynamic process using an oblivious adversary.
-The adversary provides a sequence $\sigma = \sigma_1 \sigma_2 \dots$, where $\sigma_t$ represents a request for a ball $b_d$ to be deleted and a different ball $b_a$ to be added to the bins at time step $t$.
+Although several different variants of the BiB problem exist, we will focus on a basic model to highlight the crucial ideas in the placement strategies and their analyses.
+They are based on a sequential, infinite and dynamic process that uses an oblivious adversary, which provides a sequence $\sigma = \sigma_1 \sigma_2 \dots$, where $\sigma_t = (d_t, a_t)$ represents a request for a ball $d_t$ to be deleted and a different ball $a_t$ to be added to the bins at time step $t$.
 In our model, the ball $b_d$ associated with some request $\sigma_t$ is deleted at the beginning of time step $t$.
@@
 A multiple-choice BiB strategy generalizes this idea by  choosing $d \geq 2$ bins and placing the ball into the least full bin.
-Depending on how the $d$ locations are sampled, we distinguish three classes of placement algorithms \cite{VOC03}.
+We distinguish placement algorithms depending on how the $d$ locations are sampled \cite{VOC03}.
 
@@
 \label{sec:classesOfPlacement}
-Let $B$ be the set of bins $\{1,2...n\}$ and $B^{d}$ be the sample space. The three classes of algorithms are distinguished by the way the $d$ candidate bins are chosen for each ball:
-\begin{compactitem}
-\item  Class 1: \emph{Uniform and independent.} Choose each of the $d$ bins for a ball uniformly and independently at random from the set of bins $B$.
-This implies the probability distribution for the selection of the i-th balls being
+Let $B$ be the set of bins $\{1,2,...,n\}$ and $B^{d}$ be the sample space. Placement algorithms are distinguished by the way the $d$ candidate bins are chosen for each ball:
+\begin{description}
+\item [Uniform and independent.] Choose each of the $d$ bins for a ball uniformly and independently at random from the set of bins $B$.
+Thus, the probability that any specifc bin $b$ is choosen for the $i$-th balls is
 \begin{align*}
-D_i: B \rightarrow \frac{1}{n}
+\Pr[\mathrm{bin}\;b\;\mathrm{choosen}] = \frac{1}{n}
 \end{align*}
-\item Class 2: \emph{Non-uniform and independent.} Choose the i-th bin from $B$ independently at random according to some probability distribution: 
-\begin{align*}
-D_i: B \rightarrow \left[0,1\right]
-\end{align*}
+\item [Non-uniform and independent.] Choose the $i$-th bin from $B$ independently at random according to some probability distribution: 
+\item [Non-uniform and dependent.] Choose the $d$ bins for a ball from the sample space according to some probability probability distribution:
+\end{description}
 
-\item Class 3: \emph{Non-uniform and dependent.} Choose the $d$ bins for a ball from $\Omega$ according to some probability probability distribution:
-\begin{align*}
-D: B^{d} \rightarrow \left[0,1\right]
-\end{align*}
-\end{compactitem}
-
-We will discuss two different multiple-choice placement strategies: The Greedy algorithm (Section \ref{sec:greedy}) belongs to class 1, whereas the Always-go-left scheme (Section \ref{sec:AlwaysGoLeft}) uses a non-uniform and independent selection of $d$ bins and thus belongs to class 2.
+We will discuss two different multiple-choice placement strategies: The Greedy algorithm (Section \ref{sec:greedy}), belonging to class 1, and the Always-go-left scheme (Section \ref{sec:AlwaysGoLeft}), which uses a non-uniform and independent selection of $d$ bins and thus belongs to class 2.
 
@@
 \end{align*}
-w.h.p.~for all $t$ \cite{ABKU99} \cite{BCSV06}.
+w.h.p.~for all $t$ \cite{ABKU99, BCSV06}.
 \end{theorem}
@@
 
-Throughout this section, location will denote a specific place in one of the $n$ bins.
+Throughout this section, ``location'' will denote a specific place in one of the $n$ bins.
 A location of a ball $b$ refers to a specific bin and to the number of balls already in it (at the time of insertion of ball $b$).
@@
 Section \ref{sec:analysis:nondistinctBalls} will extend these proofs to remove the simplifying assumption we make initially.
-The witness tree method can be used to show that the same bounds hold for case where $m > h n$ for some constant $h > 1$ too, but we do not give such a proof in this survey.
-Instead, we refer the interested reader to V\"ocking \cite{VOC03}.
-Berenbrink et al.~\cite{BCSV06} and Talwar et al.~\cite{TW13} provide proofs for the heavily loaded case using different approaches.
+
+We will use a proof technique called witness trees, which allows for a clean analysis of both problems and is applicable to a wide variety of BiB scenarios.
+Other proof techniques such as layered induction and a differential equations approach have been used.
+See, e.g., \cite{MRS01} for an overview of different proof techniques for the BiB problem.
+The proof for arbitrarily large $m$ is more complicated and we refer the interested reader to \cite{BCSV06, TW13}.
 
@@
 Each of the $d$ locations of this ball points to a bin with at least $L+3$ balls in it, at the time of its insertion.
-These are the $d$ bins that were sampled by the Always-go-left scheme and thus bin $i$ belongs to the $i$'th group of bins, where each group consists of $n/d$ bins\footnote{For the sake of simplicity, we assume that $d$ divides $n$ evenly, i.e., that all groups have the exact same size.
-It can be shown that the analysis still holds if each group is of size $\lfloor\frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$.}.
+These are the $d$ bins that were sampled by the Always-go-left scheme and thus bin $i$ belongs to the $i$'th group of bins, where each group consists of $n/d$ bins.
 The child node of the root with label $(L-1, i)$ is assigned the topmost ball of the bin the $i$' location points to.
@@
 Berenbrink at.
-all \cite{BFZR08} were the first to proof that the same bound indeed holds for the heavily loaded case.
+all \cite{BCSV06} were the first to proof that the same bound indeed holds for the heavily loaded case.
 Their proof is fairly complex and a simpler analysis may be found in \cite{TW13}, which also provides a tight lower bound for the heavily loaded case.

### 2014-03-11 article/article.tex
enforce no line breaks before references and citations

@@
 Equivalently, one can look at the \emph{additive gap} of the allocation, i.e., the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
-These balls-into-bins games (BiB) have been studied extensively (see, e.g., \cite{JK77}).
+These balls-into-bins games (BiB) have been studied extensively (see, e.g.,~\cite{JK77}).
 In the context of Computer Science, the BiB model is useful to study, e.g., strategies for distributing tasks (balls) to processors (bins) as evenly as possible without having to query all the processors for the number of tasks currently assigned to them.
@@
 \begin{comment}
-\footnote{A more accurate result exists: The maximum load is $\load = \Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$ \cite{G91}.}
+\footnote{A more accurate result exists: The maximum load is $\load = \Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$~\cite{G91}.}
 \end{comment}
@@
 For tie-breaking, we assume a fixed ordering of bins $i$ for $1 \leq i \leq n$ and put the ball into the bin with smallest index $i$ among those with same minimal load.
-Somewhat surprisingly, V\"ocking \cite{VOC03} showed that such a strategy employing asymmetric tie-breaking in combination with non-uniform bin selection, reduces the bound for the gap to\
+Somewhat surprisingly, V\"ocking~\cite{VOC03} showed that such a strategy employing asymmetric tie-breaking in combination with non-uniform bin selection, reduces the bound for the gap to\
 \begin{align*}
@@
 (The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers.
-Refer to Appendix \ref{sec:fibonacci} for the definition.) Not only is this a significant reduction of the asymptotic bound, V\"ocking was able to show that the Always-go-left strategy is optimal up to additive constants \cite{VOC03}.
+Refer to Appendix \ref{sec:fibonacci} for the definition.) Not only is this a significant reduction of the asymptotic bound, V\"ocking was able to show that the Always-go-left strategy is optimal up to additive constants~\cite{VOC03}.
 
@@
 \item In a \emph{sequential} process, the balls arrive one after another and ball is dealt with only after the previous one has been placed.
-Conversely, in a \emph{parallel} process, balls arrive in sets and the balls in one set are dealt with simultaneously \cite{ABS98}.
+Conversely, in a \emph{parallel} process, balls arrive in sets and the balls in one set are dealt with simultaneously~\cite{ABS98}.
 \item In a \emph{static} problem, the balls are placed into the bins and no balls are ever removed, whereas in a \emph{dynamic} problem balls may also be removed from the system.
-In the random deletions model, for each newly inserted ball another one is chosen uniformly and independently at random and removed from the system \cite{ABKU99, MRS01}.
-More elaborate schemes exists: An oblivious adversary might specify the sequence of insertions and deletions of balls in advance, without any knowledge of their placement \cite{CFM+98}.
+In the random deletions model, for each newly inserted ball another one is chosen uniformly and independently at random and removed from the system~\cite{ABKU99, MRS01}.
+More elaborate schemes exists: An oblivious adversary might specify the sequence of insertions and deletions of balls in advance, without any knowledge of their placement~\cite{CFM+98}.
 Adler et al.~\cite{ABS98} studied another approach in the context of the parallel BiB problem, where the balls are stored in FIFO-order and at every time step the first ball in each of the bins is deleted.
-\item A \emph{finite} process is analyzed for a given time interval that is known beforehand, whereas for an \emph{infinite} one, placements and deletions are given for an infinite time span \cite{ABS98}.
+\item A \emph{finite} process is analyzed for a given time interval that is known beforehand, whereas for an \emph{infinite} one, placements and deletions are given for an infinite time span~\cite{ABS98}.
 \item In the lightly loaded case, $m$ is in the order of $n$, whereas in the heavily loaded case, $m$ is arbitrarily large. W.l.o.g.~we will assume that $m = n$ for the lightly loaded case. 
@@
 Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries, results in a look-up time of $\Theta\left(\ln \ln n\right)$, w.h.p.
-This approach is easy to parallelize, it does not require re-hashing of data and requires only two hash functions \cite{ABKU99, MRS01}.
+This approach is easy to parallelize, it does not require re-hashing of data and requires only two hash functions~\cite{ABKU99, MRS01}.
 
@@
 As with hashing, the BiB model can be applied easily: we need to assign tasks (the balls) to servers (the bins).
-Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is at most $O\left(\ln \ln n\right)$, w.h.p., \cite{KLM92,RS98, MRS01}.
+Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is at most $O\left(\ln \ln n\right)$, w.h.p.,~\cite{KLM92,RS98, MRS01}.
 
@@
 Each memory module can handle only one memory access at a time step and more accesses to the same module will result in memory contention.
-The two-choice paradigm is useful for managing communication between the processors and the memory modules such that memory contention is minimal \cite{KLM92}.
+The two-choice paradigm is useful for managing communication between the processors and the memory modules such that memory contention is minimal~\cite{KLM92}.
 
@@
 
-In this context, Valiant's paradigm \cite{V82} states that any such permutation routing problem can be solved by transforming it into two problems: First, find a path for each request to an intermediate destination chosen uniformly and independently at random.
+In this context, Valiant's paradigm~\cite{V82} states that any such permutation routing problem can be solved by transforming it into two problems: First, find a path for each request to an intermediate destination chosen uniformly and independently at random.
 Next, find a path from the intermediate to the actual destination.
 This routing technique is analogous to the single-choice BiB problem where each communication request (or ball) chooses an intermediate destination (a bin) at random.
-Extensive research \cite{CMM+98,MRS01} shows that by applying the two-choice paradigm, the link congestion can be bounded by $\Theta\left(\log \log n\right)$ , w.h.p.
+Extensive research~\cite{CMM+98,MRS01} shows that by applying the two-choice paradigm, the link congestion can be bounded by $\Theta\left(\log \log n\right)$ , w.h.p.
 
@@
 A multiple-choice BiB strategy generalizes this idea by  choosing $d \geq 2$ bins and placing the ball into the least full bin.
-We distinguish placement algorithms depending on how the $d$ locations are sampled \cite{VOC03}.
+We distinguish placement algorithms depending on how the $d$ locations are sampled~\cite{VOC03}.
 
@@
 \end{align*}
-w.h.p.~for all $t$ \cite{ABKU99, BCSV06}.
+w.h.p.~for all $t$~\cite{ABKU99, BCSV06}.
 \end{theorem}
@@
 \item The apparently small change made for the Greedy algorithm compared to the single-choice BiB scheme results in an exponential decrease of the gap, even for $d=2$. 
-\item Each additional choice ($d > 2$) decreases the gap by only a constant factor \cite{MRS01}.
+\item Each additional choice ($d > 2$) decreases the gap by only a constant factor~\cite{MRS01}.
 \end{compactitem}
@@
 \label{sec:AlwaysGoLeft}
-The Greedy strategy presented in Section \ref{sec:greedy} assumes that $d$ bins are chosen uniformly and independently at random.
-V\"ocking \cite{VOC03} was the first to study the impact of non-uniform and dependent selections of the $d$ bins.
-Surprisingly, a multiple-choice BiB strategy with a non-uniform and independent selection of the $d$ bins, such as the Always-go-left algorithm introduced in \cite{VOC03}, achieves better \cite{BCSV06} load balancing than the Greedy scheme.
+The Greedy strategy presented in Section~\ref{sec:greedy} assumes that $d$ bins are chosen uniformly and independently at random.
+V\"ocking~\cite{VOC03} was the first to study the impact of non-uniform and dependent selections of the $d$ bins.
+Surprisingly, a multiple-choice BiB strategy with a non-uniform and independent selection of the $d$ bins, such as the Always-go-left algorithm introduced in~\cite{VOC03}, achieves better~\cite{BCSV06} load balancing than the Greedy scheme.
 
@@
 \end{align*}
-w.h.p.~for all $t$ \cite{VOC03}.
+w.h.p.~for all $t$~\cite{VOC03}.
 \end{theorem}
@@
 In this section, we will prove Theorems \ref{theorem:greedy} and \ref{theorem:aglm} in detail, using witness trees to provide an upper bound for the event that a bin contains too many balls.
-Our analysis closely follows the proofs in \cite{VOC03}.
+Our analysis closely follows the proofs in~\cite{VOC03}.
 
@@
 
-We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections \ref{sec:analysis:greedy} and \ref{sec:analysis:alg}, respectively.
-However, we use a simplifying assumption (Assumption \ref{assumption:independence}) in those two sections to provide a more clear presentation of the proof technique.
-Section \ref{sec:analysis:nondistinctBalls} will extend these proofs to remove the simplifying assumption we make initially.
+We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections~\ref{sec:analysis:greedy} and~\ref{sec:analysis:alg}, respectively.
+However, we use a simplifying assumption (Assumption~\ref{assumption:independence}) in those two sections to provide a more clear presentation of the proof technique.
+Section~\ref{sec:analysis:nondistinctBalls} will extend these proofs to remove the simplifying assumption we make initially.
 
@@
 Other proof techniques such as layered induction and a differential equations approach have been used.
-See, e.g., \cite{MRS01} for an overview of different proof techniques for the BiB problem.
-The proof for arbitrarily large $m$ is more complicated and we refer the interested reader to \cite{BCSV06, TW13}.
+See, e.g.,~\cite{MRS01} for an overview of different proof techniques for the BiB problem.
+The proof for arbitrarily large $m$ is more complicated and we refer the interested reader to~\cite{BCSV06, TW13}.
 
@@
 A correct analysis will have to consider balls assigned to more than one node, but this simplifying assumption is useful to highlight the differences between the symmetric and asymmetric allocation schemes.
-Section \ref{sec:analysis:nondistinctBalls} will extend the analysis to witness trees with redundant balls.
+Section~\ref{sec:analysis:nondistinctBalls} will extend the analysis to witness trees with redundant balls.
 
@@
 \end{lemma}
-Speaking less formaly, Lemma \ref{lemma:swt:activation} states that for large enough $L$, the activation of a symmetric witness tree is unlikely.
+Speaking less formaly, Lemma~\ref{lemma:swt:activation} states that for large enough $L$, the activation of a symmetric witness tree is unlikely.
 
@@
 The number of different witness trees $w$ is given by the number of ways to assign balls to its nodes.
-The ball assigned to the root node can be chosen from at most $n$ balls, since by definition (Theorem \ref{theorem:greedy}), at most $n$ balls are in the bins at any time $t$ and a ball assigned to a node has to be in some bin at time $t$.
+The ball assigned to the root node can be chosen from at most $n$ balls, since by definition (Theorem~\ref{theorem:greedy}), at most $n$ balls are in the bins at any time $t$ and a ball assigned to a node has to be in some bin at time $t$.
 By the same argument, each child node of the root is assigned one out of at most $n$ balls too.
@@
 for the event that all $k-1$ edges of a witness tree are activated.
-Note that  the assumption that all balls are distinct (Assumption \ref{assumption:independence}) implies that all edge events are independent.
+Note that  the assumption that all balls are distinct (Assumption~\ref{assumption:independence}) implies that all edge events are independent.
 
@@
 \end{align*}
-Thus, for any constant $\alpha$ and large enough $L$ as specified in Equation \ref{eqn:swt:chooseL} the probability that a symmetric witness tree of order $L$ is activated is bounded above by $p_{swt }\leq n^{-\alpha}$.
+Thus, for any constant $\alpha$ and large enough $L$ as specified in Equation~\ref{eqn:swt:chooseL} the probability that a symmetric witness tree of order $L$ is activated is bounded above by $p_{swt }\leq n^{-\alpha}$.
 \end{proof}
 
-Since the probabilty for the activation of a symmetric witness tree of order $L$ upper-bounds the probability of the bad event that some bin contains more than $L+3$ balls, Equation \ref{eqn:swt:result} is also an upper bound for the latter.
-Note that Equation \ref{eqn:swt:chooseL} can be written as
+Since the probabilty for the activation of a symmetric witness tree of order $L$ upper-bounds the probability of the bad event that some bin contains more than $L+3$ balls, Equation~\ref{eqn:swt:result} is also an upper bound for the latter.
+Note that Equation~\ref{eqn:swt:chooseL} can be written as
 \begin{align*}
@@
 
-This proves Theorem \ref{theorem:greedy} for the case $m=n$, under the assumption that all events are stochastically independent (Assumption \ref{assumption:independence}). \qed
+This proves Theorem~\ref{theorem:greedy} for the case $m=n$, under the assumption that all events are stochastically independent (Assumption~\ref{assumption:independence}). \qed
 
@@
 \label{sec:analysis:alg}
-Similarly to Section \ref{sec:analysis:greedy}, where we used symmetric witness trees to upper-bound the probability that a bin contains too many balls when using the Greedy scheme, we will use asymmetric witness trees to get an analogous upper bound for the Always-go-left scheme.
-We will make the same simplification as stated in Assumption \ref{assumption:independence}, namely that all the events represented by the edges and nodes of a witness tree are stochastically independent.
+Similarly to Section~\ref{sec:analysis:greedy}, where we used symmetric witness trees to upper-bound the probability that a bin contains too many balls when using the Greedy scheme, we will use asymmetric witness trees to get an analogous upper bound for the Always-go-left scheme.
+We will make the same simplification as stated in Assumption~\ref{assumption:independence}, namely that all the events represented by the edges and nodes of a witness tree are stochastically independent.
 
@@
 
-A $d$-ary asymmetric witness tree of order $L$ has the structure of the Fibonacci tree $T_d(d  L+1)$ and by Equation \ref{eqn:goldenFib} has $F_d(d L + 1) \geq \Phi_d^{d L-1}$ leaves.
+A $d$-ary asymmetric witness tree of order $L$ has the structure of the Fibonacci tree $T_d(d  L+1)$ and by Equation~\ref{eqn:goldenFib} has $F_d(d L + 1) \geq \Phi_d^{d L-1}$ leaves.
 
@@
 \begin{proof}
-Analogously to the symmetric witness tree, the possible number of asymmetric witness trees is upper-bounded by $w=n^k$ and the probability that all edge events occur is\footnote{As in Section \ref{sec:analysis:greedy}, $k$ denotes the number of nodes of the witness tree and $l$ the number of leaves.}
+Analogously to the symmetric witness tree, the possible number of asymmetric witness trees is upper-bounded by $w=n^k$ and the probability that all edge events occur is\footnote{As in Section~\ref{sec:analysis:greedy}, $k$ denotes the number of nodes of the witness tree and $l$ the number of leaves.}
 \begin{align*}
@@
 
-The above results imply the same probability for the activation of an asymmetric witness tree as we derived in Section \ref{sec:analysis:probabilitySymWT} for the activation of a symmetric witness tree, that is
+The above results imply the same probability for the activation of an asymmetric witness tree as we derived in Section~\ref{sec:analysis:probabilitySymWT} for the activation of a symmetric witness tree, that is
 \begin{align*}
@@
 \end{align*}
-holds\footnote{see Section \ref{sec:analysis:definitionAsymWT} for details}.
+holds\footnote{see Section~\ref{sec:analysis:definitionAsymWT} for details}.
 
@@
 \end{align*}
-Thus, for any constant $\alpha$ and large enough $L$ as specified in Equation \ref{eqn:awt:chooseL} the probability that a symmetric witness tree of order $L$ is activated is bounded above by $p_{awt }\leq n^{-\alpha}$.
+Thus, for any constant $\alpha$ and large enough $L$ as specified in Equation~\ref{eqn:awt:chooseL} the probability that a symmetric witness tree of order $L$ is activated is bounded above by $p_{awt }\leq n^{-\alpha}$.
 \end{proof}
 
-Since the probabilty for the activation of an asymmetric witness tree of order $L$ upper-bounds the probability of the bad event that some bin contains more than $L+3$ balls, Equation \ref{eqn:awt:result} also is an upper bound for the latter.
-Note that Equation \ref{eqn:awt:chooseL} can be written as
+Since the probabilty for the activation of an asymmetric witness tree of order $L$ upper-bounds the probability of the bad event that some bin contains more than $L+3$ balls, Equation~\ref{eqn:awt:result} also is an upper bound for the latter.
+Note that Equation~\ref{eqn:awt:chooseL} can be written as
 \begin{align*}
@@
 \end{align*}
-This proves Theorem \ref{theorem:aglm}, under the assumption that all events are stochastically independent (Assumption \ref{assumption:independence}). \qed
+This proves Theorem~\ref{theorem:aglm}, under the assumption that all events are stochastically independent (Assumption~\ref{assumption:independence}). \qed
 
@@
 \label{sec:analysis:nondistinctBalls}
-So far, we have assumed that all balls represented by nodes in a witness tree are stochastically independent (Assumption \ref{assumption:independence}).
+So far, we have assumed that all balls represented by nodes in a witness tree are stochastically independent (Assumption~\ref{assumption:independence}).
 Obviously, to get a valid upper bound on the probability for the activation of a witness tree, the possibility of non-distinct balls has to be accounted for.
@@
 Let this symmetric or asymmetric witness tree be denoted by $T$.
-Then, the probability for the activation of the full witness tree is upper-bounded by the probability for the activation of $T$, implying a probability for activation of this witness tree of $p = n^{- \alpha}$, as shown in Sections \ref{sec:analysis:probabilitySymWT} and \ref{sec:analysis:probabilityAsymWT}, respectively.
+Then, the probability for the activation of the full witness tree is upper-bounded by the probability for the activation of $T$, implying a probability for activation of this witness tree of $p = n^{- \alpha}$, as shown in Sections~\ref{sec:analysis:probabilitySymWT} and~\ref{sec:analysis:probabilityAsymWT}, respectively.
 
@@
 Let $k$ be the number of balls referenced by the pruned witness tree and $l$ be the number of its leaf nodes.
-Then, by the same arguments as in Sections \ref{sec:analysis:probabilitySymWT} and \ref{sec:analysis:probabilityAsymWT}, we can conclude that there are at most $w=n^k$ assignments of balls to the nodes of the witness tree.
+Then, by the same arguments as in Sections~\ref{sec:analysis:probabilitySymWT} and~\ref{sec:analysis:probabilityAsymWT}, we can conclude that there are at most $w=n^k$ assignments of balls to the nodes of the witness tree.
 Furthermore, the probability for the activation of all edge events is bounded from above by
@@
 The number of edges $k$ and the number of leaves $l$ in a pruned witness tree may be small compared to a symmetric or asymmetric witness tree, but we will use the cut-off edges to compensate.
-By definition (Section \ref{sec:analysis:pruningFullWT}), a cut-off edge $e=(v, u)$ represents the event that the ball $b_u$ represented by node $u$ shares a common location $i$ with another ball $b_{u'}$, represented by node $u'$.
+By definition (Section~\ref{sec:analysis:pruningFullWT}), a cut-off edge $e=(v, u)$ represents the event that the ball $b_u$ represented by node $u$ shares a common location $i$ with another ball $b_{u'}$, represented by node $u'$.
 Since $u'$ has already been visited by the time the breadth-first traversal reaches node $v$, it is part of the pruned witness tree and was assigned one of $k\leq K$ balls.
@@
 The single-choice BiB strategy chooses a bin uniformly and independently at random for each ball.
-The classical result for this scheme upper-bounds the gap with \cite{RS98}
+The classical result for this scheme upper-bounds the gap with~\cite{RS98}
 \begin{align*}
@@
 \end{align*}
-w.h.p.~and the Always-go-left scheme due to V\"ocking \cite{VOC03}, which achieves a slightly better gap of
+w.h.p.~and the Always-go-left scheme due to V\"ocking~\cite{VOC03}, which achieves a slightly better gap of
 \begin{align*}
@@
 \end{align*}
-Notice that $F_2$ corresponds to the usual Fibonacci numbers, while $\Phi_2 $ corresponds to the golden ratio \cite{Knuth73}, which is generalized to $\Phi_d$, for which 
+Notice that $F_2$ corresponds to the usual Fibonacci numbers, while $\Phi_2 $ corresponds to the golden ratio~\cite{Knuth73}, which is generalized to $\Phi_d$, for which 
 \begin{align*}
@@
 \label{sec:asymptotic}
-We use the following definitions for the asymptotic notations \cite{CLRS01}, where $f(\cdot)$ and $g(\cdot)$ are positive functions with one argument $n\in \mathbb{N}$ and $n_0 \in \mathbb{N}$, $c \in \mathbb{R}_{\geq 0}$ are suitable constants.
+We use the following definitions for the asymptotic notations~\cite{CLRS01}, where $f(\cdot)$ and $g(\cdot)$ are positive functions with one argument $n\in \mathbb{N}$ and $n_0 \in \mathbb{N}$, $c \in \mathbb{R}_{\geq 0}$ are suitable constants.
 
@@
 \label{sec:historyOfGreedy}
-The history of proving the remarkable property of the two-choice paradigm (see Theorem \ref{theorem:greedy}) is in itself quite interesting.
-It was first proven by Azar et al.
-\cite{ABKU99} for the case $m = n$.
+The history of proving the remarkable property of the two-choice paradigm (see Theorem~\ref{theorem:greedy}) is in itself quite interesting.
+It was first proven by Azar et al.~\cite{ABKU99} for the case $m = n$.
 In the same paper, they claimed that the bound holds for arbitrary $m$.
@@
 Berenbrink at.
-all \cite{BCSV06} were the first to proof that the same bound indeed holds for the heavily loaded case.
-Their proof is fairly complex and a simpler analysis may be found in \cite{TW13}, which also provides a tight lower bound for the heavily loaded case.
+all~\cite{BCSV06} were the first to proof that the same bound indeed holds for the heavily loaded case.
+Their proof is fairly complex and a simpler analysis may be found in~\cite{TW13}, which also provides a tight lower bound for the heavily loaded case.
 Curiously, the case $d = 2$ was implicitly proven by Karp et al.
@@
 
-As discussed in Section \ref{sec:AlwaysGoLeft}, V\"ocking \cite{VOC03} was able to slightly improve the bound as well as to show that the Always-go-left strategy that he introduces is optimal up to additive constants.
+As discussed in Section~\ref{sec:AlwaysGoLeft}, V\"ocking~\cite{VOC03} was able to slightly improve the bound as well as to show that the Always-go-left strategy that he introduces is optimal up to additive constants.
 

### 2014-03-11 article/article.tex
add comma after w.h.p. if sentence continues

@@
 \end{align*}
-Note that, since the average load is 1, $\delta \in \Theta\left( L_{max} \right)$ in this case. For the heavily loaded case, w.h.p.~the maximum gap is ~\cite{RS98}
+Note that, since the average load is 1, $\delta \in \Theta\left( L_{max} \right)$ in this case. For the heavily loaded case, w.h.p., the maximum gap is ~\cite{RS98}
 \begin{align*}
@@
 As with hashing, the BiB model can be applied easily: we need to assign tasks (the balls) to servers (the bins).
-Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is at most $O\left(\ln \ln n\right)$, w.h.p.,~\cite{KLM92,RS98, MRS01}.
+Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is at most $O\left(\ln \ln n\right)$, w.h.p., \cite{KLM92,RS98, MRS01}.
 
@@
 \end{align*}
-w.h.p.~for all $t$~\cite{ABKU99, BCSV06}.
+w.h.p., for all $t$~\cite{ABKU99, BCSV06}.
 \end{theorem}
@@
 \end{align*}
-w.h.p.~for all $t$~\cite{VOC03}.
+w.h.p.,~for all $t$~\cite{VOC03}.
 \end{theorem}
@@
 \end{align*}
-w.h.p.~and the Always-go-left scheme due to V\"ocking~\cite{VOC03}, which achieves a slightly better gap of
+w.h.p., and the Always-go-left scheme due to V\"ocking~\cite{VOC03}, which achieves a slightly better gap of
 \begin{align*}
@@
 \end{align*}
-w.h.p.~for all $m$. We used witness trees, a very useful technique in the BiB context, to prove these upper-bounds for the case $m=n$ in detail.
+w.h.p., for all $m$. We used witness trees, a very useful technique in the BiB context, to prove these upper-bounds for the case $m=n$ in detail.
 

### 2014-03-12 article/article.tex
fix punctuation after equations
reviewing first sections once more

@@
 \title{The Power of Two Random Choices\\ 
-\large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Tr\"aff and Francesco Versaci, lecture ``Parallel Algorithms'', 2012 winter term at TU Wien.}}
+\large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Tr\"aff and Francesco Versaci, ``Parallel Algorithms'' lecture, 2012 winter term at TU Wien.}}
 \author{Martin Kalany, 0825673}
+\maketitle
 
-\maketitle
 \begin{abstract}
@@
 These balls-into-bins games (BiB) have been studied extensively (see, e.g.,~\cite{JK77}).
-In the context of Computer Science, the BiB model is useful to study, e.g., strategies for distributing tasks (balls) to processors (bins) as evenly as possible without having to query all the processors for the number of tasks currently assigned to them.
+In the context of computer science, the BiB model is useful to study, e.g., strategies for distributing tasks (balls) to processors (bins) as evenly as possible without having to query all the processors for the number of tasks currently assigned to them.
 
@@
 \begin{align*}
-\delta = \load - \frac{m}{n}.
+\delta = \load - \frac{m}{n} \quad .
 \end{align*}
@@
 Using this strategy, the maximum load of the fullest bin will be
-\begin{comment}
-\footnote{A more accurate result exists: The maximum load is $\load = \Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$~\cite{G91}.}
-\end{comment}
 \begin{align*}
@@
 \begin{align*}
-\load\in \Theta\left(\frac{\ln n}{\ln \ln n}\right).
+\load\in \Theta\left(\frac{\ln n}{\ln \ln n}\right)\quad .
 \end{align*}
-Note that, since the average load is 1, $\delta \in \Theta\left( L_{max} \right)$ in this case. For the heavily loaded case, w.h.p., the maximum gap is ~\cite{RS98}
+Note that, since the average load is 1, $\delta \in \Theta\left( L_{max} \right)$ in this case. For the heavily loaded case~\cite{RS98}, the maximum gap is, w.h.p.,
 \begin{align*}
-\delta = \Theta\left(\sqrt{\frac{m  \ln n}{n}}\right).
+\delta = \Theta\left(\sqrt{\frac{m  \ln n}{n}}\right)\quad .
 \end{align*}
@@
 \subsection{Multiple-choice BiB}
-Consider a variation of the  single-choice BiB strategy: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin.
-If there are multiple bins with the same minimal load, chose one among them uniformly and independently at random.
-This simple multiple-choice scheme, called the \emph{Greedy} algorithm, achieves a gap
+Consider a variation of the  single-choice BiB strategy: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin.
+If there are multiple bins with the same minimal load, choose one among them arbitrarily.
+This simple multiple-choice scheme, called the \emph{Greedy} algorithm, achieves a maximum load of
 \begin{align*}
-\delta = \frac{\ln \ln n}{\ln d} + O(1),
+\load = \frac{m}{n}+\frac{\ln \ln n}{\ln d} + O(1) \Rightarrow \delta = \frac{\ln \ln n}{\ln d} + O(1)
 \end{align*}
-w.h.p., for all $m$~\cite{ABKU99, BCSV06}, i.e., this additional choice leads to an exponential reduction in the maximum load. 
+w.h.p., for all $m$~\cite{ABKU99, BCSV06}, i.e., this additional choice leads to an exponential reduction of the additional gap. 
 \begin{comment}
@@
 \end{comment}
-Extensive research following the influential paper by Azar et al.~\cite{ABKU99}, who were the first ones to explicitly prove this phenomenon\footnote{For a brief history of the proof, which is quite interesting by itself, refer to Appendix \ref{sec:historyOfGreedy}.}, suggests that this \emph{two-choice paradigm} is not just an artifact of the simple BiB model, but a general principle applicable to a variety of problems.
+Extensive research following the influential paper by Azar et al.~\cite{ABKU99}, who were the first ones to explicitly prove this phenomenon\footnote{For a brief history of the proof, which is quite interesting by itself, refer to Appendix \ref{sec:historyOfGreedy}.}, suggests that this \emph{two-choice paradigm} is not just an artefact of the simple BiB model, but a general principle applicable to a variety of problems.
 
-Consider the slightly more involved \emph{Always-go-left} scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$ (To simplify the presentation, we will assume that $d$ divides $n$ evenly in the remainder of this survey).
+Consider the slightly more involved \emph{Always-go-left} scheme: Suppose that the $n$ bins are partitioned evenly into $d$ groups, each of size $\lfloor \frac{n}{d}\rfloor$ or $\left\lceil\frac{n}{d}\right\rceil$ (to simplify the presentation, we will assume that $d$ divides $n$ evenly in the remainder of this survey).
 We select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
-For tie-breaking, we assume a fixed ordering of bins $i$ for $1 \leq i \leq n$ and put the ball into the bin with smallest index $i$ among those with same minimal load.
-Somewhat surprisingly, V\"ocking~\cite{VOC03} showed that such a strategy employing asymmetric tie-breaking in combination with non-uniform bin selection, reduces the bound for the gap to\
+For tie-breaking, we assume a fixed ordering of bins $i$ for $1 \leq i \leq n$ and put the ball into the bin with smallest index $i$ among those with the same minimal load.
+Somewhat surprisingly, V\"ocking~\cite{VOC03} showed that such a strategy employing asymmetric tie-breaking in combination with non-uniform bin selection reduces the bound for the gap to\
 \begin{align*}
@@
 (The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers.
-Refer to Appendix \ref{sec:fibonacci} for the definition.) Not only is this a significant reduction of the asymptotic bound, V\"ocking was able to show that the Always-go-left strategy is optimal up to additive constants~\cite{VOC03}.
+Refer to Appendix \ref{sec:fibonacci} for the definition.) V\"ocking was also able to show that the Always-go-left strategy is optimal up to additive constants~\cite{VOC03}.
 
-\emph{Outline:}
+\subsubsection*{Outline}
 In Section \ref{sec:classification}, we give a basic classification of BiB games to provide an overview of the research, followed by some applications in Section \ref{sec:applications}.
-Section \ref{sec:model} gives a detailed definition of the model we use, which can be classified as a sequential, finite and dynamic process with random deletions.
-In Section \ref{sec:multiple-choice} we provide a brief definition of three different classes of random placement strategies and present two placement algorithms outlined in the introduction: The Greedy algorithm (Section \ref{sec:greedy}) and the Always-go-left algorithm (Section \ref{sec:AlwaysGoLeft}).
+Section \ref{sec:model} contains a detailed definition of the model we use, which can be classified as a sequential, finite and dynamic process with random deletions.
+In Section \ref{sec:multiple-choice} we provide a brief definition of three different classes of random placement strategies and present two placement algorithms outlined in the introduction: Greedy (Section \ref{sec:greedy}) and Always-go-left (Section \ref{sec:AlwaysGoLeft}).
 Section \ref{sec:analysis} contains a detailed analysis of both strategies for the case $m=n$.
 
-\section{Classification}
+\section{Classification of BiB games}
 \label{sec:classification}
@@
 \begin{compactitem}
-\item In a \emph{sequential} process, the balls arrive one after another and ball is dealt with only after the previous one has been placed.
+\item In a \emph{sequential} process, the balls arrive one after another and a ball is dealt with only after the previous one has been placed.
 Conversely, in a \emph{parallel} process, balls arrive in sets and the balls in one set are dealt with simultaneously~\cite{ABS98}.
@@
 In the random deletions model, for each newly inserted ball another one is chosen uniformly and independently at random and removed from the system~\cite{ABKU99, MRS01}.
-More elaborate schemes exists: An oblivious adversary might specify the sequence of insertions and deletions of balls in advance, without any knowledge of their placement~\cite{CFM+98}.
-Adler et al.~\cite{ABS98} studied another approach in the context of the parallel BiB problem, where the balls are stored in FIFO-order and at every time step the first ball in each of the bins is deleted.
-\item A \emph{finite} process is analyzed for a given time interval that is known beforehand, whereas for an \emph{infinite} one, placements and deletions are given for an infinite time span~\cite{ABS98}.
-\item In the lightly loaded case, $m$ is in the order of $n$, whereas in the heavily loaded case, $m$ is arbitrarily large. W.l.o.g.~we will assume that $m = n$ for the lightly loaded case. 
+More elaborate schemes exists: An oblivious adversary might specify a sequence of insertions and deletions of balls in advance, without any knowledge of their placement~\cite{CFM+98}.
+Adler et al.~\cite{ABS98} studied another approach in the context of the parallel BiB problem, where the balls are stored in FIFO-order and at every time step the first ball in each bin is deleted.
+\item A \emph{finite} process is analysed for a given time interval that is known beforehand, whereas for an \emph{infinite} one, placements and deletions are given for an infinite time span~\cite{ABS98}.
+\item In the lightly loaded case, $m$ is in the order of $n$, whereas in the heavily loaded case, $m$ is arbitrarily large. W.l.o.g., we will assume that $m = n$ for the lightly loaded case. 
 \end{compactitem}
@@
 \label{sec:applications}
-The two-choice paradigm applied to the BiB model has several useful applications, some of which we will outline here.
-Hashing (Section \ref{sec:hashing}), online load balancing (Section \ref{sec:loadbalancing}) and the emulation of shared memory machines on distributed memory machines (Section \ref{sec:DMM}) share a close relation to load balancing.
-Section \ref{sec:circuitrouting} applies the BiB model and the two-choice paradigm to a different context, namely routing circuits in multi-stage interconnection networks with minimal congestion.
+The two-choice paradigm applied to the BiB model has several useful applications, e.g., hashing (Section \ref{sec:hashing}), online load balancing (Section \ref{sec:loadbalancing}), the emulation of shared memory machines on distributed memory machines (Section \ref{sec:DMM}) and routing circuits in multi-stage interconnection networks with minimal congestion (Section~\ref{sec:circuitrouting}).
 
@@
 \label{sec:hashing}
-A simple hash table implementation typically uses a single hash function to map elements to entries in a table.
-If more than one element is mapped to the same entry, all elements mapped to this table entry are stored in a chain, usually implemented as a linked list.
+A basic hash table implementation typically uses a single hash function to map elements to entries in a table.
+If several elements are mapped to the same entry, they are stored in a chain, usually implemented as a linked list.
 The maximum time to search for an element is then proportional to the length of the longest chain in the hash table.
 
-It is easy to see how the BiB model can be used: Assuming a perfect hash function\footnote{I.e., each element is mapped to one of the $n$ table entries uniformly and independently at random}, the elements can be seen as balls to be mapped to table entries (bins), implying a worst case look-up time of $\Theta\left(\frac{\ln n}{\log \ln n}\right)$, w.h.p., where $n$ denotes the number of elements.
-Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries, results in a look-up time of $\Theta\left(\ln \ln n\right)$, w.h.p.
-This approach is easy to parallelize, it does not require re-hashing of data and requires only two hash functions~\cite{ABKU99, MRS01}.
+It is easy to see how the BiB model can be applied: Assuming a perfect hash function\footnote{I.e., each element is mapped to one of the $n$ table entries uniformly and independently at random}, the elements can be seen as balls to be mapped to table entries (bins), thus implying a worst case look-up time of $\Theta\left(\frac{\ln n}{\ln \ln n}\right)$, w.h.p., where $n$ denotes the number of elements.
+Applying the two-choice paradigm by using two hash functions and mapping elements to the least full of two table entries results in a look-up time of $\Theta\left(\ln \ln n\right)$, w.h.p.
+This approach is easy to parallelize, it does not require re-hashing of data and uses only two hash functions~\cite{ABKU99, MRS01}.
 
@@
 While a central dispatcher can easily achieve this, it also poses a bottleneck within a distributed system, since querying $m$ servers for their current load involves significant overhead (e.g., sending a message to each server and waiting for the reply).
-As with hashing, the BiB model can be applied easily: we need to assign tasks (the balls) to servers (the bins).
-Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is at most $O\left(\ln \ln n\right)$, w.h.p., \cite{KLM92,RS98, MRS01}.
+The BiB model can be applied easily: we need to assign tasks (the balls) to servers (the bins).
+Assuming tasks are assigned to servers uniformly and independently at random, the two-choice paradigm is directly applicable: If each client samples two random servers for their load and sends its request to the least loaded server, the overhead is comparatively small and the difference in load on the $n$ servers is $\Theta\left(\ln \ln n\right)$, w.h.p., \cite{KLM92,RS98, MRS01}.
 
@@
 Thus, ways to emulate a PRAM on a DMM have been studied extensively.
-Such an emulation distributes the processors and memory cells of the PRAM to the processors and memory modules (typically one per processor) of the DMM, with the goal to minimize the slowdown of the emulation, i.e., the time the DMM requires to simulate one step of the PRAM execution.
-Each memory module can handle only one memory access at a time step and more accesses to the same module will result in memory contention.
-The two-choice paradigm is useful for managing communication between the processors and the memory modules such that memory contention is minimal~\cite{KLM92}.
+Such an emulation distributes the processors and memory cells of the PRAM to the processors and memory modules of the DMM, with the goal to minimize the slowdown of the emulation, i.e., the time the DMM requires to simulate one step of the PRAM execution.
+Each memory module can handle only one memory access at a time step and more accesses to the same module result in memory contention.
+The two-choice paradigm is useful for managing communication between the processors and the memory modules such that memory contention is minimized~\cite{KLM92}.
 
@@
 In multi-stage interconnection networks, virtual circuit-switching is used to route communication.
-For simplicity, we consider the permutation routing problem, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes and we assume a butterfly network as the underlying interconnection network~\cite{CLR09}.
-The congestion is the maximum number of paths that are routed through a network link and the goal of a circuit routing algorithm is to to allocate paths for all $n$ communication requests such that link congestion is minimized.
+For simplicity, we consider the permutation routing problem, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes.
+Furthermore, we assume a butterfly network as the underlying interconnection network~\cite{CLR09}.
+The link congestion is the maximum number of paths that are routed through a network link and the goal of a circuit routing algorithm is to allocate paths for all $n$ communication requests such that the link congestion is minimized.
 
-In this context, Valiant's paradigm~\cite{V82} states that any such permutation routing problem can be solved by transforming it into two problems: First, find a path for each request to an intermediate destination chosen uniformly and independently at random.
+In this context, Valiant's paradigm~\cite{V82} states that any such permutation routing problem can be solved by transforming it into two sub-problems: First, find a path for each request to an intermediate destination chosen uniformly and independently at random.
 Next, find a path from the intermediate to the actual destination.
@@
 They are based on a sequential, infinite and dynamic process that uses an oblivious adversary, which provides a sequence $\sigma = \sigma_1 \sigma_2 \dots$, where $\sigma_t = (d_t, a_t)$ represents a request for a ball $d_t$ to be deleted and a different ball $a_t$ to be added to the bins at time step $t$.
-In our model, the ball $b_d$ associated with some request $\sigma_t$ is deleted at the beginning of time step $t$.
-Next, the ball $b_a$ has to be placed into one of the $n$ bins without knowledge of the following requests $\sigma_{t+1}$, $\sigma_{t+2}\dots$.
-This infinite sequence of insertions and deletions is fixed in advance and is thus independent of previously allocated balls and any decisions made by the allocation algorithm.
+In our model, the ball $d_t$ associated with some request $\sigma_t$ is deleted at the beginning of time step $t$, next, the ball $a_t$ has to be placed into one of the $n$ bins, without any knowledge of the following requests $\sigma_{t+1}$, $\sigma_{t+2},\dots$ This infinite sequence of insertions and deletions is fixed in advance and is thus independent of previously allocated balls and any decisions made by the allocation algorithm.
 We assume that at any time $t$ at most $m = n$ balls are in the bins.
@@
 The single-choice BiB strategy (Section \ref{sec:single-choice}) chooses one bin uniformly and independently at random for each ball.
-A multiple-choice BiB strategy generalizes this idea by  choosing $d \geq 2$ bins and placing the ball into the least full bin.
-We distinguish placement algorithms depending on how the $d$ locations are sampled~\cite{VOC03}.
+A multiple-choice BiB strategy generalizes this idea by  choosing $d \geq 2$ bins and placing the ball into the least full.
+We distinguish placement algorithms depending on how the $d$ locations (i.e, bins) are sampled~\cite{VOC03}.
 
@@
 \begin{description}
-\item [Uniform and independent.] Choose each of the $d$ bins for a ball uniformly and independently at random from the set of bins $B$.
-Thus, the probability that any specifc bin $b$ is choosen for the $i$-th balls is
+\item [Uniform and independent:] Choose each of the $d$ bins for a ball uniformly and independently at random from the set of bins $B$.
+Thus, the probability that any specific bin $b$ is chosen for the $i$-th balls is
 \begin{align*}
-\Pr[\mathrm{bin}\;b\;\mathrm{choosen}] = \frac{1}{n}
+\Pr[\mathrm{bin}\;b\;is\;\mathrm{chosen}] = \frac{1}{n} \quad .
 \end{align*}
-\item [Non-uniform and independent.] Choose the $i$-th bin from $B$ independently at random according to some probability distribution: 
-\item [Non-uniform and dependent.] Choose the $d$ bins for a ball from the sample space according to some probability probability distribution:
+\item [Non-uniform and independent:] Choose the $i$-th bin from $B$ independently at random, according to some probability distribution.
+\item [Non-uniform and dependent:] Choose the $d$ bins for a ball from the sample space according to some probability probability distribution.
 \end{description}
 
-We will discuss two different multiple-choice placement strategies: The Greedy algorithm (Section \ref{sec:greedy}), belonging to class 1, and the Always-go-left scheme (Section \ref{sec:AlwaysGoLeft}), which uses a non-uniform and independent selection of $d$ bins and thus belongs to class 2.
+We will discuss two different multiple-choice placement strategies: The Greedy algorithm (Section \ref{sec:greedy}), belonging to class 1, and the Always-go-left scheme (Section \ref{sec:AlwaysGoLeft}), which belongs to class 2.
 
@@
 \label{sec:greedy}
-Consider the Greedy scheme, a variation of the one-choice BiB strategy: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin.
+Consider the Greedy scheme: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin.
 If there are multiple bins with the same minimal load, chose one among them arbitrarily.
@@
 \label{theorem:greedy}
-Consider any sequence of insertions and deletions s.t.~at most $n$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, the maximum load is at most
+Consider any sequence of insertions and deletions s.t.~at most $n$ balls are in the bins at any time. If the balls are placed into bins by the Greedy algorithm, then the maximum load is at most
 \begin{align*}
-\load = \frac{\ln \ln n}{\ln d}+ O(1)
+\load \leq \frac{\ln \ln n}{\ln d}+ O(1) \quad ,
 \end{align*}
@@
 \end{theorem}
-
 This result is remarkable in several ways:
@@
 \label{sec:AlwaysGoLeft}
-The Greedy strategy presented in Section~\ref{sec:greedy} assumes that $d$ bins are chosen uniformly and independently at random.
-V\"ocking~\cite{VOC03} was the first to study the impact of non-uniform and dependent selections of the $d$ bins.
-Surprisingly, a multiple-choice BiB strategy with a non-uniform and independent selection of the $d$ bins, such as the Always-go-left algorithm introduced in~\cite{VOC03}, achieves better~\cite{BCSV06} load balancing than the Greedy scheme.
+The Greedy strategy presented in Section~\ref{sec:greedy} selects the $d$ bins uniformly and independently at random.
+Surprisingly, a multiple-choice BiB strategy with a non-uniform and independent selection of the bins, such as the Always-go-left algorithm introduced by V\"ocking in~\cite{VOC03}, achieves better load balancing than the Greedy scheme~\cite{BCSV06}.
 
 The Always-go-left algorithm works as follows.
-Partition the $n$ bins into $d$ groups, each of size $\frac{n}{d}$.
-We select one from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
-For tie-breaking, we assume a fixed ordering of bins $b_i$ for $1 \leq i \leq n$ and put the ball into the bin with smallest index $i$ among those with same minimal load.
-
-\begin{comment}
-\begin{theorem}
-\label{theorem:agln}
-If $n$ balls are placed into $n$ bins using the Always-go-left scheme, the maximum load will be
-\begin{align*}
-\load \leq \frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
-\end{align*}
-w.h.p.
-\end{theorem}
-\end{comment}
+We partition the $n$ bins into $d$ groups (each of size $n/d$), select one bin from each group uniformly and independently at random and put the ball into the least full of those $d$ bins.
+For tie-breaking, we assume a fixed ordering of bins and put the ball into the bin with smallest index among those with same minimal load.
 
@@
 \begin{align*}
-\load = \frac{\ln\ln n}{d  \ln \Phi_d} + O(1)
+\load \leq \frac{\ln\ln n}{d  \ln \Phi_d} + O(1) \quad ,
 \end{align*}
@@
 In this section, we will prove Theorems \ref{theorem:greedy} and \ref{theorem:aglm} in detail, using witness trees to provide an upper bound for the event that a bin contains too many balls.
-Our analysis closely follows the proofs in~\cite{VOC03}.
+(Our analysis closely follows~\cite{VOC03}).
 
-When the maximum load exceeds some given threshold value we say that a bad event occurs and construct an activated witness tree.
+When the maximum load exceeds some given threshold we say that a bad event occurs and construct an activated witness tree.
 Thus, the activation of a witness tree is implied by a bad event.
@@
 
-Throughout this section, ``location'' will denote a specific place in one of the $n$ bins.
-A location of a ball $b$ refers to a specific bin and to the number of balls already in it (at the time of insertion of ball $b$).
-
 We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections~\ref{sec:analysis:greedy} and~\ref{sec:analysis:alg}, respectively.
-However, we use a simplifying assumption (Assumption~\ref{assumption:independence}) in those two sections to provide a more clear presentation of the proof technique.
-Section~\ref{sec:analysis:nondistinctBalls} will extend these proofs to remove the simplifying assumption we make initially.
+However, we will adopt a simplifying assumption to provide a more clear presentation of the proof technique.
+Section~\ref{sec:analysis:nondistinctBalls} will extend these proofs to remove this assumption.
 
 We will use a proof technique called witness trees, which allows for a clean analysis of both problems and is applicable to a wide variety of BiB scenarios.
-Other proof techniques such as layered induction and a differential equations approach have been used.
-See, e.g.,~\cite{MRS01} for an overview of different proof techniques for the BiB problem.
-The proof for arbitrarily large $m$ is more complicated and we refer the interested reader to~\cite{BCSV06, TW13}.
+Other proof techniques such as layered induction and a differential equation approaches have been used in the literature; see, e.g.,~\cite{MRS01} for an overview of different proof techniques for the BiB problem.
+The proof for arbitrarily large $m$ is more involved and we refer the interested reader to~\cite{BCSV06, TW13}.
 
@@
 \label{sec:analysis:definitionSymWT}
-A symmetric witness tree of order $L$ is a complete $d$-ary tree with $d^{L}$ leaves, where each node $v$ references a ball $b_v = \mathrm{ball}(v)$ and the nodes and edges of the witness tree represent events that may occur or not.
+A symmetric witness tree (SWT) of order $L$ is a complete $d$-ary tree with $d^{L}$ leaves, where each node $v$ references a ball $b_v = \mathrm{ball}(v)$ and the nodes and edges of the SWT represent events that may occur or not.
 In general, the same ball may be referenced by several nodes.
@@
 \label{assumption:independence}
-All events represented by the nodes and edges of a witness tree are stochastically independent. 
+All the events represented by the nodes and edges of a witness tree are stochastically independent. 
 \end{assumption} 
-In other words, we assume that no ball is referenced by more than one node of a witness tree.
+\noindent
+This implies that no ball is referenced by more than one node of a witness tree.
 A correct analysis will have to consider balls assigned to more than one node, but this simplifying assumption is useful to highlight the differences between the symmetric and asymmetric allocation schemes.
@@
 We define edge and leaf events:
-\begin{compactitem}
-\item \emph{Edge event:} Consider some edge $e = (u,v)$ of the witness tree, s.t.~$v$ is the $i$-th child of $u$.
-Then $e$ represents the event that the $i$-th location of the ball $b_u$ to the same bin as one of the locations of the ball $b_v$.
-\item \emph{Leaf event:} A leaf node $v$ of the witness tree represents the event that each of the $d$ locations of the ball $b_v$ points to a bin that contains at least three additional balls (i.e., balls that are not represented by a node of the tree), at the time of the insertion of the ball $b_v$.
-
-\end{compactitem}
-The definition for leaf events refers to bins containing some number of balls at some specified time.
-To avoid dependencies, the crucial trick is to defone edge events in terms of the possible locations of balls, not by their finally assigned bin.
+\begin{description}
+\item [Edge event:] Consider some edge $e = (u,v)$ of the witness tree, s.t.~$v$ is the $i$-th child of $u$.
+The edge $e$ represents the event that the $i$-th location of the ball $b_u$ points to the same bin as one of the locations of the ball $b_v$.
+\item [Leaf event:] A leaf node $v$ of the witness tree represents the event that each of the $d$ locations of the ball $b_v$ points to a bin that contains at least three additional balls (i.e., balls that are not represented by a node of the tree), at the time of the insertion of the ball $b_v$.
+\end{description}
+To avoid dependencies, the crucial step is to define edge events in terms of the possible locations of balls, not by their finally assigned bin.
 
 The following constraints have to be fulfilled by a witness tree. 
-The ball $b_r$ that is represented by the root node $r$ has to exist at time step $t$ and each ball $b_v$ represented by a node $v \neq r$ and parent node $u$ has to exist when the ball $b_u$ is inserted.
-
+\begin{enumerate}
+\item The ball $b_r$ that is represented by the root node $r$ has to exist at time step $t$. 
+\item Each ball $b_v$ represented by a node $v \neq r$ with parent node $u$ has to exist when the ball $b_u$ is inserted.
+\end{enumerate}
 We say that an edge or leaf event is activated if the associated event occurs. A witness tree is activated when all its edge and leaf events are activated.
 
-Assume the bad event that some bin contains more than $L+3$ balls at time $t$ occurs.
-We will use an activated symmetric witness tree of order $L$ to witness this event and provide an upper-bound for the occurrence of the event.
-Thus, we have to show that the existence of a bin $x$ containing at least $L+4$ balls implies the existence of an activated witness tree of order $L$.
+We say that a ``bad event'' occurs when a bin contains more than $L+3$ balls at time $t$.
+We will use an activated SWT of order $L$ to witness this event and provide an upper bound for its occurrence.
+Thus, we have to show that the existence of a bin containing at least $L+4$ balls implies the existence of an activated SWT of order $L$.
 
 \begin{lemma}
-When using the Greedy algorithm to put balls into bins and arbitrary $L > 0$, a bin containing more than $L+3$ balls implies the activation of a symmetric witness tree of order $L$.
+When using the Greedy algorithm to put balls into bins, a bin containing more than $L+3$ balls activates an SWT of order $L$.
 \end{lemma}
@@
 \begin{proof}
-We construct a witness tree for the event that some bin contains $L+4$ balls as follows:
+We construct an SWT for the event that some bin $x$ contains $L+4$ balls as follows:
 \begin{compactenum}
-\item The topmost ball $b_1$ in bin $x$ is assigned to the root node $v_r$. Note that $b_1$ is the last ball that was inserted into the bin $x$.
-\item Since the Greedy-scheme puts ball $b_1$ into bin $x$ only if each of the $d$ sampled locations containd at least the same number of balls as bin $x$, we can conclude that each of the $d$ locations containd at least $L+3$ balls at the time ball $b_1$ was inserted.
+\item The topmost ball $b_1$ in bin $x$ is assigned to the root node $v_r$. Note that $b_1$ is the latest ball that was inserted into the bin $x$.
+\item Since the Greedy scheme puts ball $b_1$ into bin $x$ only if each of the $d$ sampled locations contains at least the same number of balls as bin $x$, we can conclude that each of the $d$ locations contains at least $L+3$ balls at the time ball $b_1$ was inserted.
 Thus, the $d$ children of the root node are assigned the topmost balls of these $d$ bins.
@@
 It follows that each ball assigned to a leaf node is on top of at least 3 other balls, implying that the $d$ locations of each leaf node point to bins with at least 3 balls in each of them.
-This implies the activation of all leaf nodes of the witness tree.
+This implies the activation of all the leaf nodes of the witness tree.
 \end{proof}
@@
 \label{sec:analysis:probabilitySymWT}
-In this section, we derive the probability that some witness tree of order $L$ is activated, which gives an upper bound for the probability that the maximum load is greater than $L+3$ at some time $t$.
+In this section, we derive the probability that some SWT of order $L$ is activated, which gives an upper bound for the probability that the maximum load is greater than $L+3$ at some time $t$.
 
 \begin{lemma}\label{lemma:swt:activation}
-For some arbitrary $\alpha$ and large enough $L$, the activation of a symmetric witness tree of order $L$ occurs with probability at most $n^{-\alpha}$.
+For any $\alpha>0$ and large enough $L$, the activation of an SWT of order $L$ occurs with probability at most $n^{-\alpha}$.
 \end{lemma}
-Speaking less formaly, Lemma~\ref{lemma:swt:activation} states that for large enough $L$, the activation of a symmetric witness tree is unlikely.
-
 \begin{proof}
-The probability $p_{swt}$ that some symmetric witness tree of order $L$ is activated is bounded by the number of witness trees $w$ times the upper bound on the probability $p_1$ that a specific witness tree is activated.
+The probability $P_\mathrm{S}$ that some SWT of order $L$ is activated is bounded by the number of witness trees $w$ times the upper bound on the probability that a specific witness tree is activated.
 
 The number of different witness trees $w$ is given by the number of ways to assign balls to its nodes.
-The ball assigned to the root node can be chosen from at most $n$ balls, since by definition (Theorem~\ref{theorem:greedy}), at most $n$ balls are in the bins at any time $t$ and a ball assigned to a node has to be in some bin at time $t$.
-By the same argument, each child node of the root is assigned one out of at most $n$ balls too.
-Applying this argument for each level of the witness tree, there are at most $w = n^k$ possibilities to assign balls to a witness tree, with $k = d^{L+1}-1$ being the number of nodes in a $d$-ary witness tree of order $L$.
+The ball assigned to the root node can be chosen from at most $n$ balls, since by definition, at most $n$ balls are in the bins at any time $t$ and a ball assigned to a node has to be in some bin at time $t$.
+By applying this argument recursively to all levels of the witness tree, we get that there are at most $w \leq n^k$ possibilities to assign balls to a witness tree, with
+\begin{align*}
+k = \frac{d^{L+1}-1}{d-1}
+\end{align*}
+being the number of nodes in a $d$-ary witness tree of order $L$.
 
-An edge\footnote{where $v$ is the $i$-th child of $u$} $(u, v)$ is activated with probability at most $d/n$, which is the probability that the $i$-th location of the ball $b_u$ points to a particular location of the ball $b_v$ times the number of locations $d$.
+An edge $(u, v_i)$ (where $v_i$ is the $i$-th child of $u$) is activated with probability at most $d/n$, which is the probability that the $i$-th location of the ball $b_u$ points to a particular location of the ball $b_{v_i}$ times the number of locations $d$.
 This implies a probability of at most
 \begin{align*}
-p_e = \left(\frac{d}{n}\right)^{k-1}
+p_\mathrm{e} = \left(\frac{d}{n}\right)^{k-1} \quad ,
 \end{align*} 
-for the event that all $k-1$ edges of a witness tree are activated.
-Note that  the assumption that all balls are distinct (Assumption~\ref{assumption:independence}) implies that all edge events are independent.
+for the event that all $k-1$ edges of a witness tree are activated. Note that since all the balls are distinct (Assumption~\ref{assumption:independence}), all the edge events are independent.
 
-The probability that a leaf event is activated is bounded above by $3^{-d}$.
-Recall that by the definition of leaf events, each of the $d$ locations of a ball $b_{v_l}$, where $v_l$ is a leaf node, has to point to a bin with at least 3 additional balls in it.
-The number of those bins is bounded by $\lfloor n/3 \rfloor$ at all time steps $t$, implying a probability of at most
+The probability that a leaf event is activated is bounded from above by $3^{-d}$: Recall that by the definition of leaf events, each of the $d$ locations of a ball $b_v$, where $v$ is a leaf node, has to point to a bin with at least 3 additional balls in it.
+The number of those bins is bounded by $\lfloor n/3 \rfloor$ at all time steps $t$, implying a probability of at most $1/3$ that one such bin is chosen.
+This implies a probability of at most $3^{-d}$ that all the $d$ bins are chosen from among those with at least 3 balls in them.
+Thus, the probability that all leaf events are activated is at most $p_\mathrm{l} = 3^{-d l}$, where $l = d^{L}$ is the number of leaves in an SWT.
+
+Thus, the probability that an activated SWT exists at time $t$ is at most 
 \begin{align*}
-\frac{n/3}{n} = \frac{1}{3}
-\end{align*}
-that one such bin is chosen.
-This implies a probability of at most $3^{-d}$ that all $d$ bins are chosen from among those with at least 3 additional balls in them.
-Thus, the probability that all leaf events are activated is at most $p_l = 3^{-d l}$, where $l = d^{L}$ is the number of leaves in a witness tree.
-
-Thus, the probability that an activated symmetric witness tree exists at time $t$ is at most 
-\begin{align*}
-p_{swt} &= w  p_e  p_l \\
-		&= n^{k}  \left(\frac{d}{n}\right)^{k-1}  3^{-d  l} \\
-		&= n  d^{k-1}  3^{-d  l}
+P_\mathrm{s} &\leq w  p_\mathrm{e}  p_\mathrm{l} 
+		= n^{k}  \left(\frac{d}{n}\right)^{k-1}  3^{-d  l} 
+		= n  d^{k-1}  3^{-d  l}\quad .
 \end{align*}
@@
 \begin{align*}
-p_{swt} &\leq n  d^{2 l}  3^{-d  l} \\
-		&= n  \left(d^2\right)^l  \left( 3^{-d} \right)^l
+P_\mathrm{s} &\leq n  d^{2 l}  3^{-d  l} 
+		= n  \left(d^2\right)^l  \left( 3^{-d} \right)^l\quad ,
 \end{align*}
-which we simplify by using $2d^2 \leq 3^d$
+which we simplify by using $2d^2 \leq 3^d$ as
 \begin{align*}
-p_{swt} &\leq n  \left(\frac{3^d}{2}\right)^l  \left( 3^{-d}\right)^l \\
-		&= n  2^{-l}  3^{d l}  3^{-d  l} \\
-		&= n 2^{-l}		\\
-		&= n 2^{-d^L}		
+P_\mathrm{s} &\leq n  \left(\frac{3^d}{2}\right)^l  \left( 3^{-d}\right)^l 
+		= n  2^{-l}  3^{d l}  3^{-d  l} 
+		= n 2^{-l}		
+		= n 2^{-d^L} \quad .		
 \end{align*}
-
 For any constant $\alpha > 0 $, let
 \begin{align*}
-L &\geq \log_d\log_2 n + \log_d\left(1+\alpha\right) \neqn{swt:chooseL}
+L &\geq \log_d\log_2 n + \log_d\left(1+\alpha\right) \neqn{swt:chooseL} \quad .
 \end{align*}
-
 Then, 
 \begin{align*}
-p_{swt} &\leq n  2^{{-d}^{\log_d\log_2 n + \log_d\left(1+\alpha\right)}} \\
-		&= n  2^{-{d}^{\log_d\log_2 n} d^{\log_d \left(1+\alpha\right)}} \\
-		&= n  \left( 2^{-\log_2 n  \left(1+ \alpha \right)} \right) \\
-		&= n  n^{-\left(1+\alpha\right)} = n^{-\alpha} \neqn{swt:result}
+P_\mathrm{s} &\leq n  2^{{-d}^{\log_d\log_2 n + \log_d\left(1+\alpha\right)}} 
+		= n  2^{-{d}^{\log_d\log_2 n} d^{\log_d \left(1+\alpha\right)}} \\
+		&= n  \left( 2^{-\log_2 n  \left(1+ \alpha \right)} \right) 
+		= n  n^{-\left(1+\alpha\right)} = n^{-\alpha} \neqn{swt:result}\quad .
 \end{align*}
-Thus, for any constant $\alpha$ and large enough $L$ as specified in Equation~\ref{eqn:swt:chooseL} the probability that a symmetric witness tree of order $L$ is activated is bounded above by $p_{swt }\leq n^{-\alpha}$.
+Thus, for any $\alpha>0$ and large enough $L$ as specified in Eq.~\ref{eqn:swt:chooseL} the probability that an SWT of order $L$ is activated is bounded from above as $P_\mathrm{s}\leq n^{-\alpha}$.
 \end{proof}
 
-Since the probabilty for the activation of a symmetric witness tree of order $L$ upper-bounds the probability of the bad event that some bin contains more than $L+3$ balls, Equation~\ref{eqn:swt:result} is also an upper bound for the latter.
-Note that Equation~\ref{eqn:swt:chooseL} can be written as
+Since the probability for the activation of an SWT of order $L$ upper-bounds the probability of the bad event that some bin contains more than $L+3$ balls, Eq.~\ref{eqn:swt:result} is also an upper bound for the latter.
+Note that Eq.~\ref{eqn:swt:chooseL} can be rewritten as
 \begin{align*}
-L &\geq  \frac{\ln\log_2 n}{\ln d} + \frac{\ln(1+\alpha)}{\ln d} \\
-   &= \frac{\ln \log_2 n}{\ln d}  \left(1 + \frac{\ln\left(1+\alpha\right)}{\ln \log_2 n}\right) \\
-  &=  \frac{\ln \ln n}{\ln d}  \left( 1+ o\left( 1\right)\right) \neqn{swt:boundL}
+L &\geq  \frac{\ln\log_2 n}{\ln d} + \frac{\ln(1+\alpha)}{\ln d} 
+   = \frac{\ln \log_2 n}{\ln d}  \left(1 + \frac{\ln\left(1+\alpha\right)}{\ln \log_2 n}\right) \\
+  &=  \frac{\ln \ln n}{\ln d}  \left( 1+ o\left( 1\right)\right) \neqn{swt:boundL} \quad ,
 \end{align*}
-
-This proves Theorem~\ref{theorem:greedy} for the case $m=n$, under the assumption that all events are stochastically independent (Assumption~\ref{assumption:independence}). \qed
+thus proving Theorem~\ref{theorem:greedy}, under Assumption~\ref{assumption:independence} (i.e., that all events are stochastically independent). \qed
 
@@
 \label{sec:analysis:alg}
-Similarly to Section~\ref{sec:analysis:greedy}, where we used symmetric witness trees to upper-bound the probability that a bin contains too many balls when using the Greedy scheme, we will use asymmetric witness trees to get an analogous upper bound for the Always-go-left scheme.
-We will make the same simplification as stated in Assumption~\ref{assumption:independence}, namely that all the events represented by the edges and nodes of a witness tree are stochastically independent.
+Similarly to Section~\ref{sec:analysis:greedy}, where we used symmetric witness trees to upper-bound the probability that a bin contains too many balls when using the Greedy scheme, we will use asymmetric witness trees (AWT) to get an analogous upper bound for the Always-go-left scheme.
+Again, we will use Assumption~\ref{assumption:independence}, i.e., we assume that all the events represented by the edges and nodes of a witness tree are stochastically independent.
 
@@
 \label{sec:analysis:definitionAsymWT}
-With one exception, the definition of an asymmetric witness tree is the same as the definition of a symmetric witness tree, including the definition of activation.
-Only its structure differs from that of a symmetric tree.
-An asymmetric witness tree has the structure of a $d$-ary Fibonacci tree $T_d\left(k \right)$, which is defined as follows.
+An AWT is defined in the same way as an SWT  (Section~\ref{sec:analysis:definitionSymWT}), except for its structure, which is a $d$-ary Fibonacci tree $T_d\left(k \right)$:
+
 \begin{compactitem}
@@
 \end{compactitem}
-%\todo{illustrate and note difference to usual Fibonacci trees}
-
-A $d$-ary asymmetric witness tree of order $L$ has the structure of the Fibonacci tree $T_d(d  L+1)$ and by Equation~\ref{eqn:goldenFib} has $F_d(d L + 1) \geq \Phi_d^{d L-1}$ leaves.
+A $d$-ary AWT of order $L$ has the structure of the Fibonacci tree $T_d(d  L+1)$ and thus has $F_d(d L + 1) \geq \Phi_d^{d L-1}$ leaves (Eq.~\ref{eqn:goldenFib}).
 
 \begin{lemma}
-When using the Always-go-left algorithm to put balls into bins and arbitrary $L > 0$, a bin containing more than $L+3$ balls implies the activation of an asymmetric witness tree of order $L$.
+When using the Always-go-left algorithm to put balls into bins, a bin containing more than $L+3$ balls implies the activation of an AWT of order $L$.
 \end{lemma}
@@
 \item A node that is the root of a Fibonacci tree $T_d(1)$ is assigned the label $(0,1)$.
-\item Analogously, a node that is the root of a Fibonacci tree $T_d(2)$ is assigned the label $(0,2)$. Thus, nodes labeled $(0,1)$ or $(0,2)$ do not have children.
-\item A node with label $(0, i)$, where $i>2$, has $i-1$ children with labels $(0,i-1)\dots (0,1)$
-\item A node with label $(h,i)$ , where $h>0$, has $d$ children with the labels $(h,i-1)\dots(h,1)$,$(h-1,d)\dots(h-1, i)$.
+\item Analogously, a node that is the root of a Fibonacci tree $T_d(2)$ is assigned the label $(0,2)$. Thus, nodes labelled $(0,1)$ or $(0,2)$ do not have any children.
+\item A node with label $(0, i)$, where $i>2$, has $i-1$ children with labels $(0,i-1),\dots,(0,1)$
+\item A node with label $(h,i)$ , where $h>0$, has $d$ children with the labels $(h,i-1),\dots,(h,1)$,$(h-1,d),\dots,(h-1, i)$.
 \item The root of a Fibonacci tree $T_d(d h+i)$ has the label $(h, i)$. 
 \end{compactitem}
-Thus, our the root node of our witness tree will be labeled $(L, 1)$.
-Note that the labels are unique within the set of children of a node $v$, but not for all nodes of the tree.
+Thus, the root node of the witness tree will be labelled $(L, 1)$.
+Note that the labels are unique for the set of children of any node $v$, but not necessarily for all nodes of the tree.
 The labels are useful to state the following invariant, which will be maintained while assigning balls to the nodes of the witness tree.
@@
 \end{invariant}
-
 Let $x$ be the bin that contains $L+4$ balls.
-The root node gets assigned the topmost ball in bin $x$.
-Each of the $d$ locations of this ball points to a bin with at least $L+3$ balls in it, at the time of its insertion.
+The root node gets assigned the topmost ball $b_t$ in bin $x$, thus each of the $d$ locations of this ball points to a bin with at least $L+3$ balls in it, at the time of insertion of $b_t$.
 These are the $d$ bins that were sampled by the Always-go-left scheme and thus bin $i$ belongs to the $i$'th group of bins, where each group consists of $n/d$ bins.
@@
 The assignment of balls to nodes than proceeds recursively.
-Given a node $v$ with label $(h, i)$ whose ball $b_v$ was put into some bin belonging to group $i$ and containing at least $l+3$ other balls, we can conclude the following.
+Given a node $v$ with label $(h, i)$ and referencing ball $b_v$ that was put into some bin belonging to group $i$ and containing at least $l+3$ other balls, we can conclude the following:
 \begin{compactitem}
@@
 \item For $i < j \leq d$, the $j$-th location contains at least $l+3$ balls.
-If location $j$ contained less, the Always-go-left scheme would have put the ball $b_v$ into the bin it points to.
-The topmost ball in the bin that location $j$ points to is assigned to the child node $u$ of $v$ labeled with $(l-1, j)$.
+If location $j$ contained less, the Always-go-left scheme would have put the ball $b_v$ into the bin location $j$ points to.
+The topmost ball in the bin that location $j$ points to is assigned to the child node $u$ of $v$ labelled with $(l-1, j)$.
 \todo{the paper claims that the locations $j$ point to bins with $l+2$ balls.
@@
 \begin{lemma}\label{lemma:awt:activation}
-For some arbitrary $\alpha$ and large enough $L$, the activation of an asymmetric witness tree of order $L$ occurs with probability at most $n^{-\alpha}$.
+For some arbitrary $\alpha$ and large enough $L$, the activation of an AWT of order $L$ occurs with probability at most $n^{-\alpha}$.
 \end{lemma}
@@
 \begin{proof}
-Analogously to the symmetric witness tree, the possible number of asymmetric witness trees is upper-bounded by $w=n^k$ and the probability that all edge events occur is\footnote{As in Section~\ref{sec:analysis:greedy}, $k$ denotes the number of nodes of the witness tree and $l$ the number of leaves.}
+Analogously to the SWT, the possible number of AWTs is upper-bounded by $w\leq	n^k$, and the probability that all edge events occur is\footnote{As in Section~\ref{sec:analysis:greedy}, $k$ denotes the number of nodes of the witness tree and $l$ the number of leaves.}
 \begin{align*}
-p_e = \left(\frac{d}{n}\right)^{k-1}
+p_\mathrm{e} = \left(\frac{d}{n}\right)^{k-1}\quad .
 \end{align*}
-
-The probability that all leaf events occur too is the same as for the symmetric witness tree, but its derivation differs.
-Again, at any time at most $b'=n/3$ bins contain 3 or more balls, but we do not know how those bins are distributed among the groups.
+The probability that all leaf events occur is the same as for the SWT, but its derivation differs.
+Again, at any time at most $n/3$ bins contain 3 or more balls, but we do not know how those bins are distributed among the groups.
 Let $b_i$ be the number of bins in group $i$ that contain 3 or more balls.
@@
 \begin{align*}
-\frac{n}{d}\sum_{i=1}^d \beta_i \leq \frac{n}{3}
+\frac{n}{d}\sum_{i=1}^d \beta_i \leq \frac{n}{3}\quad ,
 \end{align*}
@@
 \begin{align*}
-\sum_{i=1}^d \beta_i \leq \frac{d}{3}
+\sum_{i=1}^d \beta_i \leq \frac{d}{3}\quad .
 \end{align*}
-
 The probability that all $d$ locations of a ball point to a bin containing 3 or more balls is then 
 \begin{align*}
-p_{l_1} = \prod_{i=1}^{d} \beta_i
+p_\mathrm{l}' = \prod_{i=1}^{d} \beta_i \quad ,
 \end{align*}
-In other words, a leaf event occurs with probability $p_{l_1}$, which is maximized when all $b_i$ are equal.
-We thus set $b_i \leq 1/3$ for all $i$ and get as upper bound for the probability that all leaf events occur
+i.e., a leaf event occurs with probability $p_\mathrm{l}'$, which is maximized when all $b_i$ are equal.
+We thus set $b_i = 1/3$ for all $i$ and get as upper bound for the probability that all leaf events occur
 \begin{align*}
-p_l \leq 3^{-d l}
+p_\mathrm{l} \leq 3^{-d l}\quad .
 \end{align*}
-
-The above results imply the same probability for the activation of an asymmetric witness tree as we derived in Section~\ref{sec:analysis:probabilitySymWT} for the activation of a symmetric witness tree, that is
+The above results imply the same probability for the activation of an AWT as for the activation of an SWT (Eq. \ref{eqn:swt:result}); that is
 \begin{align*}
-p_{awt} &\leq n  2^{-l}
+P_\mathrm{a} &\leq n  2^{-l}\quad .
 \end{align*}
@@
 \end{align*}
-holds\footnote{see Section~\ref{sec:analysis:definitionAsymWT} for details}.
-
-For any constant $\alpha > 0$, let 
+for the AWT (Section~\ref{sec:analysis:definitionAsymWT}) whereas $l = d^L$ for the SWT.
+For any $\alpha > 0$, let 
 \begin{align*}
-L &\geq \left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}\right\rceil+\frac{1}{d} \neqn{awt:chooseL}
+L &\geq \left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}\right\rceil+\frac{1}{d} \neqn{awt:chooseL}\quad .
 \end{align*}
@@
 \begin{align*}
-p_{awt} &\leq n  2^{-\Phi_d^{d L -1}} \\
-        &= n  2^{-\Phi_d^{d \left(\left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}\right\rceil+\frac{1}{d}\right) -1}} \\
-        &\leq n  2^{-\Phi_d^{d \left({\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}+\frac{1}{d}\right) -1}} \\
-        &= n  2^{-\Phi_d^{{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}}} \\
-        &= n  2^{-\Phi_d^{\log_{\Phi_d}\log_2 n + \log_{\Phi_d}\left(1+\alpha\right)}} \\
-         &= n  2^{-\Phi_d^{\log_{\Phi_d}\log_2 n} \Phi_d^{ \log_{\Phi_d}\left(1+\alpha\right)}} \\
-         &= n  2^{-\log_2 n \left(1+\alpha\right)} \\
-         &= n  n ^{-(1+\alpha)} = n^{-\alpha} \neqn{awt:result}
+P_\mathrm{a} &\leq n  2^{-\Phi_d^{d L -1}} 
+        = n  2^{-\Phi_d^{d \left(\left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}\right\rceil+\frac{1}{d}\right) -1}} 
+        \leq n  2^{-\Phi_d^{d \left({\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}+\frac{1}{d}\right) -1}} \\
+        &= n  2^{-\Phi_d^{{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}}} 
+        = n  2^{-\Phi_d^{\log_{\Phi_d}\log_2 n + \log_{\Phi_d}\left(1+\alpha\right)}} 
+        = n  2^{-\Phi_d^{\log_{\Phi_d}\log_2 n} \Phi_d^{ \log_{\Phi_d}\left(1+\alpha\right)}} \\
+         &= n  2^{-\log_2 n \left(1+\alpha\right)} 
+         = n  n ^{-(1+\alpha)} = n^{-\alpha} \neqn{awt:result} \quad .
 \end{align*}
-Thus, for any constant $\alpha$ and large enough $L$ as specified in Equation~\ref{eqn:awt:chooseL} the probability that a symmetric witness tree of order $L$ is activated is bounded above by $p_{awt }\leq n^{-\alpha}$.
+Thus, for any $\alpha>0$ and large enough $L$ as specified in Eq.~\ref{eqn:awt:chooseL}, the probability that an AWT of order $L$ is activated is bounded from above as $P_\mathrm{a}\leq n^{-\alpha}$.
 \end{proof}
-
-Since the probabilty for the activation of an asymmetric witness tree of order $L$ upper-bounds the probability of the bad event that some bin contains more than $L+3$ balls, Equation~\ref{eqn:awt:result} also is an upper bound for the latter.
-Note that Equation~\ref{eqn:awt:chooseL} can be written as
+Since the probability for the activation of an AWT of order $L$ upper-bounds the probability of the bad event that some bin contains more than $L+3$ balls, Eq.~\ref{eqn:awt:result} is also an upper bound for the latter.
+Note that Eq.~\ref{eqn:awt:chooseL} can be written as
 \begin{align*}
-L &\geq \frac{\ln\log_2 n}{d \ln \Phi_d} + \frac{\ln\left(1+\alpha\right)}{d \ln \Phi_d} + \frac{1}{d}\\
-  &\geq \frac{\ln\ln n}{d \ln \Phi_d} + O\left(1\right)
+L &\geq \frac{\ln\log_2 n}{d \ln \Phi_d} + \frac{\ln\left(1+\alpha\right)}{d \ln \Phi_d} + \frac{1}{d}
+  \geq \frac{\ln\ln n}{d \ln \Phi_d} + O\left(1\right)\quad ,
 \end{align*}
-This proves Theorem~\ref{theorem:aglm}, under the assumption that all events are stochastically independent (Assumption~\ref{assumption:independence}). \qed
+thus proving Theorem~\ref{theorem:aglm}, under Assumption~\ref{assumption:independence} (i.e., that all events are stochastically independent). \qed
 
@@
 Obviously, to get a valid upper bound on the probability for the activation of a witness tree, the possibility of non-distinct balls has to be accounted for.
-In other words, the edge and leaf events of the witness tree are not independent anymore.
+In other words, the edge and leaf events of the witness tree are not independent any more.
 Will start with a witness tree referencing non-distinct balls and prune it such that dependencies between edge and leaf events are removed.
@@
 Let $\gamma \geq 2$ be a suitable constant.
-The root node of a full witness tree has $\gamma$ children, denoted by $v_{c_i}$ with $1\leq i \leq \gamma$.
+The root node of a full witness tree (FWT) has $\gamma$ children, denoted by $v_{c_i}$ with $1\leq i \leq \gamma$.
 The nodes $v_{c_i}$ have exactly one child each and the balls assigned to them are pairwise distinct.
 The root thus has $\gamma$ grandchildren, each of which is the root of a witness tree of order $L$.
-The time constraints and activation of a full witness tree is defined analogously to the previous variants.
+The time constraints and activation of a FWT is defined analogously to the previous variants.
 
@@
 \label{sec:analysis:constructionFullWT}
-Assume a bin $x$ contains at least $L+3+\gamma$ balls at time $t$. We construct a witness tree as follows.
+Assume a bin $x$ contains at least $L+3+\gamma$ balls at time $t$. We construct an FWT as follows.
 \begin{compactitem}
@@
 When we reach a node $v$ that represents a ball $b$ already encountered before, let $e = (v,u)$ be the edge connecting $v$ and its parent node $u$.
-We call $e$ a cutoff edge and use it as a witness that one of the $d$ locations of the ball $b_u$ is the same as one of the locations of some other ball $b_{u'}$, represented by a node $u'$, which was already traversed.
+We call $e$ a cut-off edge and use it as a witness that one of the $d$ locations of the ball $b_u$ is the same as one of the locations of some other ball $b_{u'}$, represented by a node $u'$, which was already traversed.
 Furthermore, we remove the whole tree rooted at node $v$ from our witness tree.
@@
 \begin{lemma}\label{lemma:fwt:activation}
-For some suitable $\gamma \geq 2$ and large enough $L$, the activation of a prunded witness tree occurs with probability at most $n^{-\gamma +1 +o(1)}$.
+For some suitable $\gamma \geq 2$ and large enough $L$, the activation of a pruned witness tree occurs with probability at most $n^{-\gamma +1 +o(1)}$.
 \end{lemma}
@@
 Let this symmetric or asymmetric witness tree be denoted by $T$.
-Then, the probability for the activation of the full witness tree is upper-bounded by the probability for the activation of $T$, implying a probability for activation of this witness tree of $p = n^{- \alpha}$, as shown in Sections~\ref{sec:analysis:probabilitySymWT} and~\ref{sec:analysis:probabilityAsymWT}, respectively.
+Then, the probability for the activation of the FWT is upper-bounded by the probability for the activation of $T$, implying a probability for activation of this witness tree of $p = n^{- \alpha}$, as shown in Sections~\ref{sec:analysis:probabilitySymWT} and~\ref{sec:analysis:probabilityAsymWT}, respectively.
 
-Thus, it remains to find an upper-bound for the activation of a full witness tree with $\gamma$ cut-off edges.
-Let $K$ be the number of references to balls by the full witness tree\footnote{Since no ball was assigned to the root node, the full witness tree consists of $K+1$ nodes} and assume that $K  \leq 2 \gamma\left(\alpha+1\right) \log_2 n$, for some arbitrary $\alpha > 0$.
-Then, there are at most $w_p=K^\gamma$ possibilities to transform a full witness tree with $K+1$ nodes to a pruned witness tree with $\gamma$ cut-off edges.
+Thus, it remains to find an upper-bound for the activation of an FWT with $\gamma$ cut-off edges.
+Let $K$ be the number of references to balls by the FWT\footnote{Since no ball was assigned to the root node, the FWT consists of $K+1$ nodes} and assume that $K  \leq 2 \gamma\left(\alpha+1\right) \log_2 n$, for some arbitrary $\alpha > 0$.
+Then, there are at most $w_p=K^\gamma$ possibilities to transform an FWT with $K+1$ nodes to a pruned witness tree with $\gamma$ cut-off edges.
 Fix the shape of the tree to be one of those.
@@
 \begin{align*}
-p_e = \left(\frac{d}{n}\right)^{k}
+p_e = \left(\frac{d}{n}\right)^{k}\quad ,
 \end{align*}
@@
 \begin{align*}
-p_l \leq 3^{-d l}
+p_l \leq 3^{-d l}\quad .
 \end{align*}
@@
 \begin{align*}
-p_c &= \left(k \frac{d}{n}\right)^\gamma \\
-    & \leq \left(K \frac{d}{n}\right)^\gamma
+p_c &= \left(k \frac{d}{n}\right)^\gamma 
+    \leq \left(K \frac{d}{n}\right)^\gamma \quad .
 \end{align*}
@@
 \begin{align*}
-p_{pwt} &= w_p w  p_e  p_l  p_c \\
-		&= K\gamma n^{k+1} \left(\frac{d}{n}\right)^k3^{-d l}  \left(K \frac{d}{n}\right)^\gamma\\
-		&= n  d^k 3^{-d l}  \left(K^2 \frac{d}{n} \right)^\gamma 
+p_{pwt} &= w_p w  p_e  p_l  p_c 
+		= K\gamma n^{k+1} \left(\frac{d}{n}\right)^k3^{-d l}  \left(K \frac{d}{n}\right)^\gamma
+		= n  d^k 3^{-d l}  \left(K^2 \frac{d}{n} \right)^\gamma \quad .
 \end{align*}
@@
 \begin{align*}
-p_{pwt} &\leq n \left(\frac{\left(2\gamma  \left(\alpha +1 \right)  \log_2 n\right)^2 d}{n}\right)^\gamma \\
-        &= n^{-\gamma +1}  n ^{\log_n\left( \left(2\gamma \left(\alpha+1\right) \log_2n \right)^{2\gamma}  d^\gamma \right)}\\
-        &= n^{-\gamma+1+o\left(1\right)}
+p_{pwt} &\leq n \left(\frac{\left(2\gamma  \left(\alpha +1 \right)  \log_2 n\right)^2 d}{n}\right)^\gamma 
+        = n^{-\gamma +1}  n ^{\log_n\left( \left(2\gamma \left(\alpha+1\right) \log_2n \right)^{2\gamma}  d^\gamma \right)} \\
+        &= n^{-\gamma+1+o\left(1\right)} \quad .
 \end{align*}
@@
 \begin{align*}
-\delta =  \Theta\left(\sqrt{\frac{m  \log n}{n}}\right).
+\delta =  \Theta\left(\sqrt{\frac{m  \log n}{n}}\right) \quad ,
 \end{align*}
@@
 \begin{align*}
-\delta = \frac{\ln\ln n}{\ln d} + O(1)
+\delta = \frac{\ln\ln n}{\ln d} + O(1) \quad ,
 \end{align*}
@@
 \begin{align*}
-\delta = \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
+\delta = \frac{\ln \ln n}{d \ln \Phi_d} + O(1) \quad ,
 \end{align*}
@@
 \begin{align*}
-F_d(k) = \sum_{i=1}^{d}F_d(k-i)
+F_d(k) = \sum_{i=1}^{d}F_d(k-i) \quad .
 \end{align*}
@@
 \begin{align*}
-\Phi_d = \lim_{k \rightarrow \infty} \sqrt[k]{F_d(k)} \neqn{goldenFib}
+\Phi_d = \lim_{k \rightarrow \infty} \sqrt[k]{F_d(k)} \neqn{goldenFib} \quad .
 \end{align*}

### 2014-03-17 article/article.tex
rework second proof

@@
 L &\geq  \frac{\ln\log_2 n}{\ln d} + \frac{\ln(1+\alpha)}{\ln d} 
-   = \frac{\ln \log_2 n}{\ln d}  \left(1 + \frac{\ln\left(1+\alpha\right)}{\ln \log_2 n}\right) \\
-  &=  \frac{\ln \ln n}{\ln d}  \left( 1+ o\left( 1\right)\right) \neqn{swt:boundL} \quad ,
+   = \frac{\ln \log_2 n}{\ln d} +O(1) 
+  \geq \frac{\ln \ln n}{\ln d} +O(1) \neqn{swt:boundL} \quad ,
 \end{align*}
@@
 \label{sec:analysis:alg}
-Similarly to Section~\ref{sec:analysis:greedy}, where we used symmetric witness trees to upper-bound the probability that a bin contains too many balls when using the Greedy scheme, we will use asymmetric witness trees (AWT) to get an analogous upper bound for the Always-go-left scheme.
+Similarly to Section~\ref{sec:analysis:greedy}, where we used symmetric witness trees to upper-bound the probability that a bin contains too many balls when using Greedy as allocation algorithm, we will use asymmetric witness trees (AWT) to obtain an analogous upper bound for the Always-go-left scheme.
 Again, we will use Assumption~\ref{assumption:independence}, i.e., we assume that all the events represented by the edges and nodes of a witness tree are stochastically independent.
@@
 \label{sec:analysis:definitionAsymWT}
-An AWT is defined in the same way as an SWT  (Section~\ref{sec:analysis:definitionSymWT}), except for its structure, which is a $d$-ary Fibonacci tree $T_d\left(k \right)$:
+An AWT differes from an SWT (Section~\ref{sec:analysis:definitionSymWT}) only by its structure, which is that of a $d$-ary Fibonacci tree $T_d\left(k \right)$:
 
 \begin{compactitem}
-\item $T_d(1)$ is a single node.
-\item $T_d(2) = T_d(1)$.
+\item $T_d(1)$ is a single node and  $T_d(2) = T_d(1)$.
 \item For $3\leq k \leq d$, $T_d(k)$ is a tree with $k-1$ children, which are the roots of the trees $T_d(k-1),\ldots,T_d(1)$.
@@
 \end{compactitem}
-A $d$-ary AWT of order $L$ has the structure of the Fibonacci tree $T_d(d  L+1)$ and thus has $F_d(d L + 1) \geq \Phi_d^{d L-1}$ leaves (Eq.~\ref{eqn:goldenFib}).
-
+A $d$-ary AWT of order $L$ has the structure of a Fibonacci tree $T_d(d  L+1)$ and thus has $l=F_d(d L + 1) \geq \Phi_d^{d L-1}$ leaves (Eq.~\ref{eqn:goldenFib})\todo{Can we just rewrite Eq.~\ref{eqn:goldenFib} like this?
+what about the limes?}.
+Furthermore, for each node we define a label $(h, i)$, where $h, i \in \mathbb{N}$, $0 \leq h \leq L$ and  $1\leq i \leq d$:
+\begin{compactitem}
+\item The root nodes of the Fibonacci trees $T_d(1)$ and $T_d(2)$ are assigned the labels $(0,1)$ and $(0,2)$, respectively. Thus, nodes labelled $(0,1)$ or $(0,2)$ do not have any children.
+\item A node with label $(0, i)$, where $i>2$, has $i-1$ children with the labels $(0,i-1),\dots,(0,1)$
+\item A node with label $(h,i)$, where $h>0$, has $d$ children labelled $(h,i-1),\dots,(h,1)$,$(h-1,d),\dots,(h-1, i)$.
+\item The root of a Fibonacci tree $T_d(d h+i)$ is labelled $(h, i)$. 
+\end{compactitem}
+Thus, the root node of an AWT of order $L$ is assigned the label $(L, 1)$. Note that the labels are unique for the set of children of any node $v$, but not necessarily for all nodes of the tree. 
 \begin{lemma}
-When using the Always-go-left algorithm to put balls into bins, a bin containing more than $L+3$ balls implies the activation of an AWT of order $L$.
+When using the Always-go-left algorithm to put balls into bins, a bin containing more than $L+3$ balls activates an AWT of order $L$.
 \end{lemma}
 \begin{proof}
-For each node we define a label $(h, i)$, where $h, i \in \mathbb{N}$, $0 \leq h \leq L$ and  $1\leq i \leq d$. 
+The following invariant will be maintained while assigning balls to the nodes of the AWT:
+\begin{invariant}
+A ball $b_v$ referenced by a node $v$ with label $(h,i)$ was placed in a bin $x$ belonging to group $i$ that, at the time of insertion of the ball $b_v$, contained at least $h+3$ other ball.
+The ball represented by the root node may be put in a bin belonging to any group.
+\end{invariant}
+\noindent
+Let $x$ be the bin that contains $L+4$ balls and assign the topmost ball $b_t$ in $x$ to the root node of the AWT.
+Thus, at the time of insertion of $b_t$, each of the $d$ locations of this ball points to a bin containing at least $L+3$ balls.
+These are the $d$ bins that were sampled by the Always-go-left scheme, i.e., bin $i$ belongs to the $i$'th group of bins, where each group consists of $n/d$ bins.
+The child node of the root with label $(L-1, i)$ is assigned the topmost ball of the $i$'th location.
+The assignment of balls to nodes then proceeds recursively: Given a node $v$ labelled $(h, i)$ that references a ball $b_v$ (which was put into some bin of the $i$-th group that, at the time of insertion of ball $b_v$, contained at least $h+3$ other balls), we can conclude the following:
 \begin{compactitem}
-\item A node that is the root of a Fibonacci tree $T_d(1)$ is assigned the label $(0,1)$.
-\item Analogously, a node that is the root of a Fibonacci tree $T_d(2)$ is assigned the label $(0,2)$. Thus, nodes labelled $(0,1)$ or $(0,2)$ do not have any children.
-\item A node with label $(0, i)$, where $i>2$, has $i-1$ children with labels $(0,i-1),\dots,(0,1)$
-\item A node with label $(h,i)$ , where $h>0$, has $d$ children with the labels $(h,i-1),\dots,(h,1)$,$(h-1,d),\dots,(h-1, i)$.
-\item The root of a Fibonacci tree $T_d(d h+i)$ has the label $(h, i)$. 
-\end{compactitem}
-Thus, the root node of the witness tree will be labelled $(L, 1)$.
-Note that the labels are unique for the set of children of any node $v$, but not necessarily for all nodes of the tree.
-The labels are useful to state the following invariant, which will be maintained while assigning balls to the nodes of the witness tree.
-
-\begin{invariant}
-Ball $b_v$ referenced by a node $v$ with label $(h,i)$ was placed in a bin belonging to group $i$. The bin contained at least $h+3$ other balls at the time of insertion of the ball $b_v$.
-\end{invariant}
-Let $x$ be the bin that contains $L+4$ balls.
-The root node gets assigned the topmost ball $b_t$ in bin $x$, thus each of the $d$ locations of this ball points to a bin with at least $L+3$ balls in it, at the time of insertion of $b_t$.
-These are the $d$ bins that were sampled by the Always-go-left scheme and thus bin $i$ belongs to the $i$'th group of bins, where each group consists of $n/d$ bins.
-The child node of the root with label $(L-1, i)$ is assigned the topmost ball of the bin the $i$' location points to.
-The assignment of balls to nodes than proceeds recursively.
-Given a node $v$ with label $(h, i)$ and referencing ball $b_v$ that was put into some bin belonging to group $i$ and containing at least $l+3$ other balls, we can conclude the following:
-\begin{compactitem}
-\item For $1\leq j < i$, the $j$'th location of the ball $b_v$ points to a bin containing at least $l+4$ balls at the insertion time of the ball $b_v$.
+\item For $1\leq j < i$, the $j$'th location of the ball $b_v$ contains at least $h+4$ balls, at the insertion time of the ball $b_v$.
 If the $j$-th location contained less balls than the $i$-th location, the Always-go-left scheme would have put the ball $b_v$ into the bin pointed to by location $j$.
-The topmost ball in location $j$ is assigned to the  child $u$ of $v$ with label $(l,j)$.
-\item For $i < j \leq d$, the $j$-th location contains at least $l+3$ balls.
+The topmost ball in location $j$ is assigned to the  child $u$ of $v$ with label $(h,j)$.
+\item For $i < j \leq d$, the $j$-th location contains at least $h+3$ balls.
 If location $j$ contained less, the Always-go-left scheme would have put the ball $b_v$ into the bin location $j$ points to.
-The topmost ball in the bin that location $j$ points to is assigned to the child node $u$ of $v$ labelled with $(l-1, j)$.
-\todo{the paper claims that the locations $j$ point to bins with $l+2$ balls.
-This can not be true, since the topmost, i.e., the $l+2$ ball - call it $b_t$ - gets assigned to a node with label $(l'=l-1, j)$.
-Since there are $l+1=l'+2$ balls below it, at the time of the placement of ball $b_t$, the bin $j$ it gets put into has only $l'+2$ balls in it, violating the invariant.
+The topmost ball in the bin that location $j$ points to is assigned to the child node $u$ of $v$ labelled with $(h-1, j)$.
+\todo{the paper claims that the locations $j$ point to bins with $h+2$ balls.
+This can not be true, since the topmost, i.e., the $h+2$-th ball - call it $b_t$ - gets assigned to a node with label $(h'=h-1, j)$.
+Since there are $h+1=h'+2$ balls below it, at the time of the placement of ball $b_t$, the bin $j$ it gets put into has only $h'+2$ balls in it, violating the invariant.
 Same with the case $j<i$.}
@@
 \begin{lemma}\label{lemma:awt:activation}
-For some arbitrary $\alpha$ and large enough $L$, the activation of an AWT of order $L$ occurs with probability at most $n^{-\alpha}$.
+For any $\alpha>0$ and large enough $L$, the activation of an AWT of order $L$ occurs with probability at most $n^{-\alpha}$.
 \end{lemma}
@@
 \begin{proof}
-Analogously to the SWT, the possible number of AWTs is upper-bounded by $w\leq	n^k$, and the probability that all edge events occur is\footnote{As in Section~\ref{sec:analysis:greedy}, $k$ denotes the number of nodes of the witness tree and $l$ the number of leaves.}
+Analogously to the SWT, the possible number of AWTs is upper-bounded as $w\leq	n^k$, and the probability that all edge events occur is
 \begin{align*}
-p_\mathrm{e} = \left(\frac{d}{n}\right)^{k-1}\quad .
+p_\mathrm{e} = \left(\frac{d}{n}\right)^{k-1}\quad ,
 \end{align*}
+where $k$ is the number of nodes in a Fibonacci tree.
 The probability that all leaf events occur is the same as for the SWT, but its derivation differs.
-Again, at any time at most $n/3$ bins contain 3 or more balls, but we do not know how those bins are distributed among the groups.
+Again, at any time at most $n/3$ bins contain 3 or more balls, but we do not know how those bins are distributed among the $d$ groups.
 Let $b_i$ be the number of bins in group $i$ that contain 3 or more balls.
@@
 \begin{align*}
-\frac{n}{d}\sum_{i=1}^d \beta_i \leq \frac{n}{3}\quad ,
-\end{align*}
-and thus 
-\begin{align*}
+\frac{n}{d}\sum_{i=1}^d \beta_i \leq \frac{n}{3} \Rightarrow
 \sum_{i=1}^d \beta_i \leq \frac{d}{3}\quad .
@@
 \begin{align*}
-P_\mathrm{a} &\leq n  2^{-l}\quad .
+P_\mathrm{a} \leq wp_\mathrm{e}p_\mathrm{l} \leq n  2^{-l}\quad .
 \end{align*}
@@
         \leq n  2^{-\Phi_d^{d \left({\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}+\frac{1}{d}\right) -1}} \\
-        &= n  2^{-\Phi_d^{{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}}} 
+        &= n  2^{-\Phi_d^{{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{\ln \Phi_d}}}} 
         = n  2^{-\Phi_d^{\log_{\Phi_d}\log_2 n + \log_{\Phi_d}\left(1+\alpha\right)}} 
@@
 \end{align*}
-Thus, for any $\alpha>0$ and large enough $L$ as specified in Eq.~\ref{eqn:awt:chooseL}, the probability that an AWT of order $L$ is activated is bounded from above as $P_\mathrm{a}\leq n^{-\alpha}$.
+Thus, for any $\alpha>0$ and large enough $L$ as specified in Eq.~\ref{eqn:awt:chooseL}, the probability that an AWT of order $L$ is activated is bounded from above by $P_\mathrm{a}\leq n^{-\alpha}$.
 \end{proof}

### 2014-03-17 article/article.tex
improve appendix

@@
 for the case $m=n$, with high probability\footnote{We say that an event $\mathcal E_n$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E_n \right] \geq 1 - n^{-\alpha}$ for some arbitrary constant $\alpha > 0$.}~\cite{RS98}.
-This implies (see Theorem \ref{theorem:asymptotic} in Appendix \ref{sec:asymptotic})
+This implies
 \begin{align*}
@@
 \label{sec:conclusion}
-BiB games have many applications in Computer Science and have been studied extensively.
+BiB games have many applications in computer science and have been studied extensively.
 The single-choice BiB strategy chooses a bin uniformly and independently at random for each ball.
@@
 w.h.p.
-In this paper we have presented two different multiple-choice BiB placement algorithms: The Greedy-scheme due to Azar et al.\cite{ABKU99}, which achieves a gap of 
+In this paper we have presented two different multiple-choice BiB placement algorithms: The Greedy scheme due to Azar et al.\cite{ABKU99}, which achieves a gap of 
 \begin{align*}
@@
 \end{align*}
-w.h.p., and the Always-go-left scheme due to V\"ocking~\cite{VOC03}, which achieves a slightly better gap of
+w.h.p., and the Always-go-left scheme due to V\"ocking~\cite{VOC03}, which achieves a gap of
 \begin{align*}
@@
 \label{sec:fibonacci}
-We define $d$-ary Fibonacci numbers as follows. For $k \leq0$, $F_d(k) = 0$, $F_d(1) = 1$ and for $k \geq 1$,
+We define $d$-ary Fibonacci numbers as follows:
 
 \begin{align*}
-F_d(k) = \sum_{i=1}^{d}F_d(k-i) \quad .
+F_d(k) := \begin{cases}
+                0               & k \leq 0\\
+                1               & k = 1\\
+                \sum_{i=1}^{d}F_d(k-i) & k > 1
+            \end{cases}
+\quad .
 \end{align*}
-Furthermore, define 
+Furthermore, we define $\Phi_d$ as
 \begin{align*}
-\Phi_d = \lim_{k \rightarrow \infty} \sqrt[k]{F_d(k)} \neqn{goldenFib} \quad .
+\Phi_d := \lim_{k \rightarrow \infty} \sqrt[k]{F_d(k)} \neqn{goldenFib} \quad ,
 \end{align*}
-Notice that $F_2$ corresponds to the usual Fibonacci numbers, while $\Phi_2 $ corresponds to the golden ratio~\cite{Knuth73}, which is generalized to $\Phi_d$, for which 
+which has the following properties:
 \begin{align*}
-i, j \in \mathbb{N}, i < j \Rightarrow \Phi_i < \Phi_j
+\forall i\; \forall j \in \mathbb{N}: i < j \quad \Rightarrow \quad \Phi_i < \Phi_j \quad , \quad \lim_{d\rightarrow \infty} \Phi_d = 2 \quad .
 \end{align*}
-and 
-\begin{align*}
-\lim_{d\rightarrow \infty} \Phi_d = 2
-\end{align*}
-hold.
+Notice that $F_2$ corresponds to the usual Fibonacci numbers, while $\Phi_d$ is a generalization of the golden ratio~\cite{Knuth73} and $\Phi_2 $ corresponds to the golden ratio.
 
@@
 \label{sec:asymptotic}
-We use the following definitions for the asymptotic notations~\cite{CLRS01}, where $f(\cdot)$ and $g(\cdot)$ are positive functions with one argument $n\in \mathbb{N}$ and $n_0 \in \mathbb{N}$, $c \in \mathbb{R}_{\geq 0}$ are suitable constants.
+We use the following definitions for the asymptotic notations~\cite{CLRS01}, where $f$ and $g$ are non-negative functions with one argument $n\in \mathbb{N}$ and $n_0 \in \mathbb{N}$, $c \in \mathbb{R}_{\geq 0}$ are suitable constants.
 
 \begin{align*}
-f(n) \in O(g(n)) &\Leftrightarrow \exists c, n_0 \;\forall n> n_0: f(n)\leq c  g(n) \\
-f(n) \in \Omega(g(n)) &\Leftrightarrow \exists c, n_0\;\forall n> n_0: cg(n) \leq f(n) \\
-f(n) \in \Theta(g(n)) &\Leftrightarrow f(n) \in \Omega(g(n)) \land f(n)\in O(g(n)) \\
-f(n) \in o(g(n)) &\Leftrightarrow \forall c \; \exists n_0\;\forall n> n_0: f(n) < cg(n)
+f \in O(g) &\Leftrightarrow \exists c\; \exists n_0 \;\forall n> n_0: f(n)\leq c g(n) \\
+f \in \Omega(g) &\Leftrightarrow \exists c \; \exists n_0\;\forall n> n_0: cg(n) \leq f(n) \\
+f \in \Theta(g) &\Leftrightarrow f \in \Omega(g) \land f\in O(g) \\
+f \in o(g) &\Leftrightarrow \forall c \; \exists n_0\;\forall n> n_0: f(n) < cg(n) \\
+f \in \omega(g) &\Leftrightarrow \forall c \; \exists n_0\;\forall n> n_0: cg(n) < f(n) \\
 \end{align*}
-
-The following theorem is useful for understanding the asymptotic bounds presented throughout the paper.
-\begin{theorem}\label{theorem:asymptotic}
-\begin{align*}
-f(n) = g(n)  (1+o(1)) \Rightarrow f(n) \in \Theta(g(n)) 
-\end{align*}
-\end{theorem}
-\begin{proof}
-\begin{align*}
-f(n) &= g(n)+o(1) g(n) \\
-&\Rightarrow f(n) \in \Omega(g(n)) \land f(n) \in O(g(n)) \\
-&\Rightarrow f(n) \in \Theta(g(n))
-\end{align*}
-However, the reverse
-\begin{align*}
-f(n) \in \Theta(g(n)) \Rightarrow f(n) = g(n)  (1+o(1)) 
-\end{align*}
-is not true. As a counter example, consider $f(n) = n$ and $g(n) = 2n$.
-\end{proof}
 
@@
 Berenbrink at.
-all~\cite{BCSV06} were the first to proof that the same bound indeed holds for the heavily loaded case.
+all~\cite{BCSV06} were the first to prove that the same bound indeed holds for the heavily loaded case.
 Their proof is fairly complex and a simpler analysis may be found in~\cite{TW13}, which also provides a tight lower bound for the heavily loaded case.
@@
 in the context of simulating PRAMs on distributed memory machines\cite{KLM92}.
-
-As discussed in Section~\ref{sec:AlwaysGoLeft}, V\"ocking~\cite{VOC03} was able to slightly improve the bound as well as to show that the Always-go-left strategy that he introduces is optimal up to additive constants.
+As discussed in Section~\ref{sec:AlwaysGoLeft}, V\"ocking~\cite{VOC03} was able to improve the bound as well as to show that the Always-go-left strategy that he introduces is optimal up to additive constants.
 

### 2014-03-18 article/article.tex
use inline enumeration for 2 short requirements

@@
 
-The following constraints have to be fulfilled by a witness tree. 
-\begin{enumerate}
-\item The ball $b_r$ that is represented by the root node $r$ has to exist at time step $t$. 
-\item Each ball $b_v$ represented by a node $v \neq r$ with parent node $u$ has to exist when the ball $b_u$ is inserted.
-\end{enumerate}
+The following constraints have to be fulfilled by a witness tree:
+\begin{inparaenum}[(i)]
+\item The ball $b_r$ that is represented by the root node $r$ has to exist at time step $t$ and 
+\item each ball $b_v$ represented by a node $v \neq r$ with parent node $u$ has to exist when the ball $b_u$ is inserted.
+\end{inparaenum}
 We say that an edge or leaf event is activated if the associated event occurs. A witness tree is activated when all its edge and leaf events are activated.

### 2014-03-20 article/article.tex
use max load instead of gap

@@
 \end{align*}
-Note that, since the average load is 1, $\delta \in \Theta\left( L_{max} \right)$ in this case. For the heavily loaded case~\cite{RS98}, the maximum gap is, w.h.p.,
+Note that, since the average load is 1, $\delta \in \Theta\left( L_{max} \right)$ in this case. For the heavily loaded case~\cite{RS98}, the maximum load is, w.h.p.,
 \begin{align*}
-\delta = \Theta\left(\sqrt{\frac{m  \ln n}{n}}\right)\quad .
+\load = \frac{m}{n} + \Theta\left(\sqrt{\frac{m  \ln n}{n}}\right)\quad .
 \end{align*}
@@
 For tie-breaking, we assume a fixed ordering of bins $i$ for $1 \leq i \leq n$ and put the ball into the bin with smallest index $i$ among those with the same minimal load.
-Somewhat surprisingly, V\"ocking~\cite{VOC03} showed that such a strategy employing asymmetric tie-breaking in combination with non-uniform bin selection reduces the bound for the gap to\
+Somewhat surprisingly, V\"ocking~\cite{VOC03} showed that such a strategy employing asymmetric tie-breaking in combination with non-uniform bin selection reduces the bound for the maximum load to
 \begin{align*}
-\delta = \frac{\ln\ln n}{d  \ln \Phi_d} + O(1),
+\load = \frac{m}{n} + \frac{\ln\ln n}{d  \ln \Phi_d} + O(1),
 \end{align*}

### 2014-03-20 article/article.tex
small fixes

@@
 \end{align*}
-for the case $m=n$, with high probability\footnote{We say that an event $\mathcal E_n$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E_n \right] \geq 1 - n^{-\alpha}$ for some arbitrary constant $\alpha > 0$.}~\cite{RS98}.
+for the case $m=n$, with high probability\footnote{We say that an event $\mathcal E_n$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E_n \right] \geq 1 - n^{-\alpha}$ for some $\alpha > 0$.}~\cite{RS98}.
 This implies
@@
 \end{align*}
-For any constant $\alpha > 0 $, let
+For any $\alpha > 0 $, let
 \begin{align*}

### 2014-03-20 presentation/presentation.tex
use getargsC from readarray (faster tehn getargs from stringstrings)
basic structure of the presentation
remove subtitle
some more usefull packages!
first slides

@@
 \usecolortheme{dolphin}
+\usecolortheme{orchid}
 
-\usepackage{stringstrings}
+\usepackage{fixltx2e} %some latex fixes
+\usepackage{microtype} %Subliminal refinements towards typographical perfection
+\setlength{\emergencystretch}{2em}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage{mathrsfs} %Support use of the Raph Smith’s Formal Script font in mathematics
+\usepackage[sc]{mathpazo} %mathematical fonts
+\usepackage{amsfonts}
+\usepackage{color}
+\newcommand\todo[1]{\textcolor{red}{(TODO: #1)}}
+\newcommand\load{L_{\mathrm{max}}}
+\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
+% within an align* environment, number and label the equation. #1 is the label text to be appended to eqn:
+\newcommand\neqn[1]{\numberthis\label{eqn:#1}}
+
+%TODO: why are the following defined when using beamer but not for an article?
+%\newtheorem{lemma}{Lemma}
+%\newtheorem{corollary}{Corollary}
+%\newtheorem{theorem}{Theorem}
+\newtheorem{claim}{Claim}
+\newtheorem{assumption}{Assumption}
+\newtheorem{invariant}{Invariant}
 
 %graphs, pictures
+\usepackage{readarray}
 \usepackage{tikz}
@@
 \title{The Power of Two Random Choices}
-\subtitle{ A follow up seminar to the lecture "Parallel Algorithms"}
+%\subtitle{ A follow up seminar to the "Parallel Algorithms" lecture}
 \author{Martin Kalany}
@@
 
+% Create Color definition From Template: 
+% #1 template name, 
+% #2 foreground color name
+% #3 background color name
+\newcommand{\ccft}[3]{
+\usebeamercolor{#1}
+\definecolor{#2}{named}{fg}
+\definecolor{#3}{named}{bg}
+}
+%get some colors from the scheme
+\ccft{block body}{myblock body fg}{myblock body bg}
+\ccft{block title}{myblock title fg}{myblock title bg}
+
 %---------------------------------------------------------------------
@@
 %scale factor
-\newcommand\scalefac{0.5}
+\newcommand\scalefac{0.55}
 %diameter of balls
@@
 %height of a bin
-\newcommand\binheight{6*\balldiameter}
+\newcommand\binheight{5*\balldiameter}
 %width of a bin
@@
 %color of balls
-
 \tikzstyle{ballstyle} = [ball color=black!30!red]
@@
 	minimum width=(\binwidth+2*\padding)*\scalefac, 
-	minimum height=(\binheight+5*\padding)*\scalefac,
+	minimum height=(\binheight+4*\padding)*\scalefac,
 	append after command={
     		\pgfextra
-        		\fill[fill=blue!40] 
+        		\fill[fill=myblock body bg] 
         			(\tikzlastnode.north east) [rounded corners] |-
@@
 \newcommand\balls[1]{%
-	\getargs{#1}%
+	\getargsC{#1}%
   	\setcounter{index}{0}%
@@
 %---------------------------------------------------------------------
+
+% Structure:
+% 30 min
+% 5 min easy stuff (so that everyone understands)
+% 5 min more advanced stuff 
+% 15 min difficult stuff (the proof, either of the symmetric or asymmetric allocation scheme using the infamous assumption 1
+% 5 min further results and wrap up
+
 \begin{document}
@@
 
+\section{Balls into bins}
 \begin{frame}
+    \frametitle{Table of Contents}
+    \tableofcontents
+\end{frame}
+
+\begin{frame}
+\frametitle{The Problem}
+\begin{block}{Balls-into-bins games}
+Suppose we have $n$ initially empty bins and need to distribute $m \geq n$ balls among them, as evenly as possible.
+We are interested in the maximum load $\load$, i.e, how many balls are in the fullest bin?
+\end{block}
+\pause
+Applications:
+\begin{itemize}
+	\item Hashing
+	\item Online load balancing
+	\item Emulating PRAMs on DMMs
+	\item Low congestion circuit routing
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Single choice}
+\framesubtitle{A basic allocation scheme}
+\begin{block}{Single choice strategy}
+For each ball, choose one of the $n$ bins uniformly and independently at random.
+\end{block}
+\bigskip
+\begin{center}
 \begin{tikzpicture}[scale=\scalefac]
+	\bab{2 1 5 2 3 0}
+\end{tikzpicture}
+\end{center}
+\end{frame}
 
+\begin{frame}
+\frametitle{Randomized allocation process}
+
+Since we deal with a randomized process, we cannot give upper bounds for the maximum load $\load$ with absolute certainty.
+
+\bigskip
+
+\begin{block}{Definition}
+We say that an event $\mathcal E_n$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E_n \right] \geq 1 - n^{-\alpha}$ for some $\alpha > 0$.
+\end{block}
+\end{frame}
+
+\begin{frame}
+\frametitle{Single choice: Upper bound for the maximum load}
+\begin{claim}
+When using the single choice strategy to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
+\begin{align}
+\load = 
+	\begin{cases}
+    \frac{\ln n}{\ln\ln n} + O(1)              & m = n \\
+    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m > n
+    \end{cases}
+\end{align}
+\end{claim}
+\end{frame}
+
+\begin{comment}
+\begin{frame}
+\frametitle{Applications}
+\begin{itemize}
+	\item Hashing
+	\item Online load balancing
+	\item Emulating PRAMs on DMMs
+	\item Low congestion circuit routing
+\end{itemize}
+\end{frame}
+\end{comment}
+
+%Table of contents at the start of each section
+\AtBeginSection[]
+{
+  \begin{frame}
+    \frametitle{Table of Contents}
+    \tableofcontents[currentsection]
+  \end{frame}
+}
+
+\section{Allocation shemes}
+\begin{frame}
+\frametitle{BiB Model}
+\end{frame}
+
+\begin{frame}
+\frametitle{Multiple choice strategies}
+\end{frame}
+
+\begin{frame}
+\frametitle{Greedy}
+\end{frame}
+
+\begin{frame}
+\frametitle{Always-go-left}
+\end{frame}
+
+\begin{frame}
+\frametitle{Proof of either Greedy or Always-go-left}
+\end{frame}
+
+\begin{comment}
+\section{Test}
+\begin{frame}
+\frametitle{Ignore the following}
+\begin{tikzpicture}[scale=\scalefac]
 	\bab{1 2 3 4 5 6}
-
 \end{tikzpicture}
 \end{frame}
+\end{comment}
 

### 2014-03-20 presentation/presentation.tex
change title
add presentation template

--- 2014-01-15 presentation/presentation.tex
@@
-\documentclass{beamer}
-\let\Tiny=\tiny %to avoid warnings related to font size and beamer 
-\usetheme{Amsterdam}
-\usecolortheme{dolphin}
+\documentclass[serif,professionalfonts]{beamer}
 
-\usepackage{stringstrings}
+%\documentclass[a4paper]{article}
+%\usepackage{beamerarticle}
 
-%graphs, pictures
+%\usepackage{pgfpages}
+%\setbeameroption{show only notes}
+%\setbeameroption{show notes on second screen=top}
+
+\usetheme{Forcellini}
+
+\usepackage{amsmath}
+
+% \vec makes bold symbols
+\def\vec#1{\mathchoice{\mbox{\boldmath$\displaystyle#1$}}
+  {\mbox{\boldmath$\textstyle#1$}}
+  {\mbox{\boldmath$\scriptstyle#1$}}
+  {\mbox{\boldmath$\scriptscriptstyle#1$}}}
+
+% Hermann Zapf Optima clone (wonderful sanserif font) -- URW Classico
+\renewcommand{\sfdefault}{uop}
+\renewcommand{\rmdefault}{ppl}
+\usepackage{eulervm}
+\usepackage[right]{eurosym}
+% \usepackage{paralist}
+\usepackage{algorithm2e}
+
+% professional tables
+\usepackage{booktabs}
+\usepackage{pdfpages}
+\usepackage{multirow}
+\usepackage{array}
+\usepackage{dcolumn}
+\usepackage{subfigure}
+\usepackage{epsdice}
+
+\usepackage{fixltx2e}
+\usepackage{microtype}
+
+\usepackage{dot2texi}
 \usepackage{tikz}
-\usepackage{pgf}
-\usepackage{xxcolor}
-\usetikzlibrary{positioning}
+\usetikzlibrary{decorations.pathmorphing,arrows,shapes}
+
+\setbeamertemplate{navigation symbols}{}
+\usepackage{hyperref}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 
 \usepackage{comment}
+% description with short indentation
+\setbeamersize{description width of=x}
 
-\usepackage[style=numeric-comp,backend=bibtex]{biblatex}
-\addbibresource{../sources}
+% some math operators
+\DeclareMathOperator{\har}{H}
+\DeclareMathOperator{\E}{E}
+\DeclareMathOperator{\lag}{lag}
+\DeclareMathOperator{\var}{Var}
+\DeclareMathOperator{\mf}{\enspace .}
+\DeclareMathOperator{\mc}{\enspace ,}
+\newcommand{\eye}{\mathbb{I}}
+\newcommand{\Z}{\mathbb{Z}}
+\newcommand{\dom}{\{0,1\}^n}
+\newcommand{\reals}{\mathbb{R}}
+\newcommand{\nats}{\mathbb{N}}
+\newcommand{\deq}{\triangleq}
+\newcommand{\pd}{\cdot_p}
+\newcommand{\pp}{+_p}
+\newcommand{\pmn}{-_p}
+\newcommand{\ed}{\epsdice}
+\newcommand{\bled}{\epsdice[black]}
+\newcommand{\bitup}{\vspace*{-4mm}}
+\newcommand{\nospace}{\abovedisplayskip=0pt}
 
-%reduce font size of footnotes
-\let\oldfootnotesize\footnotesize
-\renewcommand*{\footnotesize}{\oldfootnotesize\tiny}
+\setbeamertemplate{caption}
+{
+  \raggedright%
+  \insertcaption\par%
+}
 
-%---------------------------------------------------------------------
-%Front matter
-%---------------------------------------------------------------------
+\setbeamercolor{green}{fg=green}
+
 \title{The Power of Two Random Choices}
-\subtitle{ A follow up seminar to the lecture "Parallel Algorithms"}
-\author{Martin Kalany}
-\institute
+%\subtitle{ A follow up seminar to the "Parallel Algorithms" lecture}
+\author[M. Kalany]{Martin Kalany}
+\institute[TU Wien]
 {
@@
 
-%---------------------------------------------------------------------
-%parameters for balls into bins graphs
-%---------------------------------------------------------------------
-%scale factor
-\newcommand\scalefac{0.5}
-%diameter of balls
-\newcommand\ballsize{5mm}
-%number of bins
-\newcommand\nrbins{6}
-%space between ball and bin
-\newcommand\padding{0.1*\ballsize}
-%height of a bin
-\newcommand\binheight{6*\balldiameter}
-%width of a bin
-\newcommand\binwidth{\balldiameter}
-%gap from bin to bin
-\newcommand\bingap{1.6*\balldiameter}
-%color of balls
-
-\tikzstyle{ballstyle} = [ball color=black!30!red]
-
-%---------------------------------------------------------------------
-%tikz styles and commands
-%---------------------------------------------------------------------
-\newcommand\balldiameter{2*\ballsize}
-
-%bin style
-\tikzstyle{topflat} = [
-	minimum width=(\binwidth+2*\padding)*\scalefac, 
-	minimum height=(\binheight+5*\padding)*\scalefac,
-	append after command={
-    		\pgfextra
-        		\fill[fill=blue!40] 
-        			(\tikzlastnode.north east) [rounded corners] |-
-        			(\tikzlastnode.south)  -| 
-        			(\tikzlastnode.west) [sharp corners] |- 
-        			(\tikzlastnode.north) -- cycle;
-	         \draw[rounded corners] 
-	         	(\tikzlastnode.north east) |- 
-	         	(\tikzlastnode.south) -| 
-	         	(\tikzlastnode.north west);
-    		\endpgfextra
-    }
-]
-
-%draw i'th bin
-\newcommand\bin[1]{
-	\path node[topflat, xshift=#1*\bingap*\scalefac, above, yshift=-\padding*\scalefac]  {};
+\AtBeginSubsubsection[]
+{
+  \begin{frame}<beamer>
+    \frametitle{Outline}
+    \tableofcontents[subsubsectionstyle=show/shaded,currentsection,currentsubsection]
+  \end{frame}
 }
 
-%draw all bins
-\newcommand\bins{
-	\foreach \ibin in {1,...,\nrbins}
-		\bin{\ibin};
+\AtBeginSubsection[]
+{
+  \begin{frame}<beamer>
+    \frametitle{Outline}
+    \tableofcontents[subsubsectionstyle=show/shaded,currentsection,currentsubsection]
+  \end{frame}
 }
 
-%paint a ball in bin #1, position #2 
-\newcommand\ball[2]{
-	\shade[ballstyle] (#1*\bingap,#2*\balldiameter-\ballsize) circle (\ballsize)
+\AtBeginSection[]
+{
+  \begin{frame}<beamer>
+    \frametitle{Outline}
+    \tableofcontents[subsubsectionstyle=show/shaded,currentsection,currentsubsection]
+  \end{frame}
 }
 
-%put #2 balls into bin 1 <= #1 <= 6
-\newcommand\putinbin[2]{
-	\ifnum #2 > 0
-		\foreach \nrballs in {1,...,#2}
- 			\ball{#1}{\nrballs};
- 	\fi
-}
+\begin{document}
+\begin{frame}
+  \titlepage
+\end{frame}
+\note{}
 
-%put balls into bins. #i: number of balls to put in i'th bin
-\newcounter{index}
-\newcommand\balls[1]{%
-	\getargs{#1}%
-  	\setcounter{index}{0}%
-  	\whiledo{\theindex < \narg}{%
-    		\stepcounter{index}%
-    		\putinbin{\theindex}{\csname arg\romannumeral\theindex\endcsname}%
-  	}%
-}
+\begin{comment}
+\begin{frame}[label=references]
+  \frametitle{References}
+  %\nocite{*}
+  %\setbeamertemplate{bibliography item}[book]
+  \bibliographystyle{acm}
+  \bibliography{distrib}
+\end{frame}
+\note{}
+\end{comment}
 
-%draw balls and bins. #i: number of balls to put in i'th bin
-\newcommand\bab[1]{%
-	\bins
-	\balls{#1}
-}
+\begin{frame}{Introduction}
+\end{frame}
+\note{}
 
-%---------------------------------------------------------------------
-%document
-%---------------------------------------------------------------------
-\begin{document}
-\frame{\titlepage}
+%\againframe{references}
 
-\begin{frame}
-\begin{tikzpicture}[scale=\scalefac]
+\end{document}
 
-	\bab{1 2 3 4 5 6}
+\begin{frame}{Title}
+\end{frame}
+\note{}
 
-\end{tikzpicture}
-\end{frame}
-
-%---------------------------------------------------------------------
-%bibliography
-%---------------------------------------------------------------------
-\begin{frame}[allowframebreaks]
-\frametitle<presentation>{Literature}    
-\printbibliography
-\end{frame} 	 
-\end{document}
+%  LocalWords:

### 2014-03-20 presentation/presentation.tex
fix bib sources
moved presentation to new slides

--- 2014-03-20 presentation/presentation.tex
@@
-\documentclass{beamer}
-\let\Tiny=\tiny %to avoid warnings related to font size and beamer 
-\usetheme{Amsterdam}
-\usecolortheme{dolphin}
-\usecolortheme{orchid}
+\documentclass[serif,professionalfonts]{beamer}
 
-\usepackage{fixltx2e} %some latex fixes
-\usepackage{microtype} %Subliminal refinements towards typographical perfection
-\setlength{\emergencystretch}{2em}
+%\usepackage{pgfpages}
+%\setbeameroption{show only notes}
+%\setbeameroption{show notes on second screen=top}
+
+\usetheme{Forcellini}
+
 \usepackage{amsmath}
@@
 \usepackage{amsthm}
-\usepackage{mathrsfs} %Support use of the Raph Smith’s Formal Script font in mathematics
-\usepackage[sc]{mathpazo} %mathematical fonts
-\usepackage{amsfonts}
-\usepackage{color}
+
+% Hermann Zapf Optima clone (wonderful sanserif font) -- URW Classico
+%\renewcommand{\sfdefault}{uop}
+%\renewcommand{\rmdefault}{ppl}
+%\usepackage{eulervm}
+%\usepackage[right]{eurosym}
+%\usepackage{paralist}
+%\usepackage{algorithm2e}
+
+% professional tables
+%\usepackage{booktabs}
+%\usepackage{pdfpages}
+%\usepackage{multirow}
+%\usepackage{array}
+%\usepackage{dcolumn}
+%\usepackage{subfigure}
+%\usepackage{epsdice}
+
+\usepackage{fixltx2e}
+\usepackage{microtype}
+
+\usepackage{dot2texi}
+\usepackage{readarray}
+\usepackage{tikz}
+\usepackage{xxcolor}
+\usetikzlibrary{positioning}
+
+\setbeamertemplate{navigation symbols}{}
+\usepackage{hyperref}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+
+\usepackage{comment}
+% description with short indentation
+\setbeamersize{description width of=x}
+
+% some math operators
+%\DeclareMathOperator{\har}{H}
+%\DeclareMathOperator{\E}{E}
+%\DeclareMathOperator{\lag}{lag}
+%\DeclareMathOperator{\var}{Var}
+%\DeclareMathOperator{\mf}{\enspace .}
+%\DeclareMathOperator{\mc}{\enspace ,}
+%\newcommand{\eye}{\mathbb{I}}
+%\newcommand{\Z}{\mathbb{Z}}
+%\newcommand{\dom}{\{0,1\}^n}
+%\newcommand{\reals}{\mathbb{R}}
+%\newcommand{\nats}{\mathbb{N}}
+%\newcommand{\deq}{\triangleq}
+%\newcommand{\pd}{\cdot_p}
+%\newcommand{\pp}{+_p}
+%\newcommand{\pmn}{-_p}
+%\newcommand{\ed}{\epsdice}
+%\newcommand{\bled}{\epsdice[black]}
+%\newcommand{\bitup}{\vspace*{-4mm}}
+%\newcommand{\nospace}{\abovedisplayskip=0pt}
+
+\setbeamertemplate{caption}
+{
+  \raggedright%
+  \insertcaption\par%
+}
+\setbeamercolor{green}{fg=green}
+
+%reduce font size of footnotes
+%\let\oldfootnotesize\footnotesize
+%\renewcommand*{\footnotesize}{\oldfootnotesize\tiny}
+
+%
 \newcommand\todo[1]{\textcolor{red}{(TODO: #1)}}
 \newcommand\load{L_{\mathrm{max}}}
-\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
-% within an align* environment, number and label the equation. #1 is the label text to be appended to eqn:
-\newcommand\neqn[1]{\numberthis\label{eqn:#1}}
-
-%TODO: why are the following defined when using beamer but not for an article?
-%\newtheorem{lemma}{Lemma}
-%\newtheorem{corollary}{Corollary}
-%\newtheorem{theorem}{Theorem}
 \newtheorem{claim}{Claim}
@@
 \newtheorem{invariant}{Invariant}
-
-%graphs, pictures
-\usepackage{readarray}
-\usepackage{tikz}
-\usepackage{pgf}
-\usepackage{xxcolor}
-\usetikzlibrary{positioning}
-
-\usepackage{comment}
-
-\usepackage[style=numeric-comp,backend=bibtex]{biblatex}
-\addbibresource{../sources}
-
-%reduce font size of footnotes
-\let\oldfootnotesize\footnotesize
-\renewcommand*{\footnotesize}{\oldfootnotesize\tiny}
 
@@
 %\subtitle{ A follow up seminar to the "Parallel Algorithms" lecture}
-\author{Martin Kalany}
-\institute
+\author[M. Kalany]{Martin Kalany}
+\institute[TU Wien]
 {
@@
 
-% Create Color definition From Template: 
+\AtBeginSubsubsection[]
+{
+  \begin{frame}<beamer>
+    \frametitle{Outline}
+    \tableofcontents[subsubsectionstyle=show/shaded,currentsection,currentsubsection]
+  \end{frame}
+}
+
+\AtBeginSubsection[]
+{
+  \begin{frame}<beamer>
+    \frametitle{Outline}
+    \tableofcontents[subsubsectionstyle=show/shaded,currentsection,currentsubsection]
+  \end{frame}
+}
+
+\AtBeginSection[]
+{
+  \begin{frame}<beamer>
+    \frametitle{Outline}
+    \tableofcontents[subsubsectionstyle=show/shaded,currentsection,currentsubsection]
+  \end{frame}
+}
+
+%---------------------------------------------------------------------
+% Create Color definition From Template
+%---------------------------------------------------------------------
 % #1 template name, 
@@
 %---------------------------------------------------------------------
-
 % Structure:
@@
 % 5 min further results and wrap up
-
 \begin{document}
-\frame{\titlepage}
+\begin{frame}
+  \titlepage
+\end{frame}
+\note{}
 
@@
 \begin{frame}
-\frametitle{Single choice: Upper bound for the maximum load}
+\frametitle{Single choice}
+\framesubtitle{Upper bound for the maximum load}
 \begin{claim}
@@
 
-%---------------------------------------------------------------------
-%bibliography
-%---------------------------------------------------------------------
-\begin{frame}[allowframebreaks]
-\frametitle<presentation>{Literature}    
-\printbibliography
+\begin{comment}
+\begin{frame}[label=references]
+  \frametitle{References}
+  %\nocite{*}
+  %\setbeamertemplate{bibliography item}[book]
+  \bibliographystyle{acm}
+  \bibliography{../sources}
 \end{frame}
+\note{}
+\end{comment}
+
+\begin{frame}{Introduction}
+\end{frame}
+\note{}
+
+%\againframe{references}
+
 \end{document}
+
+\begin{frame}{Title}
+\end{frame}
+\note{}
+
+%  LocalWords:

### 2014-03-24 presentation/presentation.tex
some minor fixes
more slides

@@
 
+\begin{comment}
 \AtBeginSubsubsection[]
@@
 }
+\end{comment}
 
@@
 %get some colors from the scheme
-\ccft{block body}{myblock body fg}{myblock body bg}
-\ccft{block title}{myblock title fg}{myblock title bg}
-
+\ccft{block body example}{myblock body fg}{myblock body bg} 
 %---------------------------------------------------------------------
@@
 
-\section{Balls into bins}
+\begin{comment}
 \begin{frame}
@@
 \end{frame}
+\end{comment}
+
+\section{Balls into bins}
+\begin{frame}
+\frametitle{The Problem}
+\begin{exampleblock}{Balls-into-bins games}
+Suppose we have $n$ initially empty bins and need to distribute $m \geq n$ balls among them, \alert{as evenly as possible}. 
+\end{exampleblock}
+
+\begin{exampleblock}{Definition}
+We are interested in the \alert{maximum load $\load$}, i.e, how many balls are in the fullest bin?
+\end{exampleblock}
+
+\begin{exampleblock}{Definition}
+Equivalently, we can look at the \alert{additive gap} $\delta$:
+\begin{align*}
+\delta = \load - \frac{m}{n}
+\end{align*}
+\end{exampleblock}
+
+\end{frame}
 
 \begin{frame}
-\frametitle{The Problem}
-\begin{block}{Balls-into-bins games}
-Suppose we have $n$ initially empty bins and need to distribute $m \geq n$ balls among them, as evenly as possible.
-We are interested in the maximum load $\load$, i.e, how many balls are in the fullest bin?
-\end{block}
-\pause
+\frametitle{Applications}
 Applications:
@@
 \framesubtitle{A basic allocation scheme}
-\begin{block}{Single choice strategy}
+\begin{exampleblock}{Single choice strategy}
 For each ball, choose one of the $n$ bins uniformly and independently at random.
-\end{block}
+\end{exampleblock}
 \bigskip
@@
 \begin{frame}
-\frametitle{Randomized allocation process}
+\frametitle{Some definitions}
 
-Since we deal with a randomized process, we cannot give upper bounds for the maximum load $\load$ with absolute certainty.
+Since we deal with a \alert{randomized process}, we cannot give upper bounds for the maximum load $\load$ with absolute certainty.
+
+\begin{exampleblock}{Definition}
+We say that an event $\mathcal E_n$ occurs \alert{with high probability} (w.h.p.) if $\Pr\left[\mathcal E_n \right] \geq 1 - n^{-\alpha}$ for some $\alpha > 0$.
+\end{exampleblock}
 
 \bigskip
+\pause
 
-\begin{block}{Definition}
-We say that an event $\mathcal E_n$ occurs \emph{with high probability} (w.h.p.) if $\Pr\left[\mathcal E_n \right] \geq 1 - n^{-\alpha}$ for some $\alpha > 0$.
-\end{block}
+%\begin{exampleblock}{Definition}
+We distinguish:
+\begin{itemize}
+\item The \alert{lightly loaded} case: $m = n$
+\item The \alert{heavily loaded} case:	 $m \in \omega(n\ln n)$
+\end{itemize}
+%\end{exampleblock}
+
 \end{frame}
@@
 \framesubtitle{Upper bound for the maximum load}
-\begin{claim}
+\begin{theorem}[Raab and Steger, 1998]
 When using the single choice strategy to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
-\begin{align}
+\begin{align*}
 \load = 
@@
     \frac{\ln n}{\ln\ln n} + O(1)              & m = n \\
-    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m > n
+    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
     \end{cases}
-\end{align}
-\end{claim}
+\end{align*}
+\end{theorem}
 \end{frame}
 
+%Table of contents at the start of each section
 \begin{comment}
-\begin{frame}
-\frametitle{Applications}
-\begin{itemize}
-	\item Hashing
-	\item Online load balancing
-	\item Emulating PRAMs on DMMs
-	\item Low congestion circuit routing
-\end{itemize}
-\end{frame}
-\end{comment}
-
-%Table of contents at the start of each section
 \AtBeginSection[]
@@
 }
+\end{comment}
 
-\section{Allocation shemes}
-\begin{frame}
-\frametitle{BiB Model}
-\end{frame}
-
+\section{Allocation schemes}
 \begin{frame}
 \frametitle{Multiple choice strategies}
+
+\begin{exampleblock}{Idea}
+Put a ball into the \alert{least full} of $d\geq2$ randomly chosen bins.
+\end{exampleblock}
+
+\pause
+\bigskip
+We can distinguish placement algorithms by the way the $d$ \alert{bins are sampled}:
+\begin{itemize}
+\item Uniformly and independently
+\item Non-unfiromly and independently
+\item Non-uniformly and dependently
+\end{itemize}
 \end{frame}
@@
 \frametitle{Greedy}
+\begin{exampleblock}{The Greedy algorithm}
+\begin{enumerate}
+\item For each ball, choose $d$ bins independently and uniformly at random. 
+\item Put the ball into the least full bin.
+\item In case a tie occurs, choose one of the $d$ bins arbitrarily.
+\end{enumerate}
+\end{exampleblock}
+\todo{visualize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Greedy}
+\begin{theorem}[Azar, Broder, Karlin and Upfal, 1999]
+When using the \alert{Greedy} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be (w.h.p.~and for all $m$): 
+\begin{align*}
+\load = \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) \quad .
+\end{align*}
+\end{theorem}
+\pause
+\medskip
+Compared to the \alert{single choice} strategy, with 
+\begin{align*}
+\load = 
+	\begin{cases}
+    \frac{\ln n}{\ln\ln n} + O(1)              & m = n \\
+    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
+    \end{cases} \quad ,
+\end{align*}
+\alert{Greedy} with $d=2$:
+\begin{itemize}
+\item achieves an \alert{exponential decrease} of the additive gap $\delta$ and
+\item $\delta$ does \alert{not depend on} the number of balls \alert{$m$} anymore.
+\end{itemize}
 \end{frame}
@@
 \frametitle{Always-go-left}
+\begin{exampleblock}{The Always-go-left algorithm}
+\begin{enumerate}
+\item Partition the $n$ bins into $d$ groups.
+\item Assume a fixed ordering of the bins $b_i$, $1\leq i \leq n$.
+\item Choose one bin from each group uniformly and independently.
+\item Put the ball into the least full bin.
+\item In case a tie occurs, put the ball into the leftmost bin (i.e., the bin with smallest index) among those with same minimal load.
+\end{enumerate}
+\end{exampleblock}
+\todo{visualize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Always-go-left}
+\begin{theorem}[V\"ocking, 2003]
+When using the \alert{Always-go-left} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be  (w.h.p.~and for all $m$): 
+\begin{align*}
+\load = \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1) \quad .
+\end{align*}
+\end{theorem}
+\todo{$\Phi_d$ is undefined. I think a full definition only makes sense if we present the proof for Always-go-left.}
+
+\bigskip
+The combination of \alert{asymmetric tie breaking} and \alert{non-uniform} (and independent) sampling of the $d$ bins is crucial to achieve this upper bound.
+\end{frame}
+
+\begin{frame}
+\frametitle{BiB Model}
+We will assume a sequential, infinite, dynamic process.
+
+\begin{description}
+\item[\alert{Sequential:}] The balls arrive one after another.
+\item[\alert{Infinite:}] Ball placements and deletions are given for an infinite time span.
+\item[\alert{Dynamic:}] At each time step $t$, one ball is removed from the bins.
+\begin{itemize}
+\item An oblivious adversary provides a sequence $\sigma = \sigma_1 \sigma_2 \dots$, where $\sigma_t = (d_t, a_t)$.
+\item $d_t$/$a_t$: ball to delete/add at time step t. \\
+\item $\sigma$ is fixed in advance.
+\end{itemize}
+\end{description}
 \end{frame}
@@
 \frametitle{Proof of either Greedy or Always-go-left}
+\end{frame}
+
+\begin{frame}
+\frametitle{Conclusion}
 \end{frame}
@@
 
-\begin{comment}
-\begin{frame}[label=references]
-  \frametitle{References}
-  %\nocite{*}
-  %\setbeamertemplate{bibliography item}[book]
+%\againframe{references}
+
+\begin{frame}[allowframebreaks, label=references]
+\frametitle<presentation>{References}    
 \bibliographystyle{acm}
+\nocite{*}
+\setbeamertemplate{bibliography item}[text]
 \bibliography{../sources}
 \end{frame}
-\note{}
-\end{comment}
-
-\begin{frame}{Introduction}
-\end{frame}
-\note{}
-
-%\againframe{references}
 
@@
 
-\begin{frame}{Title}
-\end{frame}
-\note{}
-
 %  LocalWords:

### 2014-03-24 presentation/presentation.tex
give each ball a node label
enhance BiB pictures

@@
 \usepackage{tikz}
+\usetikzlibrary{calc} %for calculations in tikz commands
+\usetikzlibrary{positioning}
+\usetikzlibrary{arrows}
 \usepackage{xxcolor}
-\usetikzlibrary{positioning}
 
@@
 %get some colors from the scheme
-\ccft{block body example}{myblock body fg}{myblock body bg} 
+\ccft{block body example}{examplefg}{examplebg}
+
 %---------------------------------------------------------------------
@@
 
+%arrow style
+\tikzstyle{insert} = [
+	thick,->,>=stealth'
+]
+
+%draw an arrow symbolizing a location
+\newcommand\iA[2][0]{
+	\path (\nbid) edge[insert, bend left=#1] (#2);
+}
 %bin style
@@
     		\pgfextra
-        		\fill[fill=myblock body bg] 
+        		\fill[fill=examplebg] 
         			(\tikzlastnode.north east) [rounded corners] |-
@@
 
+%set a node at bin #1, position #2 with node id = #1+(#2-1)*6
+\newcommand\setNode[2]{
+	\draw let \n1 ={#1#2} in node[circle, minimum size = \ballsize](n\n1) at (#1*\bingap,#2*\balldiameter-\ballsize) {};
+}
+
+%create nodes for all possible balls
+\newcommand\nodes{
+	\foreach \i in {1,...,6}
+		\foreach \j in {1,...,5}
+			\setNode{\i}{\j};
+}
+
 %paint a ball in bin #1, position #2
 \newcommand\ball[2]{
-	\shade[ballstyle] (#1*\bingap,#2*\balldiameter-\ballsize) circle (\ballsize)
+	\shade[ballstyle] (#1*\bingap,#2*\balldiameter-\ballsize) circle (\ballsize) {};
+}
+
+%paint a ball that is to be inserted
+\newcommand\nbid{nb}
+\newcommand\newball{
+	\draw node[circle, minimum size = \ballsize](\nbid) at (0*\bingap,6*\balldiameter-\ballsize) {};
+	\ball{0}{6};
 }
@@
 	\bins
+	\nodes
 	\balls{#1}
+}
+%draw group seperator after bin #1
+\newcommand\groupSep[1]{
+	\draw[thick, color=red!40] let \n1={(#1+0.5)*\bingap} in (\n1,-\ballsize) -- (\n1,\binheight +\ballsize);
 }
@@
 \note{}
-
-\begin{comment}
-\begin{frame}
-    \frametitle{Table of Contents}
-    \tableofcontents
-\end{frame}
-\end{comment}
 
@@
 \end{exampleblock}
-
 \end{frame}
@@
 \end{exampleblock}
+
 \bigskip
+
 \begin{center}
 \begin{tikzpicture}[scale=\scalefac]
-	\bab{2 1 5 2 3 0}
+	\bab{2 1 4 2 3 0}
+	\newball
+	\iA[40]{n23};
 \end{tikzpicture}
@@
 %\end{exampleblock}
-
 \end{frame}
@@
 \end{exampleblock}
-\todo{visualize}
+
+\begin{center}
+\begin{tikzpicture}[scale=\scalefac]
+	\bab{2 1 4 2 3 0}
+	\newball
+	\iA[25]{n13};
+	\iA[20]{n35};
+	\iA[35]{n44};
+\end{tikzpicture}
+\end{center}
+
 \end{frame}
@@
 
-\begin{frame}
+\begin{frame}[shrink]
 \frametitle{Always-go-left}
@@
 \end{exampleblock}
-\todo{visualize}
+
+\begin{center}
+\begin{tikzpicture}[scale=\scalefac]
+	\bab{2 1 4 2 3 0}
+	\newball
+	\groupSep{2};
+	\groupSep{4};
+	\iA[25]{n13};
+	\iA[35]{n43};
+	\iA[30]{n54};	
+\end{tikzpicture}
+\end{center}
+
 \end{frame}

### 2014-03-25 presentation/presentation.tex
some minor fixes+#

@@
     \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
-    \end{cases}
+    \end{cases} \quad .
 \end{align*}
@@
 \item Uniformly and independently
-\item Non-unfiromly and independently
+\item Non-uniformly and independently
 \item Non-uniformly and dependently
@@
 \begin{theorem}[Azar, Broder, Karlin and Upfal, 1999]
-When using the \alert{Greedy} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be (w.h.p.~and for all $m$): 
+When using the \alert{Greedy} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
 \begin{align*}
@@
 \end{align*}
-\alert{Greedy} with $d=2$:
+\alert{Greedy}
 \begin{itemize}
@@
 \begin{theorem}[V\"ocking, 2003]
-When using the \alert{Always-go-left} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be  (w.h.p.~and for all $m$): 
+When using the \alert{Always-go-left} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
 \begin{align*}
@@
 
+\pause
 \bigskip
 The combination of \alert{asymmetric tie breaking} and \alert{non-uniform} (and independent) sampling of the $d$ bins is crucial to achieve this upper bound.
+
+\medskip
+V\"ocking further proved a matching lower bound for $\load$, i.e., the upper bound is tight.
+
 \end{frame}
@@
 \begin{description}
-\item[\alert{Sequential:}] The balls arrive one after another.
+\item[\alert{Sequential:}] The balls arrive one after another, one at each time step $t$.
 \item[\alert{Infinite:}] Ball placements and deletions are given for an infinite time span.
-\item[\alert{Dynamic:}] At each time step $t$, one ball is removed from the bins.
+\item[\alert{Dynamic:}] Balls do not stay in the bins forever.
 \begin{itemize}
@@
 \frametitle{Proof of either Greedy or Always-go-left}
+I suggest to present the Greedy proof:
+\begin{itemize}
+\item Both proofs have the same structure, but Always-go-left is a lot more technical
+\item For Always-go-left, d-ary Fibonacci trees and a generalized golden ratio need to be introduced
+\item complete d-ary trees are easier to think and reason about (number of nodes/leaves, the inductive argument used)
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Witness trees}
+\end{frame}
+
+\begin{frame}
+\frametitle{Witness trees}
+\begin{alertblock}{Assumption}
+All the events represented by the nodes and edges of a witness tree are stochastically independent. 
+\end{alertblock}
 \end{frame}
@@
 %\againframe{references}
-
+%todo: why so few references shown?
 \begin{frame}[allowframebreaks, label=references]

### 2014-03-25 presentation/presentation.tex
some more minor fixes

@@
 \begin{description}
-\item[\alert{Sequential:}] The balls arrive one after another, one at each time step $t$.
+\item[\alert{Sequential:}] The balls arrive one after another, one at each time step~$t$.
 \item[\alert{Infinite:}] Ball placements and deletions are given for an infinite time span.
@@
 \begin{itemize}
-\item Both proofs have the same structure, but Always-go-left is a lot more technical
-\item For Always-go-left, d-ary Fibonacci trees and a generalized golden ratio need to be introduced
-\item complete d-ary trees are easier to think and reason about (number of nodes/leaves, the inductive argument used)
+\item Both proofs have the same structure, but Always-go-left is a lot more technical.
+\item For Always-go-left, d-ary Fibonacci trees and a generalized golden ratio need to be introduced.
+\item Complete d-ary trees are easier to think and reason about (number of nodes/leaves, the inductive argument used).
 \end{itemize}
+\todo{The following slides are preliminary.}
 \end{frame}
@@
 \bibliographystyle{acm}
-\nocite{*}
+\nocite{ABKU99}
+%\nocite{RS98}
+\nocite{VOC03}
 \setbeamertemplate{bibliography item}[text]

### 2014-03-25 article/article.tex
some minor fixes for proof 2

@@
 \begin{align*}
-\load = \frac{m}{n}+\frac{\ln \ln n}{\ln d} + O(1) \Rightarrow \delta = \frac{\ln \ln n}{\ln d} + O(1)
+\load = \frac{m}{n}+\frac{\ln \ln n}{\ln d} + O(1) \implies \delta = \frac{\ln \ln n}{\ln d} + O(1) \quad ,
 \end{align*}
@@
 \label{sec:analysis:alg}
-Similarly to Section~\ref{sec:analysis:greedy}, where we used symmetric witness trees to upper-bound the probability that a bin contains too many balls when using Greedy as allocation algorithm, we will use asymmetric witness trees (AWT) to obtain an analogous upper bound for the Always-go-left scheme.
-Again, we will use Assumption~\ref{assumption:independence}, i.e., we assume that all the events represented by the edges and nodes of a witness tree are stochastically independent.
+Similarly to Section~\ref{sec:analysis:greedy}, where we used SWTs to upper-bound the probability that a bin contains too many balls when using Greedy as allocation algorithm, we will use asymmetric witness trees (AWT) to derive an analogous upper bound for the Always-go-left scheme.
+Again, we will adopt Assumption~\ref{assumption:independence}, i.e., we assume that all the events represented by the edges and nodes of a witness tree are stochastically independent.
 
@@
 \label{sec:analysis:definitionAsymWT}
-An AWT differes from an SWT (Section~\ref{sec:analysis:definitionSymWT}) only by its structure, which is that of a $d$-ary Fibonacci tree $T_d\left(k \right)$:
+An AWT differes from an SWT (Section~\ref{sec:analysis:definitionSymWT}) only by its structure, which is a $d$-ary Fibonacci tree $T_d\left(k \right)$:
 
@@
 A $d$-ary AWT of order $L$ has the structure of a Fibonacci tree $T_d(d  L+1)$ and thus has $l=F_d(d L + 1) \geq \Phi_d^{d L-1}$ leaves (Eq.~\ref{eqn:goldenFib})\todo{Can we just rewrite Eq.~\ref{eqn:goldenFib} like this?
-what about the limes?}.
+what about the limes?
+It's not Eq.~6.
+Try to find a reference}.
 Furthermore, for each node we define a label $(h, i)$, where $h, i \in \mathbb{N}$, $0 \leq h \leq L$ and  $1\leq i \leq d$:
@@
 \begin{invariant}
-A ball $b_v$ referenced by a node $v$ with label $(h,i)$ was placed in a bin $x$ belonging to group $i$ that, at the time of insertion of the ball $b_v$, contained at least $h+3$ other ball.
+A ball $b_v$ referenced by a node $v$ with label $(h,i)$ was placed in a bin $x$ belonging to group $i$ that, at the time of insertion of the ball $b_v$, contained at least $h+3$ other balls.
 The ball represented by the root node may be put in a bin belonging to any group.
@@
 The topmost ball in the bin that location $j$ points to is assigned to the child node $u$ of $v$ labelled with $(h-1, j)$.
-\todo{the paper claims that the locations $j$ point to bins with $h+2$ balls.
+
+\begin{comment}
+Note: The paper claims that the locations $j$ point to bins with $h+2$ balls.
 This can not be true, since the topmost, i.e., the $h+2$-th ball - call it $b_t$ - gets assigned to a node with label $(h'=h-1, j)$.
 Since there are $h+1=h'+2$ balls below it, at the time of the placement of ball $b_t$, the bin $j$ it gets put into has only $h'+2$ balls in it, violating the invariant.
-Same with the case $j<i$.}
+Same with the case $j<i$.
+\end{comment}
 \end{compactitem}
@@
 \begin{align*}
-\frac{n}{d}\sum_{i=1}^d \beta_i \leq \frac{n}{3} \Rightarrow
+\frac{n}{d}\sum_{i=1}^d \beta_i \leq \frac{n}{3} \implies
 \sum_{i=1}^d \beta_i \leq \frac{d}{3}\quad .
@@
 \end{align*}
-\todo{the paper uses $+1$, instead of $+1/d$, but I can't workout the below result then} 
 Then, 
@@
 \begin{align*}
-\forall i\; \forall j \in \mathbb{N}: i < j \quad \Rightarrow \quad \Phi_i < \Phi_j \quad , \quad \lim_{d\rightarrow \infty} \Phi_d = 2 \quad .
+\forall i\; \forall j \in \mathbb{N}: i < j \implies \Phi_i < \Phi_j \quad , \quad \lim_{d\rightarrow \infty} \Phi_d = 2 \quad .
 \end{align*}
@@
 \begin{align*}
-f \in O(g) &\Leftrightarrow \exists c\; \exists n_0 \;\forall n> n_0: f(n)\leq c g(n) \\
-f \in \Omega(g) &\Leftrightarrow \exists c \; \exists n_0\;\forall n> n_0: cg(n) \leq f(n) \\
-f \in \Theta(g) &\Leftrightarrow f \in \Omega(g) \land f\in O(g) \\
-f \in o(g) &\Leftrightarrow \forall c \; \exists n_0\;\forall n> n_0: f(n) < cg(n) \\
-f \in \omega(g) &\Leftrightarrow \forall c \; \exists n_0\;\forall n> n_0: cg(n) < f(n) \\
+f \in O(g) &\iff \exists c\; \exists n_0 \;\forall n> n_0: f(n)\leq c g(n) \\
+f \in \Omega(g) &\iff \exists c \; \exists n_0\;\forall n> n_0: cg(n) \leq f(n) \\
+f \in \Theta(g) &\iff f \in \Omega(g) \land f\in O(g) \\
+f \in o(g) &\iff \forall c \; \exists n_0\;\forall n> n_0: f(n) < cg(n) \\
+f \in \omega(g) &\iff \forall c \; \exists n_0\;\forall n> n_0: cg(n) < f(n) \\
 \end{align*}

### 2014-03-26 article/article.tex
extend slides

@@
 \newtheorem{invariant}{Invariant}
+
 \begin{document}
@@
 \label{sec:analysis:definitionAsymWT}
-An AWT differes from an SWT (Section~\ref{sec:analysis:definitionSymWT}) only by its structure, which is a $d$-ary Fibonacci tree $T_d\left(k \right)$:
+An AWT differes from an SWT (Section~\ref{sec:analysis:definitionSymWT}) only by its structure, which is a $d$-step Fibonacci tree $T_d\left(k \right)$:
 
@@
 \end{compactitem}
-A $d$-ary AWT of order $L$ has the structure of a Fibonacci tree $T_d(d  L+1)$ and thus has $l=F_d(d L + 1) \geq \Phi_d^{d L-1}$ leaves (Eq.~\ref{eqn:goldenFib})\todo{Can we just rewrite Eq.~\ref{eqn:goldenFib} like this?
-what about the limes?
-It's not Eq.~6.
-Try to find a reference}.
+A $d$-ary AWT of order $L$ has the structure of a Fibonacci tree $T_d(d  L+1)$ and thus has $l=F_d(d L + 1) \geq \Phi_d^{d L-1}$ leaves (Eq.~\ref{eqn:capocelli}).
 Furthermore, for each node we define a label $(h, i)$, where $h, i \in \mathbb{N}$, $0 \leq h \leq L$ and  $1\leq i \leq d$:
@@
 \item The root nodes of the Fibonacci trees $T_d(1)$ and $T_d(2)$ are assigned the labels $(0,1)$ and $(0,2)$, respectively. Thus, nodes labelled $(0,1)$ or $(0,2)$ do not have any children.
-\item A node with label $(0, i)$, where $i>2$, has $i-1$ children with the labels $(0,i-1),\dots,(0,1)$
+\item A node with label $(0, i)$, where $i>2$, has $i-1$ children with the labels $(0,i-1),\dots,(0,1)$.
 \item A node with label $(h,i)$, where $h>0$, has $d$ children labelled $(h,i-1),\dots,(h,1)$,$(h-1,d),\dots,(h-1, i)$.
@@
 \end{align*}
-i.e., a leaf event occurs with probability $p_\mathrm{l}'$, which is maximized when all $b_i$ are equal.
-We thus set $b_i = 1/3$ for all $i$ and get as upper bound for the probability that all leaf events occur
+i.e., a leaf event occurs with probability $p_\mathrm{l}'$, which is maximized when all $\beta_i$ are equal.
+We thus set $\beta_i = 1/3$ for all $i$ and get as upper bound for the probability that all leaf events occur
 \begin{align*}
@@
 \label{sec:fibonacci}
-We define $d$-ary Fibonacci numbers as follows:
+We define $d$-step Fibonacci numbers as follows:
 
@@
 \end{align*}
-Notice that $F_2$ corresponds to the usual Fibonacci numbers, while $\Phi_d$ is a generalization of the golden ratio~\cite{Knuth73} and $\Phi_2 $ corresponds to the golden ratio.
+Notice that $F_2$ corresponds to the usual Fibonacci numbers and $\Phi_2 $ corresponds to the golden ratio~\cite{Knuth73}, of which $\Phi_d$ is a generalization, for which
+\begin{align*}
+F_d(k) \geq \Phi_d^{k-2} \neqn{capocelli}
+\end{align*}
+holds ~\cite{capocelli1990generalized}.
 

### 2014-03-26 presentation/presentation.tex
remove some commented out stuff from the tex file
use sanserif font in presentation
extend slides

--- 2014-03-25 presentation/presentation.tex
@@
 \documentclass[serif,professionalfonts]{beamer}
-
-%\usepackage{pgfpages}
-%\setbeameroption{show only notes}
-%\setbeameroption{show notes on second screen=top}
 
@@
 % Hermann Zapf Optima clone (wonderful sanserif font) -- URW Classico
-%\renewcommand{\sfdefault}{uop}
-%\renewcommand{\rmdefault}{ppl}
-%\usepackage{eulervm}
-%\usepackage[right]{eurosym}
-%\usepackage{paralist}
-%\usepackage{algorithm2e}
-
-% professional tables
-%\usepackage{booktabs}
-%\usepackage{pdfpages}
-%\usepackage{multirow}
-%\usepackage{array}
-%\usepackage{dcolumn}
-%\usepackage{subfigure}
-%\usepackage{epsdice}
+\renewcommand{\sfdefault}{uop}
+\renewcommand{\rmdefault}{ppl}
 
@@
 \usetikzlibrary{arrows}
+\usepackage{tikz-qtree}
 \usepackage{xxcolor}
@@
 
-% some math operators
-%\DeclareMathOperator{\har}{H}
-%\DeclareMathOperator{\E}{E}
-%\DeclareMathOperator{\lag}{lag}
-%\DeclareMathOperator{\var}{Var}
-%\DeclareMathOperator{\mf}{\enspace .}
-%\DeclareMathOperator{\mc}{\enspace ,}
-%\newcommand{\eye}{\mathbb{I}}
-%\newcommand{\Z}{\mathbb{Z}}
-%\newcommand{\dom}{\{0,1\}^n}
-%\newcommand{\reals}{\mathbb{R}}
-%\newcommand{\nats}{\mathbb{N}}
-%\newcommand{\deq}{\triangleq}
-%\newcommand{\pd}{\cdot_p}
-%\newcommand{\pp}{+_p}
-%\newcommand{\pmn}{-_p}
-%\newcommand{\ed}{\epsdice}
-%\newcommand{\bled}{\epsdice[black]}
-%\newcommand{\bitup}{\vspace*{-4mm}}
-%\newcommand{\nospace}{\abovedisplayskip=0pt}
-
 \setbeamertemplate{caption}
@@
 
-%reduce font size of footnotes
-%\let\oldfootnotesize\footnotesize
-%\renewcommand*{\footnotesize}{\oldfootnotesize\tiny}
-
-%
 \newcommand\todo[1]{\textcolor{red}{(TODO: #1)}}
 \newcommand\load{L_{\mathrm{max}}}
+\newcommand\loadS{\load^{\mathrm{S}}}
+\newcommand\loadG{\load^{\mathrm{G}}}
+\newcommand\loadAgl{\load^{\mathrm{A}}}
 \newtheorem{claim}{Claim}
@@
 \newtheorem{invariant}{Invariant}
-
-%---------------------------------------------------------------------
-%Front matter
-%---------------------------------------------------------------------
-\title{The Power of Two Random Choices}
-%\subtitle{ A follow up seminar to the "Parallel Algorithms" lecture}
-\author[M. Kalany]{Martin Kalany}
-\institute[TU Wien]
-{
-  Graduate student in Computer Science\\
-  Vienna University of Technology\\
-}
-\date{\today}
-
-\begin{comment}
-\AtBeginSubsubsection[]
-{
-  \begin{frame}<beamer>
-    \frametitle{Outline}
-    \tableofcontents[subsubsectionstyle=show/shaded,currentsection,currentsubsection]
-  \end{frame}
-}
-
-\AtBeginSubsection[]
-{
-  \begin{frame}<beamer>
-    \frametitle{Outline}
-    \tableofcontents[subsubsectionstyle=show/shaded,currentsection,currentsubsection]
-  \end{frame}
-}
-
-\AtBeginSection[]
-{
-  \begin{frame}<beamer>
-    \frametitle{Outline}
-    \tableofcontents[subsubsectionstyle=show/shaded,currentsection,currentsubsection]
-  \end{frame}
-}
-\end{comment}
 
@@
 %---------------------------------------------------------------------
+% tikz stuff
+%---------------------------------------------------------------------
+
+% nicer tikz qtrees
+\tikzset{edge from parent/.style=
+     {draw, edge from parent path={(\tikzparentnode) -- (\tikzchildnode)}}}
+\tikzset{every tree node/.style={font=\tiny, thick}}
+
+%---------------------------------------------------------------------
+%Front matter
+%---------------------------------------------------------------------
+\title{The Power of Two Random Choices}
+\subtitle{TODO}
+\author[M. Kalany]{Martin Kalany}
+\institute[TU Wien]
+{
+  Graduate student in Computer Science\\
+  Vienna University of Technology\\
+}
+\date{\today}
+
+%---------------------------------------------------------------------
 %document
@@
 % 5 min more advanced stuff 
-% 15 min difficult stuff (the proof, either of the symmetric or asymmetric allocation scheme using the infamous assumption 1
+% 15 min difficult stuff (the proof, of the asymmetric allocation scheme using the infamous assumption 1)
 % 5 min further results and wrap up
@@
 
-\section{Balls into bins}
 \begin{frame}
@@
 \begin{frame}
-\frametitle{Applications}
-Applications:
-\begin{itemize}
-	\item Hashing
-	\item Online load balancing
-	\item Emulating PRAMs on DMMs
-	\item Low congestion circuit routing
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Single choice}
+\frametitle{Single choice strategy}
 \framesubtitle{A basic allocation scheme}
-\begin{exampleblock}{Single choice strategy}
-For each ball, choose one of the $n$ bins uniformly and independently at random.
+\begin{exampleblock}{Basic idea}
+For each ball, choose one of the $n$ bins \alert{uniformly and independently at random}.
 \end{exampleblock}
-
 \bigskip
@@
 \frametitle{Some definitions}
-
-Since we deal with a \alert{randomized process}, we cannot give upper bounds for the maximum load $\load$ with absolute certainty.
+\todo{merge with next slide?}
+Since we deal with a \alert{random process}, we cannot give upper bounds for the maximum load $\load$ with absolute certainty.
 
@@
 \bigskip
-\pause
+%\pause
 
-%\begin{exampleblock}{Definition}
 We distinguish:
@@
 \end{itemize}
-%\end{exampleblock}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Single choice}
+\frametitle{Single choice strategy}
 \framesubtitle{Upper bound for the maximum load}
@@
 \begin{align*}
-\load = 
+\loadS \leq 
 	\begin{cases}
@@
 
-%Table of contents at the start of each section
-\begin{comment}
-\AtBeginSection[]
-{
-  \begin{frame}
-    \frametitle{Table of Contents}
-    \tableofcontents[currentsection]
-  \end{frame}
-}
-\end{comment}
-
-\section{Allocation schemes}
 \begin{frame}
@@
 
-\pause
 \bigskip
 We can distinguish placement algorithms by the way the $d$ \alert{bins are sampled}:
-\begin{itemize}
+\begin{enumerate}
 \item Uniformly and independently
@@
 \item Non-uniformly and dependently
-\end{itemize}
+\end{enumerate}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Greedy}
-\begin{exampleblock}{The Greedy algorithm}
+\frametitle{The Greedy algorithm}
+\begin{exampleblock}{Algorithm}
 \begin{enumerate}
@@
 
+\only<1>{
 \begin{center}
@@
 \end{center}
+}
 
+\only<2->{
+\begin{theorem}[Azar, Broder, Karlin and Upfal, 1999]
+When using the \alert{Greedy} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
+\begin{align*}
+\loadG \leq \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) \quad .
+\end{align*}
+\end{theorem}
+}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Greedy}
-\begin{theorem}[Azar, Broder, Karlin and Upfal, 1999]
-When using the \alert{Greedy} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
+\frametitle{Single choice vs.~Greedy}
+%\begin{columns}[onlytextwidth]
+%\begin{column}{0.5\textwidth}
 \begin{align*}
-\load = \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) \quad .
-\end{align*}
-\end{theorem}
-\pause
-\medskip
-Compared to the \alert{single choice} strategy, with 
-\begin{align*}
-\load = 
+\loadS \leq 
 	\begin{cases}
@@
     \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
-    \end{cases} \quad ,
+    \end{cases}
 \end{align*}
-\alert{Greedy}
-\begin{itemize}
-\item achieves an \alert{exponential decrease} of the additive gap $\delta$ and
-\item $\delta$ does \alert{not depend on} the number of balls \alert{$m$} anymore.
-\end{itemize}
+%\end{column}
+%\begin{column}{0.5\textwidth}
+\begin{align*}
+\loadG \leq \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) 
+\end{align*}
+%\end{column}
+%\end{columns}
+
+\begin{alertblock}{The two-choice paradigm}
+Even for $d=2$, Greedy achieves an \alert{exponential decrease} of the additive gap $\delta$, which does \alert{not depend on} the number of balls \alert{$m$} anymore.
+\end{alertblock}
 \end{frame}
 
-\begin{frame}[shrink]
-\frametitle{Always-go-left}
-\begin{exampleblock}{The Always-go-left algorithm}
+\begin{frame}
+\frametitle{Applications}
+\begin{description}
+	\item[Hashing:] Assuming a perfect hash function, $n$ elements (balls) are mapped to $m$ table entries (bins).
+Using two hash functions: look-up time $\Theta(\ln\ln n)$, w.h.p.
+\alert{Easy to parallelize} and no \alert{re-hashing}!
+	\item[Online load balancing:] Assign tasks (balls) to servers/processors (bins).
+The two-choice paradigm helps to \alert{reduce the overhead} (of e.g., a central dispatcher), while the maximum load will be at most  $\Theta(\ln\ln n)$, w.h.p.
+	\item[Emulating PRAMs on DMMs:] Distribute processors and memory cells of the PRAM on the DMM s.t.~the overhead of the emulation is minimized.
+	\item[Low congestion circuit routing:] \alert{Valiant's paradigm} (a permuation routing problem can be split into two sub-problems: route to random intermediate destination and then on to the actual destination) is analogous to the single-choice technique.
+\end{description}
+\end{frame}
+
+\newcommand\aglAlgorithm{
+\begin{exampleblock}{Algorithm}
 \begin{enumerate}
@@
 \end{exampleblock}
+}
+
+\begin{frame}[shrink]
+\frametitle{The Always-go-left algorithm}
+\aglAlgorithm
 
@@
 \begin{frame}
-\frametitle{Always-go-left}
+\frametitle{$d$-step Fibonacci numbers and the golden ratio}
+\framesubtitle{(Capocelli, Cull, 1990)}
+
+$d$-step Fibonacci numbers:
+\begin{align*}
+F_d(k) := \begin{cases}
+                0               & k \leq 0\\
+                1               & k = 1\\
+                \sum_{i=1}^{d}F_d(k-i) & k > 1
+            \end{cases}
+\end{align*}
+Generalized golden ratio $\Phi_d$:
+\begin{align*}
+\Phi_d := \lim_{k \rightarrow \infty} \frac{F_d(k)}{F_d(k-1)}
+\end{align*}
+\begin{align*}
+\Phi_2 = 1.61 \quad , \quad \forall i\; \forall j \in \mathbb{N}: i < j \implies \Phi_i < \Phi_j \quad , \quad \lim_{d\rightarrow \infty} \Phi_d = 2 \quad .
+\end{align*}
+
+\begin{align*}
+F_d(k) \geq \Phi_d^{k-2}
+\end{align*}
+\end{frame}
+
+\newcommand\theoremVocking{
 \begin{theorem}[V\"ocking, 2003]
@@
 \begin{align*}
-\load = \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1) \quad .
+\loadAgl \leq \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1) \quad .
 \end{align*}
 \end{theorem}
-\todo{$\Phi_d$ is undefined. I think a full definition only makes sense if we present the proof for Always-go-left.}
+}
 
-\pause
+\begin{frame}[shrink]
+\frametitle{The Always-go-left algorithm}
+\aglAlgorithm
+
+\theoremVocking
+\end{frame}
+
+\begin{frame}
+\frametitle{The Always-go-left algorithm}
+\begin{columns}[onlytextwidth]
+\begin{column}{0.5\textwidth}
+\begin{align*}
+\loadG \leq \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) 
+\end{align*}
+\end{column}
+\begin{column}{0.5\textwidth}
+\begin{align*}
+\loadAgl \leq \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
+\end{align*}
+\end{column}
+\end{columns}
 \bigskip
 The combination of \alert{asymmetric tie breaking} and \alert{non-uniform} (and independent) sampling of the $d$ bins is crucial to achieve this upper bound.
-
-\medskip
-V\"ocking further proved a matching lower bound for $\load$, i.e., the upper bound is tight.
-
+\bigskip
+\begin{theorem}[V\"ocking, 2003]
+For $m=n$, using an arbitrary allocation scheme that chooses $d$ bins for each ball at random (i.e., class 3) the maximum load is, w.h.p,
+\begin{align*}
+\loadAgl \leq \frac{\ln \ln n}{d \ln \Phi_d} - O(1) \quad .
+\end{align*}
+\end{theorem}
 \end{frame}
@@
 \begin{description}
-\item[\alert{Sequential:}] The balls arrive one after another, one at each time step $t$.
-\item[\alert{Infinite:}] Ball placements and deletions are given for an infinite time span.
-\item[\alert{Dynamic:}] Balls do not stay in the bins forever.
+\item[Sequential:] The balls arrive one after another, one at each time step~$t$.
+\item[Infinite:] Ball placements and deletions are given for an infinite time span.
+\item[Dynamic:] Balls do not stay in the bins forever.
 \begin{itemize}
@@
 
+% Outline of the proof section:
+% - State what will be proved
+% - Outline the proof
+% - Witness trees 
+% - Fibonacci trees and labelling
+% - Lemma 3
+% - Prob. of activation
 \begin{frame}
-\frametitle{Proof of either Greedy or Always-go-left}
-I suggest to present the Greedy proof:
+\frametitle{The Always-go-left algorithm}
+\framesubtitle{Proof of the upper bound for $m=n$}
+\begin{theorem}[V\"ocking, 2003]
+\label{theorem:aglm}
+Consider any sequence of insertions and deletions s.t.~at most $n$ balls are in the bins at any time.
+If the balls are placed into bins by the Always-go-left algorithm, the maximum load at any time step $t$ is, w.h.p.,
+\begin{align*}
+\loadAgl \leq \frac{\ln\ln n}{d  \ln \Phi_d} + O(1) \quad .
+\end{align*}
+\end{theorem}
+
+\bigskip
+In case $\load$ exceeds a given threshold of $L+3$ balls ("a bad event occurs"), we construct an activated \alert{witness tree of order $L$}.
+The probability for its existance upper bounds the probabilty that the associated bad event occurs.
+\end{frame}
+
+\newcommand\assumptionOne{
+\begin{alertblock}{Assumption}
+All the events represented by the nodes and edges of a witness tree are stochastically independent. 
+\end{alertblock}
+}
+
+\begin{frame}
+\frametitle{Witness trees}
+A witness tree
 \begin{itemize}
-\item Both proofs have the same structure, but Always-go-left is a lot more technical
-\item For Always-go-left, d-ary Fibonacci trees and a generalized golden ratio need to be introduced
-\item complete d-ary trees are easier to think and reason about (number of nodes/leaves, the inductive argument used)
+\item is a \alert{tree-like structure},
+\item where each node $v$ references a ball $b_v$ and
+\item nodes and edges represent \alert{events} (that may occur or not),
+\item that are, in general, \alert{not independent}.
+\item It is \alert{activated} when all its leaf and edge events occur. 
 \end{itemize}
+
+\bigskip
+
+\assumptionOne
+
+I.e., we assume that a ball is referenced by at most one node.
 \end{frame}
@@
 \begin{frame}
-\frametitle{Witness trees}
+\frametitle{Asymmetric witness trees (AWT)} 
+A $d$-ary AWT of order $L$ has the structure of a Fibonacci tree $T_d(dL+1)$:
+
+\begin{center}
+\begin{tikzpicture}
+\matrix{
+ \Tree 
+	[.$T_d(6)$
+		[.$T_d(5)$
+			[.$T_d(4)$ 
+				[.$T_d(3)$ $T_d(2)$ $T_d(1)$ ]
+				$T_d(2)$ $T_d(1)$
+			]
+			[.$T_d(3)$ $T_d(2)$ $T_d(1)$ ]
+			$T_d(2)$ %$T_d(1)$
+		]
+		[.$T_d(4)$ 
+			[.$T_d(3)$ $T_d(2)$ $T_d(1)$ ]
+			$T_d(2)$ $T_d(1)$
+		]
+		[.$T_d(3)$ $T_d(2)$ $T_d(1)$ ]
+		%$T_d(2)$ 
+	]
+\\};
+\end{tikzpicture}
+\end{center}
+\end{frame}
+
+\newcommand\defEdgeEvent{
+\begin{exampleblock}{Edge event}
+Edge $e =(u,v)$, s.t.~$v$ is $i$-th child of $u$: $e$ represents the event that the $i$-th location of ball $b_u$ points to the same bin as one of the locations of ball $b_v$.
+\end{exampleblock}
+}
+
+\newcommand\defLeafEvent{
+\begin{exampleblock}{Leaf event}
+A leaf node $v$ represents the event that each of the $d$ locations of the ball $b_v$ points to a bin that contains at least three additional balls, at the time of the insertion of the ball $b_v$.
+\end{exampleblock}
+}
+
+\begin{frame}
+\frametitle{Edge and leaf events}
+\begin{columns}[onlytextwidth]
+\begin{column}{0.7\textwidth}
+\defEdgeEvent
+
+\defLeafEvent
+\end{column}
+\begin{column}{0.3\textwidth}
+\begin{tikzpicture}
+\todo{illustrate}
+\end{tikzpicture}
+\end{column}
+\end{columns}
+
+\end{frame}
+
+\begin{frame}[shrink]
+\frametitle{Asymmetric witness trees (AWT)} 
+We assign a label $(h,i)$, with $0\leq h \leq L$, $1\leq i \leq d$, to each node:
+\begin{itemize}
+\item Node $T_d(1)$/$T_d(2)$ is labelled $(0,1)$/$(0,2)$.
+\item A node labelled $(0,i)$, $i>2$, has $i-1$ children labelled $(0,i-1),\dots,(0,1)$.
+\item A node with label $(h,i)$, $h>0$, has $d$ children labelled
+$(h,i-1),\dots,(h,1)$,$(h-1,d),\dots,(h-1, i)$.
+\item Thus, the node $T_d(dh+i)$ is labelled $(h,i)$.
+\end{itemize}
+\begin{center}
+\begin{tikzpicture}
+\matrix{
+ \Tree 
+	[.$(1,2)$ 
+		[.$(1,1)$
+			[.$(0,3)$ $(0,2)$ $(0,1)$ ]
+			$(0,2)$ $(0,1)$
+		]
+		[.$(0,3)$ $(0,2)$ $(0,1)$ ]
+		$(0,2)$
+	]
+\\};
+\end{tikzpicture}
+\end{center}
 \end{frame}
@@
 \begin{frame}
-\frametitle{Witness trees}
-\begin{alertblock}{Assumption}
-All the events represented by the nodes and edges of a witness tree are stochastically independent. 
-\end{alertblock}
+\frametitle{Proof of activation}
+We assign balls to the nodes of the AWT s.t.~th
+
+\begin{exampleblock}{Invariant}
+A ball $b_v$ referenced by a node $v$ with label $(h,i)$ was placed in a bin $x$ belonging to group $i$ that, at the time of insertion of the ball $b_v$, contained at least $h+3$ other balls. 
+\end{exampleblock}
+
+\end{frame}
+
+\begin{frame}
+\frametitle{Probability of activation}
+\begin{lemma}
+For any $\alpha > 0$ and large enough $L$, the activation of an AWT of order L occurs with probability at most $n^{-\alpha}$.
+\end{lemma}
+
+\bigskip
+\begin{itemize}
+\item $k$\dots number of nodes
+\item $l$\dots number of leaves
+\item $\omega$\dots number of ways to assign balls to the nodes of an AWT
+\item $p_\mathrm{e}$\dots probability that all edge events occur
+\item $p_\mathrm{l}$\dots probability that all edge events occur
+\item $P_\mathrm{a}$\dots probability of activation of an AWT of order $L$
+\end{itemize}
+
+\begin{align*}
+P_\mathrm{a} = \omega p_e p_l \quad , \quad \omega = n^k 
+\end{align*}
+\end{frame}
+
+\begin{frame}
+\frametitle{Edge events}
+\defEdgeEvent
+\bigskip
+One edge is activated with probability at most $d/n$ and we have $k-1$ edges; thus
+\begin{align*}
+p_\mathrm{e} = \left(\frac{d}{n}\right)^{k-1} \quad .
+\end{align*}
+\end{frame}
+
+\begin{frame}
+\frametitle{Leaf events}
+\only<1>{
+\defLeafEvent
+At any time step $t$, at most $n/3$ bins contain 3 or more balls. \alert{But how are they distributed among the $d$ groups?}
+
+\medskip
+$b_i$\dots number of bins in group $i$ containing 3 or more balls.
+$\beta_i = b_i d/n$\dots fraction of bins $b_i$ in group $i$. 
+}
+\only<1-2>{
+\begin{align*}
+\frac{n}{d}\sum_{i=1}^d \beta_i \leq \frac{n}{3} \implies
+\sum_{i=1}^d \beta_i \leq \frac{d}{3}
+\end{align*}
+}
+
+\only<2>{
+The probability that all $d$ locations of a ball point to a bin containing 3 or more balls is
+\begin{align*}
+p_\mathrm{l}' = \prod_{i=1}^{d} \beta_i \quad ,
+\end{align*}
+which is maximized when all $\beta_i$ are equal and thus
+\begin{align*}
+p_\mathrm{l}' \leq  \left(\frac{1}{3}\right)^d \implies p_\mathrm{l} \leq 3^{-dl} \quad .
+\end{align*}
+}
+\end{frame}
+
+\begin{frame}
+\frametitle{Probability of activation}
+Putting it together, we get
+\begin{align*}
+P_\mathrm{a} = \omega p_e p_l \leq  n^k \left(\frac{d}{n}\right)^{k-1} 3^{-dl} \quad .
+\end{align*}
+Using $ k\leq 2l$, $2d^2\leq 3^d$ and $l \geq \Phi_d^{dL-1}$, we can simplify to
+\begin{align*}
+P_\mathrm{a} \leq n 2 ^{-\Phi_d^{dL-1}} \quad .
+\end{align*}
+For arbitrary $\alpha > 0$, let
+\begin{align*}
+L &\geq \left\lceil{\frac{\ln\log_2 n + \ln\left(1+\alpha\right)}{d \ln \Phi_d}}\right\rceil+\frac{1}{d} \implies \frac{\ln\ln n}{d \ln \Phi_d} + O\left(1\right) ,
+\end{align*}
+and we get
+\begin{align*}
+P_\mathrm{a} \leq n^{-\alpha} \quad  \qed
+\end{align*}
+\end{frame}
+
+\begin{frame}
+\frametitle{Recap.}
+\theoremVocking
+\assumptionOne
+\bigskip 
+Next steps:
+\begin{itemize}
+\item Proof \alert{without} the simplifying assumption (V\"ocking, 2003).
+\item Proof for the \alert{heavily loaded} case (Berenbrink et al.,~2006).
+\end{itemize}
+
 \end{frame}
@@
 
-\begin{comment}
-\section{Test}
-\begin{frame}
-\frametitle{Ignore the following}
-\begin{tikzpicture}[scale=\scalefac]
-	\bab{1 2 3 4 5 6}
-\end{tikzpicture}
-\end{frame}
-\end{comment}
-
-%\againframe{references}
-%todo: why so few references shown?
-\begin{frame}[allowframebreaks, label=references]
+\begin{frame}[allowframebreaks]
 \frametitle<presentation>{References}    
 \bibliographystyle{acm}
-\nocite{*}
+\nocite{VOC03}
+\nocite{ABKU99}
+\nocite{RS98}
+\nocite{BCSV06}
+\nocite{capocelli1990generalized}
 \setbeamertemplate{bibliography item}[text]

### 2014-03-30 article/article.tex
add theorem for lower bound of agl and highlight that asymmetric tie-breaking and partitioning is crucial

@@
 (The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers.
-Refer to Appendix \ref{sec:fibonacci} for the definition.) V\"ocking was also able to show that the Always-go-left strategy is optimal up to additive constants~\cite{VOC03}.
+Refer to Appendix \ref{sec:fibonacci} for the definition.) Berenbrink et al.~\cite{BCSV06} and V\"ocking~\cite{VOC03} were able to show that the Always-go-left strategy is optimal up to additive constants.
 
@@
 \end{theorem}
+\noindent
+V\"ocking~\cite{VOC03} also proved a matching lower bound for the lightly loaded case:
+\begin{theorem}
+\label{theorem:aglmLower}
+For $m=n$, using an arbitrary allocation scheme that chooses $d$ bins for each ball at random (i.e., any allocation scheme in class 3) the maximum load is
+\begin{align*}
+\load = \frac{\ln \ln n}{d \ln \Phi_d} - O(1) \quad 
+\end{align*}
+\end{theorem}
+\noindent
+w.h.p.,~for all $t$~\cite{VOC03}.
+We not that the combination of asymmetric tie-breaking and and non-uniform (but independent) sampling of the $d$ bins is crucial to achieve bounds given in Theorems~\ref{theorem:aglm} and~\ref{theorem:aglmLower}.
 
@@
 \begin{compactitem}
-\item For $1\leq j < i$, the $j$'th location of the ball $b_v$ contains at least $h+4$ balls, at the insertion time of the ball $b_v$.
-If the $j$-th location contained less balls than the $i$-th location, the Always-go-left scheme would have put the ball $b_v$ into the bin pointed to by location $j$.
+\item For $1\leq j < i$, the $j$-th location of the ball $b_v$ contains at least $h+4$ balls, at the insertion time of the ball $b_v$.
+If it contained less balls than the $i$-th location, the Always-go-left scheme would have put the ball $b_v$ into the bin pointed to by location $j$.
 The topmost ball in location $j$ is assigned to the child $u$ of $v$ with label $(h,j)$.
 \item For $i < j \leq d$, the $j$-th location contains at least $h+3$ balls.
-If location $j$ contained less, the Always-go-left scheme would have put the ball $b_v$ into the bin location $j$ points to.
+If it contained less, the Always-go-left scheme would have put the ball $b_v$ into the bin location $j$ points to.
 The topmost ball in the bin that location $j$ points to is assigned to the child node $u$ of $v$ labelled with $(h-1, j)$.
@@
 all~\cite{BCSV06} were the first to prove that the same bound indeed holds for the heavily loaded case.
-Their proof is fairly complex and a simpler analysis may be found in~\cite{TW13}, which also provides a tight lower bound for the heavily loaded case.
+Their proof is fairly complex but provides a tight analysis of the heavily loaded case; a simpler analysis may be found in~\cite{TW13}.
 Curiously, the case $d = 2$ was implicitly proven by Karp et al.

### 2014-03-30 presentation/presentation.tex
polishing slides

@@
 
+\newcommand\xqed[1]{%
+   \rlap{\hbox to#1{\hfil\llap{\ensuremath{\qed}}}}
+}
+
 %---------------------------------------------------------------------
@@
 \tikzstyle{ballstyle} = [ball color=black!30!red]
+\tikzstyle{specialBallstyle} = [ball color=black!30!red,  opacity=0.2]
 
@@
 
-%paint a ball in bin #1, position #2
+%paint a ball in bin #1, position #2 with style specialBallstyle
+\newcommand\specialBall[2]{
+	\shade[specialBallstyle] (#1*\bingap,#2*\balldiameter-\ballsize) circle (\ballsize) {};
+}
+
+%paint a ball in bin #1, position #2 with style = ballstyle
 \newcommand\ball[2]{
@@
      {draw, edge from parent path={(\tikzparentnode) -- (\tikzchildnode)}}}
-\tikzset{every tree node/.style={font=\tiny, thick}}
+\tikzset{every tree node/.style={font=\tiny, thick, align=center}}
 
@@
 \title{The Power of Two Random Choices}
-\subtitle{TODO}
+\subtitle{And the Surprising Influence of Asymmetry}
 \author[M. Kalany]{Martin Kalany}
@@
 \end{frame}
-\note{}
 
 \begin{frame}
-\frametitle{The Problem}
+\frametitle{The problem}
+\framesubtitle{...and the goal}
 \begin{exampleblock}{Balls-into-bins games}
@@
 \end{exampleblock}
-
+\pause
 \begin{exampleblock}{Definition}
@@
 \begin{frame}
-\frametitle{Some definitions}
-\todo{merge with next slide?}
-Since we deal with a \alert{random process}, we cannot give upper bounds for the maximum load $\load$ with absolute certainty.
+\frametitle{Single choice strategy}
+\framesubtitle{Upper bound for the maximum load}
+%&Since we deal with a \alert{random process}, we cannot give upper bounds for the maximum load $\load$ with absolute certainty!
 
@@
 
-\bigskip
-%\pause
+\pause
+\medskip
+We distinguish the \alert{lightly loaded} case ($m = n$) and the \alert{heavily loaded} case ($m \in \omega(n\ln n)$).
 
-We distinguish:
-\begin{itemize}
-\item The \alert{lightly loaded} case: $m = n$
-\item The \alert{heavily loaded} case:	 $m \in \omega(n\ln n)$
-\end{itemize}
-\end{frame}
-
-\begin{frame}
-\frametitle{Single choice strategy}
-\framesubtitle{Upper bound for the maximum load}
+\medskip
 \begin{theorem}[Raab and Steger, 1998]
@@
 	\begin{cases}
-    \frac{\ln n}{\ln\ln n} + O(1)              & m = n \\
+    \Theta\left(\frac{\ln n}{\ln\ln n}\right)           & m = n \\
     \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
@@
 \end{theorem}
+
 \end{frame}
@@
 \frametitle{Multiple choice strategies}
-
 \begin{exampleblock}{Idea}
@@
 	\newball
-	\iA[25]{n13};
-	\iA[20]{n35};
+	\iA[35]{n22};
 	\iA[35]{n44};
@@
 \frametitle{Single choice vs.~Greedy}
-%\begin{columns}[onlytextwidth]
-%\begin{column}{0.5\textwidth}
 \begin{align*}
@@
 \end{align*}
-%\end{column}
-%\begin{column}{0.5\textwidth}
 \begin{align*}
@@
 \end{align*}
-%\end{column}
-%\end{columns}
 
 \begin{alertblock}{The two-choice paradigm}
-Even for $d=2$, Greedy achieves an \alert{exponential decrease} of the additive gap $\delta$, which does \alert{not depend on} the number of balls \alert{$m$} anymore.
+Even for $d=2$, Greedy achieves an \alert{exponential decrease} of the additive gap $\delta$, which does \alert{not depend on} the number of balls \alert{$m$}.
 \end{alertblock}
@@
 Using two hash functions: look-up time $\Theta(\ln\ln n)$, w.h.p.
-\alert{Easy to parallelize} and no \alert{re-hashing}!
-	\item[Online load balancing:] Assign tasks (balls) to servers/processors (bins).
-The two-choice paradigm helps to \alert{reduce the overhead} (of e.g., a central dispatcher), while the maximum load will be at most  $\Theta(\ln\ln n)$, w.h.p.
+\alert{Easy to parallelize} and  \alert{no re-hashing}!
+	\item[Online load balancing:] Assign tasks (balls) to servers/processors (bins). The two-choice paradigm helps to \alert{reduce communication overhead}, %(of e.g., a central dispatcher)
+	while the maximum load will be $\Theta(\ln\ln n)$, w.h.p.
 	\item[Emulating PRAMs on DMMs:] Distribute processors and memory cells of the PRAM on the DMM s.t.~the overhead of the emulation is minimized.
-	\item[Low congestion circuit routing:] \alert{Valiant's paradigm} (a permuation routing problem can be split into two sub-problems: route to random intermediate destination and then on to the actual destination) is analogous to the single-choice technique.
+	\item[Low congestion circuit routing:] \alert{Valiant's paradigm} %(a permuation routing problem can be split into two sub-problems: route to random intermediate destination and then on to the actual destination)
+	is analogous to the single-choice technique.
 \end{description}
@@
 	\newball
-	\groupSep{2};
-	\groupSep{4};
-	\iA[25]{n13};
-	\iA[35]{n43};
+	\groupSep{3};
+	\iA[30]{n13};
 	\iA[30]{n54};	
@@
 \end{align*}
-
 \begin{align*}
@@
 \aglAlgorithm
-
 \theoremVocking
@@
 \begin{align*}
-\loadAgl \leq \frac{\ln \ln n}{d \ln \Phi_d} - O(1) \quad .
+\loadAgl = \frac{\ln \ln n}{d \ln \Phi_d} - O(1) \quad .
 \end{align*}
@@
 
-% Outline of the proof section:
-% - State what will be proved
-% - Outline the proof
-% - Witness trees 
-% - Fibonacci trees and labelling
-% - Lemma 3
-% - Prob. of activation
 \begin{frame}
@@
 
+\pause
 \bigskip
 In case $\load$ exceeds a given threshold of $L+3$ balls ("a bad event occurs"), we construct an activated \alert{witness tree of order $L$}.
-The probability for its existance upper bounds the probabilty that the associated bad event occurs.
+The probability for its existence upper bounds the probability that the associated bad event occurs.
 \end{frame}
@@
 \bigskip
-
 \assumptionOne
-
-I.e., we assume that a ball is referenced by at most one node.
 \end{frame}
@@
 \begin{center}
-\begin{tikzpicture}
-\matrix{
+\begin{tikzpicture}[scale=1]
  \Tree 
+ 	[.$T_d(7)$ \edge[draw, dotted];
 		[.$T_d(6)$
@@
 					[.$T_d(3)$ $T_d(2)$ $T_d(1)$ ]
-				$T_d(2)$ $T_d(1)$
-			]
-			[.$T_d(3)$ $T_d(2)$ $T_d(1)$ ]
 					$T_d(2)$ %$T_d(1)$
-		]
-		[.$T_d(4)$ 
-			[.$T_d(3)$ $T_d(2)$ $T_d(1)$ ]
-			$T_d(2)$ $T_d(1)$
 				]
@@
 			]
-\\};
+			[.$T_d(4)$ 
+				[.$T_d(3)$ $T_d(2)$ $T_d(1)$ ]
+				$T_d(2)$ %$T_d(1)$
+			]
+			%[.$T_d(3)$ $T_d(2)$ $T_d(1)$ ]
+		]
+		\edge[draw=none]; {}
+		%\edge[draw=none]; {}		
+	]
 \end{tikzpicture}
@@
 \begin{frame}
-\frametitle{Edge and leaf events}
-\begin{columns}[onlytextwidth]
-\begin{column}{0.7\textwidth}
+\frametitle{Edge events}
 \defEdgeEvent
+\begin{center}
+\begin{tikzpicture}[every tree node/.style={thick}]
+ \Tree 
+ 	[.\node[draw,circle]{u}; 
+		\edge[draw, densely dotted]; {}  
+		%\edge[draw, densely dotted]; {}
+		\edge[draw] node[draw=none, thick, anchor=west] {$i$};
+		[.\node[draw,circle]{v};
+		]
+	]
+	\node (bu) at (0.75,0.5) {};
+	\node (bv) at (1.1,-1.25) {};
 
+	\path node[topflat, xshift=3.5cm, above, yshift=-3cm]  {};
+	\node[circle] (bin1) at (3.5,-0.75) {};
+
+	\path node[topflat, xshift=4.5cm, above, yshift=-3cm]  {};
+	\node[circle] (bin2) at (4.5,-0.75) {};
+
+	\path node[topflat, xshift=5.5cm, above, yshift=-3cm]  {};
+	\node[circle] (bin3) at (5.5,-0.75) {};
+
+	\path node[topflat, xshift=6.5cm, above, yshift=-3cm]  {};
+	\node[circle] (bin4) at (6.5,-0.75) {};
+
+	\path node[topflat, xshift=7.5cm, above, yshift=-3cm]  {};
+	\node[circle] (bin5) at (7.5,-0.75) {};
+
+	\path node[topflat, xshift=8.5cm, above, yshift=-3cm]  {};
+	\node[circle] (bin6) at (8.5,-0.75) {};
+
+	\path (bu) edge[insert, bend left=20] node[draw=none, font=\tiny, sloped, anchor=south]{$i$-th location} (bin3);
+	\path (bu) edge[insert, dotted, bend left=30] (bin6);
+
+	\path (bv) edge[insert, bend right=30] node[draw=none, font=\tiny, sloped, anchor=south]{any location} (bin3);
+	\path (bv) edge[insert, dotted, bend right=40] (bin4);
+
+	\shade[ballstyle] (0.75,0.5) circle (\ballsize*\scalefac) {};
+	\shade[ballstyle] (1.1,-1.25) circle (\ballsize*\scalefac) {};
+\end{tikzpicture}
+\end{center}
+\end{frame}
+
+\begin{frame}
+\frametitle{Leaf events}
 \defLeafEvent
-\end{column}
-\begin{column}{0.3\textwidth}
-\begin{tikzpicture}
-\todo{illustrate}
+
+\bigskip
+\begin{center}
+\begin{tikzpicture}[every tree node/.style={thick}]
+\begin{scope}[xshift=-0.9cm, yshift=4cm]
+\Tree
+ 	[.\node[draw,circle]{u}; 
+		\edge[draw, densely dotted]; {}  
+		\edge[draw] {};
+		[.\node[draw,circle]{v};
+		]
+	]
+\end{scope}
+\begin{tikzpicture}[scale=\scalefac]
+\begin{scope}[yshift=1cm]
+	\bab{3 0 0 0 3 0}
+	\newball
+	\iA[30]{n14};
+	\iA[25]{n54};	
+\end{scope}
 \end{tikzpicture}
-\end{column}
-\end{columns}
+
+\end{tikzpicture}
+\end{center}
 
@@
 \begin{itemize}
-\item Node $T_d(1)$/$T_d(2)$ is labelled $(0,1)$/$(0,2)$.
-\item A node labelled $(0,i)$, $i>2$, has $i-1$ children labelled $(0,i-1),\dots,(0,1)$.
-\item A node with label $(h,i)$, $h>0$, has $d$ children labelled
-$(h,i-1),\dots,(h,1)$,$(h-1,d),\dots,(h-1, i)$.
-\item Thus, the node $T_d(dh+i)$ is labelled $(h,i)$.
+\item Node $T_d(1)$/$T_d(2)$ is labeled $(0,1)$/$(0,2)$.
+\item For $i>2$, a node $(0,i)$ has $i-1$ children $(0,i-1),\dots,(0,1)$.
+\item For $h>0$, a node $(h,i)$, has $d$ children
+$(h,i-1),\dots,(h,1)$, $(h-1,d),\dots,(h-1, i)$.
+\item Thus, the node $T_d(dh+i)$ is labeled $(h,i)$.
 \end{itemize}
@@
 \begin{tikzpicture}
-\matrix{
  \Tree 
-	[.$(1,2)$ 
-		[.$(1,1)$
-			[.$(0,3)$ $(0,2)$ $(0,1)$ ]
-			$(0,2)$ $(0,1)$
+	[.$T_d(6)$\\$(1,3)$ \edge[draw, dotted];
+		[.$T_d(5)$\\$(1,2)$
+			[.$T_d(4)$\\$(1,1)$
+				[.$T_d(3)$\\$(0,3)$ $T_d(2)$\\$(0,2)$ $T_d(1)$\\$(0,1)$ ]
+				$T_d(2)$\\$(0,2)$ %$T_d(1)$\\$(0,1)$
 			]
-		[.$(0,3)$ $(0,2)$ $(0,1)$ ]
-		$(0,2)$
+			[.$T_d(3)$\\$(0,3)$ $T_d(2)$\\$(0,2)$ $T_d(1)$\\$(0,1)$ ]
+			%$T_d(2)$\\$(0,2)$
 		]
-\\};
+	\edge[draw=none]; {}
+	]
 \end{tikzpicture}
@@
 
-\begin{frame}
-\frametitle{Proof of activation}
-We assign balls to the nodes of the AWT s.t.~th
-
+\newcommand\myInvariant{
 \begin{exampleblock}{Invariant}
@@
 \end{exampleblock}
+}
+
+\begin{frame}
+\frametitle{Proof of activation}
+\begin{block}{Lemma}
+The occurrence of a bad event (i.e., a bin contains more than $L+3$) implies the activation of an AWT of order $L$.
+\end{block}
+\bigskip
+\visible<2>{
+We assign balls to the nodes of the AWT s.t.:
+\myInvariant
+}
+\end{frame}
+
+\begin{frame}
+\frametitle{Proof of activation}
+
+\myInvariant
+
+Let a bin contain $L+4$ balls. We assign the topmost ball to the root of the AWT:
+\begin{columns}[onlytextwidth]
+\begin{column}{0.5\textwidth}
+\begin{center}
+\begin{tikzpicture}[scale=1, edge from parent/.style=
+     {draw, edge from parent path={(\tikzparentnode) -- (\tikzchildnode)}}]
+\Tree 
+	[.$(L,1)$ \edge[draw=black!30!red, thick];
+		[.$(L-1,3)$
+			[.$(L-1,2)$
+				[.$(L-1,1)$ ]
+				\edge[draw, densely dotted]; {}
+			]
+			\edge[draw, densely dotted]; {}
+		]
+		\edge[draw=black!30!red, thick];
+		[.$(L-1,2)$ \edge[draw, densely dotted]; {} \edge[draw, densely dotted]; {} ]
+	]
+\end{tikzpicture}
+\end{center}
+\end{column}
+\begin{column}{0.5\textwidth}
+\begin{tikzpicture}[scale=\scalefac]
+	\bab{2 4 4 2 3 4}
+	\groupSep{3};
+	\setNode{1}{7};
+	\ball{1}{7};
+	\specialBall{2}{5};
+	\path (n17) edge[insert, bend left=25] (n25);
+	\path (n17) edge[insert, bend left=25] (n65);
+\end{tikzpicture}
+\end{column}
+\end{columns}
+\end{frame}
+
+\begin{frame}
+\frametitle{Proof of activation}
+Given a node $v$ with label $(h,i)$ that references a ball $b_v$ (which was put into some bin of group $i$ that contained at least $h+3$ other balls at the time of insertion of ball $b_v$), \alert{we conclude} that:
+\begin{enumerate}
+\only<1>{
+\item For $1\leq j < i$, the $j$-th location of the ball $b_v$ contains at least $h+4$ balls, at the insertion time of the ball $b_v$.
+If it contained less balls than the $i$-th location, the Always-go-left scheme would have put the ball $b_v$ into the bin pointed to by location $j$.
+We assign the topmost ball in location $j$ to the child $u$ of $v$ with label $(h,j)$.
+}
+\only<2>{
+\item For $i < j \leq d$, the $j$-th location contains at least $h+3$ balls.
+If it contained less, the Always-go-left scheme would have put the ball $b_v$ into the bin location $j$ points to.
+We assign the topmost ball in location $j$ to the child node $u$ of $v$ labeled with $(h-1, j)$.
+\qed
+
+}
+\end{enumerate}
 
@@
 \end{lemma}
-
-\bigskip
+\medskip
 \begin{itemize}
@@
 \defLeafEvent
-At any time step $t$, at most $n/3$ bins contain 3 or more balls. \alert{But how are they distributed among the $d$ groups?}
+At any time step $t$, at most $n/3$ bins contain 3 or more balls. But \alert{how are they distributed} among the $d$ groups?
 
 \medskip
-$b_i$\dots number of bins in group $i$ containing 3 or more balls.
-$\beta_i = b_i d/n$\dots fraction of bins $b_i$ in group $i$. 
+$b_i$\dots number of bins in group $i$ containing 3 or more balls
+$\beta_i = b_i d/n$\dots fraction of bins $b_i$ in group $i$
 }
@@
 \begin{align*}
-P_\mathrm{a} \leq n^{-\alpha} \quad  \qed
+P_\mathrm{a} \leq n^{-\alpha} \xqed{5cm}
 \end{align*}
@@
 \begin{frame}
-\frametitle{Recap.}
+\frametitle{The big picture}
 \theoremVocking
 \assumptionOne
-\bigskip 
+\medskip 
 Next steps:
 \begin{itemize}
-\item Proof \alert{without} the simplifying assumption (V\"ocking, 2003).
-\item Proof for the \alert{heavily loaded} case (Berenbrink et al.,~2006).
+\item Prove \alert{without} the simplifying assumption (V\"ocking, 2003).
+\item Prove for the \alert{heavily loaded} case (Berenbrink et al.,~2006).
 \end{itemize}
-
 \end{frame}
@@
 \frametitle{Conclusion}
+\begin{enumerate}
+\item The balls-into-bins model \alert{adopts naturally} to various problems (hashing, load balancing,\dots).
+\item Applying the two-choice paradigm reduces the upper bound for the maximum load by an \alert{exponential factor} (compared to a single choice scheme).
+\item \alert{Many variations} possible: Non-uniform balls/bins; weighted balls; parallel allocation of several balls;\dots
+\item \alert{Witness trees} are a suitable proof technique for this kind of problems.
+\end{enumerate}
 \end{frame}
 
-\begin{frame}[allowframebreaks]
+\begin{frame}[shrink]
 \frametitle<presentation>{References}    

### 2014-03-30 article/article.tex
improve proof 3

@@
 \label{sec:analysis:nondistinctBalls}
-So far, we have assumed that all balls represented by nodes in a witness tree are stochastically independent (Assumption~\ref{assumption:independence}).
-Obviously, to get a valid upper bound on the probability for the activation of a witness tree, the possibility of non-distinct balls has to be accounted for.
-In other words, the edge and leaf events of the witness tree are not independent any more.
-Will start with a witness tree referencing non-distinct balls and prune it such that dependencies between edge and leaf events are removed.
-Note that pruning will reduce the size of the witness tree and thus increase the probability of its activation.
-To compensate for this loss in size, we start with a larger witness tree, which we call a full witness tree and will show that, speaking informally, the pruning of many events is unlikely.
-Note that the technique presented in this section is applicable to both symmetric and asymmetric witness trees.
+So far, we have assumed that all balls represented by nodes of a witness tree are stochastically independent (Assumption~\ref{assumption:independence}).
+To get a valid upper bound on the probability for the activation of a witness tree, the possibility of non-distinct balls has to be accounted for.
+In this section, we do not require the edge and leaf events of the witness tree to be independent any more.
+We start with a witness tree referencing non-distinct balls and prune it such that all dependencies between edge and leaf events are removed.
+Note that pruning reduces the size of the witness tree and therefore increases the probability of its activation.
+To compensate for this loss in size, we start with a larger witness tree, which we call a full witness tree (FWT) and show that, speaking informally, the pruning of many events is unlikely.
+Note that the technique presented in this section is applicable to both SWTs and AWTs.
 
@@
 Let $\gamma \geq 2$ be a suitable constant.
-The root node of a full witness tree (FWT) has $\gamma$ children, denoted by $v_{c_i}$ with $1\leq i \leq \gamma$.
-The nodes $v_{c_i}$ have exactly one child each and the balls assigned to them are pairwise distinct.
-The root thus has $\gamma$ grandchildren, each of which is the root of a witness tree of order $L$.
-The time constraints and activation of a FWT is defined analogously to the previous variants.
+The root node of an FWT has $\gamma$ children, denoted by $v_c$, with $1\leq c \leq \gamma$.
+The nodes $v_c$ have exactly one child each and the balls assigned to them are pairwise distinct.
+The root thus has $\gamma$ grandchildren, each of which is the root of an SWT or AWT of order $L$ (depending on which allocation algorithm we want to analyze).
+The time constraints and activation of an FWT are defined analogously to the previous variants.
 
@@
 \label{sec:analysis:constructionFullWT}
-Assume a bin $x$ contains at least $L+3+\gamma$ balls at time $t$. We construct an FWT as follows.
+Assume a bin $x$ contains at least $L+4+\gamma$ balls at time $t$. We construct an FWT as follows:
 \begin{compactitem}
@@
 \item Each child of the root is assigned one of the topmost $\gamma$ balls in bin $x$, which are necessarily distinct.
-\item For the assignment of nodes to the grandchildren of the root, consider a ball $b_c$ that is assigned to a child node $v_c$.
-Let the child node of $v_c$ be denoted by $v_g$.
-Since the ball $b$ was placed into bin $x$, at least one of its locations points to this bin $x$, which we call location $i$.
+\item For the assignment of nodes to the grandchildren of the root, consider a ball $b_c$ that is assigned to a child node $v_c$ and let the child node of $v_c$ be denoted by $v_g$.
+Since the ball $b_c$ was placed into bin $x$, at least one of its locations points to this bin $x$, which we call location $i$.
 Let $j = i+1 mod d$ be the ball's $j$-th location.
@@
 We assign the topmost ball in bin $y$ to node $v_g$.
+\item Below the grandchild node $v_g$, the assignment of balls proceeds according to the rules for the SWT (Section~\ref{sec:analysis:definitionSymWT}) or AWT (Section~\ref{sec:analysis:definitionAsymWT}), respectively.
 \end{compactitem}
-Notice that the $\gamma$ additional balls in bin $x$ are assigned to the child nodes of the root only.
+Note that the $\gamma$ additional balls in bin $x$ are assigned to the child nodes of the root only.
 
@@
 To remove all dependencies in the witness tree, we have to find nodes that represent the same ball and prune the tree such that each ball is referenced only once.
-Thus, traverse the witness tree in breadth-first order.
+To achieve this, we traverse the witness tree in breadth-first order.
 When we reach a node $v$ that represents a ball $b$ already encountered before, let $e = (v,u)$ be the edge connecting $v$ and its parent node $u$.
-We call $e$ a cut-off edge and use it as a witness that one of the $d$ locations of the ball $b_u$ is the same as one of the locations of some other ball $b_{u'}$, represented by a node $u'$, which was already traversed.
-Furthermore, we remove the whole tree rooted at node $v$ from our witness tree.
-We continue this pruning until either the breath-first traversal terminates or until we have obtained $\gamma$ cut-off edges.
-Note that more than $\gamma$ cut-off edges cannot exist.
-When finished, the pruned witness tree does not contain any ball twice and thus all its events are independent.
+We call $e$ a cut-off edge and use it as a witness that one of the $d$ locations of the ball $b_u$ is the same as one of the locations of some other ball $b_{u'}$, represented by a node $u'$, which was already encountered during the traversal.
+Furthermore, we remove the whole tree rooted at node $v$ from our witness tree and we continue this pruning until either the breath-first traversal terminates or we have obtained $\gamma$ cut-off edges.
+Note that more than $\gamma$ cut-off edges cannot exist, because \todo{why??}.
+When finished, the pruned witness tree (PWT) does not reference any ball twice and thus all its events are independent.
 
@@
 \begin{lemma}\label{lemma:fwt:activation}
-For some suitable $\gamma \geq 2$ and large enough $L$, the activation of a pruned witness tree occurs with probability at most $n^{-\gamma +1 +o(1)}$.
+For some suitable $\gamma \geq 2$ and large enough $L$, the activation of a PWT occurs with probability at most $n^{-\gamma +1 +o(1)}$.
 \end{lemma}
-If there are less than $\gamma$ cut-off edges, in at least one of the $\gamma$ symmetric or asymmetric witness trees contained in the full witness tree the referenced balls are distinct.
-Let this symmetric or asymmetric witness tree be denoted by $T$.
-Then, the probability for the activation of the FWT is upper-bounded by the probability for the activation of $T$, implying a probability for activation of this witness tree of $p = n^{- \alpha}$, as shown in Sections~\ref{sec:analysis:probabilitySymWT} and~\ref{sec:analysis:probabilityAsymWT}, respectively.
+If there are less than $\gamma$ cut-off edges, in at least one of the $\gamma$ SWTs or AWTs contained in the FWT, the referenced balls are distinct.
+Let this SWT or AWT be denoted by $T$.
+Then, the probability $P_\mathrm{f}$ for the activation of the FWT is upper-bounded by the probability for the activation of $T$ and thus $P_\mathrm{f} \leq n^{- \alpha}$, for some $\alpha >0$ and large enough $L$, as shown in Sections~\ref{sec:analysis:probabilitySymWT} and~\ref{sec:analysis:probabilityAsymWT}, respectively.
 
-Thus, it remains to find an upper-bound for the activation of an FWT with $\gamma$ cut-off edges.
-Let $K$ be the number of references to balls by the FWT\footnote{Since no ball was assigned to the root node, the FWT consists of $K+1$ nodes} and assume that $K  \leq 2 \gamma\left(\alpha+1\right) \log_2 n$, for some arbitrary $\alpha > 0$.
-Then, there are at most $w_p=K^\gamma$ possibilities to transform an FWT with $K+1$ nodes to a pruned witness tree with $\gamma$ cut-off edges.
+It remains to find an upper-bound for the activation of an FWT with $\gamma$ cut-off edges.
+Let $K$ be the number of references to balls by the FWT and assume that $K  \leq 2 \gamma\left(\alpha+1\right) \log_2 n$, for some arbitrary $\alpha > 0$.
+Then, there are at most $w_p=K^\gamma$ possibilities to transform an FWT with $K+1$ nodes\footnote{Since no ball was assigned to the root node, the FWT consists of $K+1$ nodes.} to a PWT with $\gamma$ cut-off edges.
 Fix the shape of the tree to be one of those.
 
-To bound the probability for the activation of all edge events, we need to specify which event is represented by an edge $e_i = (v_r, v_i)$, where $v_i$ with $1 \leq i \leq \gamma$ is a child of the root node $v_r$.
+To bound the probability for the activation of all edge events, we need to specify which event is represented by an edge $e_c = (v_r, v_c)$, where $v_c$ with $1 \leq c \leq \gamma$ is a child of the root node $v_r$.
 Assume that the root node $v_r$ represents an arbitrary bin.
-Then the edge $e_i$ represents the event that one of the locations of the ball represented by node $v_i$ points to this bin.
-This implies that the probability of activation for an each $e_i$ is $p_{e_i} = 1/n \leq d/n$.
+Then the edge $e_c$ represents the event that one of the locations of the ball represented by node $v_c$ points to this bin.
+This implies that the probability of activation for an edge $e_c$ is $p_\mathrm{e}' = 1/n \leq d/n$.
 
-Let $k$ be the number of balls referenced by the pruned witness tree and $l$ be the number of its leaf nodes.
-Then, by the same arguments as in Sections~\ref{sec:analysis:probabilitySymWT} and~\ref{sec:analysis:probabilityAsymWT}, we can conclude that there are at most $w=n^k$ assignments of balls to the nodes of the witness tree.
+Let $k$ be the number of balls referenced by the PWT and $l$ be the number of its leaf nodes.
+Then, by the same arguments as in Sections~\ref{sec:analysis:probabilitySymWT} and~\ref{sec:analysis:probabilityAsymWT}, we can conclude that there are at most $w\leq n^k$ assignments of balls to the nodes of the witness tree.
 Furthermore, the probability for the activation of all edge events is bounded from above by
 \begin{align*}
-p_e = \left(\frac{d}{n}\right)^{k}\quad ,
+p_\mathrm{e} \leq  \left(\frac{d}{n}\right)^{k}\quad ,
 \end{align*}
@@
 \begin{align*}
-p_l \leq 3^{-d l}\quad .
+p_\mathrm{l} \leq 3^{-d l}\quad .
 \end{align*}
 
-The number of edges $k$ and the number of leaves $l$ in a pruned witness tree may be small compared to a symmetric or asymmetric witness tree, but we will use the cut-off edges to compensate.
-By definition (Section~\ref{sec:analysis:pruningFullWT}), a cut-off edge $e=(v, u)$ represents the event that the ball $b_u$ represented by node $u$ shares a common location $i$ with another ball $b_{u'}$, represented by node $u'$.
-Since $u'$ has already been visited by the time the breadth-first traversal reaches node $v$, it is part of the pruned witness tree and was assigned one of $k\leq K$ balls.
+The number of edges $k$ and the number of leaves $l$ in a PWT may be small compared to an SWT or AWT, but we will use the cut-off edges to compensate.
+By definition (Section~\ref{sec:analysis:pruningFullWT}), a cut-off edge $e=(v, u)$ represents the event that the ball $b_u$ referenced by node $u$ shares a common location $i$ with another ball $b_{u'}$, represented by node $u'$.
+Since $u'$ has already been visited by the time the breadth-first traversal reaches node $v$, it is part of the PWT and was assigned one of $k\leq K$ balls.
 The probability that location $i$ of ball $b_u$ points to one of the $d$ locations of ball $b_{u'}$ is at most $d/n$.
-The probability that there are $\gamma$ cut-off edges is thus at most
+The probability that there are $\gamma$ cut-off edges thus is at most
 \begin{align*}
-p_c &= \left(k \frac{d}{n}\right)^\gamma 
+p_\mathrm{c} &\leq \left(k \frac{d}{n}\right)^\gamma 
     \leq \left(K \frac{d}{n}\right)^\gamma \quad .
 \end{align*}
-
-Putting it all together, we get 
+Note that after the pruning step all the edge, leaf and cut-off events are independent since we have removed redudent balls from the witness tree. Putting it all together, we have
 \begin{align*}
-p_{pwt} &= w_p w  p_e  p_l  p_c 
-		= K\gamma n^{k+1} \left(\frac{d}{n}\right)^k3^{-d l}  \left(K \frac{d}{n}\right)^\gamma
-		= n  d^k 3^{-d l}  \left(K^2 \frac{d}{n} \right)^\gamma \quad .
+P_\mathrm{p} &= w_\mathrm{p} w  p_\mathrm{e}  p_\mathrm{l}  p_\mathrm{c} 
+		\leq K\gamma n^{k+1} \left(\frac{d}{n}\right)^k3^{-d l}  \left(K \frac{d}{n}\right)^\gamma
+		= n  d^k 3^{-d l}  \left(K^2 \frac{d}{n} \right)^\gamma \quad ,
 \end{align*}
-
-Using $K \leq 2l$, $d^2 \leq 3^d$ and $K \leq 2\gamma  \left( \alpha +1 \right)  \log_2 n$ this becomes
+for the probability of activation of a PWT. Using $K \leq 2l$, $d^2 \leq 3^d$ and $K \leq 2\gamma  \left( \alpha +1 \right)  \log_2 n$, we get
 \begin{align*}
-p_{pwt} &\leq n \left(\frac{\left(2\gamma  \left(\alpha +1 \right)  \log_2 n\right)^2 d}{n}\right)^\gamma 
+P_\mathrm{p} &\leq n \left(\frac{\left(2\gamma  \left(\alpha +1 \right)  \log_2 n\right)^2 d}{n}\right)^\gamma 
         = n^{-\gamma +1}  n ^{\log_n\left( \left(2\gamma \left(\alpha+1\right) \log_2n \right)^{2\gamma}  d^\gamma \right)} \\
-        &= n^{-\gamma+1+o\left(1\right)} \quad .
+        &= n^{-\gamma+1+o\left(1\right)} \quad ,
 \end{align*}
-\todo{Where are our $L$ and $\alpha$ here?}
+which concludes the proof of Theorems~\ref{theorem:greedy} and~\ref{theorem:aglm}.
 \qed

### 2014-03-31 presentation/presentation.tex
group references by importance

@@
 \documentclass[serif,professionalfonts]{beamer}
+\usetheme{Forcellini}
 
-\usetheme{Forcellini}
+\usepackage[backend=bibtex]{biblatex} 
+\addbibresource{../sources.bib}
 
@@
 \title{The Power of Two Random Choices}
-\subtitle{And the Surprising Influence of Asymmetry}
+\subtitle{\dots and the Surprising Influence of Asymmetry}
 \author[M. Kalany]{Martin Kalany}
@@
 \begin{theorem}[Raab and Steger, 1998]
-When using the single choice strategy to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
+When using the \alert{single choice} strategy to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
 \begin{align*}
@@
 \end{theorem}
-
 \end{frame}
@@
 \bigskip
-We can distinguish placement algorithms by the way the $d$ \alert{bins are sampled}:
+We distinguish placement algorithms by the way the $d$ \alert{bins are sampled}:
 \begin{enumerate}
@@
 \begin{align*}
-\Phi_2 = 1.61 \quad , \quad \forall i\; \forall j \in \mathbb{N}: i < j \implies \Phi_i < \Phi_j \quad , \quad \lim_{d\rightarrow \infty} \Phi_d = 2 \quad .
+\Phi_2 = 1.61 \quad , \quad \forall i\; \forall j \in \mathbb{N}: i < j \implies \Phi_i < \Phi_j \quad , \quad \lim_{d\rightarrow \infty} \Phi_d = 2 
 \end{align*}
@@
 \begin{theorem}[V\"ocking, 2003]
-For $m=n$, using an arbitrary allocation scheme that chooses $d$ bins for each ball at random (i.e., class 3) the maximum load is, w.h.p,
+For $m=n$, using an arbitrary allocation scheme that chooses $d$ bins for each ball at random (i.e., class 3), the maximum load is, w.h.p,
 \begin{align*}
@@
 \begin{block}{Lemma}
-The occurrence of a bad event (i.e., a bin contains more than $L+3$) implies the activation of an AWT of order $L$.
+The occurrence of a bad event (i.e., a bin contains more than $L+3$ balls) implies the activation of an AWT of order $L$.
 \end{block}
@@
 \end{enumerate}
-
 \end{frame}
@@
 \begin{align*}
-p_\mathrm{e} = \left(\frac{d}{n}\right)^{k-1} \quad .
+p_\mathrm{e} \leq \left(\frac{d}{n}\right)^{k-1} \quad .
 \end{align*}
@@
 \begin{align*}
-P_\mathrm{a} = \omega p_e p_l \leq  n^k \left(\frac{d}{n}\right)^{k-1} 3^{-dl} \quad .
+P_\mathrm{a} \leq \omega p_e p_l \leq  n^k \left(\frac{d}{n}\right)^{k-1} 3^{-dl} \quad .
 \end{align*}
@@
 \item The balls-into-bins model \alert{adopts naturally} to various problems (hashing, load balancing,\dots).
-\item Applying the two-choice paradigm reduces the upper bound for the maximum load by an \alert{exponential factor} (compared to a single choice scheme).
+\item Applying the two-choice paradigm reduces the upper bound for the maximum load by an \alert{exponential factor} (compared to the single choice scheme).
 \item \alert{Many variations} possible: Non-uniform balls/bins; weighted balls; parallel allocation of several balls;\dots
@@
 
-\begin{frame}[shrink]
-\frametitle<presentation>{References}    
-\bibliographystyle{acm}
+\begin{frame}
+\frametitle{References}
 \nocite{VOC03}
@@
 \nocite{capocelli1990generalized}
-\setbeamertemplate{bibliography item}[text]
-\bibliography{../sources}
+\setbeamertemplate{bibliography item}[triangle]
+\printbibliography[keyword=main]
+\bigskip
+
+\printbibliography[notkeyword=main]
 \end{frame}

### 2014-03-31 presentation/presentation.tex
rebaseme
use symlinks for source files
some more polishing after tips from francesco

@@
-\documentclass[serif,professionalfonts]{beamer}
+\documentclass[professionalfonts]{beamer}
+\usepackage{pgfpages}
+
 \usetheme{Forcellini}
@@
 \usepackage[backend=bibtex]{biblatex} 
-\addbibresource{../sources.bib}
+\addbibresource{sources.bib}
 
@@
 % Hermann Zapf Optima clone (wonderful sanserif font) -- URW Classico
-\renewcommand{\sfdefault}{uop}
-\renewcommand{\rmdefault}{ppl}
+%\renewcommand{\sfdefault}{uop}
+%\renewcommand{\rmdefault}{ppl}
+\usepackage{eulervm}
 
@@
 
-\setbeamertemplate{caption}
-{
-  \raggedright%
-  \insertcaption\par%
-}
 \setbeamercolor{green}{fg=green}
@@
 \newcommand\todo[1]{\textcolor{red}{(TODO: #1)}}
-\newcommand\load{L_{\mathrm{max}}}
-\newcommand\loadS{\load^{\mathrm{S}}}
-\newcommand\loadG{\load^{\mathrm{G}}}
-\newcommand\loadAgl{\load^{\mathrm{A}}}
-\newtheorem{claim}{Claim}
-\newtheorem{assumption}{Assumption}
-\newtheorem{invariant}{Invariant}
+
+\DeclareMathOperator{\load}{L_{\max}} % ensuremath
+\DeclareMathOperator{\loadS}{L_{\max}^{S}} 
+\DeclareMathOperator{\loadG}{L_{\max}^{G}} 
+\DeclareMathOperator{\loadAgl}{L_{\max}^{Agl}} 
+\DeclareMathOperator{\deltaS}{S}
+\DeclareMathOperator{\deltaG}{G}
+\DeclareMathOperator{\deltaAgl}{Agl}
 
@@
 \title{The Power of Two Random Choices}
-\subtitle{\dots and the Surprising Influence of Asymmetry}
+\subtitle{\dots and the surprising influence of asymmetry}
 \author[M. Kalany]{Martin Kalany}
@@
 \framesubtitle{...and the goal}
-\begin{exampleblock}{Balls-into-bins games}
-Suppose we have $n$ initially empty bins and need to distribute $m \geq n$ balls among them, \alert{as evenly as possible}. 
-\end{exampleblock}
+\begin{alertblock}{Balls-into-bins games}
+Suppose we have $n$ initially empty bins and need to distribute $m \geq n$ balls
+among them, \alert{as evenly as possible}. 
+\end{alertblock}
 \pause
 \begin{exampleblock}{Definition}
-We are interested in the \alert{maximum load $\load$}, i.e, how many balls are in the fullest bin?
+We are interested in the \alert{maximum load $\load(m,n)$}, i.e, how many balls are in the fullest bin?
 \end{exampleblock}
@@
 \begin{theorem}[Raab and Steger, 1998]
-When using the \alert{single choice} strategy to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
+When using the \alert{single choice} strategy to distribute $m$ balls into $n$ bins, the maximum load will be, w.h.p.,
 \begin{align*}
-\loadS \leq 
+\loadS = 
 	\begin{cases}
     \Theta\left(\frac{\ln n}{\ln\ln n}\right)           & m = n \\
-    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
+    \frac{m}{n}  \, + \, \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
     \end{cases} \quad .
@@
 \bigskip
-We distinguish placement algorithms by the way the $d$ \alert{bins are sampled}:
+We distinguish placement algorithms by the way the \alert{$d$ bins are sampled}:
 \begin{enumerate}
@@
 
-\begin{frame}
+\begin{frame}[t]
 \frametitle{The Greedy algorithm}
@@
 \begin{theorem}[Azar, Broder, Karlin and Upfal, 1999]
-When using the \alert{Greedy} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
+When using the \alert{Greedy} algorithm to distribute $m$ balls into $n$ bins, the maximum load will be, w.h.p.,
 \begin{align*}
@@
 \begin{align*}
-\loadS \leq 
+\delta^{\deltaS} &= 
 	\begin{cases}
     \Theta\left(\frac{\ln n}{\ln\ln n}\right)           & m = n \\
-    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
-    \end{cases}
-\end{align*}
-\begin{align*}
-\loadG \leq \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) 
+    \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
+    \end{cases}\\[5mm]
+\delta^{\deltaG} &\leq\frac{\ln \ln n}{\ln d} + O(1) 
 \end{align*}
@@
 \begin{alertblock}{The two-choice paradigm}
-Even for $d=2$, Greedy achieves an \alert{exponential decrease} of the additive gap $\delta$, which does \alert{not depend on} the number of balls \alert{$m$}.
+  Even for $d=2$, Greedy's additive gap $\delta$
+  \begin{itemize}
+  \item decreases \alert{exponentially} compared to the single choice algorithm,
+  \item and does \alert{not depend on} the number of balls \alert{$m$}.
+  \end{itemize}
 \end{alertblock}
@@
 \begin{description}
-	\item[Hashing:] Assuming a perfect hash function, $n$ elements (balls) are mapped to $m$ table entries (bins).
-Using two hash functions: look-up time $\Theta(\ln\ln n)$, w.h.p.
+	\item[Hashing:] Assuming a perfect hash function, $m$ elements (balls) are mapped to $n$ table entries (bins).
+Using two hash functions: look-up time $O(\ln\ln n)$, w.h.p.
 \alert{Easy to parallelize} and  \alert{no re-hashing}!
 	\item[Online load balancing:] Assign tasks (balls) to servers/processors (bins). The two-choice paradigm helps to \alert{reduce communication overhead}, %(of e.g., a central dispatcher)
-	while the maximum load will be $\Theta(\ln\ln n)$, w.h.p.
-	\item[Emulating PRAMs on DMMs:] Distribute processors and memory cells of the PRAM on the DMM s.t.~the overhead of the emulation is minimized.
-	\item[Low congestion circuit routing:] \alert{Valiant's paradigm} %(a permuation routing problem can be split into two sub-problems: route to random intermediate destination and then on to the actual destination)
+	while the maximum load is $O(\ln\ln n)$, w.h.p.
+	\item[Emulating PRAMs on DMMs:] Distribute processors and memory cells of the PRAM on the DMM, s.t.~the overhead of the emulation is minimized.
+	\item[Low congestion circuit routing:] \alert{Valiant's paradigm} (first
+          route to some random intermediate
+          destination) %(a permuation routing problem can be split into two sub-problems: route to random intermediate destination and then on to the actual destination)
           is analogous to the single-choice technique.
@@
 \begin{enumerate}
-\item Partition the $n$ bins into $d$ groups.
-\item Assume a fixed ordering of the bins $b_i$, $1\leq i \leq n$.
-\item Choose one bin from each group uniformly and independently.
-\item Put the ball into the least full bin.
-\item In case a tie occurs, put the ball into the leftmost bin (i.e., the bin with smallest index) among those with same minimal load.
+\item \alert{Partition} the $n$ bins into $d$ groups.
+\item Assume a \alert{fixed ordering} of the bins $b_i$, $1\leq i \leq n$.
+\item Choose one bin \alert{from each group} uniformly and independently.
+\item Put the ball into the \alert{least full} bin.
+\item In case a \alert{tie} occurs, put the ball into the \alert{leftmost} bin (i.e., the bin with smallest index) among those with same minimal load.
 \end{enumerate}
@@
 \begin{frame}
-\frametitle{$d$-step Fibonacci numbers and the golden ratio}
-\framesubtitle{(Capocelli, Cull, 1990)}
+\frametitle{The $d$-step Fibonacci numbers and the golden ratio}
+\framesubtitle{(Capocelli and Cull, 1990)}
 
-$d$-step Fibonacci numbers:
+\begin{itemize}
+\item The \alert{$d$-step} Fibonacci numbers:
   \begin{align*}
@@
   \end{align*}
-Generalized golden ratio $\Phi_d$:
+\item Generalized \alert{golden ratio} $\Phi_d$:
   \begin{align*}
@@
   \begin{align*}
-\Phi_2 = 1.61 \quad , \quad \forall i\; \forall j \in \mathbb{N}: i < j \implies \Phi_i < \Phi_j \quad , \quad \lim_{d\rightarrow \infty} \Phi_d = 2 
+    \Phi_2 \approx 1.618 \  , \ \forall i\; \forall j \in \mathbb{N}: i < j
+    \implies \Phi_i < \Phi_j \ , \ \lim_{d\rightarrow \infty} \Phi_d = 2
   \end{align*}
   \begin{align*}
-F_d(k) \geq \Phi_d^{k-2}
+    \forall d\; \forall k\geq	2: \quad F_d(k) \geq \Phi_d^{k-2}
   \end{align*}
+\end{itemize}
 \end{frame}
@@
 \begin{theorem}[V\"ocking, 2003]
-When using the \alert{Always-go-left} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
+When using the \alert{Always-go-left} algorithm to distribute $m$ balls into $n$ bins, the maximum load will be, w.h.p.,
 \begin{align*}
@@
 
-\begin{frame}[shrink]
+\begin{frame}
 \frametitle{The Always-go-left algorithm}
@@
 \begin{align*}
-\loadG \leq \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) 
+\delta^{\deltaG} \leq \frac{\ln \ln n}{\ln d} + O(1) 
 \end{align*}
@@
 \begin{align*}
-\loadAgl \leq \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
+\delta^{\deltaAgl} \leq \frac{\ln \ln n}{d \ln \Phi_d} + O(1)
 \end{align*}
@@
 \begin{theorem}[V\"ocking, 2003]
-For $m=n$, using an arbitrary allocation scheme that chooses $d$ bins for each ball at random (i.e., class 3), the maximum load is, w.h.p,
+For $m=n$, using an arbitrary allocation scheme that chooses $d$ bins for each ball at random, the maximum load will be, w.h.p,
 \begin{align*}
@@
 
-\begin{frame}
-\frametitle{BiB Model}
+\begin{frame}{Balls-into-bins model: formal definition}
 We will assume a sequential, infinite, dynamic process.
@@
 \item An oblivious adversary provides a sequence $\sigma = \sigma_1 \sigma_2 \dots$, where $\sigma_t = (d_t, a_t)$.
-\item $d_t$/$a_t$: ball to delete/add at time step t. \\
+\item $d_t$ (and $a_t$): ball to delete (and add) at time step t. \\
 \item $\sigma$ is fixed in advance.
@@
 \end{frame}
+\end{comment}
 
@@
 Consider any sequence of insertions and deletions s.t.~at most $n$ balls are in the bins at any time.
-If the balls are placed into bins by the Always-go-left algorithm, the maximum load at any time step $t$ is, w.h.p.,
+If the balls are placed into the bins by the Always-go-left algorithm, the maximum load at any time step $t$ is, w.h.p.,
 \begin{align*}
-\loadAgl \leq \frac{\ln\ln n}{d  \ln \Phi_d} + O(1) \quad .
+\loadAgl \leq \frac{\ln \ln n}{d \ln \Phi_d} + O(1) \quad .
 \end{align*}
@@
 
-\pause
 \bigskip
@@
 \begin{exampleblock}{Edge event}
-Edge $e =(u,v)$, s.t.~$v$ is $i$-th child of $u$: $e$ represents the event that the $i$-th location of ball $b_u$ points to the same bin as one of the locations of ball $b_v$.
+Edge $e =(u,v)$, s.t.~$v$ is $i$-th child of $u$: $e$ represents the event that the $i$-th location of ball $b_u$ points to the \alert{same bin} as one of the locations of ball $b_v$.
 \end{exampleblock}
@@
 \begin{exampleblock}{Leaf event}
-A leaf node $v$ represents the event that each of the $d$ locations of the ball $b_v$ points to a bin that contains at least three additional balls, at the time of the insertion of the ball $b_v$.
+A leaf node $v$ represents the event that each of the $d$ locations of the ball $b_v$ points to a bin that contains \alert{at least three additional} balls, at the time of the insertion of the ball $b_v$.
 \end{exampleblock}
@@
 		\edge[draw, densely dotted]; {}  
-		%\edge[draw, densely dotted]; {}
 		\edge[draw] node[draw=none, thick, anchor=west] {$i$};
@@
 
-\begin{frame}[shrink]
+\begin{frame}
 \frametitle{Asymmetric witness trees (AWT)} 
@@
 \begin{block}{Lemma}
-The occurrence of a bad event (i.e., a bin contains more than $L+3$ balls) implies the activation of an AWT of order $L$.
+The occurrence of a \alert{bad event} (i.e., a bin contains more than $L+3$ balls) implies the \alert{activation} of an AWT of order $L$.
 \end{block}
@@
 Given a node $v$ with label $(h,i)$ that references a ball $b_v$ (which was put into some bin of group $i$ that contained at least $h+3$ other balls at the time of insertion of ball $b_v$), \alert{we conclude} that:
-\begin{enumerate}
-\only<1>{
+\begin{itemize}
 \item For $1\leq j < i$, the $j$-th location of the ball $b_v$ contains at least $h+4$ balls, at the insertion time of the ball $b_v$.
@@
 We assign the topmost ball in location $j$ to the child $u$ of $v$ with label $(h,j)$.
-}
-\only<2>{
-\item For $i < j \leq d$, the $j$-th location contains at least $h+3$ balls.
-If it contained less, the Always-go-left scheme would have put the ball $b_v$ into the bin location $j$ points to.
+\item For $i < j \leq d$, the $j$-th location contains at least $h+3$ balls (by similar arguments as above).
 We assign the topmost ball in location $j$ to the child node $u$ of $v$ labeled with $(h-1, j)$.
 \qed
-
-}
-\end{enumerate}
+\end{itemize}
 \end{frame}
@@
 \begin{itemize}
-\item $k$\dots number of nodes
-\item $l$\dots number of leaves
-\item $\omega$\dots number of ways to assign balls to the nodes of an AWT
-\item $p_\mathrm{e}$\dots probability that all edge events occur
-\item $p_\mathrm{l}$\dots probability that all edge events occur
-\item $P_\mathrm{a}$\dots probability of activation of an AWT of order $L$
+\item $k$: number of nodes
+\item $l$: number of leaves
+\item $\omega$: number of ways to assign balls to the nodes of an AWT
+\item $p_\mathrm{e}$: probability that all edge events occur
+\item $p_\mathrm{l}$: probability that all edge events occur
+\item $P_\mathrm{a}$: probability of activation of an AWT of order $L$
 \end{itemize}

### 2014-04-09 article/article.tex
use symlinks for source files
some minor fixes to finalize the paper

@@
 
-\title{The Power of Two Random Choices\\ 
-\large A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Tr\"aff and Francesco Versaci, ``Parallel Algorithms'' lecture, 2012 winter term at TU Wien.}}
+\title{The Power of Two Random Choices and the Surprising Influence of Asymmetry\\ 
+\normalsize A follow up seminar to the Parallel Algorithms lecture\footnote{Jesper Larsson Tr\"aff and Francesco Versaci, ``Parallel Algorithms'' lecture, 2012 winter term at TU Wien.}}
 \author{Martin Kalany, 0825673}
@@
 \begin{abstract}
+\noindent
 Balls-into-bins is a classical allocation problem, where $m$ balls have to be distributed among $n$ bins as evenly as possible.
-A well studied approach is to select a bin uniformly and independently at random for each ball.
-Another strategy is to select $d\geq 2$ bins for each ball and put the ball into the bin with the least load.
+A well studied approach is to select a bin uniformly and independently at random for each ball; another strategy is to select $d\geq 2$ bins for each ball and put the ball into the bin with the least load.
 Somewhat surprisingly, the maximum load achieved by such a multiple-choice scheme is significantly smaller, even for $d=2$.
@@
 We are interested in the number of balls in the fullest bin, also called the \emph{maximum load}.
-Equivalently, one can look at the \emph{additive gap} of the allocation, i.e., the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
+Equivalently, we can look at the \emph{additive gap} of the allocation, i.e., the difference of the number of balls in the fullest bin and the average amount of balls in the bins.
 These balls-into-bins games (BiB) have been studied extensively (see, e.g.,~\cite{JK77}).
@@
 While most of this survey and the analyses focuses on the lightly loaded case where $m=n$, we will give results for the case $m > n$ as well.
-We will denote the maximum load by $\load$ and the additive gap by $\delta$:
+We denote the maximum load by $\load$ and the additive gap by $\delta$:
 \begin{align*}
@@
 \subsection{Multiple-choice BiB}
-Consider a variation of the  single-choice BiB strategy: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin.
+Consider the following variation of the  single-choice BiB strategy: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin.
 If there are multiple bins with the same minimal load, choose one among them arbitrarily.
@@
 \end{align*}
-w.h.p., for all $m$~\cite{ABKU99, BCSV06}, i.e., this additional choice leads to an exponential reduction of the additional gap. 
+w.h.p., for all $m$~\cite{ABKU99, BCSV06}, i.e., this additional choice leads to an exponential reduction of the additive gap. 
 \begin{comment}
@@
 \end{align*}
-w.h.p., for all $m$.
-(The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers.
-Refer to Appendix \ref{sec:fibonacci} for the definition.) Berenbrink et al.~\cite{BCSV06} and V\"ocking~\cite{VOC03} were able to show that the Always-go-left strategy is optimal up to additive constants.
+w.h.p., for all $m$. (The symbol $\Phi_d$ denotes a generalization of the golden ratio, based on generalized Fibonacci numbers. Refer to Appendix \ref{sec:fibonacci} for the definition.)
 
@@
 Section \ref{sec:model} contains a detailed definition of the model we use, which can be classified as a sequential, finite and dynamic process with random deletions.
-In Section \ref{sec:multiple-choice} we provide a brief definition of three different classes of random placement strategies and present two placement algorithms outlined in the introduction: Greedy (Section \ref{sec:greedy}) and Always-go-left (Section \ref{sec:AlwaysGoLeft}).
+In Section \ref{sec:multiple-choice} we provide a brief definition of three different classes of random placement strategies and present the two placement algorithms outlined in the introduction: Greedy (Section \ref{sec:greedy}) and Always-go-left (Section \ref{sec:AlwaysGoLeft}).
 Section \ref{sec:analysis} contains a detailed analysis of both strategies for the case $m=n$.
@@
 Conversely, in a \emph{parallel} process, balls arrive in sets and the balls in one set are dealt with simultaneously~\cite{ABS98}.
-\item In a \emph{static} problem, the balls are placed into the bins and no balls are ever removed, whereas in a \emph{dynamic} problem balls may also be removed from the system.
+\item In a \emph{static} problem, the balls are placed into the bins and no balls are ever removed, whereas in a \emph{dynamic} problem balls may also be deleted from the system.
 In the random deletions model, for each newly inserted ball another one is chosen uniformly and independently at random and removed from the system~\cite{ABKU99, MRS01}.
@@
 They are based on a sequential, infinite and dynamic process that uses an oblivious adversary, which provides a sequence $\sigma = \sigma_1 \sigma_2 \dots$, where $\sigma_t = (d_t, a_t)$ represents a request for a ball $d_t$ to be deleted and a different ball $a_t$ to be added to the bins at time step $t$.
-In our model, the ball $d_t$ associated with some request $\sigma_t$ is deleted at the beginning of time step $t$, next, the ball $a_t$ has to be placed into one of the $n$ bins, without any knowledge of the following requests $\sigma_{t+1}$, $\sigma_{t+2},\dots$ This infinite sequence of insertions and deletions is fixed in advance and is thus independent of previously allocated balls and any decisions made by the allocation algorithm.
-We assume that at any time $t$ at most $m = n$ balls are in the bins.
+In our model, the ball $d_t$ associated with some request $\sigma_t$ is deleted at the beginning of time step $t$ and the ball $a_t$ has to be placed into one of the $n$ bins, without any knowledge of the following requests $\sigma_{t+1}$, $\sigma_{t+2},\dots$ This infinite sequence of insertions and deletions is fixed in advance and is thus independent of previously allocated balls and any decisions made by the allocation algorithm.
+We assume that, at any time $t$, at most $m = n$ balls are in the bins.
 
@@
 A multiple-choice BiB strategy generalizes this idea by  choosing $d \geq 2$ bins and placing the ball into the least full.
-We distinguish placement algorithms depending on how the $d$ locations (i.e, bins) are sampled~\cite{VOC03}.
+We distinguish placement algorithms depending on how the $d$ locations (i.e., bins) are sampled~\cite{VOC03}.
 
@@
 Consider the Greedy scheme: For each ball, choose $d \geq 2$ bins independently and uniformly at random and place the ball into the least full bin.
-If there are multiple bins with the same minimal load, chose one among them arbitrarily.
+If there are multiple bins with the same minimal load, choose one among them arbitrarily.
 Note that the $d$ selected bins are not necessarily distinct.
@@
 \begin{compactitem}
-\item The apparently small change made for the Greedy algorithm compared to the single-choice BiB scheme results in an exponential decrease of the gap, even for $d=2$. 
+\item The apparently small change made for the Greedy algorithm compared to the single-choice BiB scheme results in an exponential decrease of the additive gap, even for $d=2$. 
 \item Each additional choice ($d > 2$) decreases the gap by only a constant factor~\cite{MRS01}.
@@
 w.h.p.,~for all $t$~\cite{VOC03}.
-We not that the combination of asymmetric tie-breaking and and non-uniform (but independent) sampling of the $d$ bins is crucial to achieve bounds given in Theorems~\ref{theorem:aglm} and~\ref{theorem:aglmLower}.
+Berenbrink et al.~\cite{BCSV06} were able to show that the Always-go-left strategy is optimal for the heavily loaded case as well.
+We note that the combination of asymmetric tie-breaking and and non-uniform (but independent) sampling of the $d$ bins is crucial to achieve the bounds given in Theorems~\ref{theorem:aglm} and~\ref{theorem:aglmLower}.
 
@@
 We will prove the bounds for $m=n$ for both the Greedy and the Always-go-left scheme in Sections~\ref{sec:analysis:greedy} and~\ref{sec:analysis:alg}, respectively.
-However, we will adopt a simplifying assumption to provide a more clear presentation of the proof technique.
+However, we will adopt a simplifying assumption to provide a clearer presentation of the proof technique.
 Section~\ref{sec:analysis:nondistinctBalls} will extend these proofs to remove this assumption.
@@
 We will use a proof technique called witness trees, which allows for a clean analysis of both problems and is applicable to a wide variety of BiB scenarios.
-Other proof techniques such as layered induction and a differential equation approaches have been used in the literature; see, e.g.,~\cite{MRS01} for an overview of different proof techniques for the BiB problem.
+Other proof techniques such as layered induction and a differential equations approach have been used in the literature; see, e.g.,~\cite{MRS01} for an overview of different proof techniques for the BiB problem.
 The proof for arbitrarily large $m$ is more involved and we refer the interested reader to~\cite{BCSV06, TW13}.
@@
 \label{sec:analysis:definitionSymWT}
-A symmetric witness tree (SWT) of order $L$ is a complete $d$-ary tree with $d^{L}$ leaves, where each node $v$ references a ball $b_v = \mathrm{ball}(v)$ and the nodes and edges of the SWT represent events that may occur or not.
+A symmetric witness tree (SWT) of order $L$ is a complete $d$-ary tree with $l=d^{L}$ leaves, where each node $v$ references a ball $b_v = \mathrm{ball}(v)$ and the nodes and edges of the SWT represent events that may occur or not.
 In general, the same ball may be referenced by several nodes.
@@
 \end{description}
-To avoid dependencies, the crucial step is to define edge events in terms of the possible locations of balls, not by their finally assigned bin.
+To avoid dependencies, the crucial step is to define edge events in terms of the possible locations of the balls, not by the bin that a ball is finally assigned to.
 
@@
 \label{sec:analysis:probabilitySymWT}
-In this section, we derive the probability that some SWT of order $L$ is activated, which gives an upper bound for the probability that the maximum load is greater than $L+3$ at some time $t$.
+In this section, we derive an upper bound for the probability that some SWT of order $L$ is activated.
 
@@
 
-The probability that a leaf event is activated is bounded from above by $3^{-d}$: Recall that by the definition of leaf events, each of the $d$ locations of a ball $b_v$, where $v$ is a leaf node, has to point to a bin with at least 3 additional balls in it.
+The probability that a leaf event is activated is bounded from above by $3^{-d}$: By the definition of leaf events, each of the $d$ locations of a ball $b_v$, where $v$ is a leaf node, has to point to a bin with at least 3 additional balls in it.
 The number of those bins is bounded by $\lfloor n/3 \rfloor$ at all time steps $t$, implying a probability of at most $1/3$ that one such bin is chosen.
@@
 \end{align*}
-Using $k \leq 2l$ we get
+Using $k \leq 2l$, we get
 \begin{align*}
@@
 \end{align*}
-which we simplify by using $2d^2 \leq 3^d$ as
+which we simplify by using $2d^2 \leq 3^d$ to
 \begin{align*}
@@
 \end{align*}
-thus proving Theorem~\ref{theorem:greedy}, under Assumption~\ref{assumption:independence} (i.e., that all events are stochastically independent). \qed
+thus proving Theorem~\ref{theorem:greedy}, under Assumption~\ref{assumption:independence}. \qed
 
@@
 \label{sec:analysis:definitionAsymWT}
-An AWT differes from an SWT (Section~\ref{sec:analysis:definitionSymWT}) only by its structure, which is a $d$-step Fibonacci tree $T_d\left(k \right)$:
+An AWT differes from an SWT only by its structure, which is a $d$-step Fibonacci tree $T_d\left(k \right)$:
 
@@
 Thus, at the time of insertion of $b_t$, each of the $d$ locations of this ball points to a bin containing at least $L+3$ balls.
-These are the $d$ bins that were sampled by the Always-go-left scheme, i.e., bin $i$ belongs to the $i$'th group of bins, where each group consists of $n/d$ bins.
-The child node of the root with label $(L-1, i)$ is assigned the topmost ball of the $i$'th location.
+These are the $d$ bins that were sampled by the Always-go-left scheme, i.e., bin $i$ belongs to the $i$-th group of bins, where each group consists of $n/d$ bins.
+The child node of the root with label $(L-1, i)$ is assigned the topmost ball of the $i$-th location.
 The assignment of balls to nodes then proceeds recursively: Given a node $v$ labelled $(h, i)$ that references a ball $b_v$ (which was put into some bin of the $i$-th group that, at the time of insertion of ball $b_v$, contained at least $h+3$ other balls), we can conclude the following:
@@
 Since the probability for the activation of an AWT of order $L$ upper-bounds the probability of the bad event that some bin contains more than $L+3$ balls, Eq.~\ref{eqn:awt:result} is also an upper bound for the latter.
-Note that Eq.~\ref{eqn:awt:chooseL} can be written as
+Note that Eq.~\ref{eqn:awt:chooseL} can be rewritten as
 \begin{align*}
@@
 \end{align*}
-thus proving Theorem~\ref{theorem:aglm}, under Assumption~\ref{assumption:independence} (i.e., that all events are stochastically independent). \qed
+thus proving Theorem~\ref{theorem:aglm}, under Assumption~\ref{assumption:independence}. \qed
 
@@
 Note that pruning reduces the size of the witness tree and therefore increases the probability of its activation.
-To compensate for this loss in size, we start with a larger witness tree, which we call a full witness tree (FWT) and show that, speaking informally, the pruning of many events is unlikely.
+To compensate for this loss in size, we start with a larger witness tree, which we call a full witness tree (FWT), and show that the pruning of many events is unlikely.
 Note that the technique presented in this section is applicable to both SWTs and AWTs.
@@
 \item For the assignment of nodes to the grandchildren of the root, consider a ball $b_c$ that is assigned to a child node $v_c$ and let the child node of $v_c$ be denoted by $v_g$.
-Since the ball $b_c$ was placed into bin $x$, at least one of its locations points to this bin $x$, which we call location $i$.
-Let $j = i+1 mod d$ be the ball's $j$-th location.
+Since the ball $b_c$ was placed into bin $x$, at least one of its locations points to bin $x$, which we call location $i$.
+Let $j = i+1\mod d$ be the ball's $j$-th location.
 Then, at the time of insertion of ball $b_c$, its $j$-th location points to a bin $y$ with at least $L+3$ other balls in it.
@@
 \end{compactitem}
-Note that the $\gamma$ additional balls in bin $x$ are assigned to the child nodes of the root only.
+Note that the $\gamma$ additional balls in bin $x$ are assigned to the child nodes of the root only, i.e., the balls referenced by the $\gamma$ child nodes of the root node are pairwise distinct.
 
@@
 We call $e$ a cut-off edge and use it as a witness that one of the $d$ locations of the ball $b_u$ is the same as one of the locations of some other ball $b_{u'}$, represented by a node $u'$, which was already encountered during the traversal.
-Furthermore, we remove the whole tree rooted at node $v$ from our witness tree and we continue this pruning until either the breath-first traversal terminates or we have obtained $\gamma$ cut-off edges.
-Note that more than $\gamma$ cut-off edges cannot exist, because \todo{why??}.
+Furthermore, we remove the whole tree rooted at node $v$ from our witness tree and continue this pruning until either the breath-first traversal terminates or we have obtained $\gamma$ cut-off edges.
+Note that no child node of the root will be pruned, since the balls referenced by them are pairwise distinct and visitied first when traversing the tree in breadth-first order.
 When finished, the pruned witness tree (PWT) does not reference any ball twice and thus all its events are independent.
@@
 \begin{lemma}\label{lemma:fwt:activation}
-For some suitable $\gamma \geq 2$ and large enough $L$, the activation of a PWT occurs with probability at most $n^{-\gamma +1 +o(1)}$.
+For suitable $\gamma \geq 2$ and large enough $L$, the activation of a PWT occurs with probability at most $n^{-\gamma +1 +o(1)}$.
 \end{lemma}
-If there are less than $\gamma$ cut-off edges, in at least one of the $\gamma$ SWTs or AWTs contained in the FWT, the referenced balls are distinct.
+If there are less than $\gamma$ cut-off edges, the referenced balls are distinct in at least one of the $\gamma$ SWTs or AWTs contained in the FWT.
 Let this SWT or AWT be denoted by $T$.
@@
 Then the edge $e_c$ represents the event that one of the locations of the ball represented by node $v_c$ points to this bin.
-This implies that the probability of activation for an edge $e_c$ is $p_\mathrm{e}' = 1/n \leq d/n$.
+This implies that the probability of activation for an edge event $e_c$ is $p_\mathrm{e}' = 1/n \leq d/n$.
 
@@
 Then, by the same arguments as in Sections~\ref{sec:analysis:probabilitySymWT} and~\ref{sec:analysis:probabilityAsymWT}, we can conclude that there are at most $w\leq n^k$ assignments of balls to the nodes of the witness tree.
-Furthermore, the probability for the activation of all edge events is bounded from above by
+Furthermore, the probability for the activation of all edge events is bounded from above as
 \begin{align*}
@@
 		\leq K\gamma n^{k+1} \left(\frac{d}{n}\right)^k3^{-d l}  \left(K \frac{d}{n}\right)^\gamma
-		= n  d^k 3^{-d l}  \left(K^2 \frac{d}{n} \right)^\gamma \quad ,
+		= n  d^k 3^{-d l}  \left(K^2 \frac{d}{n} \right)^\gamma 
 \end{align*}
@@
 \begin{align*}
-\delta =  \Theta\left(\sqrt{\frac{m  \log n}{n}}\right) \quad ,
+\delta &= 
+    \begin{cases}
+    \Theta\left(\frac{\ln n}{\ln\ln n}\right)           & m = n \\
+    \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
+    \end{cases} \quad ,    
 \end{align*}
 w.h.p.
-In this paper we have presented two different multiple-choice BiB placement algorithms: The Greedy scheme due to Azar et al.\cite{ABKU99}, which achieves a gap of 
+In this survey we have presented two different multiple-choice BiB placement algorithms: The Greedy scheme due to Azar et al.\cite{ABKU99}, which achieves a gap of 
 \begin{align*}
@@
 \end{align*}
-w.h.p., and the Always-go-left scheme due to V\"ocking~\cite{VOC03}, which achieves a gap of
+w.h.p., for all $m$. This implies an exponential reduction of the additive gap even for $d=2$ and illustrates the two-choice paradigm.
+The Always-go-left scheme due to V\"ocking~\cite{VOC03}, which by a combination of non-uniform bin sampling and an asymmetric tie-breacking mechanism achieves a gap of
 \begin{align*}
@@
 \end{align*}
-w.h.p., for all $m$. We used witness trees, a very useful technique in the BiB context, to prove these upper-bounds for the case $m=n$ in detail.
+w.h.p., for all $m$, demonstrates the surprising influence of asymmetry. We used witness trees, a very useful technique in the BiB context, to prove these upper-bounds for the case $m=n$ in detail.
 
